### 2021-01-05 manuscriptPGPE.tex

%% ****** Start of file aiptemplate.tex ****** %
%%
%%   This file is part of the files in the distribution of AIP substyles for REVTeX4.
%%   Version 4.1 of 9 October 2009.
%%
%
% This is a template for producing documents for use with 
% the REVTEX 4.1 document class and the AIP substyles.
% 
% Copy this file to another name and then work on that file.
% That way, you always have this original template file to use.

%\documentclass[aip,graphicx]{revtex4-1}
%\documentclass[aip,reprint]{revtex4-1}

%\usepackage{graphicx}

%\draft % marks overfull lines with a black rule on the right
%\documentclass[pre,aps,floatfix,authordate1-4,twocolumn]{revtex4-1}
%\documentclass[pre,aps,floatfix,authordate1-4]{revtex4-1}

\documentclass[aps,prl,superscriptaddress,twocolumn]{revtex4}

%\documentclass[aps,prl,preprint,groupedaddress]{revtex4}

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{csquotes}
\usepackage{xr}
\externaldocument{manuscriptPGPEsuppl}

\begin{document}

% Use the \preprint command to place your local institutional report number 
% on the title page in preprint mode.
% Multiple \preprint commands are allowed.
%\preprint{}

\title{NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids} %Title of paper

% repeat the \author .. \affiliation  etc. as needed
% \email, \thanks, \homepage, \altaffiliation all apply to the current author.
% Explanatory text should go in the []'s, 
% actual e-mail address or url should go in the {}'s for \email and \homepage.
% Please use the appropriate macro for the type of information

% \affiliation command applies to all authors since the last \affiliation command. 
% The \affiliation command should follow the other information.

\author{Am{\'e}lie Bacle}
\affiliation{Paris, France}

\author{Pavel Buslaev}
\affiliation{University of Jyv{\"a}skyl{\"a}}

\author{Rebeca Garc{\'i}a Fandi{\~n}o}
\affiliation{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}
\affiliation{CIQUP, Centro de Investigação em Química, Departamento de Química e Bioquímica, Faculdade de Ciências, Universidade do Porto, Porto, Portugal}

\author{Fernando Favela-Rosales}
\affiliation{Departamento de Investigaci\'{o}n, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}

\author{Tiago Ferreira}
\affiliation{Halle, Germany}

\author{Patrick Fuchs}
\affiliation{Paris, France}

\author{Matti Javanainen}
\affiliation{Institute of Organic Chemistry and Biochemistry of the 
Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}

\author{Jesper J. Madsen}
\affiliation{Department of Chemistry, The University of Chicago, Chicago, Illinois, United States of America}
\affiliation{Department of Global Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}

\author{Josef Melcr}
\affiliation{Institute of Organic Chemistry and Biochemistry of the 
Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}
\affiliation{Groningen Biomolecular Sciences and Biotechnology Institute 
and The Zernike Institute for Advanced Materials, 
University of Groningen, 9747 AG Groningen, The Netherlands}

\author{Paula Milan Rodriguez}
\affiliation{Paris, France}

\author{Markus S. Miettinen}
% \affiliation[Max Planck Institute of Colloids and Interfaces]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
\affiliation{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}

\author{O. H. Samuli Ollila}
\email[]{samuli.ollila@helsinki.fi}
\affiliation{Institute of Biotechnology, University of Helsinki}

\author{Chris G. Papadopoulos}
\affiliation{I2BC - University Paris Sud}

\author{Antonio Pe{\'o}n}
\affiliation{Spain}

\author{Thomas J. Piggot}
\affiliation{Chemistry, University of Southampton, Highfield, Southampton SO17 1BJ, United Kingdom}

\author{{\'A}ngel Pi{\~n}eiro}
\affiliation{Departamento de F{\'i}sica Aplicada, Facultade de F{\'i}sica, Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}

% Collaboration name, if desired (requires use of superscriptaddress option in \documentclass). 
% \noaffiliation is required (may also be used with the \author command).
%\collaboration{}
%\noaffiliation

\date{\today}

\begin{abstract}
% insert abstract here
  Chemistry of lipid headgroups, the water facing components of cell membranes, varies between organisms and
  organelles and regulate cell functions via lipid-protein interactions.
  Because lipid membranes are in liquid state in physiological conditions,
  the conformational ensembles of lipids and their dependence of headgroup chemistry are difficult to resolve experimentally.
  Here, we combine solid state NMR experiments and MD simulations from NMRlipids open collaboration to
  resolve conformational ensembles of PC, PS, PG and PE lipid headgroups in liquid lamellar phase in various biologically relevant conditions.
  Interpretation of NMR experiments using the best simulations from the NMRlipids project suggests that all lipid headgroups sample wide range of
  conformations in neutral and charged cell membranes, and that the differences in headgroup chemistry
  appears in the probability weights of different conformations. Together with the analysis of protein-bound-lipids
  from the protein data bank (PDB), this suggests that lipids can bind to proteins in wide range of conformations
  independently on the headgroup chemistry. Therefore, the specific lipid-protein interactions are likely driven by
  protein-lipid interactions, rather than conformational restrictions of lipids. Our results paves the way to
  comprehensive understanding of lipid-protein interaction energetics, and understanding complex biomolecular assemblies
  such as membrane proteins.

  %The force field giving the best description for glycerol backbone and headgroup structures
  %of PC, PS, PG and PE headgroups (CHARMM36) reproduces the essential differences in order parameters
  %between these headgroups, and therefore enables the analysis of structural differences between the
  %headgroups.

\end{abstract}

%\pacs{}% insert suggested PACS numbers in braces on next line

\maketitle %\maketitle must follow title, authors, abstract and \pacs

% Body of paper goes here. Use proper sectioning commands. 
% References should be done using the \cite, \ref, and \label commands

%\label{}
\section{Introduction}

Chemical compositions of hydrophilic lipid headgroups vary between different
organelles and organisms, and different lipid types
regulate protein functions in many different ways \cite{lee03,vanmeer08}.
Lipids can directly bind to proteins or indirectly affect protein
functions by altering membrane properties or charge \cite{lee03,lemmon08}.
While the specific interactions with certain lipid headgroups
are known to be essential for the function of several proteins \cite{lee11,lemmon08},
it is not clear if the specificity is driven by the differences in accessible
conformational states between lipid types or
by specific intermolecular lipid-protein interactions.

Lipid conformational ensembles in physiologically relevant lamellar liquid phase
are typically derived from NMR experiments, particularly from 
%Attempts to resolve the conformational ensembles of lipid headgroup rely mainly on
C-H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}
which can be detected even from cells \cite{gally81,scherer87,seelig90}.
%but lipids are not in solid state in biomembranes. 
%On the other hand, NMR experiments enable detection of lipid headgroups
%in physiological liquid states also in cells \cite{gally81,scherer87,seelig90}.
These studies suggest that the glycerol backbone structures are largely similar irrespectively of the headgroup \cite{gally81}, 
and the headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
%Extensive discussion about structural details of PE, PG or PS headgroups do not exists (as far as I know), 
%In contrast to PC lipids (see \cite{botan15} and references therein).
%Some attempts to resolve conformational ensembles from NMR for PC and PE lipids have been made,
%but lesser extend for PG or PS lipids \cite{seelig77c,davis83,Semchyschyn04}.
However, these results are based only on the absolute values while 
%we have recently demonstrated that also
the necessity of order parameters signs in capturing the lipid conformational ensembles has been resently demonstrated \cite{botan15,ollila16,ferreira16}.
Furthermore, the detailed understanding of lipid conformational ensembles
is limited by the lack of universal models to interpret the experiments \cite{pezeshkian18,akutsu20}.
%do not exist and
%the proposed structures are often based on lipid crystal structures \cite{pezeshkian18,akutsu20}.
%which were solved many decades ago \cite{buldt81,pascher92}.
%determine the structural 
%Classical molecular dynamics simulations could potentially give such ensembles and therefore enable
%the detailed studies of lipid headgroup behaviour in complex biomolecular systems, but current
%force fields are not accurate enough to reproduce the correct conformational ensembles for PC and PS headgroups \cite{botan15,antila19}.

Stuctures of different lipid types in protein bound states can be extracted from the protein data bank (PDB) \cite{berman00},
but their relation to the conformational ensembles in liquid lamellar state remains unclear \cite{marsh13b}.
In addition to the changes in lipid conformational ensembles upon binding to proteins,
%PC headgroup conformations do not change upon addition of zwitterionic lipids or cholesterol \cite{scherer87,ferreira13} into a membrane,
%but are sensitive to the presence of charged molecules in membranes \cite{seelig87}.
also the experimentally measured response of lipid headgroup to membrane bound charges remains poorly understood 
%The response to charges is explained by the change in the orientation of lipid headgroup dipole vector due to electrostatic interactions,
%but the detailed changes are not yet understood
due to the lack of suitable models to interpret the lipid conformational ensembles in liquid lamellar state \cite{Semchyschyn04}.

Here, we use natural abundance $^{13}$C NMR experiments and MD simulations from the NMRlipids open collaboration
to resolve the differences in conformational ensembles between PC, PE, PG and PS lipid headgroups.
%open collaboration and order parameters of glycerol backbone and headgroup
%to evaluate the accuracy of PE and PG heagroup structures, and the cation binding affinity
Zwitterionic PC and PE are the most common lipids in eukaryotes and bacteria, respectively \cite{vanmeer08,sohlenkamp16}.
PE is also the second most abundant glycerophospholipid in eukaryotic cells
and has been related to various diseases \cite{vance15,calzada16,patel17}.
PS and PG are the most common negatively charged lipids in eukaryotes and bacteria, respectively,
and affect membrane protein functionality and signaling \cite{lemmon08,leventis10,sohlenkamp16,hariharan18}.
%PE and PG lipids are most common lipids in bacteria \cite{sohlenkamp16}.
%Anionic PG lipids are less abundant, but is also proposed to be fundamental for terrestrial life \cite{furse17}.
%PE and PG affect membrane protein functionality \cite{hariharan18} and
All the studied lipids specifically bind to various proteins \cite{yeagle14}.
%PE headgroup is also prone for negative membrane curvature and causes membrane fusion \cite{Chernomordik08,calzada16}.
%Therefore, the PE and PG headgroup structures play probably essential roles in 
%many biological processes.
Furthermore, we use our results to elucidate lipid-protein interactions and the effect of charges on lipid conformations.

The lipid conformational ensembles in liquid lamellar state paves the way toward understanding the
specific binding of different lipid types to membrane proteins and how they regulate the protein function.
Because glycerol backbone and headgroup structures of PC lipids are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90},
the results from model systems could be used to understand the biological role of lipid headgroup conformational ensembles in different lipid types.

%Several MD simulations of PE and PG lipids have been published especially in the context of modeling
%inner membrane of Gram-negative
%bacteria \cite{devries04,murzyn05,pedersen06,zhao07,gurtovenko08,zhao08,henin09,kukol09,tsai12,dickson12,venable13,dickson14,berglund15}
%\todo{There may be some relevant publication missing from here},
%but evaluation of glycerol backbone and headgroup structures against experiments is rare \cite{henin09}.

%Derivation of different lipid headgroup conformational ensembles in liquid state has been
%inconclusive due to lack of suitable experimental data and tools to interpret the conformational ensembles.

%Besides the structure, also ion binding may regulate biophysical activity of especially negatively
%charged lipid headgroups \cite{seelig90}. Monovalent cation (except Lithium) binding to zwitterionic PC and anionic PS headgroups is
%very weak, while multivalent ion binding is stronger but still weak \cite{cevc90,tocanne90,roux90,catte16,antila19}. The ion binding affinity
%data for PE is more scarce \cite{marra85}, but large differences to PC would be surprising.
%Negatively charged lipids are suggested to bear same cation binding constants than zwitterionic lipids, but
%the amount of bound ions to negatively charged membranes would still be larger because
%the concentration of cations in the vicinity of membranes would be higher  \cite{seelig90}.
%
%is stronger
%Based on the electrometer concept and other data is has been suggested that \cite{seelig90}
%\begin{displayquote}
%  {\it ''(i) Ca$^{2+}$ binds to neutral lipids (phosphatidylcholine, phosphatidylethanolamine) and negatively charged lipids
%    (phosphatidylglycerol) with approximately the same binding constant of K = 10-20 M$^{-1}$; \\
%    (ii) the free Ca$^{2+}$
%    concentration at the membrane interface is distinctly enhanced if the membrane carries a negative surface
%    charge, either due to protein or to lipid; \\
%    (iii) increased inter-facial Ca$^{2+}$ also means increased amounts
%    of bound Ca$^{2+}$ at neutral and charged lipids; \\
%    (iv) the actual binding step can be described by a Langmuir
%    adsorption isotherm with a 1 lipid:1 Ca$^{2+}$ stoichiometry, provided the interfacial concentration C$_M$, is
%    used to describe the chemical binding equilibrium.''}
%\end{displayquote}
%
%On the other hand, anionic PS lipids are proposed chelate with calcium ions \cite{hauser85,feigenson86,roux91}.
%In simulations, the cation binding affinity to PC and PS membranes is typically overestimated \cite{catte16,antila19},
%which can be improved by applying the ECC to the partial charges of the force fields \cite{melcr18,melcr19}.

%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
%  \caption{\label{lipids}
%    Chemical structures and labels for the headgroup carbons.
%  }
%\end{figure}

\section{Methods}
\subsection{Experimental C--H bond order parameters}
The headgroup and glycerol backbone C--H bond order parameters of POPE and POPG
were measured using natural abundance $^{13}$C solid state NMR spectroscopy
as described previously \cite{ferreira13,ferreira16}.
The magnitudes of order parameters were determined from the chemical-shift resolved dipolar splittings
using a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04} and
the signs from S-DROSS experiments~\cite{gross97} combined with SIMPSON simulations \cite{bak00}.
%The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
%Magic angle spinning (MAS) of the sample was used at a frequency of 5.15 kHz (R-PDLF experiment) and 5 kHz (S-DROSS experiment).
%The following experimental setups were used.
The NMR experiments were identical as in our previous work~\cite{antila19}.
POPE and POPG powder were purchased from Avanti polar lipids \todo{Which isomer of POPG we have in the experiments?}.
%\todo{Is this enough and correct, or should we repeat some methods from the NMRlipidsIVps paper?}
The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.

%Absolute values of the headgroup and glycerol backbone order parameters from PE and PG lipids are measured
%previously using $^2$H NMR~\cite{seelig76,gally81,wohlgemuth80,borle85}. Because also the order parameter
%signs bear essential information about the lipid structures \cite{botan15,ollila16}, we measured the
%magnitudes and signs of POPE and POPG C--H bond headgroup and glycerol backbone order parameter in liquid phase
%using the 2D-RPDLF and S-DROSS experiments, as described previously \cite{ferreira13,ferreira16,antila19}.

Glycerol backbone peaks from both lipids, and $\alpha$-carbon peak from POPE in the INEPT spectra 
were assigned based on previously measured POPC spectra~\cite{ferreira13}.
%For POPE, the glycerol backbone and $\alpha$-carbon peaks in INEPT spectra were assigned based on
%previously measured POPC spectra~\cite{ferreira13} and
The $\beta$-carbon peak from POPE was assigned based on $^{13}$C chemical shift table for amines available
at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm}.
%For POPG, the glycerol backbone peaks in INEPT spectra were assigned based on
%previously measured POPC spectra~\cite{ferreira13},
\todo{How were $\alpha$ and  $\gamma$-carbon peaks assigned in POPG?}.
The $\beta$-carbon peak from POPG overlapped with the g$_2$ peak from glycerol backbone
because their chemical environments are similar in POPG.
\todo{Details to be checked by Tiago}.

\subsection{Molecular dynamics simulations}

Molecular dynamics simulation data were collected using
the NMRlipids Open Collaboration \cite{botan15}, which is running at
\url{nmrlipids.blogspot.fi} and \url{github.com/NMRlipids/NMRlipidsIVotherHGs}.
The simulated systems of pure PE and PG bilayers without additional ions
are listed in Tables \ref{systemsPE} and \ref{systemsPG},
and lipid mixtures with additional ions in Table~\ref{systemsMIX}.
Further simulation details are given in the SI, and
the simulation data are indexed in a
searchable database available at \url{www.nmrlipids.fi},
and in the NMRlipids/MATCH repository (\url{github.com/NMRlipids/MATCH}).

\subsection{Analysis of molecular dynamics simulation data}
The big data set of MD simulations was analysed in the NMRlipids databank manner.
Unique naming convention for lipid atoms in each force field was defined using the mapping files
and analysis for all simulations indexed in NMRlipids databank manner were performed
using python codes.

The C--H bond order parameters were calculated directly
from the carbon and hydrogen positions using the definition
\begin{equation}
S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
\end{equation}
where $\theta$ is the angle between the C--H bond and the membrane normal
(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
Angular brackets denote average over all sampled configurations.
The order parameters were first calculated averaging over time separately
for each lipid in the system. The average and
the standard error of the mean were then calculated over different lipids.
Python programs that use the MDAnalysis library \cite{agrawal11,gowers16}
used for all atom simulations is available in Ref. \citenum{MATCHgit}
({\tt scripts/calcOrderParameters.py}). For united atom simulations, the trajectories
with hydrogens having ideal geometry were constructed first using either buildH program~\cite{buildH}
or ({\tt scratch/opAAUA\_prod.py}) in  Ref. \citenum{MATCHgit}, and the order parameters were
then calculated from these trajectories. This approach has been tested against trajectories
with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot17}.\\
\todo{BuildH program is now cited with a direct link to the GitHub repo. I think that a release to Zenodo would be nice in the final publication.}\\
\todo{Maybe we should also shortly discuss here about the reasons for slight dependence of order parameter values on the method used to reconstruct hydrogens?}\\
The ion number density profiles were calculated using the {\tt gmx density} tool
of the Gromacs sofware package \cite{gromacsMANUAL}.

%\clearpage
\begin{figure*}[]
  \centering
%  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
  %  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCPSPEPG.eps}
   \includegraphics[width=18.0cm]{./Figs/figure1.eps}
   \caption{\label{HGorderParameters}
     Chemical structure, refocused-INEPT spectrum, 2D R-PDLF spectra, and S-DROSS data (from top to bottom) of {\bf A)} POPE  and {\bf B)} POPG.
      Full NMR spectra are shown Figs. \ref{POPEspectra} and \ref{POPGspectra}.
     {\bf C)} Chemical structure of POPC and POPS.
    {\bf D)} Headgroup and glycerol backbone order parameters from different experiments in lamellar liquid disordered phase.
    The values and signs for POPE (310~K) and POPG (298~K)
    measured in this work, and for POPS (298~K) \cite{antila19} and POPC (300~K) \cite{ferreira13,ferreira16}
    previously using $^{13}$C NMR. The literature values for
    DOPS with 0.1M of NaCl (303~K) \cite{browning80},
    POPG with 10nM PIPES (298~K) \cite{borle85},
    DPPG with 10mM PIPES and 100mM NaCl (314~K) \cite{wohlgemuth80}, 
    DPPE (341~K) \cite{seelig76},
    E.coliPE and E.coliPG (310~K) \cite{gally81}
    are measured using $^2$H NMR. The signs from $^{13}$C NMR are used also for the literature values.
   }
   \todo{This is a sketch, Tiago Ferreira will make a new figure.} \\
%  \todo{D) could be clarified as Fig. 2 in the NMRlipids IVps paper.}
\end{figure*}

\subsection{Analysis of lipid conformations bound to proteins}
Lipid structures from PDB (\url{http://www.rcsb.org/})
were searched using PDBe REST API (\url{www.ebi.ac.uk/pdbe/pdbe-rest-api}).
First, all PDB entries containing PC, PE, PG, or PS lipid headgroups were searched.
The ligand names to identify the lipids were:
PLC, PX4, 6PL, LIO, HGX, PC7, PC8, P1O, 6O8, XP5, EGY, PLD, SBM, HXG, and PCW for PC;
8PE, PTY, 3PE, PEH, PEF, 6OE, 6O9, 9PE, PEV, 46E, SBJ, L9Q, PEK, EPH, ZPE, 9TL, 9Y0, 6OU, LOP, and PEE for PE;
PGT, PGK, LHG, 44G, PGV, OZ2, D3D, PGW, DR9, P6L, PG8, H3T, and GOT for PG; and
PSF, PS6, Q3G, P5S, D39, PS2, 17F, and 8SP for PS.
All PDB structures containing these ligands were downloaded and the dihedral angles of
the first lipid in the latest version of the structures within each PDB code were calculated  
using MDTraj python library \cite{mcgibbon15}.
The used Jupyter notebook is available from the project's GitHub repository ({\it scripts/pdbSEARCH.ipynb}).

\section{Results and Discussion}

%\subsection{Headgroup and glycerol backbone order parameters of POPE and POPG from $^{13}$C NMR}

\begin{figure*}[]
  \centering
   \includegraphics[width=18.0cm]{../Figs/figure2.eps}
%  \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures2.png}
%  \includegraphics[width=8.0cm]{../Figs/CHARMMfromLIPIDS.eps}
%%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
%%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
%  \includegraphics[width=8.0cm]{../Figs/PNangleNORM.eps}
%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALS.eps}
   \caption{\label{structures}
     Results from CHARMM36 simulations demonstrating the differences in conformational ensembles betwee different lipids. 
     {\bf A)} Snapshots with overlayed C$_\beta$, C$_\alpha$ and O$_\alpha$ atoms and occurence of different conformations.
     {\bf B)} Headgroup and glycerol backbone region order parameters of different different lipids.
     {\bf C)} Distributions of heavy atom dihedral angles of different lipids from CHARMM36 simulations.
     {\bf D)} Distributions of P-N vector angle with respect to membrane normal.
  }
  \todo{This is a draft and requires quite a bit of polishing.
    More detailed discussion of this figure is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9} \\
\end{figure*}

\subsection{Conformational ensembles of different lipid headgroups in bulk bilayer}

To experimentally characterize the headgroup conformational ensembles of lipids
that are not bound to proteins in electrostatically neutral cell membrane,
we measured the C-H bond order parameters and their signs from of POPG and POPE
in liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
Determination of headgroup and glycerol backbone order parameters and their signs
was straightforward from the
%R-PDLF and S-DROSS
data in Figs. \ref{HGorderParameters}, \ref{POPEspectra} and \ref{POPGspectra}
for all the C-H bonds, except the $\beta$ and g$_2$ carbons in POPG.
These carbons have overlapping peaks in the INEPT spectra due to their similar chemical environments,
and only the magnitude of the larger order parameter could be determined from the R-PDLF spectra (Fig. \ref{HGorderParameters} B).
Nevertheless, based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85},
we assigned the larger order parameter to the g$_2$ carbon
%order parameter of $\beta$-carbon is expected to be clearly smaller than for 
and used the literature value for the $\beta$-carbon in SIMPSON simulations to determine the signs.
The decrease in the beginning of the S-DROSS curve suggests that the sign of larger g$_2$ order parameter
is negative and later increase suggests that sign of smaller $\beta$ order parameter is positive (Fig. \ref{HGorderParameters} (B)).
This interpretation is confirmed by SIMPSON calculations in Fig \ref{POPGsimpson}.
%using negative value for g$_2$ and positive value for $\beta$ order parameter (Fig. \ref{POPGsimpson}).
%The signs of these order parameters were solved using SIMPSON simulations to interpre the
%S-DROSS experiments in similar fashion as we did previously for PS lipids \cite{antila19}.

%Also the headgroup order parameters of PC lipids are close PE, althought the latter gives systematically
%slightly more positive values (Fig. \ref{HGorderParameters}).
% These could be explained with slightly larger temperature in PE measurements, except for the $\alpha$-carbon with the positive sign, for which the
% more positive value is farther away from zero.
%While the headgroup order parameters are similar for PC and PE lipids,
%PG and PS lipids exhibit distinct values in comparison with other lipids.

Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and
headgroups from this and previous studies are collected in Fig. \ref{HGorderParameters} D),
where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
%The resulting order parameters are compared with our previously published results
%for POPC \cite{ferreira13,ferreira16} and POPS \cite{antila19} in figure \ref{HGorderParameters} D.
%Order parameters for POPE are in good agreement with previous $^2$H NMR experiments \cite{seelig76}
%and similar to our previous results for POPC \cite{ferreira16}.
Overall agreement of order parameters determined by different authors and different techniques
for same lipid headgroup is very good in here and previous studies \cite{botan15,ollila16,antila19},
suggesting that the observed differences between lipid types arise from differences in
headgroup chemistry rather than inaccuracies in experiments. 
The most distinct order parameters are observed for PS headgroups,
for which the $\alpha$-carbon order parameter exhibits significant forking
and the $\beta$-carbon has more negative value than in other lipids.
On the other hand, the $\beta$-carbon order parameter of PG headgroup
%the $\alpha$-carbon order parameter is similar to PE and PG, while
has positive sign in contrast to all the other lipids.
Notably, this has not been observed in traditional $^2$H NMR experiments,
%because absolute value of $\beta$-carbon order parameter is similar in PG, PE and PC lipids and
where only the absolute value of the order parameters are measured~\cite{wohlgemuth80,gally81,borle85}.
The glycerol backbone order parameters are similar for all the lipids, although they move slightly toward
positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG.
Essential differences between PC and PE headgroups are not observed.

%As discussed previously for PC and PS headgroups \cite{ollila16,antila19}, also 
%the headgroup and glycerol backbone order parameters of PE and PG are essentially independent of 
%acyl chain composition, and therefore manifest mainly headgroup chemistry (Fig. \ref{HGorderParameters}).

\begin{figure*}[bt]
  \centering
  \includegraphics[width=18.0cm]{../Figs/HGorderparametersPCvsSURFchangeDIHEDRALS.eps}
  \caption{\label{changesWITHsurf}
    {\bf A)} Modulation of PC headgroup order parameters and P-N vector angle upon addition of cationic surfactant
    from CHARMM36 simulations compared with experimental data \cite{??}.
    {\bf B)} Changes in PC headgroup conformational ensembles upon increasing amount of positive charge in bilayer,
    characterized by the heavy atom dihedral distributions, from CHARMM36 simulations.
  }
\end{figure*}

To resolve the conformational ensembles of lipid headgroups,
we compared the headgroup and glycerol backbone C-H bond order parameters
from different MD simulation force fields to the experimental data. 
None of the force fields reproduced the order parameters
within experimental accuracy for any of the lipids analysed in this work
(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}) or in our previous studies \cite{botan15,antila19}.
%has least problems in
%reproducing the experimental lipid headgroup and glycerol backbone order parameters in different
%lipids \cite{botan15,antila19} (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
Nevertheless, the CHARMM36 force field performs best for all the headgroups and
%Importantly, it captures the experimentally observed distinct order parameters for PS and PG headgroups (Fig. \ref{??}).
%To characterize the differences in lipid headgroup conformational ensembles
reproduces the distinct order parameters of PS and PG lipids (Fig. \ref{structures} B).
To understand the structural origin of these distinct order parameters,
we calculated the distributions of heavy atom dihedral angles from CHARMM36 simulations in figure \ref{structures} C).
%The dihedral distributions in figure \ref{structures} show significant similarity between different lipids for all the other
%bonds, except
Major differences between headgroups are observed only for the last two dihedrals
in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$,
which prefer {\it gauche$^-$} conformations for PG and {\it gauche$^+$} for PS.
Rest of the dihedrals are similar between different lipids, with the exception of
%However, some dihedral angles of phosphorus-oxygen bonds are less likely for PS, which
small differences for PS lipids. Therefore we suggest that the main differences between
lipid headgroups leading to distinct order parameter occur in the choline part, while
also changes in phosphate region may contribute in PS lipids, which could also
explain the more rigid headgroup structures~\cite{browning80,buldt81}.
%These differences explaing the disctinct order parameters of
%PS and PG headgroups with respenct to the other lipids.
%The differences between lipids are reflected also to
Furthermore, the angle between headgroup dipole and membrane normal
decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig. \ref{structures} B)).
However, the differences between PC and PE in P-O$_\alpha$-C$_\alpha$-C$_\beta$ dihedral
and P-N vector dipole may be an artefacts as the $\beta$-carbon order parameter in PC
is too negative the CHARMM36 force field, thereby not being equal to the order parameter
in PE as observed in experiments \cite{botan15}.

%The eclipsed (0 or 360 degrees) are not present in any other dihedrals, and anti-eclipsed (120 or 240 degrees)
%are not present in C$_\alpha$-C$_\beta$ and g$_1$-g$_2$ bonds.

%Characteristic dihedral conformations in PS headgroup are asymmetric conformations
%preferring gauche 270$^\circ$ conformations in N-C$_\beta$-C$_\alpha$-O$_\alpha$ and
%%dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
%%in , than for the PC headgroup.
%C$_\beta$-C$_\alpha$-O$_\alpha$-P dihedrals. 
%In PG headgroup, the O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other% lipids)
%is mostly in trans conformation, and C$_\beta$-C$_\alpha$-O$_\alpha$-P  has asymmetric tendency to be in
%gauche 60$^\circ$ conformation. 
%Main difference between PC and PE is the lower probability of trans state in C$_\beta$-C$_\alpha$-O$_\alpha$-P PC dihedral,
%which could be a potential reason for the too negative $\beta$ headgroup order parameter in PC.%

In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase.
%the results suggest that free rotations around O-P bonds decouple the
%headgroup and glycerol backbone structure and dynamics in similarly in all lipids,
%althought this rotation is slower in PS lipids where crossing certain barriers are slower.
Despite the differences in distributions
%of two the last bonds at the headgroup end,
%leading to the distinct order parameters for PS and PG lipids,
%lipids with
between headgroups, the sampled dihedral angles are within approximately same ranges for all headgroup types.
Therefore, we suggest that all lipid headgroups are very flexible thereby being
able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
%Furthermore, approximately same conformations are present liquid lamellar phase for different
%headgroups studied here, althought the probability weight for the conformations are slightly different.
%the order parameter experiments suggest that the glycerol backbone conformations in all lipids
%and the headgroup conformations in PC and PE lipids are relatively similar, while 
%PS and PG headgropus exhibit distinct conformational sampling.
%The details of sampled conformation are difficult to deduce from order parameters only,
%but the distinct headgroup order parameters
%of PS lipids are previously related to the more rigid structure of the headgroup~\cite{browning80,buldt81,antila19}.
%The analysis of dihedral angle distributions reveal the free rotation of dihedral around
%phosphorus-oxygen bonds in all lipids as all possible angles are observed in the distributions.
Our results support the models that propose free rotation around phosphate group for PC lipids,
which decouples the dynamics and conformatios between acyl chains and headgroups \cite{klauda08c,antila21b},
and suggests that these models could be applied for all lipids containing the phosphorus group.
The wide range of observed conformations suggest that the structures in lipid crystals \cite{buldt81,pascher92}
play only a minor role, and that the models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20}
are not sufficient to capture the large conformational space of lipids in liquid lamellar state.

\subsection{Lipid conformational ensembles in lipid bilayers with bound ions}

Charged lipids, proteins, surfactants, drugs, and ions incorporated in membranes
reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of
lipid headgroups \cite{seelig87}. However, detailed understanding on
lipid conformational ensembles in membranes bearing electric charge is still lacking \cite{Semchyschyn04}.
%Order parameters of $\alpha$ and $\beta$ carbons in PC lipid headgroup are known to depend
%linearly on the incorporation of charged molecules into a lipid bilayer,
%which is explained by tilting of the headgroup dipole orientation \cite{??}.
%%are kn affect the headgroup dipole vector tilt which can
%%be detected by measuring C-H bond order parameters of $\alpha$ and $\beta$ carbons.
%Such changes have been observed upon addition of charged lipids, proteins, surfactants,
%drugs, or ions \cite{??}, but how charges affect the lipid conformational ensemble remains unknown.

\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
  \caption{\label{changesWITHCaClPG}
    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
    The changes with respect to the systems without CaCl$_2$ are shown for other data than
    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
    Calsium density distributions are shown in figure \ref{CAdensPG}.
    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
  }
\end{figure*}

To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge,
%suggesting that these simulations can be used to interpret the charge induced changes in lipid conformational ensembles.
we first analyzed CHARMM36 simulations which correctly captures the experimentally measured decrease in PC headgroup order parameters
upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
Heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B
reveal that the decrease of trans state probabilities in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
dihedrals are the major effects of positive charge on lipid headgroup conformations.
Choline region remains essentially unchanged and only minor changes are observed in other dihedrals
even thought almost half of the molecules in membranes are cationic surfactants.
%The major changes upon addition of charge are observed in dihedrals related to the phosphate oxygens,
%while small change is observed also in the g$_2$-g$_3$ bond in the glycerol backbone.
%This result is in line with the recently proposed model suggesting that the flexible phosphate enables the headgroup
%orient according to the charge accumulated into a membrane \cite{antila21}.
We also tried to resolve PC lipid headgroup conformational ensembles in mixed membrane with
negatively charged lipids \cite{scherer87},
%Headgroup order parameters of PC lipids respond similarly also to the addition of charged lipids in experiments \cite{??}.
%To resolve the conformational ensembles of lipid headgroups also in biologically relevant
%mixtures, % of zwitterionic and charged lipids,
%we compared the response of headgroup order parameters in
%PC lipids to the addition of PE or PG (Fig. \ref{??}),
%as well as in PG lipids to the addition of PC (Fig. \ref{??}),
%between simulations and experiments. However, the
but the accuracy of currently available force fields
was not sufficient for this (Fig. \ref{HGorderparametersPCvsPEPG}).
%to capture the headgroup conformational ensembles
%in mixed membranes: 
%The best performing force field for single componen membranes, CHARMM36, overestimates
%the effect of PE to PC conformations and underestimates the response PC to the addition
%of anionic PG headgroup. Previously we explained similar result for PC/PS mixtures  
%%as we previously observed also for mixtures with PS lipids \cite{antila19}.
%%The latter may arise from
%by overbinding of the sodium counterions \cite{antila19}.

Also binding of ions may affect the lipid headgroup conformational ensembles in physiological conditions.
The resolve this effect, we compared the changes of headgroup order parameters
in POPC:POPG mixtures 
upon addition of CaCl$_2$ between different simulations and experiments \cite{borle85,macdonald87}
in figures \ref{changesWITHCaClPG} (molar ratio 1:1) and \ref{changesWITHCaClPG1PC4} (molar ratio 4:1).
The results are in line with our previous studies:
most force fields overestimate the calcium binding \cite{catte16,antila19}, but CHARMM36 with the NBfix correction
underestimates the binding affinity \cite{antila19}, and the implicit inclusion of electronic polarizability
using the electronic continuum correction (ECC) improves the results \cite{melcr18,melcr20}.
The most realistic response of PC lipid headgroup order parameters to the calcium binding
is observed in the Lipid17ecc simulations with the ECC included, where
the decreased probability for trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
is the main structural consequence of bound ions (Fig. \ref{DIHSwithCAlipid17eccPOPC}).
Similar but smaller effect is seen in CHARMM36 simulations with less bound calcium (Fig. \ref{DIHSwithCAcharmm36POPC}).

%distribution from
%to eclipsed conformations 
%In this model, the main effect of calcium to the lipid conformational ensemble is
%the slight change of 
%This is in line with the changes observed in CHARMM36 simulations upon addition of charged surgactants (Fig. \ref{changesWITHsurf}),
%despite the major differences in lipid headgroup conformational ensembles between
%these models without ions (Fig. \ref{structures} vs. Fig. \ref{DIHSwithCAlipid17eccPOPC}).

%of lipid headgroup is not 
%Changes in lipid conformational ensembles characterized by distributions of heavy atom dihedral angles
%are very small upon binding of calcium or addition of cationic surfactants to membranes (Figs. \ref{??}).
%Yet, such changes are sufficient to tilt the headgroup dipole angle and reproduce the experimentally observed
%order parameter changes in $\alpha$ and $\beta$ carbons.

%Among the available simulations, there are two models that correctly reproduce the
%PG headgroup order parameter changes upon addition of calcium, Lipid17 and Slipids,
%even though the binding affinity of calcium is overestimated in these simulations.
%The response of PC to calcium binding is correctly reproduced by Lipid17ecc model, as
%previously reported for other membranes \cite{melcr18,melcr19}.

%to Amber based lipid models
%improves the ion binding affinity leading to a good agreement with experiments.
%(Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4})
%
%indicating too strong binding affinity of calcium into the bilayers as previously
%observed for pure PC lipid bilayers and mixtures with PS lipids \cite{catte16,antila19}.
%
%The decrease of $\alpha$-carbon order parameter of PC lipids in PC:PG mixtures as a function
%of calcium concentration is close to experiments CHARMM36 simulations (Fig. \ref{changesWITHCaClPG}),
%but the decrease of $\beta$-carbon order parameter seems to be overestimated.
%However, the $\beta$-carbon order parameter was not actually measured from these samples,
%but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
%On the other hand, the simulations are not converged.
%
%The calcium binding affinity to lipid bilayers with PC and PS lipids
%was recently improved by applying the electronic continuum correction (ECC) to Amber Lipid14/17 force fields \cite{melcr18,melcr20}.
%In this approach, the electronic polarizability is implicitly included in the classical force fields
%by scaling the charges with constant factors \cite{leontyev11}. Here, we make a ECC-POPG force field
%by applying the scaling factors originally used for POPS also to POPG, i.e., we multiply
%charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions with 
%$f_q$=0.75 and $f_\sigma$=0.89, respectively \cite{melcr20}.
%ECC-POPG model gives a weaker calcium binding affinity (Fig.\ref{CAdensPG}) and better agreement with the experimental
%PC headgroup order parameter changes (Fig. \ref{changesWITHCaClPG}) for POPC:POPG mixtures than the original Lipid17 model
%\todo{to be finished when we have all the data}, indicating that the ECC improves the simulation predictions of
%calcium binding affinity as previously observed for PC and PS lipids \cite{melcr18,melcr20}.

%\begin{figure}[]
%  \centering
%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MCHARMM36.png}
%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
%  \caption{\label{DIHSwithCA}
%    Changes in POPG CHARMM36 dihedrals with increasing amount of CaCl$_2$.
%  }
%\end{figure}

%Simulations also qualitatively reproduce the experimentally observed decrease of POPG
%%Experimental data for the
%$\beta$-carbon order parameter
%%of POPG shows a rapid decrease with increasing
%upon addition of CaCl$_2$,
%%concentrations up to 10 mM
%suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
%to PS in our previous work \cite{antila19}.
%and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
%This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
%but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
%addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
%calcium are not available.
%%upon addition of calcium in experiments,
%%exhibiting 
%%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.

%Decrease in headgroup order parameters of PC lipids upon addition of charges is
%usually qualitatitely captured in simulations despite the inaccuries in lipid conformational ensembles \cite{catte16},
%but situation with PS lipids is more complicated \cite{antila19,melcr20}.
Also the headgroup order parameters of charged lipids are affected by the binding of ions,
but their response is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
In our simulations, the Lipid17 and Slipids force fields give the most realistic response of PG $\beta$-carbon to CaCl$_2$
in figure \ref{changesWITHCaClPG} even thought the calcium binding affinity was overestimated in these models.
%Even thought the
%change in headgroup orientation upon addition of calcium is smaller
%On the other hand, the
These models suggest that the upward tilting of the headgroup dipole is weaker in PG than in PC headgroup (\ref{changesWITHCaClPG}),
%Heavy atom dihedral angle distributions calculated from these simulations
%suggest that also the headgroup glycerol conformations in PG are affected by calcium
but dihedral distributions are more sensitive to the bound ions in PG 
(Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
%in contrast to PC lipids where only conformations near phosphate were affected.
%, possible due to the compensating effects from the changes in
%phosphate and glycerol regions. 
However, none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids
simultaneously, experimental data to evaluate the response of $\alpha$-carbon order parameters to
the added calcium in PG is not available, and changes in dihedrals upon addition of cacium differ between
the best models (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions
and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.

%and PG
%The response of PG $\alpha$-carbon order parameters to CaCl$_2$ differs between force fields,
%but experimental data to evaluate these predictions is not available. 
%Simulations also qualitatively reproduce the experimentally observed decrease of POPG
%%Experimental data for the
%$\beta$-carbon order parameter
%%of POPG shows a rapid decrease with increasing
%upon addition of CaCl$_2$,
%%concentrations up to 10 mM
%suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
%to PS in our previous work \cite{antila19}.
%and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
%This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
%but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
%addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
%calcium are not available.
%%upon addition of calcium in experiments,
%%exhibiting 
%%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.

% The result is similar to the $\sim$200~ns simulations with PC lipids in previous work \cite{catte16}.
%However, when simulation was continued for $\mu$s, the binding affinity substantially increased
%and interpretation was that calcium overbinds to PC lipid in CHARMM36. Therefore, the
%conclusion seems to be similar here, although the new NBfix parameters may complicate
%the situation \todo{The status of NBfix parameters in these simulations should be checked.}.

%The $\beta$-carbon order parameter of PG exhibits a rapid decrease with small CaCl$_2$ concentrations
%and a more modest decrease with larger concentrations in experiments \cite{borle85} (Fig. \ref{PSPGchangesWITHCaCl}).
%The rapid decrease with CaCl$_2$ is observed but overestimated in CHARMM36 simulation
%with POPC:POPG 1:1 mixture, but not in 4:1 mixture \todo{This is little bit weird, should be checked.}.

%\todo{We still need more data to finish the discussion. More detailed discussion is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}

Despite the limited capability of simulations to interpret some of the experimental data due to
the inaccuries in force fields, we conclude that the experimentally observed changes
in headgroup order parameters upon addition of charges into bilayer arise from relatively
small changes in conformational ensembles. These changes eventuate from
mild changes in dihedral angle distributions, rather than from restriction of lipids into
fixed conformations. Therefore, lipid headgroups remain in disordered state sampling large
space of different conformations, thereby being able to interact with different molecules
in various ways, also in membranes bearing electric charge.

\subsection{Protein bound lipid conformations}

To analyze lipid conformations in protein bound states,
we calculated the heavy atom dihedral distributions from 
all lipid conformations available within protein structures
deposited in the PDB \cite{berman00} in figure \ref{dihedralsFROMpdb}.
We found 176 PC, 198 PE, 70 PG, and 41 PS conformations, which
present lipids that are tightly bound to proteins in fixed conformations
that can be determined as a part of protein structure using crystallography, NMR, or cryo-EM.

\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSALLfromPDB.eps}
%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPGfromPDB.eps}
%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPSfromPDB.eps}
  \caption{\label{dihedralsFROMpdb}
    Dihedral distributions from simulations and lipid structures in PDB.
  }
\end{figure}

Similarly to the liquid lamellar state, lipid dihedrals have very wide range of angles
when bound to proteins. Only %restrictions seem to be in 
%Almost all dihedral angles are found from the protein bounds states for all dihedrals
%except for
P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals
seem to avoid cis conformations as observed also in liquid lamellar state.
The %range of available angles for
O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and
g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals are even less restricted in protein bound state
than in liquid lamellar phase where cis and anti ecplipsed state were not present (Fig. \ref{structures}). 
Major differences between different lipid headgroups are not observed when bound to proteins.
Slight preference for the trans state in O$_\alpha$-C$_\alpha$-C$_\beta$-/C$_\gamma$ dihedral of PC,
and for positive angles in P-O$_\alpha$-C$_\alpha$-C$_\gamma$ dihedral of PS lipids with respect
to other lipids could be present in the data, but the statistics is not sufficient for conclusions.
Therefore, the differences in conformational ensembles between different
lipids in liquid lamellar state in Fig. \ref{structures} are not seen in the protein bound states.

This suggests that lipids can bind to proteins in wide range of conformations independently
of headgroup chemistry. Thereby, lipids compromize their preferred conformations when binding
to proteins and the binding of lipids to specific locations would be driven by 
the intermolecular interactions between proteins and lipids.
However, it is important to note that lipids are often not the main target in
structures deposited in PDB and therefore their conformations may be less reliable than proteins.
Stereochemical violations and structures deviating from lipid crystals have been
previously proposed to indicate inaccuracies in lipid structures in PDB \cite{marsh13b,pezeshkian18}. 
However, we see large deviations from lipid crystals structures also in conformational ensembles
that reproduce the NMR data in liquid lamellar phase, thereby proposing that such deviations
are realistic also in protein bound states.

\section{Conclusions}

%We have measured the C-H bond order parameters with the signs for headgroup and glycerol
%backbone regions of the
Conformational ensembles resolved using solid state NMR experiments and MD simulations
from NMRlipids open collaboration revealed that headgroups of the most abundant biological
phopholipids, PC, PE, PG and PS, sample wide range of different conformations in lamellar liquid state.
%Combining this experimental data with the large amount of simulation data collected 
%using the NMRlipids open collaboration, we were able to resolve the differences between
%conformational ensembles of lipid headgroup leading to the differences observed in NMR experiments.
%Our results indicate that lipid headgroups are flexible and sample large conformational space
%in biologically relevant liquid lamellar state,
Differences in NMR order parameters between different headgroups can be explained
by the changes in dihedral angle distributions, suggesting that
similar conformations are accessed by all headgroups, but with different probabilities.
Also the changes in order parameters upon addition of charged molecules, such as cationic lipids, surfactants, or drugs,
in membranes can be explained by the changes in dihedral angle probabilities, particularly close to the phosphate region.

Wide range of conformations are observed also in lipids that are tightly bound to proteins in PDB,
%Therefore, lipids can bind to proteins and other biomolecules in multiple different conformations,
%as indeed observed in the analysis of protein bound lipid conformations available from PDB.
%The weak correlation between lipid structures in liquid lamellar phase and protein bound state
suggesting that the specific binding of lipids to proteins is dominated by the intermolecular lipid-protein
interactions, while the differences in conformational ensembles between different lipid types
play a minor role. Therefore, lipids can conform themself to multiple binding positions locating
in various proteins.

Our results pave the way to understand how lipids regulate membrane protein function.
Lipid conformational ensembles in liquid lamellar phase are necessary
for the detailed analysis of lipid-protein interaction energetics,
%the
%realistic MD simulations enables accurate
%analysis of
particularly for the entropic part. 
Furthermore, our results demonstrate how the NMRlipids databank, containing MD simulations and NMR data,
can be used to resolve
%NMR data can be combined with the large amount of
%indexed MD simulation data to find the most realistic
conformational ensembles of disordered biomolecules in membrane environment,
and how this information can be coupled with structural data in the PDB.
We believe that this combination brings us closer to the comprehensive understanding of
complex biomolecular assemblies containg both disordered and structural regions, such as membrane
protein complexes.
%This paves the way toward standardized methods to resolve the quality
%evaluated conformational ensembles of disordered biomolecules in membrane environment.
%In the NMRlipids project we aim to build a general databank of quality evaluated
%conformational ensembles of lipids and other disordered biomolecules based primarily on
%MD simulations and NMR data.

% Tables may be be put in the text as floats.
% Here is an example of the general form of a table:
% Fill in the caption in the braces of the \caption{} command. Put the label
% that you will use with \ref{} command in the braces of the \label{} command.
% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
% \begin{tabular}{} command.
%
% \begin{table}
% \caption{\label{} }
% \begin{tabular}{}
% \end{tabular}
% \end{table}

% If you have acknowledgments, this puts in the proper section head.
\begin{acknowledgments}
AP is grateful to the Centro de
Supercomputación de Galicia (CESGA) for use of the Finis
Terrae computer
    %     Put your acknowledgments here.
\end{acknowledgments}
%\newpage
%\appendix
%\begin{center}
%{\bf SUPPLEMENTARY INFORMATION}
%\end{center}

%\section{Measurements of order parameter sign}

%Fig. \ref{PShgSIGNS} summarizes the experimental results on the order parameter sign
%measurement for POPS sample. The experimental protocol is the same used in Ref. \citenum{ferreira16}.
%In (a) you see the headgroup region of the INEPT spectrum where alpha and beta are
%identified. In (b) you have the R-PDLF slices for alpha and beta where you see one single
%splitting for beta (which gives an order parameter equal to 0.12), and for alpha a superposition
%of a large splitting (order parameter equal to 0.09) and a very small splitting which cannot be
%calculated. On the bottom you have the S-DROSS slices of these two carbons. The grey lines show a
%random collection of slices from noise such that it gets clear what is significant. The S-DROSS
%slice for beta clearly shows that the order parameter is negative. The slice for alpha shows that
%the higher order parameter is positive and suggests that the smaller order parameter is negative
%(from the deviation towards negative values in the longer t1 times).
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/PShgSIGNS.pdf}
%  \caption{\label{PShgSIGNS}
%    Experimental results for sign measurement for POPS sample
%  }
%\end{figure}

%The results updated with SIMPSON simulations for the SDROSS profiles
%are shown in Fig. \ref{PShgSIGNSsimpson}. The value for the smaller
%alpha order parameter is taken from Fig 3 in Ref. \citenum{roux91},
%because resolution in 13C NMR experiments was nor high enough to determine
%numerical value for this. The plots in Fig. \ref{PShgSIGNSsimpson} (c) show
%the following. The error bars and points are the experimental SDROSS data.
%The thick lines are SIMPSON simulations. The simulations were done by using
%the order parameter for beta equal to -0.12 and for alpha one order parameter
%equal to 0.09 and the other equal to -0.02 (black) or 0.02 (grey).
%Since the black lines agree with experimental data, we conlude that
%the order parameters for $\beta$ carbon are -0.12 and for $\alpha$
%order parameters are 0.09 and -0.02.
%\begin{figure}[]
%  \centering
%  \includegraphics[width=9.0cm]{../Figs/PShgSIGNSsimpson.pdf}
%  \caption{\label{PShgSIGNSsimpson}
%    Experimental results for sign measurement for POPS sample
%  }
%\end{figure}

%\section{Dihedrals}
%\begin{figure*}[]
%  \centering
%  \includegraphics[width=8.0cm]{../Figs/dihed1.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed2.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed3.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed4.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed5.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed6.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed7.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed8.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed9.png}
%  \includegraphics[width=8.0cm]{../Figs/dihed10.png}
%  \caption{\label{dihedrals}
%    Experimental results for sign measurement for POPS sample
%  }
%\end{figure*}

% Create the reference section using BibTe
\bibliography{refs.bib}

%\newpage
%\section{APPENDIX: The NMR results reported by Tiago Ferreira}

\listoftodos

\end{document}
%
% ****** End of file aiptemplate.tex ******

### 2021-01-05 manuscriptPGPE.tex

@@
 
-  %The force field giving the best description for glycerol backbone and headgroup structures
-  %of PC, PS, PG and PE headgroups (CHARMM36) reproduces the essential differences in order parameters
-  %between these headgroups, and therefore enables the analysis of structural differences between the
-  %headgroups.
-
 \end{abstract}
@@
 are typically derived from NMR experiments, particularly from 
-%Attempts to resolve the conformational ensembles of lipid headgroup rely mainly on
 C-H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}
 which can be detected even from cells \cite{gally81,scherer87,seelig90}.
-%but lipids are not in solid state in biomembranes. 
-%On the other hand, NMR experiments enable detection of lipid headgroups
-%in physiological liquid states also in cells \cite{gally81,scherer87,seelig90}.
 These studies suggest that the glycerol backbone structures are largely similar irrespectively of the headgroup \cite{gally81}, 
 and the headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
-%Extensive discussion about structural details of PE, PG or PS headgroups do not exists (as far as I know), 
-%In contrast to PC lipids (see \cite{botan15} and references therein).
-%Some attempts to resolve conformational ensembles from NMR for PC and PE lipids have been made,
-%but lesser extend for PG or PS lipids \cite{seelig77c,davis83,Semchyschyn04}.
 However, these results are based only on the absolute values while 
-%we have recently demonstrated that also
 the necessity of order parameters signs in capturing the lipid conformational ensembles has been resently demonstrated \cite{botan15,ollila16,ferreira16}.
@@
 is limited by the lack of universal models to interpret the experiments \cite{pezeshkian18,akutsu20}.
-%do not exist and
-%the proposed structures are often based on lipid crystal structures \cite{pezeshkian18,akutsu20}.
-%which were solved many decades ago \cite{buldt81,pascher92}.
-%determine the structural 
-%Classical molecular dynamics simulations could potentially give such ensembles and therefore enable
-%the detailed studies of lipid headgroup behaviour in complex biomolecular systems, but current
-%force fields are not accurate enough to reproduce the correct conformational ensembles for PC and PS headgroups \cite{botan15,antila19}.
 
@@
 In addition to the changes in lipid conformational ensembles upon binding to proteins,
-%PC headgroup conformations do not change upon addition of zwitterionic lipids or cholesterol \cite{scherer87,ferreira13} into a membrane,
-%but are sensitive to the presence of charged molecules in membranes \cite{seelig87}.
 also the experimentally measured response of lipid headgroup to membrane bound charges remains poorly understood 
-%The response to charges is explained by the change in the orientation of lipid headgroup dipole vector due to electrostatic interactions,
-%but the detailed changes are not yet understood
 due to the lack of suitable models to interpret the lipid conformational ensembles in liquid lamellar state \cite{Semchyschyn04}.
@@
 to resolve the differences in conformational ensembles between PC, PE, PG and PS lipid headgroups.
-%open collaboration and order parameters of glycerol backbone and headgroup
-%to evaluate the accuracy of PE and PG heagroup structures, and the cation binding affinity
 Zwitterionic PC and PE are the most common lipids in eukaryotes and bacteria, respectively \cite{vanmeer08,sohlenkamp16}.
@@
 and affect membrane protein functionality and signaling \cite{lemmon08,leventis10,sohlenkamp16,hariharan18}.
-%PE and PG lipids are most common lipids in bacteria \cite{sohlenkamp16}.
-%Anionic PG lipids are less abundant, but is also proposed to be fundamental for terrestrial life \cite{furse17}.
-%PE and PG affect membrane protein functionality \cite{hariharan18} and
 All the studied lipids specifically bind to various proteins \cite{yeagle14}.
-%PE headgroup is also prone for negative membrane curvature and causes membrane fusion \cite{Chernomordik08,calzada16}.
-%Therefore, the PE and PG headgroup structures play probably essential roles in 
-%many biological processes.
 Furthermore, we use our results to elucidate lipid-protein interactions and the effect of charges on lipid conformations.
@@
 the results from model systems could be used to understand the biological role of lipid headgroup conformational ensembles in different lipid types.
-
-%Several MD simulations of PE and PG lipids have been published especially in the context of modeling
-%inner membrane of Gram-negative
-%bacteria \cite{devries04,murzyn05,pedersen06,zhao07,gurtovenko08,zhao08,henin09,kukol09,tsai12,dickson12,venable13,dickson14,berglund15}
-%\todo{There may be some relevant publication missing from here},
-%but evaluation of glycerol backbone and headgroup structures against experiments is rare \cite{henin09}.
-
-%Derivation of different lipid headgroup conformational ensembles in liquid state has been
-%inconclusive due to lack of suitable experimental data and tools to interpret the conformational ensembles.
-
-%Besides the structure, also ion binding may regulate biophysical activity of especially negatively
-%charged lipid headgroups \cite{seelig90}. Monovalent cation (except Lithium) binding to zwitterionic PC and anionic PS headgroups is
-%very weak, while multivalent ion binding is stronger but still weak \cite{cevc90,tocanne90,roux90,catte16,antila19}. The ion binding affinity
-%data for PE is more scarce \cite{marra85}, but large differences to PC would be surprising.
-%Negatively charged lipids are suggested to bear same cation binding constants than zwitterionic lipids, but
-%the amount of bound ions to negatively charged membranes would still be larger because
-%the concentration of cations in the vicinity of membranes would be higher  \cite{seelig90}.
-%
-%is stronger
-%Based on the electrometer concept and other data is has been suggested that \cite{seelig90}
-%\begin{displayquote}
-%  {\it ''(i) Ca$^{2+}$ binds to neutral lipids (phosphatidylcholine, phosphatidylethanolamine) and negatively charged lipids
-%    (phosphatidylglycerol) with approximately the same binding constant of K = 10-20 M$^{-1}$; \\
-%    (ii) the free Ca$^{2+}$
-%    concentration at the membrane interface is distinctly enhanced if the membrane carries a negative surface
-%    charge, either due to protein or to lipid; \\
-%    (iii) increased inter-facial Ca$^{2+}$ also means increased amounts
-%    of bound Ca$^{2+}$ at neutral and charged lipids; \\
-%    (iv) the actual binding step can be described by a Langmuir
-%    adsorption isotherm with a 1 lipid:1 Ca$^{2+}$ stoichiometry, provided the interfacial concentration C$_M$, is
-%    used to describe the chemical binding equilibrium.''}
-%\end{displayquote}
-%
-%On the other hand, anionic PS lipids are proposed chelate with calcium ions \cite{hauser85,feigenson86,roux91}.
-%In simulations, the cation binding affinity to PC and PS membranes is typically overestimated \cite{catte16,antila19},
-%which can be improved by applying the ECC to the partial charges of the force fields \cite{melcr18,melcr19}.
-
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=9.0cm]{../Figs/lipids.pdf}
-%  \caption{\label{lipids}
-%    Chemical structures and labels for the headgroup carbons.
-%  }
-%\end{figure}
 
@@
 %Magic angle spinning (MAS) of the sample was used at a frequency of 5.15 kHz (R-PDLF experiment) and 5 kHz (S-DROSS experiment).
-%The following experimental setups were used.
 The NMR experiments were identical as in our previous work~\cite{antila19}.
 POPE and POPG powder were purchased from Avanti polar lipids \todo{Which isomer of POPG we have in the experiments?}.
-%\todo{Is this enough and correct, or should we repeat some methods from the NMRlipidsIVps paper?}
 The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.
-
-%Absolute values of the headgroup and glycerol backbone order parameters from PE and PG lipids are measured
-%previously using $^2$H NMR~\cite{seelig76,gally81,wohlgemuth80,borle85}. Because also the order parameter
-%signs bear essential information about the lipid structures \cite{botan15,ollila16}, we measured the
-%magnitudes and signs of POPE and POPG C--H bond headgroup and glycerol backbone order parameter in liquid phase
-%using the 2D-RPDLF and S-DROSS experiments, as described previously \cite{ferreira13,ferreira16,antila19}.
 
@@
 were assigned based on previously measured POPC spectra~\cite{ferreira13}.
-%For POPE, the glycerol backbone and $\alpha$-carbon peaks in INEPT spectra were assigned based on
-%previously measured POPC spectra~\cite{ferreira13} and
 The $\beta$-carbon peak from POPE was assigned based on $^{13}$C chemical shift table for amines available
 at \url{https://www.chem.wisc.edu/areas/reich/nmr/c13-data/cdata.htm}.
-%For POPG, the glycerol backbone peaks in INEPT spectra were assigned based on
-%previously measured POPC spectra~\cite{ferreira13},
 \todo{How were $\alpha$ and  $\gamma$-carbon peaks assigned in POPG?}.
@@
 
-%\subsection{Headgroup and glycerol backbone order parameters of POPE and POPG from $^{13}$C NMR}
-
 \begin{figure*}[]
   \centering
-   \includegraphics[width=18.0cm]{../Figs/figure2.eps}
-%  \includegraphics[width=18.0cm]{../Figs/PCPEPGPSstructures2.png}
-%  \includegraphics[width=8.0cm]{../Figs/CHARMMfromLIPIDS.eps}
-%%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals1.png}
-%%  \includegraphics[width=8.0cm]{../Figs/PCPEPGPSdihedrals2.png}
-%  \includegraphics[width=8.0cm]{../Figs/PNangleNORM.eps}
-%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALS.eps}
+   \includegraphics[width=18.0cm]{./Figs/figure2.eps}
    \caption{\label{structures}
@@
 Determination of headgroup and glycerol backbone order parameters and their signs
-was straightforward from the
-%R-PDLF and S-DROSS
-data in Figs. \ref{HGorderParameters}, \ref{POPEspectra} and \ref{POPGspectra}
+was straightforward from the data in Figs. \ref{HGorderParameters}, \ref{POPEspectra} and \ref{POPGspectra}
 for all the C-H bonds, except the $\beta$ and g$_2$ carbons in POPG.
@@
 we assigned the larger order parameter to the g$_2$ carbon
-%order parameter of $\beta$-carbon is expected to be clearly smaller than for 
 and used the literature value for the $\beta$-carbon in SIMPSON simulations to determine the signs.
@@
 This interpretation is confirmed by SIMPSON calculations in Fig \ref{POPGsimpson}.
-%using negative value for g$_2$ and positive value for $\beta$ order parameter (Fig. \ref{POPGsimpson}).
-%The signs of these order parameters were solved using SIMPSON simulations to interpre the
-%S-DROSS experiments in similar fashion as we did previously for PS lipids \cite{antila19}.
-
-%Also the headgroup order parameters of PC lipids are close PE, althought the latter gives systematically
-%slightly more positive values (Fig. \ref{HGorderParameters}).
-% These could be explained with slightly larger temperature in PE measurements, except for the $\alpha$-carbon with the positive sign, for which the
-% more positive value is farther away from zero.
-%While the headgroup order parameters are similar for PC and PE lipids,
-%PG and PS lipids exhibit distinct values in comparison with other lipids.
 
@@
 where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
-%The resulting order parameters are compared with our previously published results
-%for POPC \cite{ferreira13,ferreira16} and POPS \cite{antila19} in figure \ref{HGorderParameters} D.
-%Order parameters for POPE are in good agreement with previous $^2$H NMR experiments \cite{seelig76}
-%and similar to our previous results for POPC \cite{ferreira16}.
 Overall agreement of order parameters determined by different authors and different techniques
@@
 On the other hand, the $\beta$-carbon order parameter of PG headgroup
-%the $\alpha$-carbon order parameter is similar to PE and PG, while
 has positive sign in contrast to all the other lipids.
 Notably, this has not been observed in traditional $^2$H NMR experiments,
-%because absolute value of $\beta$-carbon order parameter is similar in PG, PE and PC lipids and
 where only the absolute value of the order parameters are measured~\cite{wohlgemuth80,gally81,borle85}.
@@
 
-%As discussed previously for PC and PS headgroups \cite{ollila16,antila19}, also 
-%the headgroup and glycerol backbone order parameters of PE and PG are essentially independent of 
-%acyl chain composition, and therefore manifest mainly headgroup chemistry (Fig. \ref{HGorderParameters}).
-
 \begin{figure*}[bt]
   \centering
-  \includegraphics[width=18.0cm]{../Figs/HGorderparametersPCvsSURFchangeDIHEDRALS.eps}
+  \includegraphics[width=18.0cm]{./Figs/HGorderparametersPCvsSURFchangeDIHEDRALS.eps}
   \caption{\label{changesWITHsurf}
@@
 (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}) or in our previous studies \cite{botan15,antila19}.
-%has least problems in
-%reproducing the experimental lipid headgroup and glycerol backbone order parameters in different
-%lipids \cite{botan15,antila19} (Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}).
 Nevertheless, the CHARMM36 force field performs best for all the headgroups and
-%Importantly, it captures the experimentally observed distinct order parameters for PS and PG headgroups (Fig. \ref{??}).
-%To characterize the differences in lipid headgroup conformational ensembles
 reproduces the distinct order parameters of PS and PG lipids (Fig. \ref{structures} B).
@@
 we calculated the distributions of heavy atom dihedral angles from CHARMM36 simulations in figure \ref{structures} C).
-%The dihedral distributions in figure \ref{structures} show significant similarity between different lipids for all the other
-%bonds, except
 Major differences between headgroups are observed only for the last two dihedrals
@@
 Rest of the dihedrals are similar between different lipids, with the exception of
-%However, some dihedral angles of phosphorus-oxygen bonds are less likely for PS, which
 small differences for PS lipids. Therefore we suggest that the main differences between
@@
 explain the more rigid headgroup structures~\cite{browning80,buldt81}.
-%These differences explaing the disctinct order parameters of
-%PS and PG headgroups with respenct to the other lipids.
-%The differences between lipids are reflected also to
 Furthermore, the angle between headgroup dipole and membrane normal
@@
 
-%The eclipsed (0 or 360 degrees) are not present in any other dihedrals, and anti-eclipsed (120 or 240 degrees)
-%are not present in C$_\alpha$-C$_\beta$ and g$_1$-g$_2$ bonds.
-
-%Characteristic dihedral conformations in PS headgroup are asymmetric conformations
-%preferring gauche 270$^\circ$ conformations in N-C$_\beta$-C$_\alpha$-O$_\alpha$ and
-%%dihedral exhibits a more asymmetric angle distribution for PS, having largest probability being
-%%in , than for the PC headgroup.
-%C$_\beta$-C$_\alpha$-O$_\alpha$-P dihedrals. 
-%In PG headgroup, the O$_\beta$-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral (corresponding N-C$_\beta$-C$_\alpha$-O$_\alpha$ dihedral in other% lipids)
-%is mostly in trans conformation, and C$_\beta$-C$_\alpha$-O$_\alpha$-P  has asymmetric tendency to be in
-%gauche 60$^\circ$ conformation. 
-%Main difference between PC and PE is the lower probability of trans state in C$_\beta$-C$_\alpha$-O$_\alpha$-P PC dihedral,
-%which could be a potential reason for the too negative $\beta$ headgroup order parameter in PC.%
-
 In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase.
-%the results suggest that free rotations around O-P bonds decouple the
-%headgroup and glycerol backbone structure and dynamics in similarly in all lipids,
-%althought this rotation is slower in PS lipids where crossing certain barriers are slower.
-Despite the differences in distributions
-%of two the last bonds at the headgroup end,
-%leading to the distinct order parameters for PS and PG lipids,
-%lipids with
-between headgroups, the sampled dihedral angles are within approximately same ranges for all headgroup types.
-Therefore, we suggest that all lipid headgroups are very flexible thereby being
+between headgroups, the sampled dihedral angles are within approximately same ranges for all headgroup types. Therefore, we suggest that all lipid headgroups are very flexible thereby being
 able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
-%Furthermore, approximately same conformations are present liquid lamellar phase for different
-%headgroups studied here, althought the probability weight for the conformations are slightly different.
-%the order parameter experiments suggest that the glycerol backbone conformations in all lipids
-%and the headgroup conformations in PC and PE lipids are relatively similar, while 
-%PS and PG headgropus exhibit distinct conformational sampling.
-%The details of sampled conformation are difficult to deduce from order parameters only,
-%but the distinct headgroup order parameters
-%of PS lipids are previously related to the more rigid structure of the headgroup~\cite{browning80,buldt81,antila19}.
-%The analysis of dihedral angle distributions reveal the free rotation of dihedral around
-%phosphorus-oxygen bonds in all lipids as all possible angles are observed in the distributions.
-Our results support the models that propose free rotation around phosphate group for PC lipids,
-which decouples the dynamics and conformatios between acyl chains and headgroups \cite{klauda08c,antila21b},
-and suggests that these models could be applied for all lipids containing the phosphorus group.
-The wide range of observed conformations suggest that the structures in lipid crystals \cite{buldt81,pascher92}
-play only a minor role, and that the models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20}
-are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
+Our results support the models that propose free rotation around phosphate group for PC lipids, which decouples the dynamics and conformatios between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
+The wide range of observed conformations suggest that the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and that the models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
 
@@
 lipid conformational ensembles in membranes bearing electric charge is still lacking \cite{Semchyschyn04}.
-%Order parameters of $\alpha$ and $\beta$ carbons in PC lipid headgroup are known to depend
-%linearly on the incorporation of charged molecules into a lipid bilayer,
-%which is explained by tilting of the headgroup dipole orientation \cite{??}.
-%%are kn affect the headgroup dipole vector tilt which can
-%%be detected by measuring C-H bond order parameters of $\alpha$ and $\beta$ carbons.
-%Such changes have been observed upon addition of charged lipids, proteins, surfactants,
-%drugs, or ions \cite{??}, but how charges affect the lipid conformational ensemble remains unknown.
 
@@
   \centering
-  \includegraphics[width=18.0cm]{../Figs/CHANGESwithCaClPG1PC1.eps}
+  \includegraphics[width=18.0cm]{./Figs/CHANGESwithCaClPG1PC1.eps}
   \caption{\label{changesWITHCaClPG}
@@
 
-To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge,
-%suggesting that these simulations can be used to interpret the charge induced changes in lipid conformational ensembles.
-we first analyzed CHARMM36 simulations which correctly captures the experimentally measured decrease in PC headgroup order parameters
-upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
+To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge, we first analyzed CHARMM36 simulations which correctly captures the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
 Heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B
@@
 even thought almost half of the molecules in membranes are cationic surfactants.
-%The major changes upon addition of charge are observed in dihedrals related to the phosphate oxygens,
-%while small change is observed also in the g$_2$-g$_3$ bond in the glycerol backbone.
-%This result is in line with the recently proposed model suggesting that the flexible phosphate enables the headgroup
-%orient according to the charge accumulated into a membrane \cite{antila21}.
 We also tried to resolve PC lipid headgroup conformational ensembles in mixed membrane with
 negatively charged lipids \cite{scherer87},
-%Headgroup order parameters of PC lipids respond similarly also to the addition of charged lipids in experiments \cite{??}.
-%To resolve the conformational ensembles of lipid headgroups also in biologically relevant
-%mixtures, % of zwitterionic and charged lipids,
-%we compared the response of headgroup order parameters in
-%PC lipids to the addition of PE or PG (Fig. \ref{??}),
-%as well as in PG lipids to the addition of PC (Fig. \ref{??}),
-%between simulations and experiments. However, the
 but the accuracy of currently available force fields
 was not sufficient for this (Fig. \ref{HGorderparametersPCvsPEPG}).
-%to capture the headgroup conformational ensembles
-%in mixed membranes: 
-%The best performing force field for single componen membranes, CHARMM36, overestimates
-%the effect of PE to PC conformations and underestimates the response PC to the addition
-%of anionic PG headgroup. Previously we explained similar result for PC/PS mixtures  
-%%as we previously observed also for mixtures with PS lipids \cite{antila19}.
-%%The latter may arise from
-%by overbinding of the sodium counterions \cite{antila19}.
 
@@
 The resolve this effect, we compared the changes of headgroup order parameters
-in POPC:POPG mixtures 
-upon addition of CaCl$_2$ between different simulations and experiments \cite{borle85,macdonald87}
-in figures \ref{changesWITHCaClPG} (molar ratio 1:1) and \ref{changesWITHCaClPG1PC4} (molar ratio 4:1).
+in POPC:POPG mixtures  upon addition of CaCl$_2$ between different simulations and experiments \cite{borle85,macdonald87} in figures \ref{changesWITHCaClPG} (molar ratio 1:1) and \ref{changesWITHCaClPG1PC4} (molar ratio 4:1).
 The results are in line with our previous studies:
-most force fields overestimate the calcium binding \cite{catte16,antila19}, but CHARMM36 with the NBfix correction
-underestimates the binding affinity \cite{antila19}, and the implicit inclusion of electronic polarizability
-using the electronic continuum correction (ECC) improves the results \cite{melcr18,melcr20}.
+most force fields overestimate the calcium binding \cite{catte16,antila19}, but CHARMM36 with the NBfix correction underestimates the binding affinity \cite{antila19}, and the implicit inclusion of electronic polarizability using the electronic continuum correction (ECC) improves the results \cite{melcr18,melcr20}.
 The most realistic response of PC lipid headgroup order parameters to the calcium binding
-is observed in the Lipid17ecc simulations with the ECC included, where
-the decreased probability for trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
-is the main structural consequence of bound ions (Fig. \ref{DIHSwithCAlipid17eccPOPC}).
-Similar but smaller effect is seen in CHARMM36 simulations with less bound calcium (Fig. \ref{DIHSwithCAcharmm36POPC}).
+is observed in the Lipid17ecc simulations with the ECC included, where the decreased probability for trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral is the main structural consequence of bound ions (Fig.
+\ref{DIHSwithCAlipid17eccPOPC}).
+Similar but smaller effect is seen in CHARMM36 simulations with less bound calcium (Fig.
+\ref{DIHSwithCAcharmm36POPC}).
 
-%distribution from
-%to eclipsed conformations 
-%In this model, the main effect of calcium to the lipid conformational ensemble is
-%the slight change of 
-%This is in line with the changes observed in CHARMM36 simulations upon addition of charged surgactants (Fig. \ref{changesWITHsurf}),
-%despite the major differences in lipid headgroup conformational ensembles between
-%these models without ions (Fig. \ref{structures} vs. Fig. \ref{DIHSwithCAlipid17eccPOPC}).
-
-%of lipid headgroup is not 
-%Changes in lipid conformational ensembles characterized by distributions of heavy atom dihedral angles
-%are very small upon binding of calcium or addition of cationic surfactants to membranes (Figs. \ref{??}).
-%Yet, such changes are sufficient to tilt the headgroup dipole angle and reproduce the experimentally observed
-%order parameter changes in $\alpha$ and $\beta$ carbons.
-
-%Among the available simulations, there are two models that correctly reproduce the
-%PG headgroup order parameter changes upon addition of calcium, Lipid17 and Slipids,
-%even though the binding affinity of calcium is overestimated in these simulations.
-%The response of PC to calcium binding is correctly reproduced by Lipid17ecc model, as
-%previously reported for other membranes \cite{melcr18,melcr19}.
-
-%to Amber based lipid models
-%improves the ion binding affinity leading to a good agreement with experiments.
-%(Figs. \ref{changesWITHCaClPG} and \ref{changesWITHCaClPG1PC4})
-%
-%indicating too strong binding affinity of calcium into the bilayers as previously
-%observed for pure PC lipid bilayers and mixtures with PS lipids \cite{catte16,antila19}.
-%
-%The decrease of $\alpha$-carbon order parameter of PC lipids in PC:PG mixtures as a function
-%of calcium concentration is close to experiments CHARMM36 simulations (Fig. \ref{changesWITHCaClPG}),
-%but the decrease of $\beta$-carbon order parameter seems to be overestimated.
-%However, the $\beta$-carbon order parameter was not actually measured from these samples,
-%but they are calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
-%On the other hand, the simulations are not converged.
-%
-%The calcium binding affinity to lipid bilayers with PC and PS lipids
-%was recently improved by applying the electronic continuum correction (ECC) to Amber Lipid14/17 force fields \cite{melcr18,melcr20}.
-%In this approach, the electronic polarizability is implicitly included in the classical force fields
-%by scaling the charges with constant factors \cite{leontyev11}. Here, we make a ECC-POPG force field
-%by applying the scaling factors originally used for POPS also to POPG, i.e., we multiply
-%charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions with 
-%$f_q$=0.75 and $f_\sigma$=0.89, respectively \cite{melcr20}.
-%ECC-POPG model gives a weaker calcium binding affinity (Fig.\ref{CAdensPG}) and better agreement with the experimental
-%PC headgroup order parameter changes (Fig. \ref{changesWITHCaClPG}) for POPC:POPG mixtures than the original Lipid17 model
-%\todo{to be finished when we have all the data}, indicating that the ECC improves the simulation predictions of
-%calcium binding affinity as previously observed for PC and PS lipids \cite{melcr18,melcr20}.
-
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{../Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
-%  \caption{\label{DIHSwithCA}
-%    Changes in POPG CHARMM36 dihedrals with increasing amount of CaCl$_2$.
-%  }
-%\end{figure}
-
-%Simulations also qualitatively reproduce the experimentally observed decrease of POPG
-%%Experimental data for the
-%$\beta$-carbon order parameter
-%%of POPG shows a rapid decrease with increasing
-%upon addition of CaCl$_2$,
-%%concentrations up to 10 mM
-%suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
-%to PS in our previous work \cite{antila19}.
-%and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
-%This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
-%but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
-%addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
-%calcium are not available.
-%%upon addition of calcium in experiments,
-%%exhibiting 
-%%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.
-
-%Decrease in headgroup order parameters of PC lipids upon addition of charges is
-%usually qualitatitely captured in simulations despite the inaccuries in lipid conformational ensembles \cite{catte16},
-%but situation with PS lipids is more complicated \cite{antila19,melcr20}.
 Also the headgroup order parameters of charged lipids are affected by the binding of ions,
@@
 in figure \ref{changesWITHCaClPG} even thought the calcium binding affinity was overestimated in these models.
-%Even thought the
-%change in headgroup orientation upon addition of calcium is smaller
-%On the other hand, the
-These models suggest that the upward tilting of the headgroup dipole is weaker in PG than in PC headgroup (\ref{changesWITHCaClPG}),
-%Heavy atom dihedral angle distributions calculated from these simulations
-%suggest that also the headgroup glycerol conformations in PG are affected by calcium
-but dihedral distributions are more sensitive to the bound ions in PG 
-(Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
-%in contrast to PC lipids where only conformations near phosphate were affected.
-%, possible due to the compensating effects from the changes in
-%phosphate and glycerol regions. 
-However, none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids
+These models suggest that the upward tilting of the headgroup dipole is weaker in PG than in PC headgroup (\ref{changesWITHCaClPG}), but dihedral distributions are more sensitive to the bound ions in PG
+(Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}). However, none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids
 simultaneously, experimental data to evaluate the response of $\alpha$-carbon order parameters to
@@
 the best models (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
-Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions
-and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.
-
-%and PG
-%The response of PG $\alpha$-carbon order parameters to CaCl$_2$ differs between force fields,
-%but experimental data to evaluate these predictions is not available. 
-%Simulations also qualitatively reproduce the experimentally observed decrease of POPG
-%%Experimental data for the
-%$\beta$-carbon order parameter
-%%of POPG shows a rapid decrease with increasing
-%upon addition of CaCl$_2$,
-%%concentrations up to 10 mM
-%suggesting that the PG headgroup response is qualitatively captured by simulations, in contrast
-%to PS in our previous work \cite{antila19}.
-%and more modest decrease with larger concentrations (Fig. \ref{changesWITHCaClPG}) \cite{borle85}.
-%This behaviour is similar to that of $\beta$-carbon order parameters of POPC,
-%but essentially different than observed for POPS, where $\beta$-carbon order parameters increases with
-%addition of calcium \cite{antila19}. Experimentally measured changes of PG $\alpha$-carbon order parameters upon addition of
-%calcium are not available.
-%%upon addition of calcium in experiments,
-%%exhibiting 
-%%The behaviour is qualitatively different than in the $\beta$-carbon of PS lipids which increase upon addition of CaCl$_2$.
-
-% The result is similar to the $\sim$200~ns simulations with PC lipids in previous work \cite{catte16}.
-%However, when simulation was continued for $\mu$s, the binding affinity substantially increased
-%and interpretation was that calcium overbinds to PC lipid in CHARMM36. Therefore, the
-%conclusion seems to be similar here, although the new NBfix parameters may complicate
-%the situation \todo{The status of NBfix parameters in these simulations should be checked.}.
-
-%The $\beta$-carbon order parameter of PG exhibits a rapid decrease with small CaCl$_2$ concentrations
-%and a more modest decrease with larger concentrations in experiments \cite{borle85} (Fig. \ref{PSPGchangesWITHCaCl}).
-%The rapid decrease with CaCl$_2$ is observed but overestimated in CHARMM36 simulation
-%with POPC:POPG 1:1 mixture, but not in 4:1 mixture \todo{This is little bit weird, should be checked.}.
-
-%\todo{We still need more data to finish the discussion. More detailed discussion is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}
+Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.
 
@@
   \centering
-  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSALLfromPDB.eps}
+  \includegraphics[width=8.0cm]{./Figs/DIHEDRALSALLfromPDB.eps}
 %  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
@@
 when bound to proteins. Only %restrictions seem to be in 
-%Almost all dihedral angles are found from the protein bounds states for all dihedrals
-%except for
 P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals
@@
 
-%We have measured the C-H bond order parameters with the signs for headgroup and glycerol
-%backbone regions of the
 Conformational ensembles resolved using solid state NMR experiments and MD simulations
@@
 phopholipids, PC, PE, PG and PS, sample wide range of different conformations in lamellar liquid state.
-%Combining this experimental data with the large amount of simulation data collected 
-%using the NMRlipids open collaboration, we were able to resolve the differences between
-%conformational ensembles of lipid headgroup leading to the differences observed in NMR experiments.
-%Our results indicate that lipid headgroups are flexible and sample large conformational space
-%in biologically relevant liquid lamellar state,
 Differences in NMR order parameters between different headgroups can be explained
@@
 Wide range of conformations are observed also in lipids that are tightly bound to proteins in PDB,
-%Therefore, lipids can bind to proteins and other biomolecules in multiple different conformations,
-%as indeed observed in the analysis of protein bound lipid conformations available from PDB.
-%The weak correlation between lipid structures in liquid lamellar phase and protein bound state
 suggesting that the specific binding of lipids to proteins is dominated by the intermolecular lipid-protein
@@
 Lipid conformational ensembles in liquid lamellar phase are necessary
-for the detailed analysis of lipid-protein interaction energetics,
-%the
-%realistic MD simulations enables accurate
-%analysis of
-particularly for the entropic part. 
+for the detailed analysis of lipid-protein interaction energetics, particularly for the entropic part. 
 Furthermore, our results demonstrate how the NMRlipids databank, containing MD simulations and NMR data,
-can be used to resolve
-%NMR data can be combined with the large amount of
-%indexed MD simulation data to find the most realistic
-conformational ensembles of disordered biomolecules in membrane environment,
+can be used to resolve conformational ensembles of disordered biomolecules in membrane environment,
 and how this information can be coupled with structural data in the PDB.
@@
 protein complexes.
-%This paves the way toward standardized methods to resolve the quality
-%evaluated conformational ensembles of disordered biomolecules in membrane environment.
-%In the NMRlipids project we aim to build a general databank of quality evaluated
-%conformational ensembles of lipids and other disordered biomolecules based primarily on
-%MD simulations and NMR data.
-
-% Tables may be be put in the text as floats.
-% Here is an example of the general form of a table:
-% Fill in the caption in the braces of the \caption{} command. Put the label
-% that you will use with \ref{} command in the braces of the \label{} command.
-% Insert the column specifiers (l, r, c, d, etc.) in the empty braces of the
-% \begin{tabular}{} command.
-%
-% \begin{table}
-% \caption{\label{} }
-% \begin{tabular}{}
-% \end{tabular}
-% \end{table}
 
@@
 \end{acknowledgments}
-%\newpage
-%\appendix
-%\begin{center}
-%{\bf SUPPLEMENTARY INFORMATION}
-%\end{center}
-
-%\section{Measurements of order parameter sign}
-
-%Fig. \ref{PShgSIGNS} summarizes the experimental results on the order parameter sign
-%measurement for POPS sample. The experimental protocol is the same used in Ref. \citenum{ferreira16}.
-%In (a) you see the headgroup region of the INEPT spectrum where alpha and beta are
-%identified. In (b) you have the R-PDLF slices for alpha and beta where you see one single
-%splitting for beta (which gives an order parameter equal to 0.12), and for alpha a superposition
-%of a large splitting (order parameter equal to 0.09) and a very small splitting which cannot be
-%calculated. On the bottom you have the S-DROSS slices of these two carbons. The grey lines show a
-%random collection of slices from noise such that it gets clear what is significant. The S-DROSS
-%slice for beta clearly shows that the order parameter is negative. The slice for alpha shows that
-%the higher order parameter is positive and suggests that the smaller order parameter is negative
-%(from the deviation towards negative values in the longer t1 times).
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=9.0cm]{../Figs/PShgSIGNS.pdf}
-%  \caption{\label{PShgSIGNS}
-%    Experimental results for sign measurement for POPS sample
-%  }
-%\end{figure}
-
-%The results updated with SIMPSON simulations for the SDROSS profiles
-%are shown in Fig. \ref{PShgSIGNSsimpson}. The value for the smaller
-%alpha order parameter is taken from Fig 3 in Ref. \citenum{roux91},
-%because resolution in 13C NMR experiments was nor high enough to determine
-%numerical value for this. The plots in Fig. \ref{PShgSIGNSsimpson} (c) show
-%the following. The error bars and points are the experimental SDROSS data.
-%The thick lines are SIMPSON simulations. The simulations were done by using
-%the order parameter for beta equal to -0.12 and for alpha one order parameter
-%equal to 0.09 and the other equal to -0.02 (black) or 0.02 (grey).
-%Since the black lines agree with experimental data, we conlude that
-%the order parameters for $\beta$ carbon are -0.12 and for $\alpha$
-%order parameters are 0.09 and -0.02.
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=9.0cm]{../Figs/PShgSIGNSsimpson.pdf}
-%  \caption{\label{PShgSIGNSsimpson}
-%    Experimental results for sign measurement for POPS sample
-%  }
-%\end{figure}
-
-%\section{Dihedrals}
-%\begin{figure*}[]
-%  \centering
-%  \includegraphics[width=8.0cm]{../Figs/dihed1.png}
-%  \includegraphics[width=8.0cm]{../Figs/dihed2.png}
-%  \includegraphics[width=8.0cm]{../Figs/dihed3.png}
-%  \includegraphics[width=8.0cm]{../Figs/dihed4.png}
-%  \includegraphics[width=8.0cm]{../Figs/dihed5.png}
-%  \includegraphics[width=8.0cm]{../Figs/dihed6.png}
-%  \includegraphics[width=8.0cm]{../Figs/dihed7.png}
-%  \includegraphics[width=8.0cm]{../Figs/dihed8.png}
-%  \includegraphics[width=8.0cm]{../Figs/dihed9.png}
-%  \includegraphics[width=8.0cm]{../Figs/dihed10.png}
-%  \caption{\label{dihedrals}
-%    Experimental results for sign measurement for POPS sample
-%  }
-%\end{figure*}
 

### 2021-01-05 manuscriptPGPEsuppl.tex
Fix figure paths and remove comments

\documentclass[journal=jpcbfk]{achemso}

\usepackage[version=3]{mhchem}
\usepackage[T1]{fontenc}
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
%\newcommand{\todo}[1]{\textcolor{red}{#1}}
\usepackage[obeyFinal]{easy-todo}
\usepackage{rotating}
\usepackage{upgreek}				
\usepackage{xcolor}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{xr}
\externaldocument{manuscriptPGPE}
\usepackage{soul} % for highlights with \hl{}

\author{Pavel Buslaev}
\affiliation{University of Jyv{\"a}skyl{\"a}}

\author{Rebeca Garc{\'i}a Fandi{\~n}o}
\affiliation{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}
\affiliation{CIQUP, Centro de Investigação em Química, Departamento de Química e Bioquímica, Faculdade de Ciências, Universidade do Porto, Porto, Portugal}

\author{Fernando Favela-Rosales}
\affiliation[Tecnol\'{o}gico Nacional de M\'{e}xico]{Departamento de Investigaci\'{o}n, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}

\author{Tiago Ferreira}
\affiliation{Halle, Germany}

\author{Patrick Fuchs}
\affiliation{Paris, France}

\author{Matti Javanainen}
\affiliation[Czech Academy of Sciences]{Institute of Organic Chemistry and Biochemistry of the 
Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}

\author{Jesper J. Madsen}
\affiliation[University of Chicago]{Department of Chemistry, The University of Chicago, Chicago, Illinois, United States of America}
\alsoaffiliation[University of South Florida]{Department of Global Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}

\author{Josef Melcr}
\affiliation[Czech Academy of Sciences]{Institute of Organic Chemistry and Biochemistry of the 
Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}
\alsoaffiliation{Groningen Biomolecular Sciences and Biotechnology Institute 
and The Zernike Institute for Advanced Materials, 
University of Groningen, 9747 AG Groningen, The Netherlands}

\author{Paula Milan Rodriguez}
\affiliation{Paris, France}

\author{Markus S. Miettinen}
% \affiliation[Max Planck Institute of Colloids and Interfaces]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
\affiliation{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}

\author{O. H. Samuli Ollila}
\email{samuli.ollila@helsinki.fi}
\affiliation{Institute of Biotechnology, University of Helsinki}

\author{Chris G. Papadopoulos}
\affiliation[]{I2BC - University Paris Sud}

\author{Antonio Pe{\'o}n}
\affiliation[]{Spain}

\author{Thomas J. Piggot}
\affiliation[University of Southampton]{Chemistry, University of Southampton, Highfield, Southampton SO17 1BJ, United Kingdom}

\author{{\'A}ngel Pi{\~n}eiro}
\affiliation{Departamento de F{\'i}sica Aplicada, Facultade de F{\'i}sica, Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}

%\author{O. H. Samuli Ollila}
%\email{samuli.ollila@helsinki.fi}
%%\homepage[]{Your web page}
%\affiliation{Institute of Organic Chemistry and Biochemistry,
%Academy of Sciences of the Czech Republic, 
%Prague 6, Czech Republic}
%\affiliation{Institute of Biotechnology, University of Helsinki}

\SectionNumbersOn

\renewcommand{\thetable}{S\arabic{table}}%
\renewcommand{\thefigure}{S\arabic{figure}}%
\renewcommand{\thesection}{S\arabic{section}}%
\renewcommand{\thepage}{S\arabic{page}}%

\title{ Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids }

\begin{document}

\newpage
%\tableofcontents

\section{R-PDLF and SDROSS experiments}

\begin{figure}[]
%  \centering
  \includegraphics[width=0.8\textwidth]{./Figs/POPEexperiment.pdf}
  \caption{\label{POPEspectra}
    (A) Chemical structure of POPE with the labeling of headgroup and glycerol backbone carbons.
    (B) INEPT spectra from POPE sample with the headgroup and glycerol backbone peaks labeled.
    (C) 2D R-PDLF spectra
    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
    (E) Experimetal S-DROSS curves giving signs of the order parameters.
  }
  \todo{A, B etc. labels to be put in the figure.}
\end{figure}

\begin{figure}[]
%  \centering
  \includegraphics[width=0.8\textwidth]{./Figs/PGexpRPDLF.eps}
  \caption{\label{POPGspectra}
    (A) Chemical structure of POPG with the labeling of headgroup and glycerol backbone carbons.
    (B) INEPT spectra from POPG sample with the headgroup and glycerol backbone peaks labeled.
    (C) 2D R-PDLF spectra
    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
    (E) Experimetal S-DROSS curves giving signs of the order parameters.
  }
\end{figure}

\begin{figure}[]
%  \centering
  \includegraphics[width=0.8\textwidth]{./Figs/SIMPSON.pdf}
  \caption{\label{POPGsimpson}
    Simpson simulaton of S-DROSS curve of $\beta$-carbon of POPG.
  }
\end{figure}

\clearpage
\section{Comparison of headgroup order parameters from different force fields to experiments}

The quality of PE and PG headgroup conformational ensembles in different simulations against NMR experiments
is evaluated in figures \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}
using C-H bond order parameters as in our previous studies for PC and PS lipids \cite{botan15,antila19}.
Conclusions are the same for all lipids: None of the force fields correctly captures the lipid headgroup
conformational ensembles, but CHARMM36 gives results closest to experiments.

%The poor performance of headgroup order parameters in Berger model can be probably explained by
%ring like structures seen in Fig. 6 in Ref. \citenum{mukhopadhyay04}, which is a typical feature
%for Berger based lipid force fields containing explicit hydrogen atoms in the head group \cite{zhao08,henin09,dahlberg10}.
%The poor performance of glycerol backbone of Slipids simulations is systemically observed also
%for other lipids in previous studies \cite{botan15,antila19}.
%\todo{Should we comment more the relative quality of different force fields and/or make the subjective force field ranking figures?
%https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/8}

\begin{figure}[]
  \centering
  \includegraphics[width=9.0cm]{./Figs/HGorderparametersPE.pdf}
  \caption{\label{HGorderParametersPE}
    The headgroup and glycerol backbone order parameters of PE lipids
    from experiments (POPE and signs this work, DPPE from Ref.~\citenum{seelig76}
    and E.coliPE from Ref.~\citenum{gally81}) and simulations with different force fields.
  }
  \todo{This should be clarified as in NMRlipidsI and error bars should be added.
    Probably larger error bars for united atom models based on the report by Fuchs et al.
  } \\
\end{figure}

\begin{figure}[!h]
  \centering
  \includegraphics[width=9.0cm]{./Figs/HGorderparametersPG.pdf}
  \caption{\label{HGorderParametersPOPG}
    The headgroup and glycerol backbone order parameters of PG lipids
    from experiments (POPG and signs from this work and from Ref.~\citenum{borle85}, %contains 10mM of PIPES,
    DPPG with 100mM NaCl from Ref.~\citenum{wohlgemuth80},% contains 10mM PIPES and,
    and E.Coli PG results from Ref.~\citenum{gally81})
    and simulations with different force fields.
  }
\end{figure}

\clearpage

\section{PC headgroup in mixtures with PE or PG lipids}
\begin{figure*}[]
  \centering
  \includegraphics[width=16.0cm]{./Figs/HGorderparametersPCvsPEPG.eps}
  \caption{\label{HGorderparametersPCvsPEPG}
    Modulation of POPC headgroup order parameters with increasing amount of POPE (left) and POPG (right) in bilayer
    from experiments at 298 K \cite{scherer87,macdonald87} and simulations with different force fields
    (temperatures listed in tables \ref{systemsMIX} and \ref{systemsMIX2} are between 298-310 K).
    Signs are determined as discussed in Refs. \citenum{botan15,ollila16}.
  }
\end{figure*}

Headgroup order parameters of PC lipids are unchanged upon addition
of zwitterionic lipids or cholesterol in experiments, but increase
upon addition of negatively charged PG or PS lipids because
headgroup dipole tilts more parallel to the membrane plane after incorporation
of negative charges into the membrane~\cite{seelig87, scherer87,antila18}.
The response of PC headgroup order parameters to the addition of PE or PG lipids
from different simulations is compared with experiments in figure~\ref{HGorderparametersPCvsPEPG}.
None of the simulations reproduce neither the experimentally observed increase in PC headgroup order parameters
with increasing amount of PG nor the related tilting of the headgroup more parallel with the membrane.
Similar observations in our previous work for PS lipids were explained by the overestimated counterion
binding affinity that neturalizes the effect of added negative charge \cite{antila19}.
\todo{Difference between Åqvist and Dang in Slipids results probably supports this,
but this has to checked from density profiles.} 
All simulations except Berger-OPLS predict tilting of P-N headgroup outwards from the membrane and
decrease of PC headgroup order parameters upon addition of PE lipids.
These results are not in line with experiments where the PC headgroup order parameters are not affected by zwitterionic lipids \cite{scherer87}.
The good performance of Berger-OPLS simulations in here is surprising because
headgroup conformational enemble is not very close to experiments in this model and
the response of headgroup order parameters
to cholesterol was significantly overestimated by the Berger/H{\"o}ltje force field in our previous work \cite{botan15}.

\clearpage
\section{PG headgroup in mixtures with PC lipids}
Changes in other than PC lipid headgroup with changing membrane composition are less
extensively characterized in the literature.
The $\beta$-carbon order parameter in PG headgroup increases
mildly \cite{macdonald87} or is unchanged~\cite{borle85} upon increasing amount
of PC lipids (Fig. \ref{HGorderparametersPGvsPCchange}), but experimental data from
$\alpha$-carbon is not available. Also the tested force fields predict very small changes
for the $\beta$-carbon order parameter, while the P-N vector tilt and its response to the
increased amount of PC varies significantly between force fields in figure \ref{HGorderparametersPGvsPCchange}. 
Therefore, more experimental data and more accurate force fields are still required
to resolve the PG conformational ensembles in mixtures with other lipids.
\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{./Figs/HGorderparametersPGvsPCchange.eps}
  \caption{\label{HGorderparametersPGvsPCchange}
    Modulation of PG lipid headgroup order parameters with the increasing amount of PC in lipid bilayer
    from experiments at 298 K \cite{borle85,macdonald87} and simulations with different force fields at 310 K.
  }
\end{figure}

\clearpage
\section{Calcium binding to POPC:POPG mixtures}

\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{./Figs/CAdensPG1PC1.eps}
  \caption{\label{CAdensPG}
    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (1:1) mixtures with different force fields.
    The changes in the order parameters upon addition of CaCl$_2$ are compared with experiments in figure \ref{changesWITHCaClPG} in the main text.
  }
\end{figure}

\begin{figure*}[]
  \centering
  \includegraphics[width=18.0cm]{./Figs/CHANGESwithCaClPG1PC4.eps}
  \caption{\label{changesWITHCaClPG1PC4}
    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (4:1)
    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{macdonald87} and simulations.
    The changes with respect to the systems without CaCl$_2$ are shown for other data than
    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
  }
\end{figure*}

\begin{figure}[]
  \centering
  \includegraphics[width=8.0cm]{./Figs/CAdensPG1PC4.eps}
  \caption{\label{CAdensPG1PC4}
    Calcium ion density profiles along membrane normal from simulations of POPC:POPG (4:1) mixtures with different force fields.
  }
\end{figure}

\clearpage
\section{Changes in headgroup conformations upon addition of CaCl$_2$}

\begin{figure}[]
  \centering
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17ecc.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
  \caption{\label{DIHSwithCAlipid17eccPOPC}
    Changes in POPC lipid17ecc dihedrals with increasing amount of CaCl$_2$.
  }
\end{figure}

\begin{figure}[]
  \centering
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
  \caption{\label{DIHSwithCAcharmm36POPC}
    Changes in POPC CHARMM36 dihedrals with increasing amount of CaCl$_2$.
  }
\end{figure}

%\begin{figure}[]
%  \centering
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MSlipids.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
%  \caption{\label{DIHSwithCA}
%    Changes in POPC Slipids dihedrals with increasing amount of CaCl$_2$.
%  }
%\end{figure}

\begin{figure}[]
  \centering
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MSlipids.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
  \caption{\label{DIHSwithCAslipidsPOPG}
    Changes in POPG Slipids dihedrals with increasing amount of CaCl$_2$.
  }
\end{figure}

%\begin{figure}[]
%  \centering
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
%  \caption{\label{DIHSwithCA}
%    Changes in POPC lipid17 dihedrals with increasing amount of CaCl$_2$.
%  }
%\end{figure}

\begin{figure}[]
  \centering
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
  \caption{\label{DIHSwithCAlipid17POPG}
    Changes in POPG lipid17 dihedrals with increasing amount of CaCl$_2$.
  }
\end{figure}

%\begin{figure}[]
%  \centering
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17ecc.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
%  \caption{\label{DIHSwithCA}
%    Changes in POPG lipid17ecc dihedrals with increasing amount of CaCl$_2$.
%  }
%\end{figure}

\clearpage
\section{Simulated systems}

%\begin{table*}[htb]
%
\begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with PE lipids.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systemsPE}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      POPE  & CHARMM36 \cite{??}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
      POPE  & CHARMM36 \cite{??}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
      POPE  & CHARMM36 \cite{??}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
      POPE  & CHARMM36ua \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
      \hline
      DPPE  & Slipids \cite{jambeck12b}    &0    & 288 	& 9386  &0    & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
      POPE  & Slipids \cite{jambeck12b} &0    & 336	& ?     &0    & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
      POPE  & Slipids \cite{jambeck12b}            & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
      \hline
      DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3655  &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
      POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
      POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
      \hline
      POPE  & GROMOS 43A1-S3 \cite{??}     &0    & 128	& 3552     &0    & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles}  \\
      \hline
      POPE  & OPLS-UA vdW on H \cite{??}   &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
      POPE  & OPLS-UA \cite{??}            &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \\
      \hline
      POPE  & OPLS-MacRog \cite{rog16}     &0    & 144	& 5760     &0    & 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
      \hline
      POPE  & Berger-Vries \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
      POPE  & Berger-largeH \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
      DOPE  & Berger-Vries \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
      DOPE  & Berger-largeH \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
      \hline
      POPE             & LIPID17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
    \end{tabular}
  \end{minipage}
  \todo{Simulations with added NaCl are not currently used here, maybe should be removed from the table?} \\
  \todo{Citation for CHARMM36 PE?} \\
  \todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
  \todo{Citation for GROMOS-CKP?} \\
  \todo{Citation for GROMOS 43A1-S3?} \\
  \todo{Citation for OPLS-UA models?} \\
  \todo{Citations for Berger-* simulations?} \\
  \todo{LIPID17 simulations with correct dihedrals still coming}
%\end{table*}
  \end{sidewaystable*}

  %    \begin{table*}[htb]
  \begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with PG lipids.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systemsPG}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}    &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
      POPG             & CHARMM36 \cite{??}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
      POPG             & CHARMM36 \cite{??}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
      \hline
      POPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 10664   &0     & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232  &0     & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232   &0     & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0    & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGslipids} \\
      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
      \hline
      POPG             & LIPID17 / Dang \cite{gould18,smith94}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
      POPG             & LIPID17 \cite{??}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
      \hline
      POPG             & GROMOS-CKP \cite{??}         & 0              & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP} \\
      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 49 &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
    \end{tabular}
  \end{minipage}
    \todo{Simulations with added NaCl are not currently used here, maybe should be removed from the table?} \\
    \todo{Citations and ion model for CHARMM36?} \\
  \todo{Lipid17 simulation with ions with correct dihedral potentials still coming?} \\
  \todo{Citation and ion model for GROMOS-CKP?}
%\end{table*}
 \end{sidewaystable*}

%\begin{table*}[htb]
  \begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with PE and PG lipids mixed with PC.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systemsMIX}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
%      POPC                   & CHARMM36 \cite{??}        & 0.11      & 0  & 500     & 25000 & 48  &  310  & 500 & 100 & \cite{POPCcharmm150mMNaCl301K}  \\
%      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
      POPC                   & CHARMM36 \cite{??}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
      \hline
      POPC             & CHARMM36 \cite{??}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
      POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
      \hline
      POPC             & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPCmacrog300K} \\
      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1macrogT300K} \\
      \hline
      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  298  & 170 & 100 & \cite{POPCslipid298K} \\
      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120  & 0  &  298  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\
     \hline
      POPC                   & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCgromosCKPT310K}  \\
      POPC:POPG (7:3)        & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG3gromosCKPT310K} \\
     \hline
      POPC                   & Slipid \cite{jambeck12b}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
      POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
    \end{tabular}
  \end{minipage}
    \todo{Citation and ion model for GROMOS-CKP?} \\
  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
%\end{table*}
\end{sidewaystable*}

  \begin{sidewaystable*}[!p]
  \centering
  \caption{List of MD simulations with PE and PG lipids mixed with PC.
    % The salt concentrations calculated as [salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M.
    % these correspond the concentrations reported in the experiments by Akutsu et al.~\cite{akutsu81}.
    % The lipid force fields named as in our previous work~\cite{botan15}.
  }\label{systemsMIX2}
  \begin{minipage}[t]{\textwidth}
    \begin{tabular}{l c c r r r r r r c c}
      %\hline
      % some footnotes are not visible in typeset-MS (pdf)
      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
      \hline
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 150:150 & 31401 & 57 &  298  & 718 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
      \hline
      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020} \\
      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020100mMCaCl} \\
      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 350:88 & 24840 & 475 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG80201000mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG5050100mMCaCl} \\
      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 150:150 & 29865 & 569 &  298  & 600 & 400 & \cite{Lipid17eccPOPCPOPG50501000mMCaCl} \\
      \hline
      POPC             & Berger \cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  &0  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
      POPC:POPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
      POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
     \hline
      DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
      DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}  &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
    \end{tabular}
  \end{minipage}
  \todo{Citation and description for ''Berger'' model?} \\
  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
%\end{table*}
\end{sidewaystable*}

\subsection{CHARMM36}

\noindent {\it POPE} \todo{Simulation details by M. Javanainen}.

\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.}

\noindent {\it POPG} \todo{Simulation details by Ollila}.

\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.}

\noindent {\it POPC:POPE mixtures} Data is available at \cite{POPCcharmm300K,POPC1POPE1charmm36}.
300 K with v-rescale (tau=0.1 ps),
1 bar with PR semiisotropic (tau=4 ps, compressibility=4.5e-5 bar$^{-1}$),
PME order 4 and space 0.12,
rcoulomb and rvdw 1.0,
128 lipids per leaflet,
no ion \todo{Full simulation details by Fuchs et al.}

\noindent {\it POPC:POPG mixture with additional calcium} \todo{Simulation details by A. Kiirikki.}

\noindent {\it POPC and POPC:POPG (7:3) mixture} \todo{Simulation details by A. Peon.}

\subsection{CHARMM36ua}

\noindent {\it POPE} Data is available at \cite{charmm36uaPOPEfiles}. \todo{Simulation details by  T. Piggot.}

%\subsection{MacRog}

%\subsection{Lipid17}

\subsection{Slipids}
\noindent {\it POPE} Data is available at \cite{slipidsPOPEfiles}. \todo{Simulation details by  T. Piggot.}

\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.
  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.
}

\noindent {\it DPPE} Data is available at \cite{slipidsDPPEfiles}. \todo{Simulation details by F. Favela.}

\noindent {\it POPG} Data is available at \cite{slipidsPOPGfiles}. \todo{Simulation details by F. Favela.
  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}

\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.
  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}

\noindent {\it DPPG} Data in 298~K is available at \cite{slipidsDPPGfilesT298K} and in 314~K at \cite{slipidsDPPGfiles}.
\todo{Simulation details by F. Favela.
  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}

\noindent {\it POPC:POPG mixture with additional NaCl} \todo{Simulation details by A. Peon.
  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}

\noindent {\it POPC:POPG mixture with additional CaCl} \todo{Simulation details by M. Javanainen.}

\subsection{Berger}

\noindent {\it POPE} Data is available at \cite{bergerPOPEfiles,berger2POPEfiles}. \todo{Simulation details by  T. Piggot.}

\noindent {\it DOPE} Data is available at \cite{bergerDOPEfiles,berger2DOPEfiles}. \todo{Simulation details by  T. Piggot.}

\noindent {\it POPC:POPE, POPC:DOPE and DOPC:DOPE mixtures} Data is available at \cite{POPCberger300K,POPC1POPE1berger}. 
300 K with v-rescale (tau=0.1 ps),
1 bar with PR semiisotropic (tau=4 ps, compressibility=4.5e-5 bar$^{-1}$),
PME order 4 and space 0.12,
rcoulomb and rvdw 1.0,
128 lipids per leaflet,
no ion \todo{Simulation details by Fuchs et al.}

\subsection{GROMOS 43A1-S3}

\noindent {\it POPE} Data is available at \cite{gromos43a1s3POPEfiles}. \todo{Simulation details by  T. Piggot.}

\subsection{OPLS-UA}

\noindent {\it POPE} Data is available at \cite{OPLSuaPOPEfiles}. \todo{Simulation details by  T. Piggot.}

\noindent {\it POPE with vdW interaction in H} Data is available at \cite{OPLSuaWvdWPOPEfiles}. \todo{Simulation details by  T. Piggot.}

\subsection{GROMOS-CKP and GROMOS-CKPM}

\noindent {\it POPE} Data is available at \cite{gromosCKPpope}. \todo{Simulation details by  T. Piggot.}

\noindent {\it DOPE} Data is available at \cite{gromosCKPdope}. \todo{Simulation details by  T. Piggot.}

\noindent {\it DPPE} Data is available at \cite{gromosCKPdppe}. \todo{Simulation details by  T. Piggot.}

\noindent {\it POPG} \todo{Simulation details by A. Peon.}

\noindent {\it POPC:POPG mixture} \todo{Simulation details by A. Peon.}

\subsection{OPLS-MacRog}

\noindent {\it POPE} \todo{Simulation details by M. Javanainen and P. Fuchs.}

\noindent {\it POPC:POPE mixtures} \todo{Simulation details by P. Fuchs.}

\subsection{Lipid17}

\noindent {\it POPE} \todo{Simulation details by A. Peon.}

\noindent {\it POPG} \todo{Simulation details by A. Peon.}

\noindent {\it POPC:POPG} \todo{Simulation details by S. Virtanen or O. H. S. Ollila.}

\subsection{Lipid17ecc}

\todo{This is to be finished and POPC:POPG mixtures to be described}
In ECC-lipid models, electronic continuum correction (ECC) is applied to 
implicitly include the missing electronic polarizability into the force field description \cite{melcr18,melcr19}.
In practise, this is implemented by scaling the charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
of Amber Lipid14/17 models are scaled by constant factors.
Here, we follow the approach that previously improved ion binding to bilayers containing negatively charged PS lipids \cite{melcr19}:
ECC-POPC parameters (scaling factors $f_q$=0.8 and $f_\sigma$=0.89 applied to Lipid14 POPC parameters) \cite{melcr18} were used for POPC
and scaling factors of $f_q$=0.75 and $f_\sigma$=0.89 were applied to the charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
of Amber Lipid17 POPG parameters. The Lipid17 parameters (described above) and initial configurations
were taken from Ref. \cite{POPGlipid17} with the correct dihedral type,
and the resulting parameters are available from Ref. \citenum{??}. ECC-ion parameters with the scaled charges \cite{pluharova14,kohagen16,martinek18},
downloaded from \url{bitbucket.org/hseara/ions/src/master/}, were used in these simulations.

\clearpage
\bibliography{refs.bib}

\end{document}

### 2021-01-05 manuscriptPGPE.tex
Small updates in the text

@@
 % insert abstract here
-  Chemistry of lipid headgroups, the water facing components of cell membranes, varies between organisms and
-  organelles and regulate cell functions via lipid-protein interactions.
-  Because lipid membranes are in liquid state in physiological conditions,
-  the conformational ensembles of lipids and their dependence of headgroup chemistry are difficult to resolve experimentally.
-  Here, we combine solid state NMR experiments and MD simulations from NMRlipids open collaboration to
-  resolve conformational ensembles of PC, PS, PG and PE lipid headgroups in liquid lamellar phase in various biologically relevant conditions.
-  Interpretation of NMR experiments using the best simulations from the NMRlipids project suggests that all lipid headgroups sample wide range of
-  conformations in neutral and charged cell membranes, and that the differences in headgroup chemistry
-  appears in the probability weights of different conformations. Together with the analysis of protein-bound-lipids
-  from the protein data bank (PDB), this suggests that lipids can bind to proteins in wide range of conformations
-  independently on the headgroup chemistry. Therefore, the specific lipid-protein interactions are likely driven by
-  protein-lipid interactions, rather than conformational restrictions of lipids. Our results paves the way to
-  comprehensive understanding of lipid-protein interaction energetics, and understanding complex biomolecular assemblies
-  such as membrane proteins.
+  Chemistry of lipid headgroups, the water facing components of cell membranes that regulate cell functions via lipid-protein interactions, varies between organisms and organelles.
+Because lipid membranes are in liquid state in physiological conditions, the conformational ensembles of lipid headgroups and their dependence on chemistry are difficult to resolve experimentally.
+Here, we combine solid state NMR experiments and MD simulations from NMRlipids open collaboration to resolve conformational ensembles of PC, PS, PG and PE lipid headgroups in liquid lamellar phase in various biologically relevant conditions.
+Interpretation of NMR experiments using the best simulations from the NMRlipids project suggests that all lipid headgroups sample wide range of conformations in neutral and charged cell membranes, and that the differences in headgroup chemistry appears in the probability weights of different conformations.
+Together with the analysis of protein-bound-lipids from the protein data bank (PDB), this suggests that lipids can bind to proteins in wide range of conformations independently on the headgroup chemistry.
+Therefore, the specific lipid-protein interactions are likely driven by protein-lipid interactions, rather than conformational restrictions of lipids.
+Our results paves the way to comprehensive understanding of lipid-protein interaction energetics, and understanding complex biomolecular assemblies such as membrane proteins.
 
@@
 The NMR experiments were identical as in our previous work~\cite{antila19}.
-POPE and POPG powder were purchased from Avanti polar lipids \todo{Which isomer of POPG we have in the experiments?}.
+POPE and POPG powder were purchased from Avanti polar lipids 
+\todo{POPG is racemic in experiments. How much we should discuss this in relation with simulations?}.
 The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.
@@
 The $\beta$-carbon peak from POPG overlapped with the g$_2$ peak from glycerol backbone
-because their chemical environments are similar in POPG.
+because their chemical environments are similar.
 \todo{Details to be checked by Tiago}.
@@
 are listed in Tables \ref{systemsPE} and \ref{systemsPG},
-and lipid mixtures with additional ions in Table~\ref{systemsMIX}.
+and lipid mixtures with additional ions in Tables~\ref{systemsMIX} and~\ref{systemsMIX2}.
 Further simulation details are given in the SI, and
@@
 \subsection{Analysis of molecular dynamics simulation data}
-The big data set of MD simulations was analysed in the NMRlipids databank manner.
+The large set of MD simulation data was analysed using the development version of NMRlipids databank.
 Unique naming convention for lipid atoms in each force field was defined using the mapping files
@@
 with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot17}.\\
-\todo{BuildH program is now cited with a direct link to the GitHub repo. I think that a release to Zenodo would be nice in the final publication.}\\
-\todo{Maybe we should also shortly discuss here about the reasons for slight dependence of order parameter values on the method used to reconstruct hydrogens?}\\
+%\todo{BuildH program is now cited with a direct link to the GitHub repo. I think that a release to Zenodo would be nice in the final publication.}\\
 The ion number density profiles were calculated using the {\tt gmx density} tool
@@
 PSF, PS6, Q3G, P5S, D39, PS2, 17F, and 8SP for PS.
-All PDB structures containing these ligands were downloaded and the dihedral angles of
-the first lipid in the latest version of the structures within each PDB code were calculated  
-using MDTraj python library \cite{mcgibbon15}.
+All PDB structures containing these ligands were downloaded and the dihedral angles of the first lipid in the latest version of the structures within each PDB code were calculated using the MDTraj python library \cite{mcgibbon15}.
 The used Jupyter notebook is available from the project's GitHub repository ({\it scripts/pdbSEARCH.ipynb}).
@@
 
-\begin{figure*}[]
+\begin{figure*}[bt]
   \centering
@@
 
-\subsection{Conformational ensembles of different lipid headgroups in bulk bilayer}
+\subsection{Differences between lipid headgroups in bulk bilayer from $^{13}$C NMR experiments}
 
@@
 
-Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and
-headgroups from this and previous studies are collected in Fig. \ref{HGorderParameters} D),
-where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
-Overall agreement of order parameters determined by different authors and different techniques
-for same lipid headgroup is very good in here and previous studies \cite{botan15,ollila16,antila19},
-suggesting that the observed differences between lipid types arise from differences in
-headgroup chemistry rather than inaccuracies in experiments. 
-The most distinct order parameters are observed for PS headgroups,
-for which the $\alpha$-carbon order parameter exhibits significant forking
-and the $\beta$-carbon has more negative value than in other lipids.
-On the other hand, the $\beta$-carbon order parameter of PG headgroup
-has positive sign in contrast to all the other lipids.
-Notably, this has not been observed in traditional $^2$H NMR experiments,
-where only the absolute value of the order parameters are measured~\cite{wohlgemuth80,gally81,borle85}.
-The glycerol backbone order parameters are similar for all the lipids, although they move slightly toward
-positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG.
+Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.
+\ref{HGorderParameters} D), where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
+Overall agreement of order parameters determined by different authors and different techniques for same lipid headgroup is very good in here and previous studies \cite{botan15,ollila16,antila19},
+suggesting that the observed differences between lipid types arise from differences in headgroup chemistry rather than inaccuracies in experiments, or differences in acyl or experimental conditions.
+The most distinct order parameters are observed for PS headgroups, for which the $\alpha$-carbon order parameter exhibits significant forking and the $\beta$-carbon has more negative value than in other lipids.
+On the other hand, the $\beta$-carbon order parameter of PG headgroup has positive sign in contrast to all the other lipids.
+Notably, this has not been observed in traditional $^2$H NMR experiments, where only the absolute value of the order parameters are measured~\cite{wohlgemuth80,gally81,borle85}.
+The glycerol backbone order parameters are similar for all the lipids, although they move slightly toward positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG.
 Essential differences between PC and PE headgroups are not observed.
+
+\subsection{Conformational ensembles of different lipid headgroups in bulk bilayer from MD simulations}
 
@@
     {\bf A)} Modulation of PC headgroup order parameters and P-N vector angle upon addition of cationic surfactant
-    from CHARMM36 simulations compared with experimental data \cite{??}.
-    {\bf B)} Changes in PC headgroup conformational ensembles upon increasing amount of positive charge in bilayer,
+    from CHARMM36 simulations compared with the experimental data \cite{scherer89}.
+    {\bf B)} Changes in PC headgroup conformational ensembles upon increasing the amount of positive charge in bilayer,
     characterized by the heavy atom dihedral distributions, from CHARMM36 simulations.
@@
 
-To resolve the conformational ensembles of lipid headgroups,
-we compared the headgroup and glycerol backbone C-H bond order parameters
-from different MD simulation force fields to the experimental data. 
-None of the force fields reproduced the order parameters
-within experimental accuracy for any of the lipids analysed in this work
-(Figs. \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}) or in our previous studies \cite{botan15,antila19}.
-Nevertheless, the CHARMM36 force field performs best for all the headgroups and
-reproduces the distinct order parameters of PS and PG lipids (Fig. \ref{structures} B).
-To understand the structural origin of these distinct order parameters,
-we calculated the distributions of heavy atom dihedral angles from CHARMM36 simulations in figure \ref{structures} C).
-Major differences between headgroups are observed only for the last two dihedrals
-in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$,
-which prefer {\it gauche$^-$} conformations for PG and {\it gauche$^+$} for PS.
-Rest of the dihedrals are similar between different lipids, with the exception of
-small differences for PS lipids. Therefore we suggest that the main differences between
-lipid headgroups leading to distinct order parameter occur in the choline part, while
-also changes in phosphate region may contribute in PS lipids, which could also
-explain the more rigid headgroup structures~\cite{browning80,buldt81}.
-Furthermore, the angle between headgroup dipole and membrane normal
-decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig. \ref{structures} B)).
+To resolve the conformational ensembles of lipid headgroups, we compared the headgroup and glycerol backbone C-H bond order parameters from different MD simulation force fields to the experimental data.
+None of the force fields reproduced the order parameters within experimental accuracy for any of the lipids analysed in this work (Figs.
+\ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}) or in our previous studies \cite{botan15,antila19}.
+Nevertheless, the CHARMM36 force field performs best for all the headgroups and reproduces the experimentally observed distinct order parameters of PS and PG lipids (Fig.
+\ref{structures} B).
+To understand the structural origin of these distinct order parameters, we calculated the distributions of heavy atom dihedral angles from CHARMM36 simulations in figure \ref{structures} C).
+Major differences between headgroups are observed only for the last two dihedrals in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$, which prefer {\it gauche$^-$} conformations for PG and {\it gauche$^+$} for PS.
+Rest of the dihedrals are similar between different lipids, with the exception of small differences for PS lipids.
+Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids, which could also explain the more rigid headgroup structures~\cite{browning80,buldt81}.
+Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.
+\ref{structures} B)).
 However, the differences between PC and PE in P-O$_\alpha$-C$_\alpha$-C$_\beta$ dihedral
-and P-N vector dipole may be an artefacts as the $\beta$-carbon order parameter in PC
-is too negative the CHARMM36 force field, thereby not being equal to the order parameter
-in PE as observed in experiments \cite{botan15}.
+and P-N vector dipole may be artificial as the $\beta$-carbon order parameter in PC is too negative in the CHARMM36 force field, thereby not being equal to the order parameter in PE as observed in experiments \cite{botan15}.
 
@@
 
-Also the headgroup order parameters of charged lipids are affected by the binding of ions,
-but their response is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
-In our simulations, the Lipid17 and Slipids force fields give the most realistic response of PG $\beta$-carbon to CaCl$_2$
-in figure \ref{changesWITHCaClPG} even thought the calcium binding affinity was overestimated in these models.
-These models suggest that the upward tilting of the headgroup dipole is weaker in PG than in PC headgroup (\ref{changesWITHCaClPG}), but dihedral distributions are more sensitive to the bound ions in PG
-(Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}). However, none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids
-simultaneously, experimental data to evaluate the response of $\alpha$-carbon order parameters to
-the added calcium in PG is not available, and changes in dihedrals upon addition of cacium differ between
-the best models (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+Also the headgroup order parameters of charged lipids are affected by the binding of ions, but their response is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
+In our simulations, the Lipid17 and Slipids force fields give the most realistic response of PG $\beta$-carbon to CaCl$_2$ in figure \ref{changesWITHCaClPG} even thought the calcium binding affinity was overestimated in these models.
+These models suggest that the upward tilting of the headgroup dipole is weaker in PG than in PC headgroup (\ref{changesWITHCaClPG}), but dihedral distributions are more sensitive to the bound ions in PG (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+However, none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids simultaneously, experimental data to evaluate the response of $\alpha$-carbon order parameters to the added calcium in PG is not available, and the changes in dihedrals upon addition of cacium differ between the best models (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
 Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.

### 2021-01-08 manuscriptPGPEsuppl.tex
Update density profiles

@@
 binding affinity that neturalizes the effect of added negative charge \cite{antila19}.
-\todo{Difference between Åqvist and Dang in Slipids results probably supports this,
-but this has to checked from density profiles.} 
 All simulations except Berger-OPLS predict tilting of P-N headgroup outwards from the membrane and

### 2021-01-12 manuscriptPGPEsuppl.tex
Updates from Overleaf
Update dihedral figure

@@
 Czech Academy of Sciences, Flemingovo n\'{a}m. 542/2, CZ-16610 Prague 6, Czech Republic}
+
+\author{Anne M. Kiirikki}
+\affiliation{Institute of Biotechnology, University of Helsinki}
 
@@
   } \\
+  \todo{Lipid17 data should be updated to the one with correct dihedrals reported here https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12\#issuecomment-756641407}
 \end{figure}
@@
   }
+   \todo{Lipid17 data should be updated to the one with correct dihedrals reported here https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12\#issuecomment-756641407}
 \end{figure}
@@
   }
+  \todo{There is something wrong the Lipid17ecc data at 1M, a new simulation is coming.} \\
 \end{figure*}
@@
   }
+ \todo{Density profiles from Lipid17ecc data will be added when we have the new simulations at 1M.} \\
 \end{figure}
@@
   \centering
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MSlipids.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
+  \includegraphics[width=18.0cm]{../Figs/DIHEDRALSslipidsWITHCaCl.eps}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MSlipids.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
   \caption{\label{DIHSwithCAslipidsPOPG}

### 2021-01-12 manuscriptPGPE.tex
Updates from Overleaf

@@
 
+\author{Anne M. Kiirikki}
+\affiliation{Institute of Biotechnology, University of Helsinki}
+
 \author{Jesper J. Madsen}
@@
 % insert abstract here
-  Chemistry of lipid headgroups, the water facing components of cell membranes that regulate cell functions via lipid-protein interactions, varies between organisms and organelles.
-Because lipid membranes are in liquid state in physiological conditions, the conformational ensembles of lipid headgroups and their dependence on chemistry are difficult to resolve experimentally.
-Here, we combine solid state NMR experiments and MD simulations from NMRlipids open collaboration to resolve conformational ensembles of PC, PS, PG and PE lipid headgroups in liquid lamellar phase in various biologically relevant conditions.
-Interpretation of NMR experiments using the best simulations from the NMRlipids project suggests that all lipid headgroups sample wide range of conformations in neutral and charged cell membranes, and that the differences in headgroup chemistry appears in the probability weights of different conformations.
-Together with the analysis of protein-bound-lipids from the protein data bank (PDB), this suggests that lipids can bind to proteins in wide range of conformations independently on the headgroup chemistry.
-Therefore, the specific lipid-protein interactions are likely driven by protein-lipid interactions, rather than conformational restrictions of lipids.
-Our results paves the way to comprehensive understanding of lipid-protein interaction energetics, and understanding complex biomolecular assemblies such as membrane proteins.
+  Chemistry of lipid headgroups, the water facing components of cell membranes that regulate cell functions via lipid--protein interactions, varies between organisms and organelles.
+Because lipid membranes are in liquid state under physiological conditions, the conformational ensembles of lipid headgroups and their dependence on chemistry are difficult to resolve experimentally.
+Here, we combine solid state NMR experiments and molecular dynamics simulations from NMRlipids open collaboration to resolve the conformational ensembles of the headgroups of key lipid types in liquid lamellar phase under various biologically relevant conditions.
+Interpretation of NMR experiments using the plethora of simulation data collected in the NMRlipids project suggests that all lipid headgroups sample a wide range of conformations in neutral and charged cellular membranes.
+However, the populations of different conformations are dictated by the headgroup chemistry.
+Together with the analysis of protein-bound lipids from the protein data bank (PDB), this suggests that lipids can bind to proteins in wide range of conformations independently on the headgroup chemistry.
+Therefore, the selective adsorption of proteins to membranes is likely regulated by specific protein--lipid interactions, rather than conformational restrictions of lipids.
+Our results pave the way to comprehensive understanding of lipid--protein interaction energetics, and the understanding of complex biomolecular assemblies such as membrane proteins.
 
@@
 conformational states between lipid types or
-by specific intermolecular lipid-protein interactions.
+by specific intermolecular lipid--protein interactions.
 
-Lipid conformational ensembles in physiologically relevant lamellar liquid phase
+The conformational ensembles of lipids in the physiologically relevant lamellar liquid phase
 are typically derived from NMR experiments, particularly from 
-C-H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}
-which can be detected even from cells \cite{gally81,scherer87,seelig90}.
-These studies suggest that the glycerol backbone structures are largely similar irrespectively of the headgroup \cite{gally81}, 
-and the headgroup structures are similar in PC, PE and PG lipids, while headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
-However, these results are based only on the absolute values while 
-the necessity of order parameters signs in capturing the lipid conformational ensembles has been resently demonstrated \cite{botan15,ollila16,ferreira16}.
+C--H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}. Notably, such measurements can be performed also on living cells \cite{gally81,scherer87,seelig90}.
+These studies suggest that the glycerol backbone conformations are largely similar irrespectively of the headgroup \cite{gally81}, 
+and the headgroup conformations are similar in PC, PE and PG lipids, while the headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
+However, these tendencies are based only on the absolute values, whereas
+the necessity of order parameter signs in capturing the conformational ensembles of lipids has been recently demonstrated \cite{botan15,ollila16,ferreira16}.
 Furthermore, the detailed understanding of lipid conformational ensembles
-is limited by the lack of universal models to interpret the experiments \cite{pezeshkian18,akutsu20}.
+is limited by the lack of universal models that would map order parameters to structural ensembles / that would help interpret the experiments \cite{pezeshkian18,akutsu20}.
 

### 2021-01-12 manuscriptPGPE.tex
Update dihedral figures, move ion binding to SI, and update discussion

--- 2021-01-05 manuscriptPGPE.tex
@@
 
-\begin{figure*}[]
-  \centering
-  \includegraphics[width=18.0cm]{./Figs/CHANGESwithCaClPG1PC1.eps}
-  \caption{\label{changesWITHCaClPG}
-    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
-    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
-    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
-    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
-    The changes with respect to the systems without CaCl$_2$ are shown for other data than
-    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
-    Calsium density distributions are shown in figure \ref{CAdensPG}.
-    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
-    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
-  }
-\end{figure*}
-
 To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge, we first analyzed CHARMM36 simulations which correctly captures the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
@@
 Also binding of ions may affect the lipid headgroup conformational ensembles in physiological conditions.
-The resolve this effect, we compared the changes of headgroup order parameters
-in POPC:POPG mixtures  upon addition of CaCl$_2$ between different simulations and experiments \cite{borle85,macdonald87} in figures \ref{changesWITHCaClPG} (molar ratio 1:1) and \ref{changesWITHCaClPG1PC4} (molar ratio 4:1).
-The results are in line with our previous studies:
-most force fields overestimate the calcium binding \cite{catte16,antila19}, but CHARMM36 with the NBfix correction underestimates the binding affinity \cite{antila19}, and the implicit inclusion of electronic polarizability using the electronic continuum correction (ECC) improves the results \cite{melcr18,melcr20}.
-The most realistic response of PC lipid headgroup order parameters to the calcium binding
-is observed in the Lipid17ecc simulations with the ECC included, where the decreased probability for trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral is the main structural consequence of bound ions (Fig.
-\ref{DIHSwithCAlipid17eccPOPC}).
-Similar but smaller effect is seen in CHARMM36 simulations with less bound calcium (Fig.
-\ref{DIHSwithCAcharmm36POPC}).
-
-Also the headgroup order parameters of charged lipids are affected by the binding of ions, but their response is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
-In our simulations, the Lipid17 and Slipids force fields give the most realistic response of PG $\beta$-carbon to CaCl$_2$ in figure \ref{changesWITHCaClPG} even thought the calcium binding affinity was overestimated in these models.
-These models suggest that the upward tilting of the headgroup dipole is weaker in PG than in PC headgroup (\ref{changesWITHCaClPG}), but dihedral distributions are more sensitive to the bound ions in PG (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
-However, none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids simultaneously, experimental data to evaluate the response of $\alpha$-carbon order parameters to the added calcium in PG is not available, and the changes in dihedrals upon addition of cacium differ between the best models (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+In line with the results from cationic surfactants, 
+the decreased probability for the trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral is the main structural consequence of
+bound ion to PC headgroup (Fig. \ref{DIHSwithCAlipid17eccPOPC} and \ref{DIHSwithCAcharmm36POPC}) in
+Lipid17ecc (having the most realistic response of PC lipid headgroup order parameters to the calcium binding in Fig. \ref{changesWITHCaClPG})
+and CHARMM36 simulations.
+%simulations with the ECC included, where
+%Similar but smaller effect is seen in CHARMM36 simulations with less bound calcium (Fig. \ref{DIHSwithCAcharmm36POPC}).
+%Also the headgroup order parameters of
+Response of charged lipids to the ion binding is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
+The most realistic simulations (Lipid17 and Slipids in figure \ref{changesWITHCaClPG}) suggest
+%istic response of PG $\beta$-carbon to CaCl$_2$  even thought the calcium binding affinity was overestimated in these models.
+%These models suggest
+that the upward tilting of the headgroup dipole upon addition of CaCl$_2$ is weaker in PG than in PC headgroup,
+%(\ref{changesWITHCaClPG}),
+but dihedral distributions are more sensitive to the bound ions in PG (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+However, the changes in dihedrals upon addition of calcium differ between the best models (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}),
+none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added calcium in PG is not available.
 Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.

### 2021-01-12 manuscriptPGPEsuppl.tex
Update dihedral figures, move ion binding to SI, and update discussion

@@
 
+The changes of headgroup order parameters in POPC:POPG mixtures upon addition of CaCl$_2$ between different simulations
+and experiments \cite{borle85,macdonald87} are compared in figures \ref{changesWITHCaClPG} (molar ratio 1:1) and \ref{changesWITHCaClPG1PC4} (molar ratio 4:1).
+The results are in line with our previous studies:
+most force fields overestimate the calcium binding \cite{catte16,antila19}, but CHARMM36 with the NBfix correction underestimates the binding affinity \cite{antila19}, and the implicit inclusion of electronic polarizability using the electronic continuum correction (ECC) improves the results \cite{melcr18,melcr20}.
+
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=18.0cm]{./Figs/CHANGESwithCaClPG1PC1.eps}
+  \caption{\label{changesWITHCaClPG}
+    Modulation of headgroup order parameters of POPC ({\it left}) and POPG ({\it right}) in POPC:POPG (1:1)
+    mixture upon addition of CaCl$_2$ in 298 K temperature from experiments \cite{borle85,macdonald87} and simulations.
+    The $\beta$-carbon order parameter of POPC (dashed line on top left) is not directly measured but
+    calculated from empirical relation $\Delta S_{\beta}=0.43\Delta S_{\alpha}$ \cite{akutsu81}.
+    The changes with respect to the systems without CaCl$_2$ are shown for other data than
+    for the $\alpha$-carbon of POPG for which experimental order parameter is not available.
+    Calsium density distributions are shown in figure \ref{CAdensPG}.
+    %(right) The headgroup order parameters of PG from PC:PG mixtures as a function CaCl$_2$
+    %concentration from experiments \cite{borle85} and CHARMM36 simulations.
+  }
+\end{figure*}
+
 \begin{figure}[]
@@
   \centering
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
+  \includegraphics[width=12.0cm]{../Figs/DIHEDRALSlipid17eccWITHCaClPOPC.eps}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
   \caption{\label{DIHSwithCAlipid17eccPOPC}
@@
   \centering
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
+   \includegraphics[width=12.0cm]{../Figs/DIHEDRALScharmm36WITHCaClPOPC.eps}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
+%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
   \caption{\label{DIHSwithCAcharmm36POPC}
@@
   \centering
-  \includegraphics[width=18.0cm]{../Figs/DIHEDRALSslipidsWITHCaCl.eps}
+  \includegraphics[width=16.0cm]{../Figs/DIHEDRALSslipidsWITHCaClPOPG.eps}
   %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MSlipids.png}
@@
   \centering
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
-  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
+  \includegraphics[width=16.0cm]{../Figs/DIHEDRALSlipid17WITHCaClPOPG.eps}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
+  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
   \caption{\label{DIHSwithCAlipid17POPG}

### 2021-01-13 manuscriptPGPE.tex
Update figure 2
Updates from Overleaf

@@
 All the studied lipids specifically bind to various proteins \cite{yeagle14}.
-Furthermore, we use our results to elucidate lipid-protein interactions and the effect of charges on lipid conformations.
+Furthermore, we use our results to elucidate lipid--protein interactions and the effect of charges on lipid conformations.
 
@@
 The simulated systems of pure PE and PG bilayers without additional ions
-are listed in Tables \ref{systemsPE} and \ref{systemsPG},
-and lipid mixtures with additional ions in Tables~\ref{systemsMIX} and~\ref{systemsMIX2}.
+are listed in Tables~\ref{systemsPE} and \ref{systemsPG},
+and lipid mixtures with additional ions in Tables~\ref{systemsMIX} and \ref{systemsMIX2}.
 Further simulation details are given in the SI, and
@@
 and analysis for all simulations indexed in NMRlipids databank manner were performed
-using python codes.
+using Python codes.
 
@@
 \begin{equation}
-S_{\rm CH}=\frac{1}{2}\langle 3\cos^2\theta -1 \rangle,
+S_{\rm CH}=\frac{1}{2}\left\langle 3\cos^2\theta -1 \right\rangle,
 \end{equation}
@@
 Python programs that use the MDAnalysis library \cite{agrawal11,gowers16}
-used for all atom simulations is available in Ref. \citenum{MATCHgit}
-({\tt scripts/calcOrderParameters.py}). For united atom simulations, the trajectories
-with hydrogens having ideal geometry were constructed first using either buildH program~\cite{buildH}
-or ({\tt scratch/opAAUA\_prod.py}) in  Ref. \citenum{MATCHgit}, and the order parameters were
+used for all atom simulations are available in Ref.~\citenum{MATCHgit}
+(\texttt{scripts/calcOrderParameters.py}). For united atom simulations, the trajectories
+with hydrogens having ideal geometry were constructed first using either \texttt{buildH} program~\cite{buildH}
+or (\texttt{scratch/opAAUA\_prod.py}) in  Ref.~\citenum{MATCHgit}, and the order parameters were
 then calculated from these trajectories. This approach has been tested against trajectories
@@
 %\todo{BuildH program is now cited with a direct link to the GitHub repo. I think that a release to Zenodo would be nice in the final publication.}\\
-The ion number density profiles were calculated using the {\tt gmx density} tool
-of the Gromacs sofware package \cite{gromacsMANUAL}.
+The number density profiles of ions were calculated using the \texttt{gmx density} tool
+of the GROMACS sofware package \cite{gromacsMANUAL}.
 
@@
    \caption{\label{HGorderParameters}
-     Chemical structure, refocused-INEPT spectrum, 2D R-PDLF spectra, and S-DROSS data (from top to bottom) of {\bf A)} POPE  and {\bf B)} POPG.
+     Chemical structure, refocused-INEPT spectrum, 2D R-PDLF spectra, and S-DROSS data (from top to bottom) of \textbf{A)} POPE  and \textbf{B)} POPG.
       Full NMR spectra are shown Figs. \ref{POPEspectra} and \ref{POPGspectra}.
-     {\bf C)} Chemical structure of POPC and POPS.
-    {\bf D)} Headgroup and glycerol backbone order parameters from different experiments in lamellar liquid disordered phase.
+     \textbf{C)} Chemical structure of POPC and POPS.
+    \textbf{D)} Headgroup and glycerol backbone order parameters from different experiments in lamellar liquid disordered phase.
     The values and signs for POPE (310~K) and POPG (298~K)
@@
 
-\subsection{Analysis of lipid conformations bound to proteins}
-Lipid structures from PDB (\url{http://www.rcsb.org/})
+\subsection{Analysis of conformations of protein-bound lipids}
+Lipid structures from Protein Data Bank (PDB, \url{http://www.rcsb.org/})
 were searched using PDBe REST API (\url{www.ebi.ac.uk/pdbe/pdbe-rest-api}).
-First, all PDB entries containing PC, PE, PG, or PS lipid headgroups were searched.
+First, all PDB entries containing PC, PE, PG, or PS lipid headgroups were collected.
 The ligand names to identify the lipids were:
@@
 PSF, PS6, Q3G, P5S, D39, PS2, 17F, and 8SP for PS.
-All PDB structures containing these ligands were downloaded and the dihedral angles of the first lipid in the latest version of the structures within each PDB code were calculated using the MDTraj python library \cite{mcgibbon15}.
-The used Jupyter notebook is available from the project's GitHub repository ({\it scripts/pdbSEARCH.ipynb}).
+Secondly, all PDB structures containing these ligands were downloaded and the dihedral angles of the first lipid in the latest version of the structures within each PDB code were calculated using the MDTraj python library \cite{mcgibbon15}.
+The used Jupyter notebook is available from the project's GitHub repository (\textit{ scripts/pdbSEARCH.ipynb}).
 
@@
    \caption{\label{structures}
-     Results from CHARMM36 simulations demonstrating the differences in conformational ensembles betwee different lipids. 
-     {\bf A)} Snapshots with overlayed C$_\beta$, C$_\alpha$ and O$_\alpha$ atoms and occurence of different conformations.
-     {\bf B)} Headgroup and glycerol backbone region order parameters of different different lipids.
-     {\bf C)} Distributions of heavy atom dihedral angles of different lipids from CHARMM36 simulations.
-     {\bf D)} Distributions of P-N vector angle with respect to membrane normal.
+     Results from CHARMM36 simulations demonstrating the differences in conformational ensembles between different lipids. 
+     \textbf{A)} Snapshots with overlayed C$_\beta$, C$_\alpha$ and O$_\alpha$ atoms and occurence of different conformations.
+     \textbf{B)} Headgroup and glycerol backbone region order parameters of different different lipids.
+     \textbf{C)} Distributions of heavy atom dihedral angles of different lipids from CHARMM36 simulations.
+     \textbf{D)} Distributions of P-N vector angle with respect to membrane normal.
   }
@@
 To experimentally characterize the headgroup conformational ensembles of lipids
-that are not bound to proteins in electrostatically neutral cell membrane,
-we measured the C-H bond order parameters and their signs from of POPG and POPE
-in liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
+that are not bound to proteins in an electrostatically neutral cell membrane,
+we measured the C--H bond order parameters and their signs from of POPG and POPE
+in the liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
 Determination of headgroup and glycerol backbone order parameters and their signs
-was straightforward from the data in Figs. \ref{HGorderParameters}, \ref{POPEspectra} and \ref{POPGspectra}
-for all the C-H bonds, except the $\beta$ and g$_2$ carbons in POPG.
+was straightforward from the data in Figs.~\ref{HGorderParameters}, \ref{POPEspectra} and \ref{POPGspectra}
+for all the C--H bonds, except for the $\beta$ and g$_2$ carbons in POPG.
 These carbons have overlapping peaks in the INEPT spectra due to their similar chemical environments,
-and only the magnitude of the larger order parameter could be determined from the R-PDLF spectra (Fig. \ref{HGorderParameters} B).
+and only the magnitude of the larger order parameter could be determined from the R-PDLF spectra (Fig.~\ref{HGorderParameters}B)).
 Nevertheless, based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85},
@@
 The decrease in the beginning of the S-DROSS curve suggests that the sign of larger g$_2$ order parameter
-is negative and later increase suggests that sign of smaller $\beta$ order parameter is positive (Fig. \ref{HGorderParameters} (B)).
-This interpretation is confirmed by SIMPSON calculations in Fig \ref{POPGsimpson}.
+is negative and later increase suggests that sign of smaller $\beta$ order parameter is positive (Fig.~\ref{HGorderParameters}B)).
+This interpretation is confirmed by SIMPSON calculations in Fig.~\ref{POPGsimpson}.
 
-Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.
-\ref{HGorderParameters} D), where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
-Overall agreement of order parameters determined by different authors and different techniques for same lipid headgroup is very good in here and previous studies \cite{botan15,ollila16,antila19},
-suggesting that the observed differences between lipid types arise from differences in headgroup chemistry rather than inaccuracies in experiments, or differences in acyl or experimental conditions.
-The most distinct order parameters are observed for PS headgroups, for which the $\alpha$-carbon order parameter exhibits significant forking and the $\beta$-carbon has more negative value than in other lipids.
-On the other hand, the $\beta$-carbon order parameter of PG headgroup has positive sign in contrast to all the other lipids.
+Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.~\ref{HGorderParameters}D), where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
+The overall agreement of order parameters determined by different research teams and different techniques for the same lipid headgroup is very good here and in previous studies \cite{botan15,ollila16,antila19}.
+This suggests that the observed differences between lipid types arise from differences in headgroup chemistry rather than inaccuracies in experiments, or differences in the acyl chains or in experimental conditions.
+The most distinct order parameters are observed for PS headgroups, for which the $\alpha$-carbon order parameter exhibits significant forking and the $\beta$-carbon has more negative value than other studied lipid types.
+On the other hand, the $\beta$-carbon order parameter of PG headgroup has a positive sign, in contrast to all the other lipid types.
 Notably, this has not been observed in traditional $^2$H NMR experiments, where only the absolute value of the order parameters are measured~\cite{wohlgemuth80,gally81,borle85}.
-The glycerol backbone order parameters are similar for all the lipids, although they move slightly toward positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG.
+The glycerol backbone order parameters are similar for all the lipid types, although they move slightly toward positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG.
 Essential differences between PC and PE headgroups are not observed.
@@
   \caption{\label{changesWITHsurf}
-    {\bf A)} Modulation of PC headgroup order parameters and P-N vector angle upon addition of cationic surfactant
+    \textbf{A)} Modulation of PC headgroup order parameters and P--N vector angle upon addition of cationic surfactant
     from CHARMM36 simulations compared with the experimental data \cite{scherer89}.
-    {\bf B)} Changes in PC headgroup conformational ensembles upon increasing the amount of positive charge in bilayer,
+    \textbf{B)} Changes in PC headgroup conformational ensembles upon increasing the amount of positive charge in bilayer,
     characterized by the heavy atom dihedral distributions, from CHARMM36 simulations.
@@
 
-To resolve the conformational ensembles of lipid headgroups, we compared the headgroup and glycerol backbone C-H bond order parameters from different MD simulation force fields to the experimental data.
-None of the force fields reproduced the order parameters within experimental accuracy for any of the lipids analysed in this work (Figs.
-\ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}) or in our previous studies \cite{botan15,antila19}.
-Nevertheless, the CHARMM36 force field performs best for all the headgroups and reproduces the experimentally observed distinct order parameters of PS and PG lipids (Fig.
-\ref{structures} B).
-To understand the structural origin of these distinct order parameters, we calculated the distributions of heavy atom dihedral angles from CHARMM36 simulations in figure \ref{structures} C).
-Major differences between headgroups are observed only for the last two dihedrals in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$, which prefer {\it gauche$^-$} conformations for PG and {\it gauche$^+$} for PS.
+To resolve the conformational ensembles of lipid headgroups, we compared the headgroup and glycerol backbone C--H bond order parameters from different MD force fields to the experimental data.
+None of the force fields reproduced the order parameters within experimental accuracy for any of the lipids analysed in this work (Figs.~\ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}) or in our previous studies \cite{botan15,antila19}.
+Nevertheless, the CHARMM36 force field performs best for all the headgroups and reproduces the experimentally observed distinct order parameters of PS and PG lipids (Fig.~\ref{structures}B).
+To understand the structural origin of these distinct order parameters, we calculated the distributions of heavy atom dihedral angles from CHARMM36 simulations in Fig.~\ref{structures}C).
+Major differences between headgroups are observed only for the last two dihedrals in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$, which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS.
 Rest of the dihedrals are similar between different lipids, with the exception of small differences for PS lipids.
 Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids, which could also explain the more rigid headgroup structures~\cite{browning80,buldt81}.
-Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.
-\ref{structures} B)).
+Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.~\ref{structures}B)).
 However, the differences between PC and PE in P-O$_\alpha$-C$_\alpha$-C$_\beta$ dihedral
-and P-N vector dipole may be artificial as the $\beta$-carbon order parameter in PC is too negative in the CHARMM36 force field, thereby not being equal to the order parameter in PE as observed in experiments \cite{botan15}.
+and P--N vector dipole may be artificial as the $\beta$-carbon order parameter in PC is too negative in the CHARMM36 force field, thereby not being equal to the order parameter in PE as observed in experiments \cite{botan15}.
 
@@
 the decreased probability for the trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral is the main structural consequence of
-bound ion to PC headgroup (Fig. \ref{DIHSwithCAlipid17eccPOPC} and \ref{DIHSwithCAcharmm36POPC}) in
-Lipid17ecc (having the most realistic response of PC lipid headgroup order parameters to the calcium binding in Fig. \ref{changesWITHCaClPG})
+bound ion to PC headgroup (Fig.~\ref{DIHSwithCAlipid17eccPOPC} and \ref{DIHSwithCAcharmm36POPC}) in
+Lipid17ecc (having the most realistic response of PC lipid headgroup order parameters to the calcium binding in Fig.~\ref{changesWITHCaClPG})
 and CHARMM36 simulations.
@@
 Response of charged lipids to the ion binding is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
-The most realistic simulations (Lipid17 and Slipids in figure \ref{changesWITHCaClPG}) suggest
+The most realistic simulations (Lipid17 and Slipids in Fig.~\ref{changesWITHCaClPG}) suggest
 %istic response of PG $\beta$-carbon to CaCl$_2$  even thought the calcium binding affinity was overestimated in these models.
@@
 %(\ref{changesWITHCaClPG}),
-but dihedral distributions are more sensitive to the bound ions in PG (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
-However, the changes in dihedrals upon addition of calcium differ between the best models (Figs \ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}),
+but dihedral distributions are more sensitive to the bound ions in PG (Figs.~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+However, the changes in dihedrals upon addition of calcium differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}),
 none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added calcium in PG is not available.
@@
 all lipid conformations available within protein structures
-deposited in the PDB \cite{berman00} in figure \ref{dihedralsFROMpdb}.
+deposited in the PDB \cite{berman00} in Fig.~\ref{dihedralsFROMpdb}.
 We found 176 PC, 198 PE, 70 PG, and 41 PS conformations, which
@@
 P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals
-seem to avoid cis conformations as observed also in liquid lamellar state.
+seem to avoid $cis$ conformations as observed also in liquid lamellar state.
 The %range of available angles for
@@
 g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals are even less restricted in protein bound state
-than in liquid lamellar phase where cis and anti ecplipsed state were not present (Fig. \ref{structures}). 
+than in liquid lamellar phase where $cis$ and $anti$ eclipsed state were not present (Fig.~\ref{structures}). 
 Major differences between different lipid headgroups are not observed when bound to proteins.
@@
 Therefore, the differences in conformational ensembles between different
-lipids in liquid lamellar state in Fig. \ref{structures} are not seen in the protein bound states.
+lipids in liquid lamellar state in Fig.~\ref{structures} are not seen in the protein bound states.
 
 This suggests that lipids can bind to proteins in wide range of conformations independently
-of headgroup chemistry. Thereby, lipids compromize their preferred conformations when binding
+of headgroup chemistry. Thereby, lipids compromise their preferred conformations when binding
 to proteins and the binding of lipids to specific locations would be driven by 
@@
 However, it is important to note that lipids are often not the main target in
-structures deposited in PDB and therefore their conformations may be less reliable than proteins.
+structures deposited in PDB and therefore their conformations may be less reliable than those of proteins.
 Stereochemical violations and structures deviating from lipid crystals have been
@@
 that reproduce the NMR data in liquid lamellar phase, thereby proposing that such deviations
-are realistic also in protein bound states.
+are realistic also in protein-bound states.
 
@@
 from NMRlipids open collaboration revealed that headgroups of the most abundant biological
-phopholipids, PC, PE, PG and PS, sample wide range of different conformations in lamellar liquid state.
+phopholipids, PC, PE, PG and PS, sample a wide range of different conformations in the lamellar liquid state.
 Differences in NMR order parameters between different headgroups can be explained
@@
 similar conformations are accessed by all headgroups, but with different probabilities.
-Also the changes in order parameters upon addition of charged molecules, such as cationic lipids, surfactants, or drugs,
-in membranes can be explained by the changes in dihedral angle probabilities, particularly close to the phosphate region.
+Also the changes in order parameters upon the addition of charged molecules---such as cationic lipids, surfactants, or drugs---in membranes can be explained by the changes in dihedral angle probabilities, particularly close to the phosphate region.
 
 Wide range of conformations are observed also in lipids that are tightly bound to proteins in PDB,
-suggesting that the specific binding of lipids to proteins is dominated by the intermolecular lipid-protein
+suggesting that the specific binding of lipids to proteins is dominated by the intermolecular lipid--protein
 interactions, while the differences in conformational ensembles between different lipid types
-play a minor role. Therefore, lipids can conform themself to multiple binding positions locating
+play a minor role. Therefore, lipids can conform themselves to multiple binding positions located
 in various proteins.
 
-Our results pave the way to understand how lipids regulate membrane protein function.
+Our results pave the way to understanding how lipids regulate membrane protein function.
 Lipid conformational ensembles in liquid lamellar phase are necessary
-for the detailed analysis of lipid-protein interaction energetics, particularly for the entropic part. 
+for the detailed analysis of lipid--protein interaction energetics, particularly for its entropic component. 
 Furthermore, our results demonstrate how the NMRlipids databank, containing MD simulations and NMR data,
-can be used to resolve conformational ensembles of disordered biomolecules in membrane environment,
+can be used to resolve conformational ensembles of disordered biomolecules in a membrane environment,
 and how this information can be coupled with structural data in the PDB.
 We believe that this combination brings us closer to the comprehensive understanding of
-complex biomolecular assemblies containg both disordered and structural regions, such as membrane
+complex biomolecular assemblies containing both disordered and structural regions, such as membrane
 protein complexes.

### 2021-01-13 manuscriptPGPEsuppl.tex
Updates from Overleaf

@@
   \centering
-  \includegraphics[width=12.0cm]{../Figs/DIHEDRALSlipid17eccWITHCaClPOPC.eps}
+  \includegraphics[width=12.0cm]{./Figs/DIHEDRALSlipid17eccWITHCaClPOPC.eps}
 %  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17ecc.png}
@@
   \centering
-   \includegraphics[width=12.0cm]{../Figs/DIHEDRALScharmm36WITHCaClPOPC.eps}
+   \includegraphics[width=12.0cm]{./Figs/DIHEDRALScharmm36WITHCaClPOPC.eps}
 %  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
@@
   \centering
-  \includegraphics[width=16.0cm]{../Figs/DIHEDRALSslipidsWITHCaClPOPG.eps}
+  \includegraphics[width=16.0cm]{./Figs/DIHEDRALSslipidsWITHCaClPOPG.eps}
   %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MSlipids.png}
@@
   \centering
-  \includegraphics[width=16.0cm]{../Figs/DIHEDRALSlipid17WITHCaClPOPG.eps}
+  \includegraphics[width=16.0cm]{./Figs/DIHEDRALSlipid17WITHCaClPOPG.eps}
   %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17.png}

### 2021-01-20 manuscriptPGPE.tex
Update information on enantiomers.

@@
 \usepackage{xr}
-\externaldocument{manuscriptPGPEsuppl}
+
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}
+  \@addtofilelist{#1}
+  \IfFileExists{#1}{}{\typeout{No file #1.}}
+}
+\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+    \externaldocument{#1}%
+    \addFileDependency{#1.tex}%
+    \addFileDependency{#1.aux}%
+}
+
+\myexternaldocument{manuscriptPGPEsuppl}
 
@@
 Furthermore, the detailed understanding of lipid conformational ensembles
-is limited by the lack of universal models that would map order parameters to structural ensembles / that would help interpret the experiments \cite{pezeshkian18,akutsu20}.
+is limited by the lack of universal models that would map order parameters to structural ensembles \cite{pezeshkian18,akutsu20}.
 
@@
 The NMR experiments were identical as in our previous work~\cite{antila19}.
-POPE and POPG powder were purchased from Avanti polar lipids 
-\todo{POPG is racemic in experiments. How much we should discuss this in relation with simulations?}.
+POPE (1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine) and POPG (1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-(1'-rac-glycerol)) powder were purchased from Avanti polar lipids.
 The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.
@@
   }
-  \todo{This is a draft and requires quite a bit of polishing.
-    More detailed discussion of this figure is in https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/9} \\
 \end{figure*}
@@
 Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids, which could also explain the more rigid headgroup structures~\cite{browning80,buldt81}.
-Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.~\ref{structures}B)).
+Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.~\ref{structures}D)).
 However, the differences between PC and PE in P-O$_\alpha$-C$_\alpha$-C$_\beta$ dihedral
@@
 In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase.
-between headgroups, the sampled dihedral angles are within approximately same ranges for all headgroup types. Therefore, we suggest that all lipid headgroups are very flexible thereby being
+Between headgroups, the sampled dihedral angles are within approximately same ranges for all headgroup types. Therefore, we suggest that all lipid headgroups are very flexible---thereby being
 able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
-Our results support the models that propose free rotation around phosphate group for PC lipids, which decouples the dynamics and conformatios between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
+Our results support the models that propose free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
 The wide range of observed conformations suggest that the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and that the models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
@@
 
-Charged lipids, proteins, surfactants, drugs, and ions incorporated in membranes
+Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes
 reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of
@@
 
-To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge, we first analyzed CHARMM36 simulations which correctly captures the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
+To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge, we first analyzed CHARMM36 simulations which correctly capture the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
 Heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B
@@
 Choline region remains essentially unchanged and only minor changes are observed in other dihedrals
-even thought almost half of the molecules in membranes are cationic surfactants.
+even though almost half of the molecules in membranes are cationic surfactants.
 We also tried to resolve PC lipid headgroup conformational ensembles in mixed membrane with
@@
 
-\subsection{Protein bound lipid conformations}
+\subsection{Protein-bound lipid conformations}
 
-To analyze lipid conformations in protein bound states,
+To analyze lipid conformations in protein-bound states,
 we calculated the heavy atom dihedral distributions from 
@@
 from NMRlipids open collaboration revealed that headgroups of the most abundant biological
-phopholipids, PC, PE, PG and PS, sample a wide range of different conformations in the lamellar liquid state.
+phospholipids, PC, PE, PG and PS, sample a wide range of different conformations in the lamellar liquid state.
 Differences in NMR order parameters between different headgroups can be explained

### 2021-01-20 manuscriptPGPEsuppl.tex
Update information on enantiomers.

@@
 
-The quality of PE and PG headgroup conformational ensembles in different simulations against NMR experiments
-is evaluated in figures \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}
-using C-H bond order parameters as in our previous studies for PC and PS lipids \cite{botan15,antila19}.
-Conclusions are the same for all lipids: None of the force fields correctly captures the lipid headgroup
-conformational ensembles, but CHARMM36 gives results closest to experiments.
+The quality of PE and PG headgroup conformational ensembles in different simulations against NMR experiments is evaluated in figures \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG} using C-H bond order parameters as in our previous studies for PC and PS lipids \cite{botan15,antila19}.
+Conclusions are the same for all lipids: None of the force fields correctly captures the lipid headgroup conformational ensembles, but CHARMM36 gives results closest to experiments.
+
+The PG headgroup is biologically abundant R enantiomer in all simulations, while our $^{13}$C NMR experiments has a racemic mixture.
+Nevertheless, previous $^{2}$H NMR experiments comparing results between different enantiomers concluded that the structural differences between these are minor \cite{wohlgemuth80}.
 

### 2021-01-20 manuscriptPGPE.tex

@@
 To understand the structural origin of these distinct order parameters, we calculated the distributions of heavy atom dihedral angles from CHARMM36 simulations in Fig.~\ref{structures}C).
-Major differences between headgroups are observed only for the last two dihedrals in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$, which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS.
-Rest of the dihedrals are similar between different lipids, with the exception of small differences for PS lipids.
+Major differences between headgroups are observed for the last two dihedrals in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$, which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS.
+Rest of the dihedrals are similar between different lipids, with the exception of PS lipids for which differences distinct distributions are observed particularly around phosphate group.
 Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids, which could also explain the more rigid headgroup structures~\cite{browning80,buldt81}.

### 2021-01-20 manuscriptPGPE.tex
FIx typo

@@
 Major differences between headgroups are observed for the last two dihedrals in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$, which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS.
-Rest of the dihedrals are similar between different lipids, with the exception of PS lipids for which differences distinct distributions are observed particularly around phosphate group.
+Rest of the dihedrals are similar between different lipids, with the exception of PS lipids for which distinct distributions are observed particularly around phosphate group.
 Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids, which could also explain the more rigid headgroup structures~\cite{browning80,buldt81}.

### 2021-01-26 manuscriptPGPE.tex
Update MD methods and results and discussion.

@@
   Chemistry of lipid headgroups, the water facing components of cell membranes that regulate cell functions via lipid--protein interactions, varies between organisms and organelles.
-Because lipid membranes are in liquid state under physiological conditions, the conformational ensembles of lipid headgroups and their dependence on chemistry are difficult to resolve experimentally.
+Because lipid membranes are in liquid state under physiological conditions, the conformational ensembles of different lipid headgroups are difficult to resolve experimentally.
 Here, we combine solid state NMR experiments and molecular dynamics simulations from NMRlipids open collaboration to resolve the conformational ensembles of the headgroups of key lipid types in liquid lamellar phase under various biologically relevant conditions.
@@
 Lipids can directly bind to proteins or indirectly affect protein
-functions by altering membrane properties or charge \cite{lee03,lemmon08}.
+functions by altering membrane properties such as charge or elasticity \cite{lee03,lemmon08}.
 While the specific interactions with certain lipid headgroups
@@
 These studies suggest that the glycerol backbone conformations are largely similar irrespectively of the headgroup \cite{gally81}, 
-and the headgroup conformations are similar in PC, PE and PG lipids, while the headgroup is more rigid in PS lipids \cite{wohlgemuth80,buldt81}. 
+and the headgroup conformations are similar in  phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG) lipids, while the headgroup is more rigid in phosphatidylserine (PS) lipids \cite{wohlgemuth80,buldt81}.
 However, these tendencies are based only on the absolute values, whereas
@@
 and affect membrane protein functionality and signaling \cite{lemmon08,leventis10,sohlenkamp16,hariharan18}.
-All the studied lipids specifically bind to various proteins \cite{yeagle14}.
-Furthermore, we use our results to elucidate lipid--protein interactions and the effect of charges on lipid conformations.
+All the studied lipids specifically bind to various proteins \cite{yeagle14}. We use our results to elucidate also lipid--protein interactions and the effect of charges on lipid conformations.
 
@@
 \subsection{Experimental C--H bond order parameters}
-The headgroup and glycerol backbone C--H bond order parameters of POPE and POPG
-were measured using natural abundance $^{13}$C solid state NMR spectroscopy
-as described previously \cite{ferreira13,ferreira16}.
-The magnitudes of order parameters were determined from the chemical-shift resolved dipolar splittings
-using a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04} and
-the signs from S-DROSS experiments~\cite{gross97} combined with SIMPSON simulations \cite{bak00}.
+The headgroup and glycerol backbone C--H bond order parameters of 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine (POPE) and 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (POPG), purchased from Avanti polar lipids, were measured using natural abundance $^{13}$C solid state NMR spectroscopy as described previously \cite{ferreira13,ferreira16}.
+The magnitudes of order parameters were determined from the chemical-shift resolved dipolar splittings using a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04} and the signs from S-DROSS experiments~\cite{gross97} combined with SIMPSON simulations \cite{bak00}.
 %The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
@@
 The NMR experiments were identical as in our previous work~\cite{antila19}.
-POPE (1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine) and POPG (1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-(1'-rac-glycerol)) powder were purchased from Avanti polar lipids.
 The POPE experiments were recorded at 310~K and POPG experiments at 298~K, where the bilayers are in the liquid disordered phase \cite{marsh13}.
@@
 
-Molecular dynamics simulation data were collected using
-the NMRlipids Open Collaboration \cite{botan15}, which is running at
-\url{nmrlipids.blogspot.fi} and \url{github.com/NMRlipids/NMRlipidsIVotherHGs}.
-The simulated systems of pure PE and PG bilayers without additional ions
-are listed in Tables~\ref{systemsPE} and \ref{systemsPG},
-and lipid mixtures with additional ions in Tables~\ref{systemsMIX} and \ref{systemsMIX2}.
-Further simulation details are given in the SI, and
-the simulation data are indexed in a
-searchable database available at \url{www.nmrlipids.fi},
-and in the NMRlipids/MATCH repository (\url{github.com/NMRlipids/MATCH}).
+Molecular dynamics simulation data were collected and analyzed using
+the methods from the NMRlipids Open Collaboration project (\url{nmrlipids.blogspot.fi}) \cite{botan15,catte16,ollila16,antila19}.
+%\url{github.com/NMRlipids/NMRlipidsIVotherHGs}. 
+Simulation and accessibility details of more than 70 systems simulated for this work are given in the supplementary information.
+Quality evaluation of these simulations enables us to select the best models for the interpretation of lipid conformational ensembles from the experimental data.
 
-\subsection{Analysis of molecular dynamics simulation data}
-The large set of MD simulation data was analysed using the development version of NMRlipids databank.
-Unique naming convention for lipid atoms in each force field was defined using the mapping files
-and analysis for all simulations indexed in NMRlipids databank manner were performed
-using Python codes.
-
-The C--H bond order parameters were calculated directly
-from the carbon and hydrogen positions using the definition
-\begin{equation}
-S_{\rm CH}=\frac{1}{2}\left\langle 3\cos^2\theta -1 \right\rangle,
-\end{equation}
-where $\theta$ is the angle between the C--H bond and the membrane normal
-(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
-Angular brackets denote average over all sampled configurations.
-The order parameters were first calculated averaging over time separately
-for each lipid in the system. The average and
-the standard error of the mean were then calculated over different lipids.
-Python programs that use the MDAnalysis library \cite{agrawal11,gowers16}
-used for all atom simulations are available in Ref.~\citenum{MATCHgit}
-(\texttt{scripts/calcOrderParameters.py}). For united atom simulations, the trajectories
-with hydrogens having ideal geometry were constructed first using either \texttt{buildH} program~\cite{buildH}
-or (\texttt{scratch/opAAUA\_prod.py}) in  Ref.~\citenum{MATCHgit}, and the order parameters were
-then calculated from these trajectories. This approach has been tested against trajectories
-with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot17}.\\
-%\todo{BuildH program is now cited with a direct link to the GitHub repo. I think that a release to Zenodo would be nice in the final publication.}\\
-The number density profiles of ions were calculated using the \texttt{gmx density} tool
-of the GROMACS sofware package \cite{gromacsMANUAL}.
+The quality of lipid headgroup and glycerol backbone conformational ensembles in PE and PG simulations were evaluated using the C-H bond order parameters \cite{botan15}.
+Interactions between different headgroups in simulations of mixed bilayers were evaluated monitoring the changes in headgroup order parameters upon mixing the lipids \cite{antila19}.
+The ion binding affinities and response of lipids to bound charge were evaluated monitoring the changes of lipid headgroup order parameters \cite{catte16,antila19}.
 
@@
 \subsection{Analysis of conformations of protein-bound lipids}
+\todo{This is to be finished once the final analysis is done as discussed in this issue: https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/40}
+
 Lipid structures from Protein Data Bank (PDB, \url{http://www.rcsb.org/})
@@
    \caption{\label{structures}
-     Results from CHARMM36 simulations demonstrating the differences in conformational ensembles between different lipids. 
+     Results from the best simulation model (CHARMM36) simulations demonstrating the differences in conformational ensembles between different lipids. 
      \textbf{A)} Snapshots with overlayed C$_\beta$, C$_\alpha$ and O$_\alpha$ atoms and occurence of different conformations.
      \textbf{B)} Headgroup and glycerol backbone region order parameters of different different lipids.
-     \textbf{C)} Distributions of heavy atom dihedral angles of different lipids from CHARMM36 simulations.
+     \textbf{C)} Distributions of heavy atom dihedral angles of different lipids from simulations.
      \textbf{D)} Distributions of P-N vector angle with respect to membrane normal.
@@
 
-To experimentally characterize the headgroup conformational ensembles of lipids
-that are not bound to proteins in an electrostatically neutral cell membrane,
-we measured the C--H bond order parameters and their signs from of POPG and POPE
-in the liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
+To experimentally characterize the headgroup conformational ensembles of lipids that are not bound to proteins in an electrostatically neutral cell membrane, we measured the C--H bond order parameters and their signs of POPG and POPE in the liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
 Determination of headgroup and glycerol backbone order parameters and their signs
@@
 
-To resolve the conformational ensembles of lipid headgroups, we compared the headgroup and glycerol backbone C--H bond order parameters from different MD force fields to the experimental data.
-None of the force fields reproduced the order parameters within experimental accuracy for any of the lipids analysed in this work (Figs.~\ref{HGorderParametersPE} and \ref{HGorderParametersPOPG}) or in our previous studies \cite{botan15,antila19}.
-Nevertheless, the CHARMM36 force field performs best for all the headgroups and reproduces the experimentally observed distinct order parameters of PS and PG lipids (Fig.~\ref{structures}B).
-To understand the structural origin of these distinct order parameters, we calculated the distributions of heavy atom dihedral angles from CHARMM36 simulations in Fig.~\ref{structures}C).
+%To resolve the conformational ensembles of lipid headgroups, we compared the headgroup and glycerol backbone C--H bond order parameters from different MD force fields to the experimental data. 
+%Although none of the simulations reproduce the headgroup order parameters within experimental accuracy for any of the analysed lipids (Figs.~\ref{HGorderParametersPE} and \ref{HGorderParametersPOPG} in the SI and Refs.
+\cite{botan15,antila19}), the CHARMM36
+%force field performs best for all the headgroups and 
+%simulations reproduce the experimentally observed distinct order parameters of PS and PG lipids (Fig.~\ref{structures}B). 
+To understand the structural origin of distinct order parameters for PG and PS lipids, we selected the simulations that best reproduce the differences between headgroups (based on the quality evaluation in the SI and Refs.
+\cite{botan15,antila19}) and calculated the distributions of heavy atom dihedral angles from simulations, shown in Fig.~\ref{structures}C.
 Major differences between headgroups are observed for the last two dihedrals in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$, which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS.
 Rest of the dihedrals are similar between different lipids, with the exception of PS lipids for which distinct distributions are observed particularly around phosphate group.
-Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids, which could also explain the more rigid headgroup structures~\cite{browning80,buldt81}.
+Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids.
+This may explain the more rigid headgroup structures in PS~\cite{browning80,buldt81}.
 Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.~\ref{structures}D)).
@@
 
-In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase.
-Between headgroups, the sampled dihedral angles are within approximately same ranges for all headgroup types. Therefore, we suggest that all lipid headgroups are very flexible---thereby being
-able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
+All lipid headgroups sample very broad conformational ensembles in liquid lamellar phase and sampled dihedral angles are within approximately same ranges for all headgroup types (Fig.
+\ref{structures}).
+The results suggest that all lipid headgroups are very flexible---thereby being able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
 Our results support the models that propose free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
@@
 
-Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes
-reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of
-lipid headgroups \cite{seelig87}. However, detailed understanding on
-lipid conformational ensembles in membranes bearing electric charge is still lacking \cite{Semchyschyn04}.
+Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of lipid headgroups \cite{seelig87}.
+However, detailed understanding on lipid conformational ensembles in membranes bearing electric charge is still lacking \cite{Semchyschyn04}.
 
-To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge, we first analyzed CHARMM36 simulations which correctly capture the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
-Heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B
-reveal that the decrease of trans state probabilities in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
+To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge, we analyzed the simulations (CHARMM36) that correctly capture the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
+Heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B reveal that the decrease of trans state probabilities in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
 dihedrals are the major effects of positive charge on lipid headgroup conformations.
-Choline region remains essentially unchanged and only minor changes are observed in other dihedrals
-even though almost half of the molecules in membranes are cationic surfactants.
-We also tried to resolve PC lipid headgroup conformational ensembles in mixed membrane with
-negatively charged lipids \cite{scherer87},
-but the accuracy of currently available force fields
-was not sufficient for this (Fig. \ref{HGorderparametersPCvsPEPG}).
+Choline region remains essentially unchanged and only minor changes are observed in other dihedrals even though almost half of the molecules in membranes are cationic surfactants.
 
 Also binding of ions may affect the lipid headgroup conformational ensembles in physiological conditions.
-In line with the results from cationic surfactants, 
-the decreased probability for the trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral is the main structural consequence of
-bound ion to PC headgroup (Fig.~\ref{DIHSwithCAlipid17eccPOPC} and \ref{DIHSwithCAcharmm36POPC}) in
-Lipid17ecc (having the most realistic response of PC lipid headgroup order parameters to the calcium binding in Fig.~\ref{changesWITHCaClPG})
-and CHARMM36 simulations.
-%simulations with the ECC included, where
-%Similar but smaller effect is seen in CHARMM36 simulations with less bound calcium (Fig. \ref{DIHSwithCAcharmm36POPC}).
-%Also the headgroup order parameters of
-Response of charged lipids to the ion binding is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
-The most realistic simulations (Lipid17 and Slipids in Fig.~\ref{changesWITHCaClPG}) suggest
+In line with the results from cationic surfactants, the decreased probability for the trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral is the main structural consequence of
+bound Ca$^{2+}$ ion to PC headgroup (Fig.~\ref{DIHSwithCAlipid17eccPOPC} and \ref{DIHSwithCAcharmm36POPC}) in the most realistic MD simulations models (lipid17ecc and CHARMM36 in Fig.~\ref{changesWITHCaClPG}).
+%Lipid17ecc (having the most realistic response of PC lipid headgroup order parameters to the calcium binding in Fig.~\ref{changesWITHCaClPG}) and CHARMM36 simulations. 
+%Response of charged lipids to the ion binding is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
+The most realistic simulations suggest
 %istic response of PG $\beta$-carbon to CaCl$_2$  even thought the calcium binding affinity was overestimated in these models.
@@
 %(\ref{changesWITHCaClPG}),
-but dihedral distributions are more sensitive to the bound ions in PG (Figs.~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
-However, the changes in dihedrals upon addition of calcium differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}),
-none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added calcium in PG is not available.
-Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.
+but dihedral distributions are more sensitive to the bound ions in PG (Lipid17 and Slipids in Figs.~\ref{changesWITHCaClPG},~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+However, the changes in dihedrals upon addition of calcium differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and~\ref{DIHSwithCAlipid17POPG}), none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added calcium in PG is not available.
+Also the lipid headgroup conformational ensembles in mixed PC:PE and PC:PG membranes were not possible to resolve with the currently available force fields and experimental data (Figs.~\ref{HGorderparametersPCvsPEPG} and~\ref{HGorderparametersPGvsPCchange}).
 
-Despite the limited capability of simulations to interpret some of the experimental data due to
-the inaccuries in force fields, we conclude that the experimentally observed changes
-in headgroup order parameters upon addition of charges into bilayer arise from relatively
-small changes in conformational ensembles. These changes eventuate from
-mild changes in dihedral angle distributions, rather than from restriction of lipids into
-fixed conformations. Therefore, lipid headgroups remain in disordered state sampling large
-space of different conformations, thereby being able to interact with different molecules
-in various ways, also in membranes bearing electric charge.
+%Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.
+
+Despite the 
+%limited capability of simulations to interpret some of the experimental data due to the 
+inaccuries in simulations in some cases, we can conclude that the experimentally observed changes in headgroup order parameters upon addition of charges into bilayer arise from relatively small changes in conformational ensembles.
+These changes eventuate from mild changes in dihedral angle distributions, rather than from restriction of lipids into
+fixed conformations.
+Therefore, lipid headgroups remain in disordered state sampling large space of different conformations, thereby being able to interact with different molecules in various ways, also in membranes bearing electric charge.
 
@@
 
-To analyze lipid conformations in protein-bound states,
-we calculated the heavy atom dihedral distributions from 
-all lipid conformations available within protein structures
-deposited in the PDB \cite{berman00} in Fig.~\ref{dihedralsFROMpdb}.
+To analyze the alterations in lipid conformations when bound to proteins,
+we calculated the heavy atom dihedral distributions from lipid conformations within protein structures deposited in the PDB \cite{berman00} in Fig.~\ref{dihedralsFROMpdb}.
+\todo{This part is to be finished once the analysis if finalized}
 We found 176 PC, 198 PE, 70 PG, and 41 PS conformations, which
 present lipids that are tightly bound to proteins in fixed conformations
-that can be determined as a part of protein structure using crystallography, NMR, or cryo-EM.
+that can be determined as a part of protein structure using crystallography or cryo-EM.
 

### 2021-01-26 manuscriptPGPEsuppl.tex
Update MD methods and results and discussion.

@@
 \clearpage
-\section{Comparison of headgroup order parameters from different force fields to experiments}
+\section{Evaluation of simulations against NMR experiments}
+\subsection{Conformational ensembles of headgroup and glycerol backbone in PE and PG lipids}
 
@@
 
-The PG headgroup is biologically abundant R enantiomer in all simulations, while our $^{13}$C NMR experiments has a racemic mixture.
+It should be noted that the PG headgroup is biologically abundant R enantiomer in all simulations, while our $^{13}$C NMR experiments has a racemic mixture.
 Nevertheless, previous $^{2}$H NMR experiments comparing results between different enantiomers concluded that the structural differences between these are minor \cite{wohlgemuth80}.
@@
 
-\section{PC headgroup in mixtures with PE or PG lipids}
+\subsection{PC headgroup in mixtures with PE or PG lipids}
 \begin{figure*}[]
@@
 
+In conclusion, more accurate force fields are needed to correctly simulate the interactions between different headgroups.
+
 \clearpage
-\section{PG headgroup in mixtures with PC lipids}
+\subsection{PG headgroup in mixtures with PC lipids}
 Changes in other than PC lipid headgroup with changing membrane composition are less
@@
 \clearpage
-\section{Calcium binding to POPC:POPG mixtures}
+\subsection{Calcium binding to POPC:POPG mixtures}
 
@@
 \section{Simulated systems}
+
+The simulated systems of pure PE and PG bilayers without additional ions
+are listed in Tables~\ref{systemsPE} and \ref{systemsPG},
+and lipid mixtures with additional ions in Tables~\ref{systemsMIX} and \ref{systemsMIX2}.
+the simulation data are indexed in a
+searchable database available at \url{www.nmrlipids.fi},
+and in the NMRlipids/MATCH repository (\url{github.com/NMRlipids/MATCH}).
+The large set of MD simulation data was analysed using the development version of NMRlipids databank.
+Unique naming convention for lipid atoms in each force field was defined using the mapping files
+and analysis for all simulations indexed in NMRlipids databank manner were performed
+using Python codes.
+
+The C--H bond order parameters were calculated directly
+from the carbon and hydrogen positions using the definition
+\begin{equation}
+S_{\rm CH}=\frac{1}{2}\left\langle 3\cos^2\theta -1 \right\rangle,
+\end{equation}
+where $\theta$ is the angle between the C--H bond and the membrane normal
+(taken to align with $z$, with bilayer periodicity in the $xy$-plane).
+Angular brackets denote average over all sampled configurations.
+The order parameters were first calculated averaging over time separately
+for each lipid in the system. The average and
+the standard error of the mean were then calculated over different lipids.
+Python programs that use the MDAnalysis library \cite{agrawal11,gowers16}
+used for all atom simulations are available in Ref.~\citenum{MATCHgit}
+(\texttt{scripts/calcOrderParameters.py}).
+For united atom simulations, the trajectories with hydrogens having ideal geometry were constructed first using either \texttt{buildH} program~\cite{buildH} or (\texttt{scratch/opAAUA\_prod.py}) in  Ref.~\citenum{MATCHgit}, and the order parameters were then calculated from these trajectories.
+This approach has been tested against trajectories with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot17}.\\
+%\todo{BuildH program is now cited with a direct link to the GitHub repo. I think that a release to Zenodo would be nice in the final publication.}\\
+The number density profiles of ions were calculated using the \texttt{gmx density} tool of the GROMACS sofware package \cite{gromacsMANUAL}.
 

### 2021-02-24 manuscriptPGPEsuppl.tex

@@
     Calcium ion density profiles along membrane normal from simulations of POPC:POPG (1:1) mixtures with different force fields.
-    The changes in the order parameters upon addition of CaCl$_2$ are compared with experiments in figure \ref{changesWITHCaClPG} in the main text.
+    The changes in the order parameters upon addition of CaCl$_2$ are compared with experiments in figure \ref{changesWITHCaClPG}.
   }
@@
   }
-  \todo{There is something wrong the Lipid17ecc data at 1M, a new simulation is coming.} \\
 \end{figure*}
@@
   }
- \todo{Density profiles from Lipid17ecc data will be added when we have the new simulations at 1M.} \\
 \end{figure}
@@
     \todo{Citation and ion model for GROMOS-CKP?} \\
-  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
 %\end{table*}
@@
   \todo{Citation and description for ''Berger'' model?} \\
-  \todo{Upcoming Lipid17ecc with POPC:POPS (4:1) mixture simulations to be added.}
 %\end{table*}

### 2021-02-24 manuscriptPGPEsuppl.tex
Updates from Overleaf

--- 2021-01-26 manuscriptPGPEsuppl.tex
@@
 \author{Fernando Favela-Rosales}
-\affiliation[Tecnol\'{o}gico Nacional de M\'{e}xico]{Departamento de Investigaci\'{o}n, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}
+\affiliation[Tecnol\'{o}gico Nacional de M\'{e}xico]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}
 
@@
 
-\noindent {\it DPPE} Data is available at \cite{slipidsDPPEfiles}. \todo{Simulation details by F. Favela.}
+\noindent {\it DPPE with 288 lipids.} The starting structure for simulation was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}: 288 DPPE lipids and 9386 water molecules.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
+Simulation was performed for 200~ns, and the last 100~ns were used for the analysis.
+Simulation was carried out within the NPT ensemble using the GROMACS 5.0.4 package~\cite{abraham2015gromacs}.
+Timestep of 2~fs was used with the leapfrog integrator.
+The Nos\'{e}--Hoover thermostat~\cite{nose84,hoover85} was used with reference temperature of 336~K and a relaxation time constant of 0.5~ps; lipids and water were coupled separately to the heat bath.
+Pressure was kept constant at 1.013~bar using a semi--isotropic Parrinello--Rahman
+barostat~\cite{parrinello81} with a time constant of 10.0~ps.
+Long-range electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}.
+A real space cut-off of 1.0~nm was employed with grid spacing of 0.12~nm in the reciprocal space.
+Lennard-Jones potentials were cut off at 1.4~nm, with a dispersion correction applied to both energy and pressure.
+All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess97}, whereas water molecules were constrained using SETTLE~\cite{miyamoto92}.
+Twin-range cutoffs, 1.0~nm and 1.6~nm, were used for the neighbor lists with the long-range neighbor list updated every 10 steps.
 
-\noindent {\it POPG} Data is available at \cite{slipidsPOPGfiles}. \todo{Simulation details by F. Favela.
-  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
+\noindent {\it POPG with 288 lipids.} The starting structure for simulation was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}: 288 POPG lipids, 10664 water molecules and 288 Na ions.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by Aqvist~\cite{aqvist90}.
+Simulation was performed for 250~ns, and the last 100~ns were used for the analysis. Same simulation conditions as DPPE with reference temperature of 298~K.
 
@@
 
-\noindent {\it DPPG} Data in 298~K is available at \cite{slipidsDPPGfilesT298K} and in 314~K at \cite{slipidsDPPGfiles}.
-\todo{Simulation details by F. Favela.
-  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
+\noindent {\it DPPG with 288 lipids.} The starting structure for simulation was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}: 288 DPPG lipids, 11232 water molecules and 288 Na ions.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by Aqvist~\cite{aqvist90}.
+For the 298~K temperature, simulation was performed for 400~ns, and the last 100~ns were used for the analysis.
+For the 314~K temperature, simulation was performed for 200~ns, and the last 100~ns were used for the analysis.
+Same simulation conditions as DPPE for both temperatures.
 

### 2021-02-25 manuscriptPGPE.tex
Update text related to the relative energies

@@
 \author{Fernando Favela-Rosales}
-\affiliation{Departamento de Investigaci\'{o}n, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}
+\affiliation{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}
 
@@
      \textbf{B)} Headgroup and glycerol backbone region order parameters of different different lipids.
-     \textbf{C)} Distributions of heavy atom dihedral angles of different lipids from simulations.
+     \textbf{C)} Relative energies for individual dihedral angles estimated from inverse Boltzmann distributions of heavy atom dihedral angles.
+     The states corresponding energies higher than 7 kT are not shown because they are not observed in simulations.
      \textbf{D)} Distributions of P-N vector angle with respect to membrane normal.
@@
 
-To experimentally characterize the headgroup conformational ensembles of lipids that are not bound to proteins in an electrostatically neutral cell membrane, we measured the C--H bond order parameters and their signs of POPG and POPE in the liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
+To experimentally characterize the headgroup conformational ensembles of lipids that are not
+bound to proteins in an electrostatically neutral cell membrane, we measured the C--H bond order parameters
+and their signs of POPG and POPE in the liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
 Determination of headgroup and glycerol backbone order parameters and their signs
@@
 
-%To resolve the conformational ensembles of lipid headgroups, we compared the headgroup and glycerol backbone C--H bond order parameters from different MD force fields to the experimental data. 
-%Although none of the simulations reproduce the headgroup order parameters within experimental accuracy for any of the analysed lipids (Figs.~\ref{HGorderParametersPE} and \ref{HGorderParametersPOPG} in the SI and Refs.
-\cite{botan15,antila19}), the CHARMM36
-%force field performs best for all the headgroups and 
-%simulations reproduce the experimentally observed distinct order parameters of PS and PG lipids (Fig.~\ref{structures}B). 
-To understand the structural origin of distinct order parameters for PG and PS lipids, we selected the simulations that best reproduce the differences between headgroups (based on the quality evaluation in the SI and Refs.
-\cite{botan15,antila19}) and calculated the distributions of heavy atom dihedral angles from simulations, shown in Fig.~\ref{structures}C.
-Major differences between headgroups are observed for the last two dihedrals in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$, which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS.
-Rest of the dihedrals are similar between different lipids, with the exception of PS lipids for which distinct distributions are observed particularly around phosphate group.
+To understand the structural origin of distinct order parameters for PG and PS lipids,
+we first calculated the heavy atom dihedral angle distributions from simulations
+that best reproduce the differences between headgroups according to the quality
+evaluation in the SI and Refs. \cite{botan15,antila19} (Fig. \ref{DIHdists}).
+Then, we used the inverse Boltzmann distributions to estimate the
+energy costs for different dihedral angle orientations.
+The results in Fig.~\ref{structures}C) suggest that all lipid headgroups are very flexible and
+energy costs for rotating individual dihedrals to almost any angle is low (below 7 kT).
+Only trans states of P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P have larger relative
+energies and are not observed for any lipids during the simulations.
+
+Major differences between headgroups are observed for the last two dihedrals in the headgroup end,
+O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$,
+which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS,
+while PC and PE exhibit symmetric distributions.
+Larger barriers for crossings between \textit{gauche} and \textit{trans} states were observed for
+the last dihedral in PS and PG than for PC and PE lipids. 
+Rest of the dihedrals are similar between different lipids,
+with the exception of PS lipids for which slightly larger energy for ecplised anti conformation next to phosphate group
+was observed.
 Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids.
-This may explain the more rigid headgroup structures in PS~\cite{browning80,buldt81}.
+The increased barriers for dihedral rotations may explain the more rigid headgroup structures in PS~\cite{browning80,buldt81}.
 Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.~\ref{structures}D)).
@@
 
-All lipid headgroups sample very broad conformational ensembles in liquid lamellar phase and sampled dihedral angles are within approximately same ranges for all headgroup types (Fig.
+In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase and sampled dihedral angles are within approximately same ranges for all headgroup types.
+% (Fig.
 \ref{structures}).
-The results suggest that all lipid headgroups are very flexible---thereby being able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
-Our results support the models that propose free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
+Because the rotation of dihedral angles to almost any position bears relatively low energy cost,
+%results suggest that all lipid headgroups are very flexible---
+%thereby being
+the lipid headgroups are able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
+Our results support the models proposing free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
 The wide range of observed conformations suggest that the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and that the models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
@@
 Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of lipid headgroups \cite{seelig87}.
-However, detailed understanding on lipid conformational ensembles in membranes bearing electric charge is still lacking \cite{Semchyschyn04}.
+However, the detailed understanding on structural and energetic response of lipids to membrane bound electric charge is still lacking \cite{Semchyschyn04}.
 

### 2021-03-02 manuscriptPGPE.tex
Update manuscript

@@
   \centering
-  \includegraphics[width=18.0cm]{./Figs/HGorderparametersPCvsSURFchangeDIHEDRALS.eps}
+  \includegraphics[width=18.0cm]{./Figs/HGorderparametersPCvsSURFchangeDIHEDRALSlog.eps}
   \caption{\label{changesWITHsurf}
@@
 while PC and PE exhibit symmetric distributions.
-Larger barriers for crossings between \textit{gauche} and \textit{trans} states were observed for
-the last dihedral in PS and PG than for PC and PE lipids. 
+Also, the energy barriers for O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ dihedral
+rotations between \textit{gauche} and \textit{trans} states are larger for
+PS and PG lipids than for PC and PE lipids. 
 Rest of the dihedrals are similar between different lipids,
-with the exception of PS lipids for which slightly larger energy for ecplised anti conformation next to phosphate group
-was observed.
+with the exception of PS lipids for which slightly larger energy for ecplised anti conformation
+in g$_3$-O$_{{\mathrm g}3}$-P-O$_\alpha$ dihedral was observed.
 Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids.
@@
 the lipid headgroups are able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
-Our results support the models proposing free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
+%Our results support the models proposing free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
 The wide range of observed conformations suggest that the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and that the models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in liquid lamellar state.

### 2021-03-03 manuscriptPGPE.tex
Update manuscript

@@
     from CHARMM36 simulations compared with the experimental data \cite{scherer89}.
-    \textbf{B)} Changes in PC headgroup conformational ensembles upon increasing the amount of positive charge in bilayer,
-    characterized by the heavy atom dihedral distributions, from CHARMM36 simulations.
+    \textbf{B)} Relative energies for individual dihedral angles estimated from inverse Boltzmann distributions of heavy atom dihedral angles
+    with different amounts of cationic surfactant from CHARMM36 simulations.
   }
@@
 
-To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge, we analyzed the simulations (CHARMM36) that correctly capture the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
-Heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B reveal that the decrease of trans state probabilities in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
-dihedrals are the major effects of positive charge on lipid headgroup conformations.
+To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge,
+we calculated the heavy atom dihedral angle distributions from 
+simulations %(CHARMM36)
+that correctly captured the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
+The dihedral angle distributions and relative energies
+in figures \ref{HGorderparametersPCvsSURFdihedrals} and \ref{changesWITHsurf} B reveal that the
+addition positive charge into a membrane 
+decrease the abundance of trans states in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
+dihedrals.
 Choline region remains essentially unchanged and only minor changes are observed in other dihedrals even though almost half of the molecules in membranes are cationic surfactants.
 
-Also binding of ions may affect the lipid headgroup conformational ensembles in physiological conditions.
-In line with the results from cationic surfactants, the decreased probability for the trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral is the main structural consequence of
-bound Ca$^{2+}$ ion to PC headgroup (Fig.~\ref{DIHSwithCAlipid17eccPOPC} and \ref{DIHSwithCAcharmm36POPC}) in the most realistic MD simulations models (lipid17ecc and CHARMM36 in Fig.~\ref{changesWITHCaClPG}).
-%Lipid17ecc (having the most realistic response of PC lipid headgroup order parameters to the calcium binding in Fig.~\ref{changesWITHCaClPG}) and CHARMM36 simulations. 
-%Response of charged lipids to the ion binding is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
-The most realistic simulations suggest
-%istic response of PG $\beta$-carbon to CaCl$_2$  even thought the calcium binding affinity was overestimated in these models.
-%These models suggest
-that the upward tilting of the headgroup dipole upon addition of CaCl$_2$ is weaker in PG than in PC headgroup,
-%(\ref{changesWITHCaClPG}),
-but dihedral distributions are more sensitive to the bound ions in PG (Lipid17 and Slipids in Figs.~\ref{changesWITHCaClPG},~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
-However, the changes in dihedrals upon addition of calcium differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and~\ref{DIHSwithCAlipid17POPG}), none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added calcium in PG is not available.
-Also the lipid headgroup conformational ensembles in mixed PC:PE and PC:PG membranes were not possible to resolve with the currently available force fields and experimental data (Figs.~\ref{HGorderparametersPCvsPEPG} and~\ref{HGorderparametersPGvsPCchange}).
+Also binding of ions to membranes may affect the lipid headgroup conformational ensembles in physiological conditions.
+The bound Ca$^{2+}$ ion to PC headgroup leads to similar decrease in trans state probability for g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
+as observed for cationic surfactants in the most realistic MD simulations models
+(lipid17ecc and CHARMM36 in Figs.~\ref{DIHSwithCAlipid17eccPOPC}, \ref{DIHSwithCAcharmm36POPC} and Fig.~\ref{changesWITHCaClPG}).
+The dihedral distributions of PG headgroup are more sensitive to the bound ions in the most realistic simulations,
+but upward tilting of the headgroup dipole upon addition of CaCl$_2$ is weaker than in PC
+(Lipid17 and Slipids in Figs.~\ref{changesWITHCaClPG},~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+However, the changes in PG lipid dihedrals upon addition of CaCl$_2$ differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and~\ref{DIHSwithCAlipid17POPG}), none of the simulations captures the Ca$^{2+}$ ion binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added CaCl$_2$ in PG is not available.
+Also the headgroup conformational ensembles in mixtures of PC and charged (PG or PS) or zwitterionic (PE)
+lipids could not be resolve with the currently available force fields and experimental data
+(Figs.~\ref{HGorderparametersPCvsPEPG} and~\ref{HGorderparametersPGvsPCchange}, and Ref. \cite{antila19,melcr20}).
 
-%Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.
-
-Despite the 
-%limited capability of simulations to interpret some of the experimental data due to the 
-inaccuries in simulations in some cases, we can conclude that the experimentally observed changes in headgroup order parameters upon addition of charges into bilayer arise from relatively small changes in conformational ensembles.
+Despite the difficulties in simulations to capture the headgroup conformations in mixtures involving charged lipids,
+we can conclude that the structural response of lipid headgroups to 
+%experimentally observed changes in headgroup order parameters upon addition of
+membrane bound charges arise from relatively small changes in conformational ensembles.
 These changes eventuate from mild changes in dihedral angle distributions, rather than from restriction of lipids into
-fixed conformations.
-Therefore, lipid headgroups remain in disordered state sampling large space of different conformations, thereby being able to interact with different molecules in various ways, also in membranes bearing electric charge.
+fixed conformations. Therefore, lipid headgroups remain in disordered state sampling large space of different conformations
+also in charged membranes, thereby being able to interact with different molecules in multiple ways.
 

### 2021-03-03 manuscriptPGPEsuppl.tex
Update text related to the relative energies
Update manuscript

@@
 \clearpage
-\section{Changes in headgroup conformations upon addition of CaCl$_2$}
+\section{Dihedral angle distributions and the analysis of relative energies}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=8.0cm]{./Figs/DIHEDRALS.eps}
+  \caption{\label{DIHdists}
+    Heavy atom dihedral angle distributions from CHARMM36 simulations that correctly capture the
+    order parameter differences between the force fields.
+  }
+\end{figure}
+
+\clearpage
+\section{Changes in headgroup conformations upon addition of charged surfactants or CaCl$_2$}
+
+\begin{figure}[]
+  \centering
+  \includegraphics[width=12.0cm]{./Figs/HGorderparametersPCvsSURFdihedrals.eps}
+  \caption{\label{HGorderparametersPCvsSURFdihedrals}
+    Changes in PC headgroup conformational ensembles upon increasing the amount of positive charge in bilayer,
+    characterized by the heavy atom dihedral distributions, from CHARMM36 simulations.
+  }
+\end{figure}
 

### 2021-03-09 manuscriptPGPE.tex
Update text

@@
 energy costs for rotating individual dihedrals to almost any angle is low (below 7 kT).
-Only trans states of P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P have larger relative
+Only $cis$ states of P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P have larger relative
 energies and are not observed for any lipids during the simulations.
@@
 
-To analyze the alterations in lipid conformations when bound to proteins,
-we calculated the heavy atom dihedral distributions from lipid conformations within protein structures deposited in the PDB \cite{berman00} in Fig.~\ref{dihedralsFROMpdb}.
-\todo{This part is to be finished once the analysis if finalized}
-We found 176 PC, 198 PE, 70 PG, and 41 PS conformations, which
-present lipids that are tightly bound to proteins in fixed conformations
-that can be determined as a part of protein structure using crystallography or cryo-EM.
+\begin{figure*}[]
+  \centering
+  \includegraphics[width=19.0cm]{./Figs/figure4.eps}
+  \caption{\label{dihedralsFROMpdb}
+    A) Dihedral distributions fromm lipid structures in PDB.
+    B) Identical structures of PE and PC, and PG and PC headgroups found from PDB.
+  }
+\end{figure*}
 
-\begin{figure}[]
-  \centering
-  \includegraphics[width=8.0cm]{./Figs/DIHEDRALSALLfromPDB.eps}
-%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
-%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPGfromPDB.eps}
-%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPSfromPDB.eps}
-  \caption{\label{dihedralsFROMpdb}
-    Dihedral distributions from simulations and lipid structures in PDB.
-  }
-\end{figure}
+Interpretation of experimental order parameters using MD simulations
+in previous sections suggest that PC, PE, PG and PS lipid headgroups are very flexible,
+thereby being able to bind proteins in various different conformations.
+To test this prediction, we analyze the protein bound lipid conformations
+from structures structures deposited in the PDB \cite{berman00}.
+We found 311 PC, 394 PE, 154 PG, and 35 PS conformations presenting
+lipids that are tightly bound to proteins in fixed conformations
+such that their structures were determined as a part of protein structure using crystallography or cryo-EM.
 
-Similarly to the liquid lamellar state, lipid dihedrals have very wide range of angles
-when bound to proteins. Only %restrictions seem to be in 
-P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals
-seem to avoid $cis$ conformations as observed also in liquid lamellar state.
-The %range of available angles for
-O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and
-g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals are even less restricted in protein bound state
-than in liquid lamellar phase where $cis$ and $anti$ eclipsed state were not present (Fig.~\ref{structures}). 
-Major differences between different lipid headgroups are not observed when bound to proteins.
-Slight preference for the trans state in O$_\alpha$-C$_\alpha$-C$_\beta$-/C$_\gamma$ dihedral of PC,
-and for positive angles in P-O$_\alpha$-C$_\alpha$-C$_\gamma$ dihedral of PS lipids with respect
-to other lipids could be present in the data, but the statistics is not sufficient for conclusions.
-Therefore, the differences in conformational ensembles between different
-lipids in liquid lamellar state in Fig.~\ref{structures} are not seen in the protein bound states.
+The heavy atom dihedral angle distributions calculated from these conformations 
+in Fig.~\ref{dihedralsFROMpdb} A) reveal that indeed lipids are bound to proteins in wide
+range of conformations independently on the heasgroup type.
+As observed also in liquid lamellar state, only conformations
+with P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals in $cis$ from are almost
+completely absent in all lipids.
+Significant differences in bound conformations between different headgroups could not be
+observed with the available data.
+%The %range of available angles for
+%O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and
+%g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals are even less restricted in protein bound state
+%than in liquid lamellar phase where $cis$ and $anti$ eclipsed state were not present (Fig.~\ref{structures}). 
+%Major differences between different lipid headgroups are not observed when bound to proteins.
+%Slight preference for the trans state in O$_\alpha$-C$_\alpha$-C$_\beta$-/C$_\gamma$ dihedral of PC,
+%and for positive angles in P-O$_\alpha$-C$_\alpha$-C$_\gamma$ dihedral of PS lipids with respect
+%to other lipids could be present in the data, but the statistics is not sufficient for conclusions.
+%Therefore, the differences in conformational ensembles between different
+%lipids in liquid lamellar state in Fig.~\ref{structures} are not seen in the protein bound states.
 
-This suggests that lipids can bind to proteins in wide range of conformations independently
-of headgroup chemistry. Thereby, lipids compromise their preferred conformations when binding
-to proteins and the binding of lipids to specific locations would be driven by 
-the intermolecular interactions between proteins and lipids.
-However, it is important to note that lipids are often not the main target in
-structures deposited in PDB and therefore their conformations may be less reliable than those of proteins.
-Stereochemical violations and structures deviating from lipid crystals have been
-previously proposed to indicate inaccuracies in lipid structures in PDB \cite{marsh13b,pezeshkian18}. 
+Altogether our results suggest that lipids can bind to proteins in wide range of conformations
+compromising their preferred structure to optimize the intermolecular interactions with proteins.
+Therefore, the specific binding of lipids to proteins is not driven by the structural differences between
+headgroups. This is demonstrated in Fig.~\ref{dihedralsFROMpdb} B) showing different lipids bound
+to different proteins with identical structures.
+%However, it is important to note that lipids are often not the main target in
+%structures deposited in PDB and therefore their conformations may be less reliable than those of proteins.
+Structures deviating from lipid crystals
+%and stereochemical violations
+have been previously proposed to indicate inaccuracies in lipid structures in PDB \cite{marsh13b,pezeshkian18}. 
 However, we see large deviations from lipid crystals structures also in conformational ensembles

### 2021-03-16 manuscriptPGPEsuppl.tex
Fitting title, authors, and affiliations on one page in SI.
Tables in readable orientation.

@@
 
-\author{Pavel Buslaev}
-\affiliation{University of Jyv{\"a}skyl{\"a}}
+\author{\scriptsize Pavel Buslaev}
+\affiliation{\tiny University of Jyv{\"a}skyl{\"a}}
 
@@
 \begin{document}
-
-\newpage
-%\tableofcontents
+\tableofcontents{}
 
@@
 
-%\begin{table*}[htb]
+\begin{table*}[htb]
 %
-\begin{sidewaystable*}[!p]
+%\begin{sidewaystable*}[!p]
   \centering
@@
   \begin{minipage}[t]{\textwidth}
+  \resizebox{\columnwidth}{!}{
     \begin{tabular}{l c c r r r r r r c c}
@@
       % some footnotes are not visible in typeset-MS (pdf)
-      lipid/counter-ions & force field for lipids / ions  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      lipid & force field for lipids / ions  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
       \hline
@@
     \end{tabular}
+    }
   \end{minipage}
@@
   \todo{LIPID17 simulations with correct dihedrals still coming}
-%\end{table*}
-  \end{sidewaystable*}
+\end{table*}
+%  \end{sidewaystable*}
 
-  %    \begin{table*}[htb]
-  \begin{sidewaystable*}[!p]
+      \begin{table*}[htb]
+  %\begin{sidewaystable*}[!p]
   \centering
@@
   \todo{Citation and ion model for GROMOS-CKP?}
-%\end{table*}
- \end{sidewaystable*}
+\end{table*}
+% \end{sidewaystable*}
 
-%\begin{table*}[htb]
-  \begin{sidewaystable*}[!p]
+\begin{table*}[htb]
+%  \begin{sidewaystable*}[!p]
   \centering
@@
   \begin{minipage}[t]{\textwidth}
+    \resizebox{\columnwidth}{!}{
     \begin{tabular}{l c c r r r r r r c c}
@@
     \end{tabular}
+    }
   \end{minipage}
     \todo{Citation and ion model for GROMOS-CKP?} \\
-%\end{table*}
-\end{sidewaystable*}
+\end{table*}
+%\end{sidewaystable*}
 
-  \begin{sidewaystable*}[!p]
+\begin{table*}
+%  \begin{sidewaystable*}[!p]
   \centering
@@
     \end{tabular}
+    }
   \end{minipage}
   \todo{Citation and description for ''Berger'' model?} \\
-%\end{table*}
-\end{sidewaystable*}
+\end{table*}
+%\end{sidewaystable*}
 

### 2021-03-16 manuscriptPGPE.tex
Update text

@@
     A) Dihedral distributions fromm lipid structures in PDB.
-    B) Identical structures of PE and PC, and PG and PC headgroups found from PDB.
+    B) PE headgroup bound to cytochrome {\it bc}$_1$ complex (1PP9 \cite{huang05})
+    with identical conformation as PC headgroup to yeast sec14 (3B7Q \cite{schaaf08}),
+    and PG headgroup bound to bovine cytochrome {\it c} oxidase (3AG3 \cite{muramoto10})
+    as PC to FraC, a pore-forming toxin (4TSQ \cite{tanaka15}).
   }
@@
 thereby being able to bind proteins in various different conformations.
-To test this prediction, we analyze the protein bound lipid conformations
-from structures structures deposited in the PDB \cite{berman00}.
+To test this prediction, we analyzed the protein bound lipid conformations
+from structures deposited in the PDB \cite{berman00}.
 We found 311 PC, 394 PE, 154 PG, and 35 PS conformations presenting
 lipids that are tightly bound to proteins in fixed conformations
-such that their structures were determined as a part of protein structure using crystallography or cryo-EM.
+that were determined as a part of protein structure using crystallography or cryo-EM.
 
 The heavy atom dihedral angle distributions calculated from these conformations 
-in Fig.~\ref{dihedralsFROMpdb} A) reveal that indeed lipids are bound to proteins in wide
-range of conformations independently on the heasgroup type.
-As observed also in liquid lamellar state, only conformations
-with P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals in $cis$ from are almost
-completely absent in all lipids.
-Significant differences in bound conformations between different headgroups could not be
-observed with the available data.
+in Fig.~\ref{dihedralsFROMpdb} A) reveal that the protein bound lipids indeed exhibit wide
+range of conformations independently on the headgroup type.
+As in bulk lipid bilayers, only $cis$ conformations
+of P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals are almost
+completely restricted in all lipids, and significant differences between
+different headgroups were not observed.
+Structures deviating from lipid crystals 
+have been previously proposed to indicate inaccuracies in lipid structures in PDB \cite{marsh13b,pezeshkian18}. 
+However, we see large deviations from lipid crystals structures also in conformational ensembles
+that reproduce the NMR data in liquid lamellar phase, thereby proposing that such deviations
+are realistic also in protein-bound states.
 %The %range of available angles for
@@
 
-Altogether our results suggest that lipids can bind to proteins in wide range of conformations
-compromising their preferred structure to optimize the intermolecular interactions with proteins.
-Therefore, the specific binding of lipids to proteins is not driven by the structural differences between
-headgroups. This is demonstrated in Fig.~\ref{dihedralsFROMpdb} B) showing different lipids bound
-to different proteins with identical structures.
+Our results suggest that flexible lipid headgroups can optimize the intermolecular interactions with proteins by binding
+in wide range of conformations.
+Therefore, the specific binding of lipids to proteins is not driven by the structural differences between headgroups.
+This is demonstrated in Fig.~\ref{dihedralsFROMpdb} B) with two examples where different lipids bound to
+different proteins have identical conformations:
+PE in cytochrome {\it bc}$_1$ complex is similar to PC bound to yeast sec14,
+and PG bound to bovine cytochrome {\it c} oxidase is similar to PC bound to pore-forming toxin (FraC).
+On the other hand, a single lipid headgroup type is cabable to adapt itself to various binding positions, 
+thereby being able to specifically bind to many different kinds of binding sites in different proteins.
+
 %However, it is important to note that lipids are often not the main target in
 %structures deposited in PDB and therefore their conformations may be less reliable than those of proteins.
-Structures deviating from lipid crystals
-%and stereochemical violations
-have been previously proposed to indicate inaccuracies in lipid structures in PDB \cite{marsh13b,pezeshkian18}. 
-However, we see large deviations from lipid crystals structures also in conformational ensembles
-that reproduce the NMR data in liquid lamellar phase, thereby proposing that such deviations
-are realistic also in protein-bound states.
 
@@
 
-Conformational ensembles resolved using solid state NMR experiments and MD simulations
-from NMRlipids open collaboration revealed that headgroups of the most abundant biological
-phospholipids, PC, PE, PG and PS, sample a wide range of different conformations in the lamellar liquid state.
-Differences in NMR order parameters between different headgroups can be explained
-by the changes in dihedral angle distributions, suggesting that
-similar conformations are accessed by all headgroups, but with different probabilities.
-Also the changes in order parameters upon the addition of charged molecules---such as cationic lipids, surfactants, or drugs---in membranes can be explained by the changes in dihedral angle probabilities, particularly close to the phosphate region.
+C-H bond order parameters from NMR experiments %in this and previous studies \cite{wohlgemuth80,buldt81,seelig87}
+suggest that lipid headgroup conformational ensembles depend on lipid type (PC, PE, PG or PS) 
+and accumulated membrane charge (cationic lipids, surfactants, ions or drugs). 
+Our interpretation of this data, using massive amount of simulations collected
+within the NMRlipids open collaboration,
+revealed that the differences in order parameters can be explained
+by relatively small changes in dihedral angle probability distributions.
+All studied headgroups (PC, PE, PG and PS) are flexible
+and access similar wide range of conformations with only very few resticted dihedral orientations
+also when charges are bound to membranes.
+The observed differences in order parameters originate from
+reweighting conformational probabilities rather than changes in accessible structures.
 
-Wide range of conformations are observed also in lipids that are tightly bound to proteins in PDB,
-suggesting that the specific binding of lipids to proteins is dominated by the intermolecular lipid--protein
-interactions, while the differences in conformational ensembles between different lipid types
-play a minor role. Therefore, lipids can conform themselves to multiple binding positions located
-in various proteins.
+The flexibility and wide conformational space of headgroups suggest that protein bound lipids can
+adapt to various binding sites to optimize the intermolecular lipid--protein interactions.
+We tested this prediction by analyzing the conformations of lipids that are tightly bound to proteins in from PDB.
+Indeed, also protein bound lipids exhibit wide range of conformations without significant 
+differences between different lipid types. Therefore, the specificify of lipid binding to proteins is not
+regulated by accessible structures of lipids, and a single lipid type can adapt to various binding sites in proteins.
 
-Our results pave the way to understanding how lipids regulate membrane protein function.
-Lipid conformational ensembles in liquid lamellar phase are necessary
-for the detailed analysis of lipid--protein interaction energetics, particularly for its entropic component. 
-Furthermore, our results demonstrate how the NMRlipids databank, containing MD simulations and NMR data,
-can be used to resolve conformational ensembles of disordered biomolecules in a membrane environment,
-and how this information can be coupled with structural data in the PDB.
-We believe that this combination brings us closer to the comprehensive understanding of
-complex biomolecular assemblies containing both disordered and structural regions, such as membrane
-protein complexes.
+Our results pave the way toward understanding of lipid mediated cell signaling and how lipids regulate membrane protein function in general.
+We suggest that the key to understand selective binding of certain lipid types to proteins
+are intermolecular lipid-protein interactions, rather than conformational restrictions of lipids.
+On the other hand, wide conformational ensembles in bulk bilayers suggest that lipid crystal structures
+play a minor role and that entropic cost of lipid binding may be significant.
+Finally, the results demonstrate the power of open access MD simulations collected in NMRlipids open collaboration
+to complement PDB in order to understand functions of complex systems containing disordered biomolecules.
 

### 2021-03-17 manuscriptPGPE.tex
Add author
Update methods for the analysis from PDB as discussed in issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/40

@@
 \affiliation{Paris, France}
+
+\author{Ivan Gushchin}
+\affiliation{Moscow Institute of Physics and Technology, Dolgoprudny, Russia}
 
@@
 \subsection{Analysis of conformations of protein-bound lipids}
-\todo{This is to be finished once the final analysis is done as discussed in this issue: https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/40}
+Lipid structures from Protein Data Bank (PDB, \url{http://www.rcsb.org/})
+were searched using PDBe REST API (\url{www.ebi.ac.uk/pdbe/pdbe-rest-api})
+using the ligand names listed in the supplementary information.
+Heavy atom dihedral distributions from the lipid structures were calculated
+using the MDAnalysis python library \cite{agrawal11,gowers16} and
+Jupyter notebook available from \url{https://github.com/pbuslaev/scr/blob/master/PDB%20analysis.ipynb}.
+Only structures from X-ray and Cryo-EM with higher than 3.2~\AA~resolution were used in the analysis.
+Lipids from same PDBID with all dihedrals angles within 3 degrees were discarded from the analysis
+to avoid overweighting of lipid structures present, e.g., due to rotational symmetries.
 
-Lipid structures from Protein Data Bank (PDB, \url{http://www.rcsb.org/})
-were searched using PDBe REST API (\url{www.ebi.ac.uk/pdbe/pdbe-rest-api}).
-First, all PDB entries containing PC, PE, PG, or PS lipid headgroups were collected.
-The ligand names to identify the lipids were:
-PLC, PX4, 6PL, LIO, HGX, PC7, PC8, P1O, 6O8, XP5, EGY, PLD, SBM, HXG, and PCW for PC;
-8PE, PTY, 3PE, PEH, PEF, 6OE, 6O9, 9PE, PEV, 46E, SBJ, L9Q, PEK, EPH, ZPE, 9TL, 9Y0, 6OU, LOP, and PEE for PE;
-PGT, PGK, LHG, 44G, PGV, OZ2, D3D, PGW, DR9, P6L, PG8, H3T, and GOT for PG; and
-PSF, PS6, Q3G, P5S, D39, PS2, 17F, and 8SP for PS.
-Secondly, all PDB structures containing these ligands were downloaded and the dihedral angles of the first lipid in the latest version of the structures within each PDB code were calculated using the MDTraj python library \cite{mcgibbon15}.
-The used Jupyter notebook is available from the project's GitHub repository (\textit{ scripts/pdbSEARCH.ipynb}).
+To demonstrate the existence of similar structures of different lipids bound to different proteins,
+we searched pairs having the last five dihedral angles (excluding g$_1$-g$_2$-g$_3$-O$_{g3}$) within
+30 degrees from structures with resolution higher than 2.5~\AA. The two most representative examples
+were handpicked from the results.
 

### 2021-03-17 manuscriptPGPEsuppl.tex
Add author
Update methods for the analysis from PDB as discussed in issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/40

--- 2021-03-03 manuscriptPGPEsuppl.tex
@@
 \end{figure}
+
+\clearpage
+\section{Lipid ligand names in PDB used in the analysis of conformations of protein-bound lipids}
+{\bf PC:} PLC, PX4, 6PL, LIO, HGX, PC7, PC8, P1O, 6O8, XP5, EGY, PLD, SBM, HXG, and PCW \\
+{\bf PE:} 8PE, PTY, 3PE, PEH, PEF, 6OE, 6O9, 9PE, PEV, 46E, SBJ, L9Q, PEK, EPH, ZPE, 9TL, 9Y0, 6OU, LOP, and PEE \\
+{\bf PG:} PGT, PGK, LHG, 44G, PGV, OZ2, D3D, PGW, DR9, P6L, PG8, H3T, and GOT \\
+{\bf PS:} PSF, PS6, Q3G, P5S, D39, PS2, 17F, and 8SP.
 

### 2021-03-17 manuscriptPGPE.tex
Updates from Overleaf

--- 2021-03-09 manuscriptPGPE.tex
@@
 \author{Pavel Buslaev}
-\affiliation{University of Jyv{\"a}skyl{\"a}}
+\affiliation{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, Finland}
+\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, Dolgoprudny, Russia}
 
@@
      \textbf{B)} Headgroup and glycerol backbone region order parameters of different different lipids.
-     \textbf{C)} Relative energies for individual dihedral angles estimated from inverse Boltzmann distributions of heavy atom dihedral angles.
-     The states corresponding energies higher than 7 kT are not shown because they are not observed in simulations.
+     \textbf{C)} Distributions of heavy atom dihedral angles of different lipids from simulations.
      \textbf{D)} Distributions of P-N vector angle with respect to membrane normal.
@@
 
-To experimentally characterize the headgroup conformational ensembles of lipids that are not
-bound to proteins in an electrostatically neutral cell membrane, we measured the C--H bond order parameters
-and their signs of POPG and POPE in the liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
+To experimentally characterize the headgroup conformational ensembles of lipids that are not bound to proteins in an electrostatically neutral cell membrane, we measured the C--H bond order parameters and their signs of POPG and POPE in the liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
 Determination of headgroup and glycerol backbone order parameters and their signs
@@
   \centering
-  \includegraphics[width=18.0cm]{./Figs/HGorderparametersPCvsSURFchangeDIHEDRALSlog.eps}
+  \includegraphics[width=18.0cm]{./Figs/HGorderparametersPCvsSURFchangeDIHEDRALS.eps}
   \caption{\label{changesWITHsurf}
@@
     from CHARMM36 simulations compared with the experimental data \cite{scherer89}.
-    \textbf{B)} Relative energies for individual dihedral angles estimated from inverse Boltzmann distributions of heavy atom dihedral angles
-    with different amounts of cationic surfactant from CHARMM36 simulations.
+    \textbf{B)} Changes in PC headgroup conformational ensembles upon increasing the amount of positive charge in bilayer,
+    characterized by the heavy atom dihedral distributions, from CHARMM36 simulations.
   }
@@
 
-To understand the structural origin of distinct order parameters for PG and PS lipids,
-we first calculated the heavy atom dihedral angle distributions from simulations
-that best reproduce the differences between headgroups according to the quality
-evaluation in the SI and Refs. \cite{botan15,antila19} (Fig. \ref{DIHdists}).
-Then, we used the inverse Boltzmann distributions to estimate the
-energy costs for different dihedral angle orientations.
-The results in Fig.~\ref{structures}C) suggest that all lipid headgroups are very flexible and
-energy costs for rotating individual dihedrals to almost any angle is low (below 7 kT).
-Only $cis$ states of P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P have larger relative
-energies and are not observed for any lipids during the simulations.
-
-Major differences between headgroups are observed for the last two dihedrals in the headgroup end,
-O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$,
-which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS,
-while PC and PE exhibit symmetric distributions.
-Also, the energy barriers for O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ dihedral
-rotations between \textit{gauche} and \textit{trans} states are larger for
-PS and PG lipids than for PC and PE lipids. 
-Rest of the dihedrals are similar between different lipids,
-with the exception of PS lipids for which slightly larger energy for ecplised anti conformation
-in g$_3$-O$_{{\mathrm g}3}$-P-O$_\alpha$ dihedral was observed.
+%To resolve the conformational ensembles of lipid headgroups, we compared the headgroup and glycerol backbone C--H bond order parameters from different MD force fields to the experimental data. 
+%Although none of the simulations reproduce the headgroup order parameters within experimental accuracy for any of the analysed lipids (Figs.~\ref{HGorderParametersPE} and \ref{HGorderParametersPOPG} in the SI and Refs.
+\cite{botan15,antila19}), the CHARMM36
+%force field performs best for all the headgroups and 
+%simulations reproduce the experimentally observed distinct order parameters of PS and PG lipids (Fig.~\ref{structures}B). 
+To understand the structural origin of distinct order parameters for PG and PS lipids, we selected the simulations that best reproduce the differences between headgroups (based on the quality evaluation in the SI and Refs.
+\cite{botan15,antila19}) and calculated the distributions of heavy atom dihedral angles from simulations, shown in Fig.~\ref{structures}C.
+Major differences between headgroups are observed for the last two dihedrals in the headgroup end,  O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and P-O$_\alpha$-C$_\alpha$-C$_\beta$, which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS.
+Rest of the dihedrals are similar between different lipids, with the exception of PS lipids for which distinct distributions are observed particularly around phosphate group.
 Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids.
-The increased barriers for dihedral rotations may explain the more rigid headgroup structures in PS~\cite{browning80,buldt81}.
+This may explain the more rigid headgroup structures in PS~\cite{browning80,buldt81}.
 Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.~\ref{structures}D)).
@@
 
-In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase and sampled dihedral angles are within approximately same ranges for all headgroup types.
-% (Fig.
+All lipid headgroups sample very broad conformational ensembles in liquid lamellar phase and sampled dihedral angles are within approximately same ranges for all headgroup types (Fig.
 \ref{structures}).
-Because the rotation of dihedral angles to almost any position bears relatively low energy cost,
-%results suggest that all lipid headgroups are very flexible---
-%thereby being
-the lipid headgroups are able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
-%Our results support the models proposing free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
+The results suggest that all lipid headgroups are very flexible---thereby being able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
+Our results support the models that propose free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
 The wide range of observed conformations suggest that the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and that the models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
@@
 Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of lipid headgroups \cite{seelig87}.
-However, the detailed understanding on structural and energetic response of lipids to membrane bound electric charge is still lacking \cite{Semchyschyn04}.
+However, detailed understanding on lipid conformational ensembles in membranes bearing electric charge is still lacking \cite{Semchyschyn04}.
 
-To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge,
-we calculated the heavy atom dihedral angle distributions from 
-simulations %(CHARMM36)
-that correctly captured the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
-The dihedral angle distributions and relative energies
-in figures \ref{HGorderparametersPCvsSURFdihedrals} and \ref{changesWITHsurf} B reveal that the
-addition positive charge into a membrane 
-decrease the abundance of trans states in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
-dihedrals.
+To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge, we analyzed the simulations (CHARMM36) that correctly capture the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
+Heavy atom dihedral angle distributions in figure \ref{changesWITHsurf} B reveal that the decrease of trans state probabilities in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
+dihedrals are the major effects of positive charge on lipid headgroup conformations.
 Choline region remains essentially unchanged and only minor changes are observed in other dihedrals even though almost half of the molecules in membranes are cationic surfactants.
 
-Also binding of ions to membranes may affect the lipid headgroup conformational ensembles in physiological conditions.
-The bound Ca$^{2+}$ ion to PC headgroup leads to similar decrease in trans state probability for g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
-as observed for cationic surfactants in the most realistic MD simulations models
-(lipid17ecc and CHARMM36 in Figs.~\ref{DIHSwithCAlipid17eccPOPC}, \ref{DIHSwithCAcharmm36POPC} and Fig.~\ref{changesWITHCaClPG}).
-The dihedral distributions of PG headgroup are more sensitive to the bound ions in the most realistic simulations,
-but upward tilting of the headgroup dipole upon addition of CaCl$_2$ is weaker than in PC
-(Lipid17 and Slipids in Figs.~\ref{changesWITHCaClPG},~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
-However, the changes in PG lipid dihedrals upon addition of CaCl$_2$ differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and~\ref{DIHSwithCAlipid17POPG}), none of the simulations captures the Ca$^{2+}$ ion binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added CaCl$_2$ in PG is not available.
-Also the headgroup conformational ensembles in mixtures of PC and charged (PG or PS) or zwitterionic (PE)
-lipids could not be resolve with the currently available force fields and experimental data
-(Figs.~\ref{HGorderparametersPCvsPEPG} and~\ref{HGorderparametersPGvsPCchange}, and Ref. \cite{antila19,melcr20}).
+Also binding of ions may affect the lipid headgroup conformational ensembles in physiological conditions.
+In line with the results from cationic surfactants, the decreased probability for the trans state of g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral is the main structural consequence of
+bound Ca$^{2+}$ ion to PC headgroup (Fig.~\ref{DIHSwithCAlipid17eccPOPC} and \ref{DIHSwithCAcharmm36POPC}) in the most realistic MD simulations models (lipid17ecc and CHARMM36 in Fig.~\ref{changesWITHCaClPG}).
+%Lipid17ecc (having the most realistic response of PC lipid headgroup order parameters to the calcium binding in Fig.~\ref{changesWITHCaClPG}) and CHARMM36 simulations. 
+%Response of charged lipids to the ion binding is less well understood and more difficult to capture in MD simulations \cite{antila19,melcr20}.
+The most realistic simulations suggest
+%istic response of PG $\beta$-carbon to CaCl$_2$  even thought the calcium binding affinity was overestimated in these models.
+%These models suggest
+that the upward tilting of the headgroup dipole upon addition of CaCl$_2$ is weaker in PG than in PC headgroup,
+%(\ref{changesWITHCaClPG}),
+but dihedral distributions are more sensitive to the bound ions in PG (Lipid17 and Slipids in Figs.~\ref{changesWITHCaClPG},~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
+However, the changes in dihedrals upon addition of calcium differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and~\ref{DIHSwithCAlipid17POPG}), none of the simulations captures the calcium binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added calcium in PG is not available.
+Also the lipid headgroup conformational ensembles in mixed PC:PE and PC:PG membranes were not possible to resolve with the currently available force fields and experimental data (Figs.~\ref{HGorderparametersPCvsPEPG} and~\ref{HGorderparametersPGvsPCchange}).
 
-Despite the difficulties in simulations to capture the headgroup conformations in mixtures involving charged lipids,
-we can conclude that the structural response of lipid headgroups to 
-%experimentally observed changes in headgroup order parameters upon addition of
-membrane bound charges arise from relatively small changes in conformational ensembles.
+%Therefore, more accurate force fields are required for the detailed analysis of the interactions between ions and charged lipids, as concluded previously also for PS lipids \cite{antila19,melcr20}.
+
+Despite the 
+%limited capability of simulations to interpret some of the experimental data due to the 
+inaccuries in simulations in some cases, we can conclude that the experimentally observed changes in headgroup order parameters upon addition of charges into bilayer arise from relatively small changes in conformational ensembles.
 These changes eventuate from mild changes in dihedral angle distributions, rather than from restriction of lipids into
-fixed conformations. Therefore, lipid headgroups remain in disordered state sampling large space of different conformations
-also in charged membranes, thereby being able to interact with different molecules in multiple ways.
+fixed conformations.
+Therefore, lipid headgroups remain in disordered state sampling large space of different conformations, thereby being able to interact with different molecules in various ways, also in membranes bearing electric charge.
 
@@
 
-\begin{figure*}[]
+To analyze the alterations in lipid conformations when bound to proteins,
+we calculated the heavy atom dihedral distributions from lipid conformations within protein structures deposited in the PDB \cite{berman00} in Fig.~\ref{dihedralsFROMpdb}.
+\todo{This part is to be finished once the analysis if finalized}
+We found 176 PC, 198 PE, 70 PG, and 41 PS conformations, which
+present lipids that are tightly bound to proteins in fixed conformations
+that can be determined as a part of protein structure using crystallography or cryo-EM.
+
+\begin{figure}[]
   \centering
-  \includegraphics[width=19.0cm]{./Figs/figure4.eps}
+  \includegraphics[width=8.0cm]{./Figs/DIHEDRALSALLfromPDB.eps}
+%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPEfromPDB.eps}
+%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPGfromPDB.eps}
+%  \includegraphics[width=8.0cm]{../Figs/DIHEDRALSPSfromPDB.eps}
   \caption{\label{dihedralsFROMpdb}
-    A) Dihedral distributions fromm lipid structures in PDB.
-    B) Identical structures of PE and PC, and PG and PC headgroups found from PDB.
+    Dihedral distributions from simulations and lipid structures in PDB.
   }
-\end{figure*}
+\end{figure}
 
-Interpretation of experimental order parameters using MD simulations
-in previous sections suggest that PC, PE, PG and PS lipid headgroups are very flexible,
-thereby being able to bind proteins in various different conformations.
-To test this prediction, we analyze the protein bound lipid conformations
-from structures structures deposited in the PDB \cite{berman00}.
-We found 311 PC, 394 PE, 154 PG, and 35 PS conformations presenting
-lipids that are tightly bound to proteins in fixed conformations
-such that their structures were determined as a part of protein structure using crystallography or cryo-EM.
+Similarly to the liquid lamellar state, lipid dihedrals have very wide range of angles
+when bound to proteins. Only %restrictions seem to be in 
+P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals
+seem to avoid $cis$ conformations as observed also in liquid lamellar state.
+The %range of available angles for
+O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and
+g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals are even less restricted in protein bound state
+than in liquid lamellar phase where $cis$ and $anti$ eclipsed state were not present (Fig.~\ref{structures}). 
+Major differences between different lipid headgroups are not observed when bound to proteins.
+Slight preference for the trans state in O$_\alpha$-C$_\alpha$-C$_\beta$-/C$_\gamma$ dihedral of PC,
+and for positive angles in P-O$_\alpha$-C$_\alpha$-C$_\gamma$ dihedral of PS lipids with respect
+to other lipids could be present in the data, but the statistics is not sufficient for conclusions.
+Therefore, the differences in conformational ensembles between different
+lipids in liquid lamellar state in Fig.~\ref{structures} are not seen in the protein bound states.
 
-The heavy atom dihedral angle distributions calculated from these conformations 
-in Fig.~\ref{dihedralsFROMpdb} A) reveal that indeed lipids are bound to proteins in wide
-range of conformations independently on the heasgroup type.
-As observed also in liquid lamellar state, only conformations
-with P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals in $cis$ from are almost
-completely absent in all lipids.
-Significant differences in bound conformations between different headgroups could not be
-observed with the available data.
-%The %range of available angles for
-%O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and
-%g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals are even less restricted in protein bound state
-%than in liquid lamellar phase where $cis$ and $anti$ eclipsed state were not present (Fig.~\ref{structures}). 
-%Major differences between different lipid headgroups are not observed when bound to proteins.
-%Slight preference for the trans state in O$_\alpha$-C$_\alpha$-C$_\beta$-/C$_\gamma$ dihedral of PC,
-%and for positive angles in P-O$_\alpha$-C$_\alpha$-C$_\gamma$ dihedral of PS lipids with respect
-%to other lipids could be present in the data, but the statistics is not sufficient for conclusions.
-%Therefore, the differences in conformational ensembles between different
-%lipids in liquid lamellar state in Fig.~\ref{structures} are not seen in the protein bound states.
-
-Altogether our results suggest that lipids can bind to proteins in wide range of conformations
-compromising their preferred structure to optimize the intermolecular interactions with proteins.
-Therefore, the specific binding of lipids to proteins is not driven by the structural differences between
-headgroups. This is demonstrated in Fig.~\ref{dihedralsFROMpdb} B) showing different lipids bound
-to different proteins with identical structures.
-%However, it is important to note that lipids are often not the main target in
-%structures deposited in PDB and therefore their conformations may be less reliable than those of proteins.
-Structures deviating from lipid crystals
-%and stereochemical violations
-have been previously proposed to indicate inaccuracies in lipid structures in PDB \cite{marsh13b,pezeshkian18}. 
+This suggests that lipids can bind to proteins in wide range of conformations independently
+of headgroup chemistry. Thereby, lipids compromise their preferred conformations when binding
+to proteins and the binding of lipids to specific locations would be driven by 
+the intermolecular interactions between proteins and lipids.
+However, it is important to note that lipids are often not the main target in
+structures deposited in PDB and therefore their conformations may be less reliable than those of proteins.
+Stereochemical violations and structures deviating from lipid crystals have been
+previously proposed to indicate inaccuracies in lipid structures in PDB \cite{marsh13b,pezeshkian18}. 
 However, we see large deviations from lipid crystals structures also in conformational ensembles

### 2021-03-17 manuscriptPGPE.tex
Updates from Overleaf

--- 2021-03-17 manuscriptPGPE.tex
@@
 \author{Ivan Gushchin}
-\affiliation{Moscow Institute of Physics and Technology, Dolgoprudny, Russia}
+\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, Dolgoprudny, Russia}
 

### 2021-03-18 manuscriptPGPEsuppl.tex
Update to correct POPE lipid17 data as discussed in issue https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12

@@
   } \\
-  \todo{Lipid17 data should be updated to the one with correct dihedrals reported here https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12\#issuecomment-756641407}
 \end{figure}
@@
   }
-   \todo{Lipid17 data should be updated to the one with correct dihedrals reported here https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12\#issuecomment-756641407}
 \end{figure}
@@
   \end{minipage}
-  \todo{Simulations with added NaCl are not currently used here, maybe should be removed from the table?} \\
   \todo{Citation for CHARMM36 PE?} \\
@@
   \todo{Citations for Berger-* simulations?} \\
-  \todo{LIPID17 simulations with correct dihedrals still coming}
 \end{table*}
@@
   \end{minipage}
-    \todo{Simulations with added NaCl are not currently used here, maybe should be removed from the table?} \\
     \todo{Citations and ion model for CHARMM36?} \\
-  \todo{Lipid17 simulation with ions with correct dihedral potentials still coming?} \\
   \todo{Citation and ion model for GROMOS-CKP?}

### 2021-03-19 manuscriptPGPEsuppl.tex

@@
   \includegraphics[width=12.0cm]{./Figs/DIHEDRALSlipid17eccWITHCaClPOPC.eps}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
   \caption{\label{DIHSwithCAlipid17eccPOPC}
@@
    \includegraphics[width=12.0cm]{./Figs/DIHEDRALScharmm36WITHCaClPOPC.eps}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MCHARMM36.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MCHARMM36.png}  
   \caption{\label{DIHSwithCAcharmm36POPC}
@@
 
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_MSlipids.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
-%  \caption{\label{DIHSwithCA}
-%    Changes in POPC Slipids dihedrals with increasing amount of CaCl$_2$.
-%  }
-%\end{figure}
-
 \begin{figure}[]
@@
   \includegraphics[width=16.0cm]{./Figs/DIHEDRALSslipidsWITHCaClPOPG.eps}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_MSlipids.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_MSlipids.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_MSlipids.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_MSlipids.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_MSlipids.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_MSlipids.png}  
   \caption{\label{DIHSwithCAslipidsPOPG}
@@
 
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3N6_Mlipid17.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPC_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
-%  \caption{\label{DIHSwithCA}
-%    Changes in POPC lipid17 dihedrals with increasing amount of CaCl$_2$.
-%  }
-%\end{figure}
-
 \begin{figure}[]
@@
   \includegraphics[width=16.0cm]{./Figs/DIHEDRALSlipid17WITHCaClPOPG.eps}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17.png}
-  %\includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17.png}  
   \caption{\label{DIHSwithCAlipid17POPG}
@@
 \end{figure}
-
-%\begin{figure}[]
-%  \centering
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O3_M_M_G3C4_M_M_G3C5_M_M_G3C6_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3P2_M_M_G3O3_M_M_G3C4_M_M_G3C5_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3O1_M_M_G3P2_M_M_G3O3_M_M_G3C4_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G3_M_M_G3O1_M_M_G3P2_M_M_G3O3_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G2_M_M_G3_M_M_G3O1_M_M_G3P2_Mlipid17ecc.png}
-%  \includegraphics[width=6.0cm]{./Figs/dihedralFIGS/POPG_M_G1_M_M_G2_M_M_G3_M_M_G3O1_Mlipid17ecc.png}  
-%  \caption{\label{DIHSwithCA}
-%    Changes in POPG lipid17ecc dihedrals with increasing amount of CaCl$_2$.
-%  }
-%\end{figure}
 
@@
 and lipid mixtures with additional ions in Tables~\ref{systemsMIX} and \ref{systemsMIX2}.
-the simulation data are indexed in a
-searchable database available at \url{www.nmrlipids.fi},
-and in the NMRlipids/MATCH repository (\url{github.com/NMRlipids/MATCH}).
-The large set of MD simulation data was analysed using the development version of NMRlipids databank.
-Unique naming convention for lipid atoms in each force field was defined using the mapping files
-and analysis for all simulations indexed in NMRlipids databank manner were performed
-using Python codes.
+The simulations were analyzed using preliminary versions of the NMRlipids databank
+(\url{www.nmrlipids.fi}, \url{github.com/NMRlipids/MATCH} and \url{https://github.com/NMRLipids/NMRlipidsIVPEandPG/tree/master/Data/Simulations})
+and unique naming convention for lipid atoms (\url{http://nmrlipids.blogspot.com/2015/03/mapping-scheme-for-lipid-atom-names-for.html}),
+which enable automatic analysis of simulations with different force fields with varying atom naming conventions.
+The automatic analyses were implemented using MDAnalysis \cite{agrawal11,gowers16} and MDTraj \cite{mcgibbon15}
+python libraries, and tools in the GROMACS sofware package \cite{gromacsMANUAL}.
+All codes are available from the project's GitHub repository \cite{NMRlipidsIVbgit}.
 
@@
 the standard error of the mean were then calculated over different lipids.
-Python programs that use the MDAnalysis library \cite{agrawal11,gowers16}
-used for all atom simulations are available in Ref.~\citenum{MATCHgit}
-(\texttt{scripts/calcOrderParameters.py}).
-For united atom simulations, the trajectories with hydrogens having ideal geometry were constructed first using either \texttt{buildH} program~\cite{buildH} or (\texttt{scratch/opAAUA\_prod.py}) in  Ref.~\citenum{MATCHgit}, and the order parameters were then calculated from these trajectories.
+Code for all atom simulations is available in Ref.~\citenum{MATCHgit} (\texttt{scripts/calcOrderParameters.py}).
+For united atom simulations, we first constructed trajectories including hydrogens with ideal geometry using either \texttt{buildH} program~\cite{buildH} or (\texttt{scratch/opAAUA\_prod.py}) in  Ref.~\citenum{MATCHgit}, and the order parameters were then calculated from these trajectories.
 This approach has been tested against trajectories with explicit hydrogens and the deviations in order parameters are small \cite{buildH,piggot17}.\\
 %\todo{BuildH program is now cited with a direct link to the GitHub repo. I think that a release to Zenodo would be nice in the final publication.}\\
-The number density profiles of ions were calculated using the \texttt{gmx density} tool of the GROMACS sofware package \cite{gromacsMANUAL}.
 
@@
 
-\noindent {\it POPG} \todo{Simulation details by Ollila}.
+\noindent {\it POPG} Lipid bilayer containing 118 POPG molecules, 4110 TIP3P water molecules, and 118 potassium ions was build using CHARMM-GUI \cite{lee16}.
+The system was simulated 100~ns, coupled to 298 K using Nose-Hoover~\cite{nose84,hoover85} thermostat and 1 bar with
+semi-isotropic Parrinello-Rahman~\cite{parrinello81} pressure coupling.
+The used default parameters and force field files from CHARMM-GUI are available from \citenum{CHARMM36popg}.
+
+\todo{Simulation details for larger simulation by A. Peon.}
 
@@
 
-\noindent {\it POPC:POPG mixture with additional calcium} \todo{Simulation details by A. Kiirikki.}
+\noindent {\it POPC:POPG 1:1 and POPC:POPG 4:1 mixtures with additional calcium} 
+The initial structures were built with CHARMM-GUI Membrane Builder \cite{lee16}.
+The TIP3P water model was used to solvate the systems.
+The simulations were run for 400~ns with timestep 2~fs and the first 100~ns were discarded as equilibration time.
+The simulations were run with GROMACS version 2020.2~\cite{pall20}.
+The Nose-Hoover thermostat~\cite{nose84,hoover85} was used with temperature of 298~K and the time constant for temperature coupling was 1.0~ps.
+The semi-isotropic Parinello-Rahman barostat~\cite{parrinello81} was used with reference pressure 1.0~bar and with a time constant of 5.0~ps with compressibility of 4.5e-5~bar$^{-1}$.
+Long range electrostatic interactions were calculated with the PME method.
+All bonds with hydrogen atoms were constrained with LINCS algorithm.
+The simulation files are available from Refs.~\citenum{CHARMM36POPCPOPG5050,CHARMM36POPCPOPG5050100mMCaCl,CHARMM36POPCPOPG50501000mMCaCl,CHARMM36POPCPOPG4010,CHARMM36POPCPOPG4010100mMCaCl,CHARMM36POPCPOPG40101000mMCaCl}.
 
@@
 
-\noindent {\it POPC:POPG} \todo{Simulation details by S. Virtanen or O. H. S. Ollila.}
+\noindent {\it POPC:POPG 4:1 and POPC:POPG 1:1 mixtures}
+Initial structures were build by removing appropriate amount of lipids from POPC:POPG 7:3 mixture available at Ref.~\cite{POPCPOPG73lipid17}.
+Force field parameters from the same reference were used \todo{We still need description from A. Peon how these were obtained},
+except that incorrect dihedrals with type 1 were changed to type 9
+(for details, see discussion in here \url{https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}).
+Simulations were performed using the Gromacs simulation package~\cite{pall20} with time step of 2~fs. 
+The non-bonded interactions were calculated directly within 1.0~nm cutoff; the Verlet scheme was used\cite{Pall13};
+and the long-range electrostatic forces were calculated using particle mesh Ewald~\cite{essman95}. 
+The bond lengths of hydrogen atoms were constrained using LINCS~\cite{hess97}.
+Temperature was coupled to the velocity rescaling thermostat \cite{bussi07} at 298~K with a coupling constant of 1~ps.
+Pressure was coupled to the Parrinello--Rahman barostat \cite{parrinello81} at 1~bar with a coupling constant of 10~ps.
 

### 2021-03-19 manuscriptPGPEsuppl.tex

@@
 
+\usepackage{hyperref} % to make TOC titles clickable
+\hypersetup{
+    colorlinks=true, %set true if you want colored links
+    linktoc=all,     %set to all if you want both sections and subsections linked
+    linkcolor=black,  %choose some color if you want links to stand out
+    citecolor=black,
+    filecolor=black,
+    urlcolor=black
+}
+
 \author{\scriptsize Pavel Buslaev}
-\affiliation{\tiny University of Jyv{\"a}skyl{\"a}}
+\affiliation{\tiny Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, Finland}
+\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, Dolgoprudny, Russia}
 

### 2021-03-19 manuscriptPGPEsuppl.tex
Removed pure DOPE data, because no exp comparison.
Update simulation details

@@
       \hline
-      POPE  & CHARMM36 \cite{??}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
-      POPE  & CHARMM36 \cite{??}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
-      POPE  & CHARMM36 \cite{??}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
+      POPE  & CHARMM36 \cite{klauda10}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
+      POPE  & CHARMM36 \cite{klauda10}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
+      POPE  & CHARMM36 \cite{klauda10}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
       POPE  & CHARMM36ua \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
@@
   \end{minipage}
-  \todo{Citation for CHARMM36 PE?} \\
+  \todo{Citation for CHARMM36ua?} \\
   \todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
@@
       \hline
-      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}    &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
-      POPG             & CHARMM36 \cite{??}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
-      POPG             & CHARMM36 \cite{??}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
+      POPG/K$^+$  & CHARMM36 \cite{venable13}       &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
+      POPG             & CHARMM36 \cite{venable13}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
+      POPG             & CHARMM36 \cite{venable13}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
       \hline
@@
       \hline
-      POPG             & LIPID17 / Dang \cite{gould18,smith94}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
-      POPG             & LIPID17 \cite{??}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
+      POPG             & LIPID17 / Dang \cite{gould18,smith94,dang06}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
+      POPG             & LIPID17 \cite{gould18}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
       \hline
@@
   \end{minipage}
-    \todo{Citations and ion model for CHARMM36?} \\
   \todo{Citation and ion model for GROMOS-CKP?}
@@
 %      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
-      POPC                   & CHARMM36 \cite{??}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
-      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
-      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
-      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
-      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
+      POPC                   & CHARMM36 \cite{klauda10}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
+      POPC:POPG (7:3)        & CHARMM36 \cite{klauda10,venable13}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
+      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
       \hline
-      POPC             & CHARMM36 \cite{??}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
-      POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
+      POPC             & CHARMM36 \cite{klauda10}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
+      POPC:POPE (1:1)  & CHARMM36 \cite{klauda10,venable13}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
       \hline
@@
 semi-isotropic Parrinello-Rahman~\cite{parrinello81} pressure coupling.
-The used default parameters and force field files from CHARMM-GUI are available from \citenum{CHARMM36popg}.
+The used default parameters and force field files from CHARMM-GUI were used.
+The used files are available from \citenum{CHARMM36popg}.
 
@@
 The Nose-Hoover thermostat~\cite{nose84,hoover85} was used with temperature of 298~K and the time constant for temperature coupling was 1.0~ps.
-The semi-isotropic Parinello-Rahman barostat~\cite{parrinello81} was used with reference pressure 1.0~bar and with a time constant of 5.0~ps with compressibility of 4.5e-5~bar$^{-1}$.
+The semi-isotropic Parrinello-Rahman barostat~\cite{parrinello81} was used with reference pressure 1.0~bar and with a time constant of 5.0~ps with compressibility of 4.5e-5~bar$^{-1}$.
 Long range electrostatic interactions were calculated with the PME method.
@@
 
-\noindent {\it DPPE with 288 lipids.} The starting structure for simulation was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}: 288 DPPE lipids and 9386 water molecules.
-The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
+\noindent {\it DPPE with 288 lipids.} The starting structure for simulation
+with 288 DPPE lipids and 9386 water molecules
+was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}. The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
 Simulation was performed for 200~ns, and the last 100~ns were used for the analysis.
@@
 
-\noindent {\it POPG with 288 lipids.} The starting structure for simulation was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}: 288 POPG lipids, 10664 water molecules and 288 Na ions.
-The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by Aqvist~\cite{aqvist90}.
+\noindent {\it POPG with 288 lipids.} The starting structure for simulation
+with 288 POPG lipids, 10664 water molecules and 288 Na ions
+was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by \AA qvist~\cite{aqvist90}.
 Simulation was performed for 250~ns, and the last 100~ns were used for the analysis. Same simulation conditions as DPPE with reference temperature of 298~K.
@@
 
-\noindent {\it DPPG with 288 lipids.} The starting structure for simulation was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}: 288 DPPG lipids, 11232 water molecules and 288 Na ions.
-The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by Aqvist~\cite{aqvist90}.
+\noindent {\it DPPG with 288 lipids.} The starting structure for simulation
+with 288 DPPG lipids, 11232 water molecules and 288 Na ions
+was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by \AA qvist~\cite{aqvist90}.
 For the 298~K temperature, simulation was performed for 400~ns, and the last 100~ns were used for the analysis.
@@
 
-\noindent {\it POPC:POPG 4:1 and POPC:POPG 1:1 mixtures}
-Initial structures were build by removing appropriate amount of lipids from POPC:POPG 7:3 mixture available at Ref.~\cite{POPCPOPG73lipid17}.
+\noindent {\it POPC:POPG 4:1 and POPC:POPG 1:1 mixtures with different CaCl$_2$ concentrations}
+Initial structures were build by removing appropriate amount of lipids from POPC:POPG 7:3 mixture available from Ref.~\citenum{POPCPOPG73lipid17}.
 Force field parameters from the same reference were used \todo{We still need description from A. Peon how these were obtained},
 except that incorrect dihedrals with type 1 were changed to type 9
-(for details, see discussion in here \url{https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}).
-Simulations were performed using the Gromacs simulation package~\cite{pall20} with time step of 2~fs. 
+(for details, see discussion in \url{https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}).
+Simulations were performed using the Gromacs simulation package~\cite{pall20} with the time step of 2~fs. 
 The non-bonded interactions were calculated directly within 1.0~nm cutoff; the Verlet scheme was used\cite{Pall13};
@@
 Pressure was coupled to the Parrinello--Rahman barostat \cite{parrinello81} at 1~bar with a coupling constant of 10~ps. 
+For simulations with CaCl$_2$, appropriate amount of ions with Dang~\cite{smith94,dang06} parameters were added into the solvent.
+The simulation files are available from
+Refs.~\citenum{Lipid17POPCPOPG8020,Lipid17POPCPOPG8020100mMCaCl,Lipid17POPCPOPG80201000mMCaCl,Lipid17POPCPOPG5050,Lipid17POPCPOPG5050100mMCaCl,Lipid17POPCPOPG50501000mMCaCl}
 
 \subsection{Lipid17ecc}
-
-\todo{This is to be finished and POPC:POPG mixtures to be described}
-In ECC-lipid models, electronic continuum correction (ECC) is applied to 
-implicitly include the missing electronic polarizability into the force field description \cite{melcr18,melcr19}.
-In practise, this is implemented by scaling the charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
-of Amber Lipid14/17 models are scaled by constant factors.
-Here, we follow the approach that previously improved ion binding to bilayers containing negatively charged PS lipids \cite{melcr19}:
-ECC-POPC parameters (scaling factors $f_q$=0.8 and $f_\sigma$=0.89 applied to Lipid14 POPC parameters) \cite{melcr18} were used for POPC
-and scaling factors of $f_q$=0.75 and $f_\sigma$=0.89 were applied to the charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
-of Amber Lipid17 POPG parameters. The Lipid17 parameters (described above) and initial configurations
-were taken from Ref. \cite{POPGlipid17} with the correct dihedral type,
-and the resulting parameters are available from Ref. \citenum{??}. ECC-ion parameters with the scaled charges \cite{pluharova14,kohagen16,martinek18},
-downloaded from \url{bitbucket.org/hseara/ions/src/master/}, were used in these simulations.
+\noindent {\it POPC:POPG 4:1 and POPC:POPG 1:1 mixtures with different CaCl$_2$ concentrations}
+Implicit inclusion of electronic polarizability by electronic continuum correction (ECC),
+implemented by scaling the partial charges in force fields, can be used to improve ion interactions with lipids
+and other biomolecules in classical MD simulations~\cite{dijon20}.
+For Amber Lipid14/17 force fields, ECC has been previously implemented by scaling the
+charges and Lennard-Jones~$\sigma$s of headgroup, glycerol backbone, and carbonyl regions
+by constant factors~\cite{melcr18,melcr20}.
+Here, we apply similar ECC approach to Amber Lipid17 PG parameters as done previously for PS \cite{melcr20}:
+charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
+of parameters POPG from Ref.~\citenum{POPCPOPG73lipid17} were scaled by factors of $f_q$=0.75 and $f_\sigma$=0.89, respectively
+(and the dihedral types were corrected to type 9 as in previous section).
+Previously introduced ECC-POPC parameters (scaling factors $f_q$=0.8 and $f_\sigma$=0.89 applied to Lipid14 POPC parameters)
+were used for POPC~\cite{melcr18}.
+ECC-ion parameters with the scaled charges \cite{pluharova14,kohagen16,martinek18}
+from \url{bitbucket.org/hseara/ions/src/master/}, and SPC/E water model~\cite{berendsen87} were used in these simulations. 
+Rest of the simulation parameters and initial configurations were taken from Lipid17
+simulations~\cite{Lipid17POPCPOPG8020,Lipid17POPCPOPG8020100mMCaCl,Lipid17POPCPOPG80201000mMCaCl,Lipid17POPCPOPG5050,Lipid17POPCPOPG5050100mMCaCl,Lipid17POPCPOPG50501000mMCaCl}.
+Simulation files of Lipid17ecc simulations are available from
+Refs.~\citenum{Lipid17eccPOPCPOPG8020,Lipid17eccPOPCPOPG8020100mMCaCl,Lipid17eccPOPCPOPG80201000mMCaCl,Lipid17eccPOPCPOPG5050,Lipid17eccPOPCPOPG5050100mMCaCl,Lipid17eccPOPCPOPG50501000mMCaCl}.
 

### 2021-03-19 manuscriptPGPEsuppl.tex
Comment out data that is not used

@@
       POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
-      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
+%      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
       \hline
@@
       POPE  & Berger-largeH \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
-      DOPE  & Berger-Vries \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
-      DOPE  & Berger-largeH \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
+%      DOPE  & Berger-Vries \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
+%      DOPE  & Berger-largeH \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
       \hline
@@
       POPC:POPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
-      POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
-     \hline
-      DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
-      DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}  &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
+      %POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
+    % \hline
+    %  DOPC             & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 256 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPCberger300K} \\
+      %DOPC:DOPE (1:1)  & Berger \cite{??}   \todoi{This is probable not plain berger, correct force filed should be described.}  &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{DOPC1DOPE1berger} \\
     \end{tabular}
@@
 
-\noindent {\it DOPE} Data is available at \cite{bergerDOPEfiles,berger2DOPEfiles}. \todo{Simulation details by  T. Piggot.}
+%\noindent {\it DOPE} Data is available at \cite{bergerDOPEfiles,berger2DOPEfiles}. \todo{Simulation details by  T. Piggot.}
 
-\noindent {\it POPC:POPE, POPC:DOPE and DOPC:DOPE mixtures} Data is available at \cite{POPCberger300K,POPC1POPE1berger}. 
+\noindent {\it POPC:POPE mixtures} Data is available at \cite{POPCberger300K,POPC1POPE1berger}. 
 300 K with v-rescale (tau=0.1 ps),
@@
 
-\noindent {\it DOPE} Data is available at \cite{gromosCKPdope}. \todo{Simulation details by  T. Piggot.}
+%\noindent {\it DOPE} Data is available at \cite{gromosCKPdope}. \todo{Simulation details by  T. Piggot.}
 

### 2021-03-23 manuscriptPGPE.tex
Updates from Overleaf
Update text

@@
 \usepackage{csquotes}
-\usepackage{xr}
+\usepackage{xr-hyper}
+\usepackage{hyperref} % to make TOC titles clickable
+\hypersetup{
+    colorlinks=true, %set true if you want colored links
+    linktoc=all,     %set to all if you want both sections and subsections linked
+    linkcolor=black,  %choose some color if you want links to stand out
+    citecolor=black,
+    filecolor=black,
+    urlcolor=black
+}
 
@@
   Chemistry of lipid headgroups, the water facing components of cell membranes that regulate cell functions via lipid--protein interactions, varies between organisms and organelles.
-Because lipid membranes are in liquid state under physiological conditions, the conformational ensembles of different lipid headgroups are difficult to resolve experimentally.
+Because membranes are in liquid state under physiological conditions,
+  individual lipids are not fixed to single conformations but they sample ensemble of conformations with the Boltzmann weighted probabilities.
+  However, these ensembles and their dependence on lipid types have not yet been experimentally determined. 
+  Therefore, it has not been clear if lipid molecules exchange between few rigid conformations,
+  or if they can freely fluctuate in all possible conformations.
   Here, we combine solid state NMR experiments and molecular dynamics simulations from NMRlipids open collaboration to resolve the conformational ensembles of the headgroups of key lipid types in liquid lamellar phase under various biologically relevant conditions.
 Interpretation of NMR experiments using the plethora of simulation data collected in the NMRlipids project suggests that all lipid headgroups sample a wide range of conformations in neutral and charged cellular membranes.
-However, the populations of different conformations are dictated by the headgroup chemistry.
+  Differences in the headgroup chemistry between different lipid types manifests
+  in probability distributions of conformations, but all lipid types can access to almost any possible conformations.
   Together with the analysis of protein-bound lipids from the protein data bank (PDB), this suggests that lipids can bind to proteins in wide range of conformations independently on the headgroup chemistry.
 Therefore, the selective adsorption of proteins to membranes is likely regulated by specific protein--lipid interactions, rather than conformational restrictions of lipids.
-Our results pave the way to comprehensive understanding of lipid--protein interaction energetics, and the understanding of complex biomolecular assemblies such as membrane proteins.
+  Our results pave the way to comprehensive understanding of
+  lipid mediated signaling and lipid--protein interactions in biomedical applications.
 
@@
 functions by altering membrane properties such as charge or elasticity \cite{lee03,lemmon08}.
-While the specific interactions with certain lipid headgroups
+Specific interactions with certain lipid headgroups
 are known to be essential for the function of several proteins \cite{lee11,lemmon08},
-it is not clear if the specificity is driven by the differences in accessible
-conformational states between lipid types or
-by specific intermolecular lipid--protein interactions.
+but it is not clear if the specificity is driven by the
+differences in accessible conformations between lipid types or
+by specific lipid--protein interactions.
 
-The conformational ensembles of lipids in the physiologically relevant lamellar liquid phase
-are typically derived from NMR experiments, particularly from 
-C--H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}. Notably, such measurements can be performed also on living cells \cite{gally81,scherer87,seelig90}.
-These studies suggest that the glycerol backbone conformations are largely similar irrespectively of the headgroup \cite{gally81}, 
-and the headgroup conformations are similar in  phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG) lipids, while the headgroup is more rigid in phosphatidylserine (PS) lipids \cite{wohlgemuth80,buldt81}.
-However, these tendencies are based only on the absolute values, whereas
-the necessity of order parameter signs in capturing the conformational ensembles of lipids has been recently demonstrated \cite{botan15,ollila16,ferreira16}.
-Furthermore, the detailed understanding of lipid conformational ensembles
-is limited by the lack of universal models that would map order parameters to structural ensembles \cite{pezeshkian18,akutsu20}.
+Stuctures of protein-bound lipids are available in the protein data bank (PDB) \cite{berman00}
+and crystal structures of lipids have been determined \cite{buldt81,pascher92},
+but their relation to the conformational ensembles in bulk membranes in liquid lamellar phase remains unclear \cite{marsh13b}.
+%In addition to the changes in lipid conformational ensembles upon binding to proteins,
+%also the experimentally measured response of lipid headgroup to membrane bound charges remains poorly understood 
+%due to the lack of suitable models to interpret the lipid conformational ensembles in liquid lamellar state \cite{Semchyschyn04}.
+Most accurate experimental information on conformational ensembles of lipids
+%in the physiologically relevant
+in liquid lamellar phase are typically derived from NMR experiments, particularly from 
+C--H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}.
+According to these experiments, the glycerol backbone conformations are largely similar irrespectively of the headgroup \cite{gally81}, 
+the headgroup conformations are similar in  phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG) lipids,
+while the headgroup of phosphatidylserine (PS) lipids is more rigid \cite{wohlgemuth80,buldt81}. 
+However, order parameters signs are not accessible in $^2$H NMR experiments \cite{ollila16}
+and universal models to map order parameters to structural ensembles are not available \cite{pezeshkian18,akutsu20}.
+Therefore, it is not clear if the conformational ensemble of lipid headgroups in liquid lamellar bilayer
+composes of exchange between few restricted conformations, or if lipid headgroups can fluctuate freely in wide range of conformations.
 
-Stuctures of different lipid types in protein bound states can be extracted from the protein data bank (PDB) \cite{berman00},
-but their relation to the conformational ensembles in liquid lamellar state remains unclear \cite{marsh13b}.
-In addition to the changes in lipid conformational ensembles upon binding to proteins,
-also the experimentally measured response of lipid headgroup to membrane bound charges remains poorly understood 
-due to the lack of suitable models to interpret the lipid conformational ensembles in liquid lamellar state \cite{Semchyschyn04}.
-
-Here, we use natural abundance $^{13}$C NMR experiments and MD simulations from the NMRlipids open collaboration
+Here, we use natural abundance $^{13}$C NMR experiments %(enabling the determination of order parameter signs)
+and MD simulations from the NMRlipids open collaboration
 to resolve the differences in conformational ensembles between PC, PE, PG and PS lipid headgroups.
+We elucidate also the effect of charges and protein binding to lipid headgroup conformations.
 Zwitterionic PC and PE are the most common lipids in eukaryotes and bacteria, respectively \cite{vanmeer08,sohlenkamp16}.
@@
 and affect membrane protein functionality and signaling \cite{lemmon08,leventis10,sohlenkamp16,hariharan18}.
-All the studied lipids specifically bind to various proteins \cite{yeagle14}. We use our results to elucidate also lipid--protein interactions and the effect of charges on lipid conformations.
+All the studied lipids specifically bind to various proteins \cite{yeagle14}.
+%We use our results to elucidate also lipid--protein interactions and the effect of charges on lipid conformations.
 
-The lipid conformational ensembles in liquid lamellar state paves the way toward understanding the
-specific binding of different lipid types to membrane proteins and how they regulate the protein function.
-Because glycerol backbone and headgroup structures of PC lipids are similar in model membranes and in bacteria \cite{gally81,scherer87,seelig90},
-the results from model systems could be used to understand the biological role of lipid headgroup conformational ensembles in different lipid types.
+Similarity of glycerol backbone and headgroup order parameters in model membranes and bacteria \cite{gally81,scherer87,seelig90}
+suggest that our results can be used to 
+%the results from model systems could be used to
+understand the biological role of lipid headgroups in lipid--protein interactions.
+These interactions are crucial, for example, in lipid mediated signaling \cite{lemmon08} and
+design of phoshoplipid-specific antibodies \cite{vigant15}. 
+%The lipid conformational ensembles in liquid lamellar state paves the way toward understanding the
+%specific binding of different lipid types to membrane proteins and how they regulate the protein function.
+
+%Notably, such measurements can be performed also on living cells \cite{gally81,scherer87,seelig90}.
 
@@
 The headgroup and glycerol backbone C--H bond order parameters of 1-palmitoyl-2-oleoyl-sn-glycero-3-phosphoethanolamine (POPE) and 1-palmitoyl-2-oleoyl-sn-glycero-3-phospho-(1'-rac-glycerol) (POPG), purchased from Avanti polar lipids, were measured using natural abundance $^{13}$C solid state NMR spectroscopy as described previously \cite{ferreira13,ferreira16}.
-The magnitudes of order parameters were determined from the chemical-shift resolved dipolar splittings using a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04} and the signs from S-DROSS experiments~\cite{gross97} combined with SIMPSON simulations \cite{bak00}.
+The magnitudes of order parameters were determined from the chemical-shift resolved dipolar splittings using a R-type Proton Detected Local Field (R-PDLF) experiment~\cite{dvinskikh04}, and the signs from S-DROSS experiments~\cite{gross97} combined with SIMPSON simulations \cite{bak00}.
 %The experiments were done in a Bruker Avance III 400 spectrometer operating at a $^1$H Larmor frequency of 400.03 MHz.
@@
 %\url{github.com/NMRlipids/NMRlipidsIVotherHGs}. 
-Simulation and accessibility details of more than 70 systems simulated for this work are given in the supplementary information.
-Quality evaluation of these simulations enables us to select the best models for the interpretation of lipid conformational ensembles from the experimental data.
+Simulation details, accessibility information and quality evaluation of more than 70 systems simulated for this work are in the supplementary information.
 
-The quality of lipid headgroup and glycerol backbone conformational ensembles in PE and PG simulations were evaluated using the C-H bond order parameters \cite{botan15}.
-Interactions between different headgroups in simulations of mixed bilayers were evaluated monitoring the changes in headgroup order parameters upon mixing the lipids \cite{antila19}.
+Best models for the interpretation of lipid headgroup conformational ensembles from the experimental data were selected
+using quality evaluation measures defined in the NMRlipids project.
+Conformational ensembles of headgroup and glycerol backbone in PE and PG simulations were evaluated using the C-H bond order parameters \cite{botan15}.
+Interactions between different headgroups were evaluated monitoring the changes in headgroup order parameters upon mixing the lipids \cite{antila19}.
 The ion binding affinities and response of lipids to bound charge were evaluated monitoring the changes of lipid headgroup order parameters \cite{catte16,antila19}.
@@
   %  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCPSPEPG.eps}
-   \includegraphics[width=18.0cm]{./Figs/figure1.eps}
+   \includegraphics[width=18.0cm]{./Figs/figure1.pdf}
    \caption{\label{HGorderParameters}
@@
    }
-   \todo{This is a sketch, Tiago Ferreira will make a new figure.} \\
+   %\todo{This is a sketch, Tiago Ferreira will make a new figure.} \\
 %  \todo{D) could be clarified as Fig. 2 in the NMRlipids IVps paper.}
@@
 
-\subsection{Analysis of conformations of protein-bound lipids}
+\subsection{Analysis of protein-bound lipid conformations}
 Lipid structures from Protein Data Bank (PDB, \url{http://www.rcsb.org/})
@@
      \textbf{B)} Headgroup and glycerol backbone region order parameters of different different lipids.
-     \textbf{C)} Relative energies for individual dihedral angles estimated from inverse Boltzmann distributions of heavy atom dihedral angles.
-     The states corresponding energies higher than 7 kT are not shown because they are not observed in simulations.
+     \textbf{C)} Relative energies for individual heavy atom dihedral angles estimated from the inverse Boltzmann formula.
+     Angles corresponding energies above 7 kT are not shown because they are not observed in simulations.
      \textbf{D)} Distributions of P-N vector angle with respect to membrane normal.
@@
 
-\subsection{Differences between lipid headgroups in bulk bilayer from $^{13}$C NMR experiments}
+\subsection{Differences between lipid headgroups from $^{13}$C NMR experiments}
 
-To experimentally characterize the headgroup conformational ensembles of lipids that are not
+To experimentally characterize the differences in headgroup conformational ensembles of lipids that are not
 bound to proteins in an electrostatically neutral cell membrane, we measured the C--H bond order parameters
@@
 These carbons have overlapping peaks in the INEPT spectra due to their similar chemical environments,
-and only the magnitude of the larger order parameter could be determined from the R-PDLF spectra (Fig.~\ref{HGorderParameters}B)).
-Nevertheless, based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85},
+and only the magnitude of the larger order parameter could be determined from the R-PDLF spectra (Fig.~\ref{HGorderParameters}B).
+Based on previous $^2$H NMR measurements \cite{wohlgemuth80,gally81,borle85},
 we assigned the larger order parameter to the g$_2$ carbon
@@
 The decrease in the beginning of the S-DROSS curve suggests that the sign of larger g$_2$ order parameter
-is negative and later increase suggests that sign of smaller $\beta$ order parameter is positive (Fig.~\ref{HGorderParameters}B)).
+is negative and later increase suggests that sign of smaller $\beta$ order parameter is positive (Fig.~\ref{HGorderParameters}B).
 This interpretation is confirmed by SIMPSON calculations in Fig.~\ref{POPGsimpson}.
@@
 The overall agreement of order parameters determined by different research teams and different techniques for the same lipid headgroup is very good here and in previous studies \cite{botan15,ollila16,antila19}.
-This suggests that the observed differences between lipid types arise from differences in headgroup chemistry rather than inaccuracies in experiments, or differences in the acyl chains or in experimental conditions.
-The most distinct order parameters are observed for PS headgroups, for which the $\alpha$-carbon order parameter exhibits significant forking and the $\beta$-carbon has more negative value than other studied lipid types.
+Therefore, the differences between lipid types are dictated by the headgroup chemistry rather than inaccuracies in experiments, or differences in the acyl chains or in experimental conditions.
+
+The most distinct order parameters are observed for PS headgroups, for which the $\alpha$-carbon order parameter exhibits significant forking (different values for different hydrogens in the same carbon) and the $\beta$-carbon has more negative value than other studied lipid types.
 On the other hand, the $\beta$-carbon order parameter of PG headgroup has a positive sign, in contrast to all the other lipid types.
-Notably, this has not been observed in traditional $^2$H NMR experiments, where only the absolute value of the order parameters are measured~\cite{wohlgemuth80,gally81,borle85}.
+This has not been previously observed  in $^2$H NMR experiments detecting only the absolute values~\cite{wohlgemuth80,gally81,borle85}.
 The glycerol backbone order parameters are similar for all the lipid types, although they move slightly toward positive values (closer to zero) in the order PC $<$ PE $<$ PS $<$ PG.
-Essential differences between PC and PE headgroups are not observed.
+Only minor differences between PC and PE headgroups are observed.
 
-\subsection{Conformational ensembles of different lipid headgroups in bulk bilayer from MD simulations}
+\subsection{Conformational ensembles of different lipid headgroups from MD simulations}
 
@@
   \caption{\label{changesWITHsurf}
-    \textbf{A)} Modulation of PC headgroup order parameters and P--N vector angle upon addition of cationic surfactant
+    \textbf{A)} Modulation of PC headgroup order parameters and P--N vector angle upon addition of cationic surfactant (dihexadecyldimethylammonium)
     from CHARMM36 simulations compared with the experimental data \cite{scherer89}.
@@
 evaluation in the SI and Refs. \cite{botan15,antila19} (Fig. \ref{DIHdists}).
-Then, we used the inverse Boltzmann distributions to estimate the
+Then, we used the inverse Boltzmann formula to estimate the
 energy costs for different dihedral angle orientations.
@@
 However, the differences between PC and PE in P-O$_\alpha$-C$_\alpha$-C$_\beta$ dihedral
-and P--N vector dipole may be artificial as the $\beta$-carbon order parameter in PC is too negative in the CHARMM36 force field, thereby not being equal to the order parameter in PE as observed in experiments \cite{botan15}.
+and P--N vector dipole may be artificial as the $\beta$-carbon order parameter in PC is too negative even in the best available force field, thereby not being equal to the order parameter in PE as observed in experiments \cite{botan15}.
 
@@
 %thereby being
-the lipid headgroups are able to adopt wide range of multiple conformations when interacting with proteins, ions or other biomolecules.
+the lipid headgroups are able to adopt wide range of conformations when interacting with proteins, ions or other biomolecules.
 %Our results support the models proposing free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
-The wide range of observed conformations suggest that the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and that the models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
+%The wide range of observed conformations suggest that
+Therefore, the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
 
-\subsection{Lipid conformational ensembles in lipid bilayers with bound ions}
+\subsection{Lipid conformational ensembles in charged membranes}
 
@@
 in figures \ref{HGorderparametersPCvsSURFdihedrals} and \ref{changesWITHsurf} B reveal that the
-addition positive charge into a membrane 
+addition of positive charge into a membrane 
 decrease the abundance of trans states in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
@@
 The bound Ca$^{2+}$ ion to PC headgroup leads to similar decrease in trans state probability for g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
-as observed for cationic surfactants in the most realistic MD simulations models
+in the most realistic MD simulations models
+as observed for cationic surfactants
 (lipid17ecc and CHARMM36 in Figs.~\ref{DIHSwithCAlipid17eccPOPC}, \ref{DIHSwithCAcharmm36POPC} and Fig.~\ref{changesWITHCaClPG}).
@@
 
-Despite the difficulties in simulations to capture the headgroup conformations in mixtures involving charged lipids,
-we can conclude that the structural response of lipid headgroups to 
+%Despite the difficulties in simulations to capture the headgroup conformations in mixtures involving charged lipids,
+In conclusion, the structural response of lipid headgroups to 
 %experimentally observed changes in headgroup order parameters upon addition of
-membrane bound charges arise from relatively small changes in conformational ensembles.
-These changes eventuate from mild changes in dihedral angle distributions, rather than from restriction of lipids into
-fixed conformations. Therefore, lipid headgroups remain in disordered state sampling large space of different conformations
-also in charged membranes, thereby being able to interact with different molecules in multiple ways.
+membrane bound charges arise
+%from relatively small changes in conformational ensembles. These changes eventuate
+from mild changes in dihedral angle distributions, rather than restriction of lipids to
+fixed conformations. Therefore, lipid headgroups sample large space of different conformations
+also in charged membranes, thereby retaining the capability to adapt multiple conformations when
+interacting with proteins or other molecules.
 
@@
   \caption{\label{dihedralsFROMpdb}
-    A) Dihedral distributions fromm lipid structures in PDB.
-    B) PE headgroup bound to cytochrome {\it bc}$_1$ complex (1PP9 \cite{huang05})
-    with identical conformation as PC headgroup to yeast sec14 (3B7Q \cite{schaaf08}),
-    and PG headgroup bound to bovine cytochrome {\it c} oxidase (3AG3 \cite{muramoto10})
-    as PC to FraC, a pore-forming toxin (4TSQ \cite{tanaka15}).
+    A) Heavy atom dihedral angle distributions calculated from lipid structures in PDB.
+    B) The structure of PE headgroup bound to cytochrome {\it bc}$_1$ complex (PDB ID: 1PP9 \cite{huang05})
+    with identical conformation as the PC headgroup bound to yeast sec14 (PDB ID: 3B7Q \cite{schaaf08}),
+    and the structure of PG headgroup bound to bovine cytochrome {\it c} oxidase (PDB ID: 3AG3 \cite{muramoto10})
+    with identical conformation as PC to FraC, a pore-forming toxin (PDB ID: 4TSQ \cite{tanaka15}).
   }
@@
 thereby being able to bind proteins in various different conformations.
-To test this prediction, we analyzed the protein bound lipid conformations
+To test this prediction, we analyzed the protein-bound lipid conformations
 from structures deposited in the PDB \cite{berman00}.
-We found 311 PC, 394 PE, 154 PG, and 35 PS conformations presenting
-lipids that are tightly bound to proteins in fixed conformations
+We found 311 PC, 394 PE, 154 PG, and 35 PS protein-bound lipid conformations
+%presenting lipids that are tightly bound to proteins in fixed conformations
 that were determined as a part of protein structure using crystallography or cryo-EM.
@@
 of P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals are almost
-completely restricted in all lipids, and significant differences between
+completely absent in all lipids, and significant differences between
 different headgroups were not observed.
@@
 and PG bound to bovine cytochrome {\it c} oxidase is similar to PC bound to pore-forming toxin (FraC).
-On the other hand, a single lipid headgroup type is cabable to adapt itself to various binding positions, 
+On the other hand, a single lipid headgroup type is capable to adapt itself to various binding positions, 
 thereby being able to specifically bind to many different kinds of binding sites in different proteins.
@@
 and accumulated membrane charge (cationic lipids, surfactants, ions or drugs). 
-Our interpretation of this data, using massive amount of simulations collected
+Our interpretation of this data, using MD simulations collected
 within the NMRlipids open collaboration,
@@
 All studied headgroups (PC, PE, PG and PS) are flexible
-and access similar wide range of conformations with only very few resticted dihedral orientations
+and access similar wide range of conformations with only very few restictions in dihedral orientations
 also when charges are bound to membranes.
@@
 
-The flexibility and wide conformational space of headgroups suggest that protein bound lipids can
+The flexibility and wide conformational space of headgroups suggest that protein--bound lipids can
 adapt to various binding sites to optimize the intermolecular lipid--protein interactions.
@@
 play a minor role and that entropic cost of lipid binding may be significant.
-Finally, the results demonstrate the power of open access MD simulations collected in NMRlipids open collaboration
-to complement PDB in order to understand functions of complex systems containing disordered biomolecules.
+Finally, we demonstrated the power of open access MD simulation data from the NMRlipids open collaboration
+to complement the data in PDB in revealing how complex systems containing disordered biomolecules behave.
 

### 2021-03-23 manuscriptPGPEsuppl.tex
Removed 300K MacRog POPE data, because exp data is at 310K, for which there is another data set.
Removed 313K Gromos-CKP POPE data, because exp data is at 310K, for which there is another data set.
Update text

@@
 Conclusions are the same for all lipids: None of the force fields correctly captures the lipid headgroup conformational ensembles, but CHARMM36 gives results closest to experiments.
+Most importantly for this work, the CHARMM36 captures the distinct headgroup order parameters for PG and PS lipids observed in NMR experiments (Figs.
+\ref{HGorderParameters} and \ref{structures} in the main text).
 
@@
 Similar observations in our previous work for PS lipids were explained by the overestimated counterion
-binding affinity that neturalizes the effect of added negative charge \cite{antila19}.
+binding affinity that neutralizes the effect of added negative charge \cite{antila19}.
 All simulations except Berger-OPLS predict tilting of P-N headgroup outwards from the membrane and
@@
 These results are not in line with experiments where the PC headgroup order parameters are not affected by zwitterionic lipids \cite{scherer87}.
-The good performance of Berger-OPLS simulations in here is surprising because
+The good performance of Berger-OPLS simulations is surprising here because
 headgroup conformational enemble is not very close to experiments in this model and
@@
 
+\subsection{Dihedral angles and relative energies of PC, PE, PG and PS headgroups}
+
+We estimated the relative energies of each dihedral angle value
+with respect to the most probable value (lowest energy) 
+from the inverse Boltzmann formula $\Delta E(\theta) = -kT \left[\ln\left[p(\theta)\right]-\ln\left[p(\theta_0)\right] \right]$,
+where $p(\theta)$ is the dihedral angle distribution and $\theta_0$ is the most probable angle from MD simulation.
+The dihedral angle distributions are shown in Fig. \ref{DIHdists}
+and relative energies in Fig. \ref{structures} in the main text.
+
 \begin{figure}[]
   \centering
-  \includegraphics[width=8.0cm]{./Figs/DIHEDRALS.eps}
+  \includegraphics[width=7.0cm]{./Figs/DIHEDRALS.eps}
   \caption{\label{DIHdists}
@@
 \clearpage
-\section{Changes in headgroup conformations upon addition of charged surfactants or CaCl$_2$}
+\subsection{Changes in headgroup conformations upon addition of charged surfactants or CaCl$_2$}
 
@@
       POPE  & CHARMM36 \cite{klauda10}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
-      POPE  & CHARMM36 \cite{klauda10}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
+%      POPE  & CHARMM36 \cite{klauda10}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
       POPE  & CHARMM36ua \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
@@
       POPE  & Slipids \cite{jambeck12b}            & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
-      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
+%      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
       \hline
@@
       POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
-      POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
+%      POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
 %      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
@@
       POPE             & LIPID17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
-      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
+%      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
     \end{tabular}
@@
       POPG/K$^+$  & CHARMM36 \cite{venable13}       &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
-      POPG             & CHARMM36 \cite{venable13}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
+%      POPG             & CHARMM36 \cite{venable13}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
       POPG             & CHARMM36 \cite{venable13}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
@@
       POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0    & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGslipids} \\
-      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
+%      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
       \hline
       POPG             & LIPID17 / Dang \cite{gould18,smith94,dang06}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
-      POPG             & LIPID17 \cite{gould18}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
+%      POPG             & LIPID17 \cite{gould18}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
       \hline
       POPG             & GROMOS-CKP \cite{??}         & 0              & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP} \\
-      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 49 &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
+%      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 49 &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
     \end{tabular}
@@
 
+\clearpage
+
 \subsection{CHARMM36}
@@
 
-\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.}
+%\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.}
 
@@
 
-\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.}
+%\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.}
 
@@
 
+\noindent {\it POPC and cationic surfactant (dihexadecyldimethylammonium) mixture}
+Intial structures were taken from similar previously published \cite{melcr18} simulations with Amber lipid14 force field,
+which are available from Ref.
+\citenum{POPClipid14T313K,POPClipid1410perCATsurfT313K,POPClipid1420perCATsurfT313K,POPClipid1430perCATsurfT313K,POPClipid1442perCATsurfT313K,POPClipid1450perCATsurfT313K}.
+Default simulations parameters and force field files from CHARMM-GUI \cite{lee16} were used, except for dihexadecyldimethylammonium
+for which the atom types and partial charges of Amber lipid14 parameters from previous work \cite{melcr18} were modified to correspond Charmm36
+force field. 
+Systems contained 50 POPC molecules, 3983 water molecules, and 12, 30, 44, or 88 dihexadecyldimethylammonium molecules.
+Chloride ions were used as counterions for dihexadecyldimethylammonium.
+Reference system without cationic surfactants contained 200 POPC and 9000 water molecules.
+Systems were simulated 200 ns (the first 20 ns was discarded as an equilibration period)
+using Gromacs 5 \cite{abraham2015gromacs} at the temperature of 313 K.
+All simulation files are available from Refs. \citenum{charmmPOPC313K,CHARMM36cationicSURF}.
+
 \subsection{CHARMM36ua}
@@
 
-\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.
-  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.
-}
+%\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.
+%  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.
+%}
 
@@
 
-\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.
-  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
+%\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.
+%  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
 
@@
 
-\noindent {\it POPC:POPG mixture with additional NaCl} \todo{Simulation details by A. Peon.
-  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
+%\noindent {\it POPC:POPG mixture with additional NaCl} \todo{Simulation details by A. Peon.
+%  I have assumed that ion parameters are default Slipids, i.e., {\AA}qvist, please correct if this is not true.}
 

### 2021-03-23 manuscriptPGPEsuppl.tex
Naming of Berger-based models to match the earlier NMRlipids convention.

--- 2021-03-19 manuscriptPGPEsuppl.tex
@@
       \hline
-      POPE  & CHARMM36 \cite{klauda10}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
-      POPE  & CHARMM36 \cite{klauda10}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
-      POPE  & CHARMM36 \cite{klauda10}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
-      POPE  & CHARMM36ua \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
+      POPE  & CHARMM36 \cite{??}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
+      POPE  & CHARMM36 \cite{??}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
+      POPE  & CHARMM36 \cite{??}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
+      POPE  & CHARMM36-UA \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
       \hline
@@
       DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3655  &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
-      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
+%      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
       POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
       POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
-      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
+%      DOPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdope} \\
       \hline
@@
       POPE  & OPLS-MacRog \cite{rog16}     &0    & 144	& 5760     &0    & 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
-      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
+%      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
       \hline
-      POPE  & Berger-Vries \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
-      POPE  & Berger-largeH \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
-      DOPE  & Berger-Vries \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
-      DOPE  & Berger-largeH \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
+      POPE  & Berger-POPE-2004 \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
+      POPE  & Berger-POPE-2018 \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
+%      DOPE  & Berger-POPE-2004 \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
+%      DOPE  & Berger-POPE-2018 \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
       \hline
-      POPE             & LIPID17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
-      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
+      POPE             & Lipid17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
+      POPE             & Lipid17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
     \end{tabular}
@@
   \end{minipage}
-  \todo{Citation for CHARMM36ua?} \\
+  \todo{Citation for CHARMM36 PE?} \\
   \todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
@@
       \hline
-      POPG/K$^+$  & CHARMM36 \cite{venable13}       &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
-      POPG             & CHARMM36 \cite{venable13}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
-      POPG             & CHARMM36 \cite{venable13}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
+      POPG/K$^+$  & CHARMM36 \cite{??} \todoi{Correct citation for CHARMM POPG}    &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
+      POPG             & CHARMM36 \cite{??}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
+      POPG             & CHARMM36 \cite{??}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
       \hline
@@
       \hline
-      POPG             & LIPID17 / Dang \cite{gould18,smith94,dang06}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
-      POPG             & LIPID17 \cite{gould18}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
+      POPG             & LIPID17 / Dang \cite{gould18,smith94}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
+      POPG             & LIPID17 \cite{??}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
       \hline
@@
   \end{minipage}
+    \todo{Citations and ion model for CHARMM36?} \\
   \todo{Citation and ion model for GROMOS-CKP?}
@@
 %      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
-      POPC                   & CHARMM36 \cite{klauda10}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
-      POPC:POPG (7:3)        & CHARMM36 \cite{klauda10,venable13}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
-      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
-      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
-      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
+      POPC                   & CHARMM36 \cite{??}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
+      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{??}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{??}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
       \hline
-      POPC             & CHARMM36 \cite{klauda10}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
-      POPC:POPE (1:1)  & CHARMM36 \cite{klauda10,venable13}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
+      POPC             & CHARMM36 \cite{??}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
+      POPC:POPE (1:1)  & CHARMM36 \cite{??}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
       \hline
@@
 semi-isotropic Parrinello-Rahman~\cite{parrinello81} pressure coupling.
-The used default parameters and force field files from CHARMM-GUI were used.
-The used files are available from \citenum{CHARMM36popg}.
+The used default parameters and force field files from CHARMM-GUI are available from \citenum{CHARMM36popg}.
 
@@
 The Nose-Hoover thermostat~\cite{nose84,hoover85} was used with temperature of 298~K and the time constant for temperature coupling was 1.0~ps.
-The semi-isotropic Parrinello-Rahman barostat~\cite{parrinello81} was used with reference pressure 1.0~bar and with a time constant of 5.0~ps with compressibility of 4.5e-5~bar$^{-1}$.
+The semi-isotropic Parinello-Rahman barostat~\cite{parrinello81} was used with reference pressure 1.0~bar and with a time constant of 5.0~ps with compressibility of 4.5e-5~bar$^{-1}$.
 Long range electrostatic interactions were calculated with the PME method.
@@
 
-\noindent {\it DPPE with 288 lipids.} The starting structure for simulation
-with 288 DPPE lipids and 9386 water molecules
-was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}. The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
+\noindent {\it DPPE with 288 lipids.} The starting structure for simulation was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}: 288 DPPE lipids and 9386 water molecules.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system.
 Simulation was performed for 200~ns, and the last 100~ns were used for the analysis.
@@
 
-\noindent {\it POPG with 288 lipids.} The starting structure for simulation
-with 288 POPG lipids, 10664 water molecules and 288 Na ions
-was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}.
-The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by \AA qvist~\cite{aqvist90}.
+\noindent {\it POPG with 288 lipids.} The starting structure for simulation was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}: 288 POPG lipids, 10664 water molecules and 288 Na ions.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by Aqvist~\cite{aqvist90}.
 Simulation was performed for 250~ns, and the last 100~ns were used for the analysis. Same simulation conditions as DPPE with reference temperature of 298~K.
@@
 
-\noindent {\it DPPG with 288 lipids.} The starting structure for simulation
-with 288 DPPG lipids, 11232 water molecules and 288 Na ions
-was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}.
-The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by \AA qvist~\cite{aqvist90}.
+\noindent {\it DPPG with 288 lipids.} The starting structure for simulation was constructed with the MEMBRANE BUILDER website~\cite{ghahremanpour13}: 288 DPPG lipids, 11232 water molecules and 288 Na ions.
+The TIP3P~\cite{jorgensen83} water model was used to solvate the system and Ions are described by the parameters derived by Aqvist~\cite{aqvist90}.
 For the 298~K temperature, simulation was performed for 400~ns, and the last 100~ns were used for the analysis.
@@
 
-\noindent {\it POPE} Data is available at \cite{bergerPOPEfiles,berger2POPEfiles}. \todo{Simulation details by  T. Piggot.}
+Following the earlier convention in the NMRlipids Project~\cite{botan15}, 
+for the Berger-based models we use the following naming convention: 
+Berger - \{{\it molecule name}\} - \{{\it year when model published first time}\} \{{\it citation}\}.
+%The reason is that there are several different molecular topologies which are using the non-bonded parameters originally
+%developed by Berger et al.~\cite{berger97}. Thus the common factor in the Berger based models are the non-bonded parameters,
+%while the molecule specific parameters might somewhat vary. However, the majority of the molecular level topologies are 
+%relying (especially for the glycerol backbone and headgroup) on the parameters originally introduced by Marrink et al.~\cite{marrink98}.
 
-\noindent {\it DOPE} Data is available at \cite{bergerDOPEfiles,berger2DOPEfiles}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE} Data are available at Ref.~\citenum{bergerPOPEfiles} for Berger-POPE-2004~\cite{??} and at Ref.~\citenum{berger2POPEfiles} for Berger-POPE-2018~\cite{??}.
+\todo{Simulation details by  T.
+Piggot.}
+
+%\noindent {\it DOPE} Data is available at \cite{bergerDOPEfiles,berger2DOPEfiles}. \todo{Simulation details by  T. Piggot.}
 
@@
 
-\noindent {\it POPE} Data is available at \cite{gromosCKPpope}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE} Data is available at \cite{gromosCKPpopeT310}. \todo{Simulation details by  A. Peon.}
 
-\noindent {\it DOPE} Data is available at \cite{gromosCKPdope}. \todo{Simulation details by  T. Piggot.}
+%\noindent {\it DOPE} Data is available at \cite{gromosCKPdope}. \todo{Simulation details by  T. Piggot.}
 
@@
 
-\noindent {\it POPE} \todo{Simulation details by M. Javanainen and P. Fuchs.}
+\noindent {\it POPE} \todo{Simulation details by M. Javanainen} % and P. Fuchs.}
 
@@
 
-\noindent {\it POPC:POPG 4:1 and POPC:POPG 1:1 mixtures with different CaCl$_2$ concentrations}
-Initial structures were build by removing appropriate amount of lipids from POPC:POPG 7:3 mixture available from Ref.~\citenum{POPCPOPG73lipid17}.
+\noindent {\it POPC:POPG 4:1 and POPC:POPG 1:1 mixtures}
+Initial structures were build by removing appropriate amount of lipids from POPC:POPG 7:3 mixture available at Ref.~\cite{POPCPOPG73lipid17}.
 Force field parameters from the same reference were used \todo{We still need description from A. Peon how these were obtained},
 except that incorrect dihedrals with type 1 were changed to type 9
-(for details, see discussion in \url{https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}).
-Simulations were performed using the Gromacs simulation package~\cite{pall20} with the time step of 2~fs. 
+(for details, see discussion in here \url{https://github.com/NMRLipids/NMRlipidsIVPEandPG/issues/12}).
+Simulations were performed using the Gromacs simulation package~\cite{pall20} with time step of 2~fs. 
 The non-bonded interactions were calculated directly within 1.0~nm cutoff; the Verlet scheme was used\cite{Pall13};
@@
 Pressure was coupled to the Parrinello--Rahman barostat \cite{parrinello81} at 1~bar with a coupling constant of 10~ps.
-For simulations with CaCl$_2$, appropriate amount of ions with Dang~\cite{smith94,dang06} parameters were added into the solvent.
-The simulation files are available from
-Refs.~\citenum{Lipid17POPCPOPG8020,Lipid17POPCPOPG8020100mMCaCl,Lipid17POPCPOPG80201000mMCaCl,Lipid17POPCPOPG5050,Lipid17POPCPOPG5050100mMCaCl,Lipid17POPCPOPG50501000mMCaCl}
 
 \subsection{Lipid17ecc}
-\noindent {\it POPC:POPG 4:1 and POPC:POPG 1:1 mixtures with different CaCl$_2$ concentrations}
-Implicit inclusion of electronic polarizability by electronic continuum correction (ECC),
-implemented by scaling the partial charges in force fields, can be used to improve ion interactions with lipids
-and other biomolecules in classical MD simulations~\cite{dijon20}.
-For Amber Lipid14/17 force fields, ECC has been previously implemented by scaling the
-charges and Lennard-Jones~$\sigma$s of headgroup, glycerol backbone, and carbonyl regions
-by constant factors~\cite{melcr18,melcr20}.
-Here, we apply similar ECC approach to Amber Lipid17 PG parameters as done previously for PS \cite{melcr20}:
-charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
-of parameters POPG from Ref.~\citenum{POPCPOPG73lipid17} were scaled by factors of $f_q$=0.75 and $f_\sigma$=0.89, respectively
-(and the dihedral types were corrected to type 9 as in previous section).
-Previously introduced ECC-POPC parameters (scaling factors $f_q$=0.8 and $f_\sigma$=0.89 applied to Lipid14 POPC parameters)
-were used for POPC~\cite{melcr18}.
-ECC-ion parameters with the scaled charges \cite{pluharova14,kohagen16,martinek18}
-from \url{bitbucket.org/hseara/ions/src/master/}, and SPC/E water model~\cite{berendsen87} were used in these simulations. 
-Rest of the simulation parameters and initial configurations were taken from Lipid17
-simulations~\cite{Lipid17POPCPOPG8020,Lipid17POPCPOPG8020100mMCaCl,Lipid17POPCPOPG80201000mMCaCl,Lipid17POPCPOPG5050,Lipid17POPCPOPG5050100mMCaCl,Lipid17POPCPOPG50501000mMCaCl}.
-Simulation files of Lipid17ecc simulations are available from
-Refs.~\citenum{Lipid17eccPOPCPOPG8020,Lipid17eccPOPCPOPG8020100mMCaCl,Lipid17eccPOPCPOPG80201000mMCaCl,Lipid17eccPOPCPOPG5050,Lipid17eccPOPCPOPG5050100mMCaCl,Lipid17eccPOPCPOPG50501000mMCaCl}.
+
+\todo{This is to be finished and POPC:POPG mixtures to be described}
+In ECC-lipid models, electronic continuum correction (ECC) is applied to 
+implicitly include the missing electronic polarizability into the force field description \cite{melcr18,melcr19}.
+In practise, this is implemented by scaling the charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
+of Amber Lipid14/17 models are scaled by constant factors.
+Here, we follow the approach that previously improved ion binding to bilayers containing negatively charged PS lipids \cite{melcr19}:
+ECC-POPC parameters (scaling factors $f_q$=0.8 and $f_\sigma$=0.89 applied to Lipid14 POPC parameters) \cite{melcr18} were used for POPC
+and scaling factors of $f_q$=0.75 and $f_\sigma$=0.89 were applied to the charges and Lennard-Jones $\sigma$s of headgroup, glycerol backbone, and carbonyl regions
+of Amber Lipid17 POPG parameters. The Lipid17 parameters (described above) and initial configurations
+were taken from Ref. \cite{POPGlipid17} with the correct dihedral type,
+and the resulting parameters are available from Ref. \citenum{??}. ECC-ion parameters with the scaled charges \cite{pluharova14,kohagen16,martinek18},
+downloaded from \url{bitbucket.org/hseara/ions/src/master/}, were used in these simulations.
 

### 2021-03-23 manuscriptPGPEsuppl.tex

--- 2021-03-23 manuscriptPGPEsuppl.tex
@@
 %      POPE  & CHARMM36 \cite{klauda10}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
-      POPE  & CHARMM36ua \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
+      POPE  & CHARMM36-UA \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
       \hline
@@
       DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3655  &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
-      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
+%      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
       POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
@@
       POPE  & OPLS-MacRog \cite{rog16}     &0    & 144	& 5760     &0    & 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
-      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
+%      POPE  & OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
       \hline
-      POPE  & Berger-Vries \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
-      POPE  & Berger-largeH \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
-%      DOPE  & Berger-Vries \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
-%      DOPE  & Berger-largeH \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
+      POPE  & Berger-POPE-2004 \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
+      POPE  & Berger-POPE-2018 \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
+%      DOPE  & Berger-POPE-2004 \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
+%      DOPE  & Berger-POPE-2018 \cite{??}      &0    & 128	& 4789  &0    & 271  & 2$\times$300 & 2$\times$100 & \cite{berger2DOPEfiles} \\ 
       \hline
-      POPE             & LIPID17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
-%      POPE             & LIPID17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
+      POPE             & Lipid17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
+%      POPE             & Lipid17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
     \end{tabular}
@@
 
-\noindent {\it POPE} Data is available at \cite{bergerPOPEfiles,berger2POPEfiles}. \todo{Simulation details by  T. Piggot.}
+Following the earlier convention in the NMRlipids Project~\cite{botan15}, 
+for the Berger-based models we use the following naming convention: 
+Berger - \{{\it molecule name}\} - \{{\it year when model published first time}\} \{{\it citation}\}.
+%The reason is that there are several different molecular topologies which are using the non-bonded parameters originally
+%developed by Berger et al.~\cite{berger97}. Thus the common factor in the Berger based models are the non-bonded parameters,
+%while the molecule specific parameters might somewhat vary. However, the majority of the molecular level topologies are 
+%relying (especially for the glycerol backbone and headgroup) on the parameters originally introduced by Marrink et al.~\cite{marrink98}.
+
+\noindent {\it POPE} Data are available at Ref.~\citenum{bergerPOPEfiles} for Berger-POPE-2004~\cite{??} and at Ref.~\citenum{berger2POPEfiles} for Berger-POPE-2018~\cite{??}.
+\todo{Simulation details by  T.
+Piggot.}
 
@@
 
-\noindent {\it POPE} Data is available at \cite{gromosCKPpope}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE} Data is available at \cite{gromosCKPpopeT310}. \todo{Simulation details by  A. Peon.}
 

### 2021-03-23 manuscriptPGPEsuppl.tex
Renamed OPLS-UAwvdW to OPLS-UA/HG-H

@@
       % some footnotes are not visible in typeset-MS (pdf)
-      lipid & force field for lipids / ions  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      lipid & force field for lipids  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
       \hline
@@
       \hline
-      POPE  & OPLS-UA vdW on H \cite{??}   &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
+      POPE  & OPLS-UA/HG-H \cite{??}   &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
       POPE  & OPLS-UA \cite{??}            &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \\
@@
       \hline
-      POPE             & Lipid17 \cite{gould18} & 0      & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17} \\
+      POPE             & Lipid17 \cite{gould18} & 0      & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPElipid17} \\
 %      POPE             & Lipid17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
@@
 \noindent {\it POPE} Data is available at \cite{OPLSuaPOPEfiles}. \todo{Simulation details by  T. Piggot.}
+%From Zenodo: The PE parameters were constructed by modifying the OPLS-UA POPC of Ulmschneider and Ulmschneider with the standard OPLS lysine parameters.
 
-\noindent {\it POPE with vdW interaction in H} Data is available at \cite{OPLSuaWvdWPOPEfiles}. \todo{Simulation details by  T. Piggot.}
+\noindent {\it POPE with vdW interaction in H (OPLS-UA/HG-H)} Data is available at \cite{OPLSuaWvdWPOPEfiles}. \todo{Simulation details by  T. Piggot.}
+%From Zenodo: The PE parameters were constructed by modifying the OPLS-UA POPC of Ulmschneider and Ulmschneider with the standard OPLS lysine parameters and the addition of LJ parameters on the hydrogen atoms in the head group (designed to increase the area per lipid).
 

### 2021-03-23 manuscriptPGPEsuppl.tex

@@
   \resizebox{\columnwidth}{!}{
-    \begin{tabular}{l c c r r r r r r c c}
+%    \begin{tabular}{l c c r r r r r r c c}
+    \begin{tabular}{l c r r r r r c c}
       %\hline
       % some footnotes are not visible in typeset-MS (pdf)
-      lipid & force field for lipids  & NaCl (M) & \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      lipid 	& force field for lipids  %& NaCl (M) 
+      		& \footnote{Number of lipid molecules}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   %& \footnote{Number of additional cations}N$_{\rm c}$   
+		& \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
       \hline
-      POPE  & CHARMM36 \cite{klauda10}           &0       & 144	& 5760  &0    & 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
-      POPE  & CHARMM36 \cite{klauda10}           & 0      & 500       & 25000 & 0   &  310  & 500 & 100 & \cite{POPEcharmm} \\
-%      POPE  & CHARMM36 \cite{klauda10}           & 0.11   & 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
-      POPE  & CHARMM36-UA \cite{??}         &0       & 336	& 15254 &0    & 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
+      POPE & CHARMM36 \cite{klauda10}           %&0       
+      		& 144	& 5760  %&0    
+		& 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
+      POPE & CHARMM36 \cite{klauda10}           %& 0      
+      		& 500       & 25000 %& 0   
+		&  310  & 500 & 100 & \cite{POPEcharmm} \\
+%   POPE	& CHARMM36 \cite{klauda10}           %& 0.11
+%		& 500       & 25000 & 50  &  310  & 500 & 100 & \cite{POPEcharmm150mMNaCl} \\
+      POPE	& CHARMM36-UA \cite{??}         %&0       
+      		& 336	& 15254 %&0    
+		& 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
       \hline
-      DPPE  & Slipids \cite{jambeck12b}    &0    & 288 	& 9386  &0    & 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
-      POPE  & Slipids \cite{jambeck12b} &0    & 336	& ?     &0    & 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
-      POPE  & Slipids \cite{jambeck12b}            & 0    & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPEslipids} \\
-%      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0.11 & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
+      DPPE	& Slipids \cite{jambeck12b}    %&0    
+      		& 288 	& 9386  %&0    
+		& 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
+      POPE	& Slipids \cite{jambeck12b} %&0    
+      		& 336	& ?     %&0    
+		& 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
+      POPE	& Slipids \cite{jambeck12b}            %& 0
+		& 500 & 25000 %& 0   
+		&  310  & 500 & 100 & \cite{POPEslipids} \\
+%      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  %& 0.11
+%		& 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPEslipids150mMNaCl} \\
       \hline
-      DPPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3655  &0    & 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
-%      POPE  & GROMOS-CKP    \cite{??}      &0    & 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
-      POPE  & GROMOS-CKP    \cite{??}      &0    & 500	& 25000 &0    & 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
+      DPPE	& GROMOS-CKP    \cite{??}      %&0
+          	& 128	& 3655  %&0    
+		& 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
+%      POPE  & GROMOS-CKP    \cite{??}      %&0
+%    		& 128	& 3552  &0    & 313  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPpope} \\
+      POPE	& GROMOS-CKP    \cite{??}      %&0
+          	& 500	& 25000 %&0    
+		& 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
 %      POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
@@
       \hline
-      POPE  & GROMOS 43A1-S3 \cite{??}     &0    & 128	& 3552     &0    & 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles}  \\
+      POPE	& GROMOS 43A1-S3 \cite{??}     %&0
+          	& 128	& 3552     %&0    
+		& 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles}  \\
       \hline
-      POPE  & OPLS-UA/HG-H \cite{??}   &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
-      POPE  & OPLS-UA \cite{??}            &0    & 128	& 3328     &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \\
+      POPE	& OPLS-UA/HG-H \cite{??}   %&0    
+      		& 128	& 3328     %&0    
+		& 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
+      POPE	& OPLS-UA \cite{??}            %&0
+      		& 128	& 3328     %&0    
+		& 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \\
       \hline
-      POPE  & OPLS-MacRog \cite{rog16}     &0    & 144	& 5760     &0    & 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
+      POPE	& OPLS-MacRog \cite{rog16}     %&0    
+      		& 144	& 5760     %&0    
+		& 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
 %   POPE	& OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
       \hline
-      POPE  & Berger-POPE-2004 \cite{??}       &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
-      POPE  & Berger-POPE-2018 \cite{??}      &0    & 128	& 3552  &0    & 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
+      POPE	& Berger-POPE-2004 \cite{??}       %&0    
+      		& 128	& 3552  %&0    
+		& 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
+      POPE	& Berger-POPE-2018 \cite{??}      %&0    
+      		& 128	& 3552  %&0    
+		& 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
 %   DOPE  & Berger-POPE-2004 \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
@@
       \hline
-      POPE             & Lipid17 \cite{gould18} & 0      & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPElipid17} \\
+      POPE	& Lipid17 \cite{gould18} %& 0      
+      		& 500 & 25000 %& 0  
+		&  310  & 500 & 100 & \cite{POPElipid17} \\
 %      POPE             & Lipid17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\

### 2021-03-24 manuscriptPGPEsuppl.tex
Update pdfs
Updates in SI

@@
 }
+
+\author{\scriptsize Am{\'e}lie Bacle}
+\affiliation{Paris, France}
 
@@
   }
-  \todo{A, B etc. labels to be put in the figure.}
+  \todo{NMR figures in SI to be updated.}
 \end{figure}
@@
   }
-  \todo{This should be clarified as in NMRlipidsI and error bars should be added.
-    Probably larger error bars for united atom models based on the report by Fuchs et al.
-  } \\
 \end{figure}
@@
   \todo{Citation for CHARMM36ua?} \\
-  \todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
+  %\todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
   \todo{Citation for GROMOS-CKP?} \\
@@
   \end{minipage}
-  \todo{Citation and ion model for GROMOS-CKP?}
+  \todo{Citation for GROMOS-CKP?}
 \end{table*}
@@
 \clearpage
+
+\section{Author contributions}
+\noindent
+{\it Am{\'e}lie Bacle}
+%% Please write description of your contribution here
+
+\noindent
+{\it Pavel Buslaev}
+%% Please write description of your contribution here
+
+\noindent
+{\it Rebeca Garc{\'i}a Fandi{\~n}o}
+%% Please write description of your contribution here
+
+\noindent
+{\it Fernando Favela-Rosales}
+%% Please write description of your contribution here
+
+\noindent
+{\it Tiago M. Ferreira}
+%% Please write description of your contribution here
+
+\noindent
+{\it Patrick Fuchs}
+%% Please write description of your contribution here
+
+\noindent
+{\it Ivan Gushchin}
+%% Please write description of your contribution here
+
+\noindent
+{\it Matti Javanainen}
+%% Please write description of your contribution here
+
+\noindent
+{\it Anne M. Kiirikki}
+%% Please write description of your contribution here
+
+\noindent
+{\it Jesper J. Madsen}
+%% Please write description of your contribution here
+
+\noindent
+{\it Josef Melcr}
+%% Please write description of your contribution here
+
+\noindent
+{\it Paula Milan Rodriguez}
+%% Please write description of your contribution here
+
+\noindent
+{\it Markus S. Miettinen}
+%% Please write description of your contribution here
+
+\noindent
+{\it O. H. Samuli Ollila}
+designed the project and managed the work.
+Ran and analysed several simulations. Wrote the manuscript.
+
+\noindent
+{\it Chris G. Papadopoulos}
+%% Please write description of your contribution here
+
+\noindent
+{\it Antonio Pe{\'o}n}
+%% Please write description of your contribution here
+
+\noindent
+{\it Thomas J. Piggot}
+%% Please write description of your contribution here
+
+\noindent
+{\it {\'A}ngel Pi{\~n}eiro}
+%% Please write description of your contribution here
+
 \bibliography{refs.bib}

### 2021-04-05 manuscriptPGPEsuppl.tex
Update manuscriptPGPEsuppl.tex

@@
 \affiliation[University of Chicago]{Department of Chemistry, The University of Chicago, Chicago, Illinois, United States of America}
-\alsoaffiliation[University of South Florida]{Department of Global Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}
+\alsoaffiliation[University of South Florida]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}
 
@@
 {\it Jesper J. Madsen}
-%% Please write description of your contribution here
+set up, performed, and analysed several of the CHARMM36 simulations.
+Provided comments on the manuscript.
 

### 2021-04-06 manuscriptPGPE.tex
Update manuscriptPGPE.tex

--- 2021-03-23 manuscriptPGPE.tex
@@
 \affiliation{Department of Chemistry, The University of Chicago, Chicago, Illinois, United States of America}
-\affiliation{Department of Global Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}
+\affiliation{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}
 
@@
 Because membranes are in liquid state under physiological conditions,
-  individual lipids are not fixed to single conformations but they sample ensemble of conformations with the Boltzmann weighted probabilities.
+  individual lipids are not fixed to single conformations but they sample an ensemble of conformations with the Boltzmann weighted probabilities.
   However, these ensembles and their dependence on lipid types have not yet been experimentally determined. 
@@
   or if they can freely fluctuate in all possible conformations.
-  Here, we combine solid state NMR experiments and molecular dynamics simulations from NMRlipids open collaboration to resolve the conformational ensembles of the headgroups of key lipid types in liquid lamellar phase under various biologically relevant conditions.
+  Here, we combine solid state NMR experiments and molecular dynamics simulations from the NMRlipids open collaboration to resolve the conformational ensembles of the headgroups of key lipid types in their liquid lamellar phase under various biologically relevant conditions.
 Interpretation of NMR experiments using the plethora of simulation data collected in the NMRlipids project suggests that all lipid headgroups sample a wide range of conformations in neutral and charged cellular membranes.
-  Differences in the headgroup chemistry between different lipid types manifests
-  in probability distributions of conformations, but all lipid types can access to almost any possible conformations.
+  Differences in the headgroup chemistry between different lipid types manifest
+  in probability distributions of conformations, but all lipid types can access almost any of the possible conformations.
   Together with the analysis of protein-bound lipids from the protein data bank (PDB), this suggests that lipids can bind to proteins in wide range of conformations independently on the headgroup chemistry.
-Therefore, the selective adsorption of proteins to membranes is likely regulated by specific protein--lipid interactions, rather than conformational restrictions of lipids.
+Therefore, the selective adsorption of proteins to membranes is likely regulated by specific protein--lipid interactions rather than conformational restrictions of the bilayer lipids.
   Our results pave the way to comprehensive understanding of
@@
 and crystal structures of lipids have been determined \cite{buldt81,pascher92},
-but their relation to the conformational ensembles in bulk membranes in liquid lamellar phase remains unclear \cite{marsh13b}.
+but their relation to the conformational ensembles in bulk membranes in the liquid lamellar phase remains unclear \cite{marsh13b}.
 %In addition to the changes in lipid conformational ensembles upon binding to proteins,
@@
 Therefore, it is not clear if the conformational ensemble of lipid headgroups in liquid lamellar bilayer
-composes of exchange between few restricted conformations, or if lipid headgroups can fluctuate freely in wide range of conformations.
+is composed of exchange between few restricted conformations, or if lipid headgroups can fluctuate freely across a wide range of conformations.
 
@@
 and MD simulations from the NMRlipids open collaboration
-to resolve the differences in conformational ensembles between PC, PE, PG and PS lipid headgroups.
+to resolve the differences in the conformational ensembles between PC, PE, PG and PS lipid headgroups.
 We elucidate also the effect of charges and protein binding to lipid headgroup conformations.
@@
 PS and PG are the most common negatively charged lipids in eukaryotes and bacteria, respectively,
-and affect membrane protein functionality and signaling \cite{lemmon08,leventis10,sohlenkamp16,hariharan18}.
+and their presence affect membrane protein functionality and signaling \cite{lemmon08,leventis10,sohlenkamp16,hariharan18}.
 All the studied lipids specifically bind to various proteins \cite{yeagle14}.
@@
 
-Glycerol backbone peaks from both lipids, and $\alpha$-carbon peak from POPE in the INEPT spectra 
+Glycerol backbone peaks from both lipids, and the $\alpha$-carbon peak from POPE in the INEPT spectra 
 were assigned based on previously measured POPC spectra~\cite{ferreira13}.
@@
 Conformational ensembles of headgroup and glycerol backbone in PE and PG simulations were evaluated using the C-H bond order parameters \cite{botan15}.
-Interactions between different headgroups were evaluated monitoring the changes in headgroup order parameters upon mixing the lipids \cite{antila19}.
-The ion binding affinities and response of lipids to bound charge were evaluated monitoring the changes of lipid headgroup order parameters \cite{catte16,antila19}.
+Interactions between different headgroups were evaluated by monitoring the changes in headgroup order parameters upon mixing the lipids \cite{antila19}.
+The ion binding affinities and response of lipids to bound charge were evaluated by monitoring the changes in lipid headgroup order parameters \cite{catte16,antila19}.
 
@@
 To demonstrate the existence of similar structures of different lipids bound to different proteins,
-we searched pairs having the last five dihedral angles (excluding g$_1$-g$_2$-g$_3$-O$_{g3}$) within
+we searched for pairs having the last five dihedral angles (excluding g$_1$-g$_2$-g$_3$-O$_{g3}$) within
 30 degrees from structures with resolution higher than 2.5~\AA. The two most representative examples
@@
 
-Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.~\ref{HGorderParameters}D), where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
+Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.~\ref{HGorderParameters}D, where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
 The overall agreement of order parameters determined by different research teams and different techniques for the same lipid headgroup is very good here and in previous studies \cite{botan15,ollila16,antila19}.
-Therefore, the differences between lipid types are dictated by the headgroup chemistry rather than inaccuracies in experiments, or differences in the acyl chains or in experimental conditions.
+Therefore, the differences between lipid types are dictated by the headgroup chemistry rather than inaccuracies in experiments, differences in the acyl chains, or in experimental conditions.
 
@@
 energy costs for different dihedral angle orientations.
-The results in Fig.~\ref{structures}C) suggest that all lipid headgroups are very flexible and
+The results in Fig.~\ref{structures}C suggest that all lipid headgroups are very flexible and
 energy costs for rotating individual dihedrals to almost any angle is low (below 7 kT).
@@
 
-In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase and sampled dihedral angles are within approximately same ranges for all headgroup types.
+In conclusion, all lipid headgroups sample very broad conformational ensembles in the liquid lamellar phase and the sampled dihedral angles are within approximately same ranges for all headgroup types.
 % (Fig.
@@
 %The wide range of observed conformations suggest that
-Therefore, the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
+Therefore, the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in the liquid lamellar state.
 
@@
 
-Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of lipid headgroups \cite{seelig87}.
-However, the detailed understanding on structural and energetic response of lipids to membrane bound electric charge is still lacking \cite{Semchyschyn04}.
+Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes can reorient the headgroup dipole in PC lipids and affect the order parameters of the lipid headgroups \cite{seelig87}.
+However, the detailed understanding of this structural and energetic response of lipids to membrane-bound electric charge is still lacking \cite{Semchyschyn04}.
 
-To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge,
+To resolve lipid headgroup conformational ensembles in cell membranes bearing positive charge,
 we calculated the heavy atom dihedral angle distributions from 
-simulations %(CHARMM36)
-that correctly captured the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
+the subset of simulations %(CHARMM36)
+that correctly captured the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer (figure \ref{changesWITHsurf} A).
 The dihedral angle distributions and relative energies
@@
 addition of positive charge into a membrane 
-decrease the abundance of trans states in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
+decreases the abundance of trans states in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
 dihedrals.
-Choline region remains essentially unchanged and only minor changes are observed in other dihedrals even though almost half of the molecules in membranes are cationic surfactants.
+The choline region remains essentially unchanged and only minor changes are observed in other dihedrals even though almost half of the molecules in the membrane are cationic surfactants.
 
-Also binding of ions to membranes may affect the lipid headgroup conformational ensembles in physiological conditions.
+In addition, binding of ions to the membrane may affect the lipid headgroup conformational ensembles under physiological conditions.
 The bound Ca$^{2+}$ ion to PC headgroup leads to similar decrease in trans state probability for g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
@@
 However, the changes in PG lipid dihedrals upon addition of CaCl$_2$ differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and~\ref{DIHSwithCAlipid17POPG}), none of the simulations captures the Ca$^{2+}$ ion binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added CaCl$_2$ in PG is not available.
-Also the headgroup conformational ensembles in mixtures of PC and charged (PG or PS) or zwitterionic (PE)
+Also, the headgroup conformational ensembles in mixtures of PC and charged (PG or PS) or zwitterionic (PE)
 lipids could not be resolve with the currently available force fields and experimental data
@@
 %experimentally observed changes in headgroup order parameters upon addition of
-membrane bound charges arise
+membrane-bound charges arise
 %from relatively small changes in conformational ensembles. These changes eventuate
 from mild changes in dihedral angle distributions, rather than restriction of lipids to
-fixed conformations. Therefore, lipid headgroups sample large space of different conformations
-also in charged membranes, thereby retaining the capability to adapt multiple conformations when
+fixed conformations. Therefore, lipid headgroups sample a large space of different conformations
+also in charged membranes, thereby retaining the capability to adopt multiple conformations when
 interacting with proteins or other molecules.
@@
 in previous sections suggest that PC, PE, PG and PS lipid headgroups are very flexible,
-thereby being able to bind proteins in various different conformations.
+allowing them to bind proteins in various different conformations.
 To test this prediction, we analyzed the protein-bound lipid conformations
@@
 The heavy atom dihedral angle distributions calculated from these conformations 
-in Fig.~\ref{dihedralsFROMpdb} A) reveal that the protein bound lipids indeed exhibit wide
+(Fig.~\ref{dihedralsFROMpdb} A) reveal that the protein-bound lipids indeed exhibit wide
 range of conformations independently on the headgroup type.
-As in bulk lipid bilayers, only $cis$ conformations
+As for bulk lipid bilayers, only $cis$ conformations
 of P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals are almost
@@
 have been previously proposed to indicate inaccuracies in lipid structures in PDB \cite{marsh13b,pezeshkian18}. 
-However, we see large deviations from lipid crystals structures also in conformational ensembles
-that reproduce the NMR data in liquid lamellar phase, thereby proposing that such deviations
+However, we observe large deviations from lipid crystals structures also in conformational ensembles
+that reproduce the NMR data in the liquid lamellar phase, indicating that such deviations
 are realistic also in protein-bound states.
@@
 Therefore, the specific binding of lipids to proteins is not driven by the structural differences between headgroups.
-This is demonstrated in Fig.~\ref{dihedralsFROMpdb} B) with two examples where different lipids bound to
+This is demonstrated in Fig.~\ref{dihedralsFROMpdb} B with two examples where different lipids bound to
 different proteins have identical conformations:
@@
 and PG bound to bovine cytochrome {\it c} oxidase is similar to PC bound to pore-forming toxin (FraC).
-On the other hand, a single lipid headgroup type is capable to adapt itself to various binding positions, 
-thereby being able to specifically bind to many different kinds of binding sites in different proteins.
+On the other hand, a single lipid headgroup type is capable of accommodating various binding positions and as such would 
+be able to specifically bind to many different kinds of binding sites in different proteins.
 
@@
 All studied headgroups (PC, PE, PG and PS) are flexible
-and access similar wide range of conformations with only very few restictions in dihedral orientations
+and access a similar wide range of conformations with only very few restictions in dihedral orientations
 also when charges are bound to membranes.
@@
 adapt to various binding sites to optimize the intermolecular lipid--protein interactions.
-We tested this prediction by analyzing the conformations of lipids that are tightly bound to proteins in from PDB.
-Indeed, also protein bound lipids exhibit wide range of conformations without significant 
+We tested this prediction by analyzing the conformations of lipids that are tightly bound to proteins in from the PDB.
+Indeed, also protein--bound lipids exhibit wide range of conformations without significant 
 differences between different lipid types. Therefore, the specificify of lipid binding to proteins is not
@@
 
-Our results pave the way toward understanding of lipid mediated cell signaling and how lipids regulate membrane protein function in general.
+Our results pave the way toward understanding of lipid-mediated cell signaling and how lipids regulate membrane protein function in general.
 We suggest that the key to understand selective binding of certain lipid types to proteins
-are intermolecular lipid-protein interactions, rather than conformational restrictions of lipids.
+are intermolecular lipid-protein interactions, rather than conformational restrictions of the lipids.
 On the other hand, wide conformational ensembles in bulk bilayers suggest that lipid crystal structures
@@
 Finally, we demonstrated the power of open access MD simulation data from the NMRlipids open collaboration
-to complement the data in PDB in revealing how complex systems containing disordered biomolecules behave.
+to complement the data in the PDB for elucidating how complex systems containing disordered biomolecules behave.
 
@@
     %     Put your acknowledgments here.
+
+J.J.M. gratefully acknowledges financial support from the Carlsberg 
+Foundation in the form of a postdoctoral fellowship while at 
+the University of Chicago (grants CF15-0552, CF16-0639, and 
+CF17-0783) and the research framework provided by the 
+Research Computing Center at the University of Chicago.
+
 \end{acknowledgments}

### 2021-04-06 manuscriptPGPEsuppl.tex

--- 2021-03-24 manuscriptPGPEsuppl.tex
@@
   \caption{\label{HGorderParametersPE}
-    The headgroup and glycerol backbone order parameters of PE lipids
-    from experiments (POPE and signs this work, DPPE from Ref.~\citenum{seelig76}
-    and E.coliPE from Ref.~\citenum{gally81}) and simulations with different force fields.
+    C--H bond order parameters, $S_{\rm CH}$, of the PE headgroup ($\beta$ and $\alpha$) and
+    glycerol backbone (g$_3$, g$_2$, g$_1$) carbons from NMR experiments (horizontal lines;
+    POPE and signs this work, DPPE from Ref.~\citenum{seelig76},
+    {\it Escherichia coli} PE from Ref.~\citenum{gally81})
+    and MD simulations with different force fields (symbols).
+    The light blue areas span 0.04 units around the average of the extremal experimental values,
+    in accordance with the expected quantitative accuracy of experiments~\cite{ollila16}.
+    The vertical bars shown for most simulation values are not error bars,
+    but demonstrate that for these systems we had at least two data sets;
+    the ends of the bars mark the extreme values from the sets,
+    the symbol marks the measurement-time-weighted average.
   }

### 2021-04-06 manuscriptPGPE.tex
Updates from Overleaf

--- 2021-03-23 manuscriptPGPE.tex
@@
 \author{Am{\'e}lie Bacle}
-\affiliation{Paris, France}
+\affiliation{Laboratoire Coopératif "Lipotoxicity and Channelopathies – ConicMeds", Université de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, Franc }
 
@@
 
-\author{Tiago Ferreira}
-\affiliation{Halle, Germany}
+\author{Tiago M. Ferreira}
+\affiliation{NMR group - Institute for Physics, Martin Luther University Halle-Wittenberg, 06120 Halle (Saale), Germany}
 
-\author{Patrick Fuchs}
-\affiliation{Paris, France}
+\author{Patrick F.J. Fuchs}
+\affiliation{Sorbonne Université, Ecole Normale Supérieure, PSL University, CNRS, Laboratoire des Biomolécules (LBM), 75005 Paris, France}
+\affiliation{Université de Paris, UFR Sciences du Vivant, 75013, Paris, France}
 
@@
 \author{Paula Milan Rodriguez}
-\affiliation{Paris, France}
+\affiliation{Sorbonne Université, Ecole Normale Supérieure, PSL University, CNRS, Laboratoire des Biomolécules (LBM), 75005 Paris, France}
 
@@
 \author{Chris G. Papadopoulos}
-\affiliation{I2BC - University Paris Sud}
+\affiliation{Université Paris-Saclay, CEA, CNRS, Institute for Integrative Biology of the Cell (I2BC), 91198 Gif-sur-Yvette, France}
 
@@
 C--H bond order parameters measured using $^2$H NMR~\cite{seelig77c,davis83,Semchyschyn04}.
-According to these experiments, the glycerol backbone conformations are largely similar irrespectively of the headgroup \cite{gally81}, 
+According to these experiments, the glycerol backbone conformations are largely similar irrespectively of the headgroup \cite{gally81} and
 the headgroup conformations are similar in phosphatidylcholine (PC), phosphatidylethanolamine (PE) and phosphatidylglycerol (PG) lipids,
-while the headgroup of phosphatidylserine (PS) lipids is more rigid \cite{wohlgemuth80,buldt81}. 
-However, order parameters signs are not accessible in $^2$H NMR experiments \cite{ollila16}
+whereas the headgroup of phosphatidylserine (PS) lipids is more rigid \cite{wohlgemuth80,buldt81}. 
+However, order parameter signs are not accessible in $^2$H NMR experiments \cite{ollila16}
 and universal models to map order parameters to structural ensembles are not available \cite{pezeshkian18,akutsu20}.
@@
 using quality evaluation measures defined in the NMRlipids project.
-Conformational ensembles of headgroup and glycerol backbone in PE and PG simulations were evaluated using the C-H bond order parameters \cite{botan15}.
+Conformational ensembles of headgroup and glycerol backbone in PE and PG simulations were evaluated using the C--H bond order parameters \cite{botan15}.
 Interactions between different headgroups were evaluated monitoring the changes in headgroup order parameters upon mixing the lipids \cite{antila19}.
@@
    \caption{\label{HGorderParameters}
-     Chemical structure, refocused-INEPT spectrum, 2D R-PDLF spectra, and S-DROSS data (from top to bottom) of \textbf{A)} POPE  and \textbf{B)} POPG.
-      Full NMR spectra are shown Figs. \ref{POPEspectra} and \ref{POPGspectra}.
+     Chemical structure, refocused-INEPT spectrum, 2D R-PDLF spectra, and S-DROSS dipolar slices (from top to bottom) for \textbf{A)} POPE  and \textbf{B)} POPG MLVs measured by $^1$H-$^{13}$C solid-state NMR.
+The S-DROSS dipolar slices show both experimental data (open circles) and the result of NMR numerical simulations (blue solid lines).
+      Full NMR spectra are shown in Figs. \ref{POPEspectra} and \ref{POPGspectra}.
      \textbf{C)} Chemical structure of POPC and POPS.
-    \textbf{D)} Headgroup and glycerol backbone order parameters from different experiments in lamellar liquid disordered phase.
-    The values and signs for POPE (310~K) and POPG (298~K)
-    measured in this work, and for POPS (298~K) \cite{antila19} and POPC (300~K) \cite{ferreira13,ferreira16}
-    previously using $^{13}$C NMR. The literature values for
-    DOPS with 0.1M of NaCl (303~K) \cite{browning80},
+    \textbf{D)} Headgroup and glycerol backbone C-H bond order parameters for different phospholipids measured in the lamellar liquid disordered phase.
+    The $S_{\rm{CH}}$ magnitudes and signs determined by $^1$H-$^{13}$C NMR spectroscopy (filled squares) for POPE (310~K) and POPG (298~K)
+    measured in this work are shown together with the data previously reported for POPS (298~K) \cite{antila19} and POPC (300~K) \cite{ferreira13,ferreira16}.
+The headgroup and glycerol backbone $S_{\rm{CH}}$ magnitudes measured previously with $^2$H NMR (empty symbols) are also plotted as real values using the signs determined in this work for
     POPG with 10nM PIPES (298~K) \cite{borle85},
@@
     DPPE (341~K) \cite{seelig76},
-    E.coliPE and E.coliPG (310~K) \cite{gally81}
-    are measured using $^2$H NMR. The signs from $^{13}$C NMR are used also for the literature values.
+    E.coli PE and E.coli PG (310~K) \cite{gally81}
+    .
    }
@@
 Jupyter notebook available from \url{https://github.com/pbuslaev/scr/blob/master/PDB%20analysis.ipynb}.
-Only structures from X-ray and Cryo-EM with higher than 3.2~\AA~resolution were used in the analysis.
-Lipids from same PDBID with all dihedrals angles within 3 degrees were discarded from the analysis
-to avoid overweighting of lipid structures present, e.g., due to rotational symmetries.
+Only structures from X-ray and Cryo-EM with a resolution higher than 3.2~\AA were used in the analysis.
+Lipids from same PDB identifier with all dihedrals angles within 3 degrees were discarded from the analysis
+to avoid overweighting of lipid structures present, \textit{e.g.}, due to rotational symmetries.
 
@@
      \textbf{B)} Headgroup and glycerol backbone region order parameters of different different lipids.
-     \textbf{C)} Relative energies for individual heavy atom dihedral angles estimated from the inverse Boltzmann formula.
-     Angles corresponding energies above 7 kT are not shown because they are not observed in simulations.
-     \textbf{D)} Distributions of P-N vector angle with respect to membrane normal.
+     \textbf{C)} Relative free energies for individual heavy atom dihedral angles estimated from the inverse Boltzmann formula.
+     Angles corresponding energies above 7~$k\mathrm{B}T$ are not shown because they are not observed in simulations.
+     \textbf{D)} Distributions of P--N vector angle with respect to membrane normal.
   }
@@
 Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.~\ref{HGorderParameters}D), where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
-The overall agreement of order parameters determined by different research teams and different techniques for the same lipid headgroup is very good here and in previous studies \cite{botan15,ollila16,antila19}.
+The overall agreement of order parameters determined by different research teams and different techniques for the same lipid headgroup is excellent here and in previous studies \cite{botan15,ollila16,antila19}.
 Therefore, the differences between lipid types are dictated by the headgroup chemistry rather than inaccuracies in experiments, or differences in the acyl chains or in experimental conditions.
 
-The most distinct order parameters are observed for PS headgroups, for which the $\alpha$-carbon order parameter exhibits significant forking (different values for different hydrogens in the same carbon) and the $\beta$-carbon has more negative value than other studied lipid types.
+The most distinct order parameters are observed for PS headgroups, for which the $\alpha$-carbon order parameter exhibits significant forking (different values for different hydrogens bound to the same carbon) and the $\beta$-carbon has a more negative value than other studied lipid types.
 On the other hand, the $\beta$-carbon order parameter of PG headgroup has a positive sign, in contrast to all the other lipid types.
@@
 The results in Fig.~\ref{structures}C) suggest that all lipid headgroups are very flexible and
-energy costs for rotating individual dihedrals to almost any angle is low (below 7 kT).
+energy costs for rotating individual dihedrals to almost any angle is low (below 7~$k_\mathrm{B}T$).
 Only $cis$ states of P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P have larger relative
@@
 which prefer \textit{gauche$^-$} conformations for PG and \textit{gauche$^+$} for PS,
-while PC and PE exhibit symmetric distributions.
+whereas PC and PE exhibit symmetric distributions.
 Also, the energy barriers for O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ dihedral
@@
 Rest of the dihedrals are similar between different lipids,
-with the exception of PS lipids for which slightly larger energy for ecplised anti conformation
+with the exception of PS lipids for which slightly larger energy for \textit{eclipsed anti} conformation
 in g$_3$-O$_{{\mathrm g}3}$-P-O$_\alpha$ dihedral was observed.
-Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameter occur in the choline part, while also changes in phosphate region may contribute in PS lipids.
+Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameters occur in the choline part, while also changes in phosphate region may contribute in PS lipids.
 The increased barriers for dihedral rotations may explain the more rigid headgroup structures in PS~\cite{browning80,buldt81}.
@@
 
-In conclusion, all lipid headgroups sample very broad conformational ensembles in liquid lamellar phase and sampled dihedral angles are within approximately same ranges for all headgroup types.
+In conclusion, all studied lipid headgroups sample very broad conformational ensembles in liquid lamellar phase and the sampled dihedral angles are within approximately same ranges for all headgroup types.
 % (Fig.
 \ref{structures}).
-Because the rotation of dihedral angles to almost any position bears relatively low energy cost,
+Since the rotation of dihedral angles to almost any position bears relatively low free energy cost,
 %results suggest that all lipid headgroups are very flexible---
 %thereby being
-the lipid headgroups are able to adopt wide range of conformations when interacting with proteins, ions or other biomolecules.
+the lipid headgroups are able to adopt a wide range of conformations when interacting with proteins, ions, or other biomolecules.
 %Our results support the models proposing free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
 %The wide range of observed conformations suggest that
-Therefore, the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and models aiming to explain NMR data using only few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in liquid lamellar state.
+Therefore, the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and models aiming to explain NMR data using only a few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in the liquid lamellar state.
 
@@
 Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of lipid headgroups \cite{seelig87}.
-However, the detailed understanding on structural and energetic response of lipids to membrane bound electric charge is still lacking \cite{Semchyschyn04}.
+However, the detailed understanding on the structural and energetic responses of lipids to membrane-bound electric charge is still lacking \cite{Semchyschyn04}.
 
-To resolve lipid headgroup conformational ensembles in cell membrane bearing positive charge,
+To resolve lipid headgroup conformational ensembles in a cell membrane bearing positive charge,
 we calculated the heavy atom dihedral angle distributions from 
 simulations %(CHARMM36)
-that correctly captured the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in figure \ref{changesWITHsurf} A).
+that correctly captured the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in (see figure \ref{changesWITHsurf} A)).
 The dihedral angle distributions and relative energies
-in figures \ref{HGorderparametersPCvsSURFdihedrals} and \ref{changesWITHsurf} B reveal that the
+in figures \ref{HGorderparametersPCvsSURFdihedrals} and \ref{changesWITHsurf} B) reveal that the
 addition of positive charge into a membrane 
-decrease the abundance of trans states in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
+decrease the abundance of \textit{trans} states in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
 dihedrals.
-Choline region remains essentially unchanged and only minor changes are observed in other dihedrals even though almost half of the molecules in membranes are cationic surfactants.
+The choline region remains essentially unchanged and only minor changes are observed in other dihedrals even at a surfactant molar fraction of 0.47.
 
-Also binding of ions to membranes may affect the lipid headgroup conformational ensembles in physiological conditions.
-The bound Ca$^{2+}$ ion to PC headgroup leads to similar decrease in trans state probability for g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
+Additionally, the binding of ions to membranes may affect the lipid headgroup conformational ensembles in physiological conditions.
+The binding of Ca$^{2+}$ ions to a PC headgroup leads to a similar decrease in the \textit{trans} state probability for g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
 in the most realistic MD simulations models
@@
 (lipid17ecc and CHARMM36 in Figs.~\ref{DIHSwithCAlipid17eccPOPC}, \ref{DIHSwithCAcharmm36POPC} and Fig.~\ref{changesWITHCaClPG}).
-The dihedral distributions of PG headgroup are more sensitive to the bound ions in the most realistic simulations,
-but upward tilting of the headgroup dipole upon addition of CaCl$_2$ is weaker than in PC
+The dihedral distributions of the PG headgroup are more sensitive to the bound ions in the most realistic simulations,
+but upward tilting of the headgroup dipole upon the addition of CaCl$_2$ is weaker than in PC
 (Lipid17 and Slipids in Figs.~\ref{changesWITHCaClPG},~\ref{DIHSwithCAslipidsPOPG} and \ref{DIHSwithCAlipid17POPG}).
-However, the changes in PG lipid dihedrals upon addition of CaCl$_2$ differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and~\ref{DIHSwithCAlipid17POPG}), none of the simulations captures the Ca$^{2+}$ ion binding affinity and conformational ensemble of PG lipids simultaneously, and experimental data to evaluate the response of $\alpha$-carbon order parameters to the added CaCl$_2$ in PG is not available.
-Also the headgroup conformational ensembles in mixtures of PC and charged (PG or PS) or zwitterionic (PE)
-lipids could not be resolve with the currently available force fields and experimental data
-(Figs.~\ref{HGorderparametersPCvsPEPG} and~\ref{HGorderparametersPGvsPCchange}, and Ref. \cite{antila19,melcr20}).
+However, the changes in PG lipid dihedrals upon the addition of CaCl$_2$ differ between the best models (Figs.~\ref{DIHSwithCAslipidsPOPG} and~\ref{DIHSwithCAlipid17POPG}), and none of the simulations captures the Ca$^{2+}$ ion binding affinity and conformational ensemble of PG lipids simultaneously.
+Moreover, experimental data to evaluate the response of $\alpha$-carbon order parameters to the added CaCl$_2$ in PG is not available.
+Additionally, the headgroup conformational ensembles in mixtures of PC and charged (PG or PS) or zwitterionic (PE)
+lipids could not be resolved with the currently available force fields and experimental data
+(Figs.~\ref{HGorderparametersPCvsPEPG} and~\ref{HGorderparametersPGvsPCchange}, and Ref.~\cite{antila19,melcr20}).
 
@@
 %experimentally observed changes in headgroup order parameters upon addition of
-membrane bound charges arise
+membrane-bound charge arises
 %from relatively small changes in conformational ensembles. These changes eventuate
 from mild changes in dihedral angle distributions, rather than restriction of lipids to
-fixed conformations. Therefore, lipid headgroups sample large space of different conformations
-also in charged membranes, thereby retaining the capability to adapt multiple conformations when
+fixed conformations. Therefore, lipid headgroups sample a large space of different conformations
+also in charged membranes, thereby retaining the capability to adopt multiple conformations when
 interacting with proteins or other molecules.
@@
 The heavy atom dihedral angle distributions calculated from these conformations 
-in Fig.~\ref{dihedralsFROMpdb} A) reveal that the protein bound lipids indeed exhibit wide
+in Fig.~\ref{dihedralsFROMpdb} A) reveal that the protein-bound lipids indeed exhibit wide
 range of conformations independently on the headgroup type.
@@
 
-C-H bond order parameters from NMR experiments %in this and previous studies \cite{wohlgemuth80,buldt81,seelig87}
+C--H bond order parameters from NMR experiments %in this and previous studies \cite{wohlgemuth80,buldt81,seelig87}
 suggest that lipid headgroup conformational ensembles depend on lipid type (PC, PE, PG or PS) 
 and accumulated membrane charge (cationic lipids, surfactants, ions or drugs). 
-Our interpretation of this data, using MD simulations collected
+Our interpretation of these data, using MD simulations collected
 within the NMRlipids open collaboration,
@@
 
-Our results pave the way toward understanding of lipid mediated cell signaling and how lipids regulate membrane protein function in general.
+Our results pave the way toward the understanding of lipid-mediated cell signaling and how lipids regulate membrane protein function in general.
 We suggest that the key to understand selective binding of certain lipid types to proteins
-are intermolecular lipid-protein interactions, rather than conformational restrictions of lipids.
-On the other hand, wide conformational ensembles in bulk bilayers suggest that lipid crystal structures
-play a minor role and that entropic cost of lipid binding may be significant.
+are intermolecular lipid--protein interactions, rather than conformational restrictions of lipids.
+On the other hand, broad conformational ensembles in bulk bilayers suggest that lipid crystal structures
+play a minor role and that the entropic cost of lipid binding may be significant.
 Finally, we demonstrated the power of open access MD simulation data from the NMRlipids open collaboration
-to complement the data in PDB in revealing how complex systems containing disordered biomolecules behave.
+to complement the data in PDB in revealing how complex systems made up of disordered biomolecules behave.
 
@@
 Terrae computer
+%
+MJ thanks CSC -- IT Center for Science for computational resources and the Emil Aaltonen foundation for financial support.s
     %     Put your acknowledgments here.

### 2021-04-06 manuscriptPGPEsuppl.tex
Updates from Overleaf

--- 2021-04-05 manuscriptPGPEsuppl.tex
@@
 \author{\scriptsize Am{\'e}lie Bacle}
-\affiliation{Paris, France}
+\affiliation{Laboratoire Coopératif "Lipotoxicity and Channelopathies – ConicMeds", Université de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }
 
@@
 
-\author{Patrick Fuchs}
-\affiliation{Paris, France}
+\author{Patrick F.J. Fuchs}
+\affiliation{Sorbonne Université, Ecole Normale Supérieure, PSL University, CNRS, Laboratoire des Biomolécules (LBM), 75005 Paris, France}
+\alsoaffiliation{Université de Paris, UFR Sciences du Vivant, 75013, Paris, France}
 
@@
 \affiliation[University of Chicago]{Department of Chemistry, The University of Chicago, Chicago, Illinois, United States of America}
-\alsoaffiliation[University of South Florida]{Global and Planetary Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}
+\alsoaffiliation[University of South Florida]{Department of Global Health, College of Public Health, University of South Florida, Tampa, Florida, United States of America}
 
@@
 
-\author{Paula Milan Rodriguez}
-\affiliation{Paris, France}
+\author{Paula Milán Rodríguez}
+\affiliation{Sorbonne Université, Ecole Normale Supérieure, PSL University, CNRS, Laboratoire des Biomolécules (LBM), 75005 Paris, France}
 
@@
 \author{Chris G. Papadopoulos}
-\affiliation[]{I2BC - University Paris Sud}
+\affiliation{Université Paris-Saclay, CEA, CNRS, Institute for Integrative Biology of the Cell (I2BC), 91198, Gif-sur-Yvette, France}
 
@@
 The good performance of Berger-OPLS simulations is surprising here because
-headgroup conformational enemble is not very close to experiments in this model and
+headgroup conformational ensemble is not very close to experiments in this model and
 the response of headgroup order parameters
@@
 \noindent {\it POPC:POPE mixtures} Data is available at \cite{POPCcharmm300K,POPC1POPE1charmm36}.
-300 K with v-rescale (tau=0.1 ps),
-1 bar with PR semiisotropic (tau=4 ps, compressibility=4.5e-5 bar$^{-1}$),
-PME order 4 and space 0.12,
-rcoulomb and rvdw 1.0,
-128 lipids per leaflet,
-no ion \todo{Full simulation details by Fuchs et al.}
+A pure POPC system and a 50:50 POPC:POPE mixture were built and equilibrated using CHARMM-GUI [https://doi.org/10.1021/acs.jctc.5b00935].
+They contained 256 lipids (for the mixture 64 lipids per leaflet for each species) and 34 water molecules per lipid.
+No ions were added.
+The production simulations were run for 300 ns with a time step of 2 fs.
+The first 50 ns were discarded for the analysis.
+The simulations were run with the GROMACS 2016.4 [https://doi.org/10.1016/j.softx.2015.06.001] version.
+The v-rescale thermostat [http://dx.doi.org/10.1063/1.2408420] was used with a temperature of 300 K and a time constant of 1 ps; lipids and water were coupled separately to the heat bath.
+Pressure was kept constant at 1 bar using a semi–isotropic Parrinello–Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 5.0 ps.
+A real space cut-off of 1.2 nm was employed for electrostatic interactions while the long-range part was evaluated using the PME method [https://doi.org/10.1063/1.464397, https://doi.org/10.1063/1.470117].
+A force-based switching function was used to switch the Lennard-Jones forces to zero over a range of 1−1.2 nm.
+All bonds with hydrogen atoms were constrained with the LINCS algorithm [https://doi.org/10.1002/(SICI)1096-987X(199709)18:12%3C1463::AID-JCC4%3E3.0.CO;2-H].
+Water molecules were kept rigid with the SETTLE algorithm [https://doi.org/10.1002/jcc.540130805].
+The simulation files are available from ref 64 [https://doi.org/10.5281/zenodo.1306800] (pure POPC) and 65 [https://doi.org/10.5281/zenodo.1306821] (POPC:POPE mixture).
 
@@
 \noindent {\it POPC:POPE mixtures} Data is available at \cite{POPCberger300K,POPC1POPE1berger}. 
-300 K with v-rescale (tau=0.1 ps),
-1 bar with PR semiisotropic (tau=4 ps, compressibility=4.5e-5 bar$^{-1}$),
-PME order 4 and space 0.12,
-rcoulomb and rvdw 1.0,
-128 lipids per leaflet,
-no ion \todo{Simulation details by Fuchs et al.}
+Two systems were simulated using the Berger force field [https://doi.org/10.1016/S0006-3495(97)78845-3], pure POPC and a mixture 50:50 POPC:POPE.
+For POPE, we additionnally used the de Vries modification implementing a repulsion of the hydrogen atoms located on the amino group of ethanolamine [https://doi.org/10.1021/jp0366926].
+Starting from a PDB file of a pure POPC system with Berger atom names, a 50:50 POPC:POPE mixture was built by mutating randomly methyl groups to hydrogens (POPC -> POPE).
+Each system contained 256 lipids (for the mixture 64 lipids per leaflet for each species) and about 40 water molecules per lipid.
+No ions were added.
+Both systems were minimized and equilibrated.
+The production simulations were run for 300 ns with a time step of 2 fs.
+The first 100 ns were discarded for the analysis (@Samuli: this sentence may be removed if the analysis was done on a different time window).
+The simulations were run with GROMACS 4.5.3 [https://doi.org/10.1093/bioinformatics/btt055] version.
+The v-rescale thermostat [http://dx.doi.org/10.1063/1.2408420] was used with a temperature of 300 K and a time constant of 0.1 ps; lipids and water were coupled separately to the heat bath.
+Pressure was kept constant at 1 bar using a semi–isotropic Parrinello–Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 4.0 ps.
+A real space cut-off of 1.0 nm was employed for van der Waals and electrostatic interactions.
+The long-range part of electrostatic interactions was evaluated using the PME method [https://doi.org/10.1063/1.464397, https://doi.org/10.1063/1.470117] with a grid spacing of 0.12 nm and an interpolation order of 4.
+All bonds with hydrogen atoms were constrained with the LINCS algorithm [https://doi.org/10.1002/(SICI)1096-987X(199709)18:12%3C1463::AID-JCC4%3E3.0.CO;2-H].
+Water molecules were kept rigid with the SETTLE algorithm [https://doi.org/10.1002/jcc.540130805].
+The simulation files are available from ref 91 [https://doi.org/10.5281/zenodo.1402417] (pure POPC) and 92 [https://doi.org/10.5281/zenodo.1402449] (POPC:POPE mixture).
 
@@
 
-\noindent {\it POPC:POPE mixtures} \todo{Simulation details by P. Fuchs.}
+\noindent {\it POPC:POPE mixtures}
+
+Force field details
+Simulations of POPC:POPE mixtures using the OPLS-MacRog force field [https://doi.org/10.1021/jp5016627] were performed.
+The initial force field parameter files in GROMACS format were taken from ref [https://doi.org/10.1016/j.dib.2016.03.067].
+However, a number of errors were detected in the published files so we fixed them in the following way.
+For POPE, the two aliphatic chains sn-1 and sn-2 were switched (in the file the lipid was in fact OPPE); one atom (named C27) was not connected to the previous atom in the aliphatic chain so bonds, angles, dihedrals, pairs were included to create the connection; two atoms were called C27 thus one of them was renamed C28.
+For POPC, some impropers were missing leading to non planar systems for double bonds (in particular the carbonyls of sn-1 and sn-2 as well as the double bond of the oleoyl chain), we thus added back those impropers.
+The resulting fixed files can be found on Zenodo (https://zenodo.org/record/3741793/files/popc_fixed.itp, https://zenodo.org/record/3725637/files/pope_fixed.itp).
+To check our fix, one simulation of pure POPC was also run and the order parameter compared to previous published results of ref [https://doi.org/10.1021/acs.jpcb.5b04878].
+We found very similar values.
+All the details of the procedure described here are described in https://www.dsimb.inserm.fr/~fuchs/project_Samuli/POPC_POPE/report_results_comparison.pdf.
+The data for pure POPC are available from [https://zenodo.org/record/3741793], for the POPC:POPE mixture from [https://zenodo.org/record/3725637].
+
+Simulation details
+Two systems of pure POPC and 50:50 POPC:POPE mixture were built using CHARMM-GUI [https://doi.org/10.1021/acs.jctc.5b00935].
+They contained 128 lipids and 5120 TIP3 water molecules.
+The initial PDB file was modified to match OPLS-MacRog nomenclature and atom order.
+No ions were added.
+The production simulations were run for 500 ns with a time step of 2 fs.
+The first 200 ns were discarded for the analysis (@Samuli: this sentence may be removed if the analysis was done on a different time window).
+The simulations were run withthe GROMACS 2018.5 [https://doi.org/10.1016/j.softx.2015.06.001] version.
+The v-rescale thermostat [http://dx.doi.org/10.1063/1.2408420] was used with a temperature of 300 K and a time constant of 0.1 ps; lipids and water were coupled separately to the heat bath.
+Pressure was kept constant at 1 bar using a semi–isotropic Parrinello–Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 4.0 ps.
+Long-range electrostatic interactions were calculated using the PME method [https://doi.org/10.1063/1.464397, https://doi.org/10.1063/1.470117].
+A real space cut-off of 1.0 nm was employed with a grid spacing of 0.1 nm in the reciprocal space.
+Lennard-Jones potentials were cut off at 1.0 nm.
+All covalent bonds in lipids were constrained using the LINCS algorithm [https://doi.org/10.1002/(SICI)1096-987X(199709)18:12%3C1463::AID-JCC4%3E3.0.CO;2-H].
+Water molecules were kept rigid with the SETTLE algorithm [https://doi.org/10.1002/jcc.540130805].
+The simulation files are available from ref 66 [https://doi.org/10.5281/zenodo.3741793] (pure POPC) and 67 [https://doi.org/10.5281/zenodo.3725637] (POPC:POPE mixture).
 
@@
 \noindent
-{\it Am{\'e}lie Bacle}
+{\it Am{\'e}lie Bacle} Amélie Bacle set up, performed and analysed POPC and POPC:POPE (1:1) simulations with the Berger Force Field.
 %% Please write description of your contribution here
@@
 \noindent
-{\it Patrick Fuchs}
+{\it Patrick F.J. Fuchs} supervised Paula Milán Rodríguez, Amélie Bacle and Chris Papadopoulos, created the buildH software, contributed to many discussions.
 %% Please write description of your contribution here
@@
 {\it Jesper J. Madsen}
-set up, performed, and analysed several of the CHARMM36 simulations.
-Provided comments on the manuscript.
+%% Please write description of your contribution here
 
@@
 \noindent
-{\it Paula Milan Rodriguez}
+{\it Paula Milán Rodríguez} set up, performed and analysed POPC and POPC:POPE (1:1) simulations with the MacRog Force Field.
+She also fixed the initial force field (itp) files prior to the simulations.
 %% Please write description of your contribution here
@@
 \noindent
-{\it Chris G. Papadopoulos}
+{\it Chris G. Papadopoulos} set up, performed and analysed POPC and POPC:POPE (1:1) simulations with the CHARMM36 Force Field.
 %% Please write description of your contribution here

### 2021-04-06 manuscriptPGPE.tex

@@
 Conformational ensembles of headgroup and glycerol backbone in PE and PG simulations were evaluated using the C--H bond order parameters \cite{botan15}.
-Interactions between different headgroups were evaluated monitoring the changes in headgroup order parameters upon mixing the lipids \cite{antila19}.
-The ion binding affinities and response of lipids to bound charge were evaluated monitoring the changes of lipid headgroup order parameters \cite{catte16,antila19}.
+Interactions between different headgroups were evaluated by monitoring the changes in headgroup order parameters upon mixing the lipids \cite{antila19}.
+The ion binding affinities and response of lipids to bound charge were evaluated by monitoring the changes in lipid headgroup order parameters \cite{catte16,antila19}.
 
@@
 
-Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.~\ref{HGorderParameters}D), where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
+Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.~\ref{HGorderParameters}D, where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
 The overall agreement of order parameters determined by different research teams and different techniques for the same lipid headgroup is excellent here and in previous studies \cite{botan15,ollila16,antila19}.
-Therefore, the differences between lipid types are dictated by the headgroup chemistry rather than inaccuracies in experiments, or differences in the acyl chains or in experimental conditions.
+Therefore, the differences between lipid types are dictated by the headgroup chemistry rather than inaccuracies in experiments, differences in the acyl chains, or in experimental conditions.
 
@@
 energy costs for different dihedral angle orientations.
-The results in Fig.~\ref{structures}C) suggest that all lipid headgroups are very flexible and
+The results in Fig.~\ref{structures}C suggest that all lipid headgroups are very flexible and
 energy costs for rotating individual dihedrals to almost any angle is low (below 7~$k_\mathrm{B}T$).
@@
 
-In conclusion, all studied lipid headgroups sample very broad conformational ensembles in liquid lamellar phase and the sampled dihedral angles are within approximately same ranges for all headgroup types.
+In conclusion, all studied lipid headgroups sample very broad conformational ensembles in the liquid lamellar phase and the sampled dihedral angles are within approximately same ranges for all headgroup types.
 % (Fig.
@@
 
-Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes reorient the headgroup dipole in PC lipids, thereby affecting the order parameters of lipid headgroups \cite{seelig87}.
-However, the detailed understanding on the structural and energetic responses of lipids to membrane-bound electric charge is still lacking \cite{Semchyschyn04}.
+Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes can reorient the headgroup dipole in PC lipids and affect the order parameters of the lipid headgroups \cite{seelig87}.
+However, the detailed understanding the structural and energetic responses of lipids to membrane-bound electric charge is still lacking \cite{Semchyschyn04}.
 
-To resolve lipid headgroup conformational ensembles in a cell membrane bearing positive charge,
+To resolve lipid headgroup conformational ensembles in cell membranes bearing positive charge,
 we calculated the heavy atom dihedral angle distributions from 
-simulations %(CHARMM36)
-that correctly captured the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer in (see figure \ref{changesWITHsurf} A)).
+the subset of simulations %(CHARMM36)
+that correctly captured the experimentally measured decrease in PC headgroup order parameters upon addition of cationic surfactants into a bilayer (figure \ref{changesWITHsurf} A).
 The dihedral angle distributions and relative energies
@@
 addition of positive charge into a membrane 
-decrease the abundance of \textit{trans} states in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
+decreases the abundance of \textit{trans} states in g$_2$-g$_3$-O$_{g_3}$-P and g$_3$-O$_{g_3}$-P-O$_\alpha$
 dihedrals.
@@
 
-Additionally, the binding of ions to membranes may affect the lipid headgroup conformational ensembles in physiological conditions.
-The binding of Ca$^{2+}$ ions to a PC headgroup leads to a similar decrease in the \textit{trans} state probability for g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
+In addition, binding of ions to the membrane may affect the lipid headgroup conformational ensembles under physiological conditions.
+The bound Ca$^{2+}$ ion to PC headgroup leads to similar decrease in trans state probability for g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
 in the most realistic MD simulations models
@@
 %experimentally observed changes in headgroup order parameters upon addition of
-membrane-bound charge arises
+membrane-bound charges arise
 %from relatively small changes in conformational ensembles. These changes eventuate
@@
 The heavy atom dihedral angle distributions calculated from these conformations 
-in Fig.~\ref{dihedralsFROMpdb} A) reveal that the protein-bound lipids indeed exhibit wide
+(Fig.~\ref{dihedralsFROMpdb} A) reveal that the protein-bound lipids indeed exhibit wide
 range of conformations independently on the headgroup type.
-As in bulk lipid bilayers, only $cis$ conformations
+As for bulk lipid bilayers, only $cis$ conformations
 of P-O$_\alpha$-C$_\alpha$-C$_\beta$ and g$_2$-g$_3$-O$_{g_3}$-P dihedrals are almost
@@
 We suggest that the key to understand selective binding of certain lipid types to proteins
-are intermolecular lipid--protein interactions, rather than conformational restrictions of lipids.
+are intermolecular lipid--protein interactions, rather than conformational restrictions of the lipids.
 On the other hand, broad conformational ensembles in bulk bilayers suggest that lipid crystal structures
@@
 Finally, we demonstrated the power of open access MD simulation data from the NMRlipids open collaboration
-to complement the data in PDB in revealing how complex systems made up of disordered biomolecules behave.
+to complement the data in the PDB for elucidating how complex systems made up of disordered biomolecules behave.
 

### 2021-04-18 manuscriptPGPEsuppl.tex
Fixed bunch of latex syntax errors.

@@
 \author{\scriptsize Am{\'e}lie Bacle}
-\affiliation{Laboratoire Coopératif "Lipotoxicity and Channelopathies – ConicMeds", Université de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }
+\affiliation{Laboratoire Coop\'{e}ratif "Lipotoxicity and Channelopathies – ConicMeds", Universit\'{e} de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }
 
@@
 \author{Patrick F.J. Fuchs}
-\affiliation{Sorbonne Université, Ecole Normale Supérieure, PSL University, CNRS, Laboratoire des Biomolécules (LBM), 75005 Paris, France}
-\alsoaffiliation{Université de Paris, UFR Sciences du Vivant, 75013, Paris, France}
+\affiliation{Sorbonne Universit\'{e}, Ecole Normale Sup\'{e}rieure, PSL University, CNRS, Laboratoire des Biomol\'{e}cules (LBM), 75005 Paris, France}
+\alsoaffiliation{Universit\'{e} de Paris, UFR Sciences du Vivant, 75013, Paris, France}
 
@@
 The v-rescale thermostat [http://dx.doi.org/10.1063/1.2408420] was used with a temperature of 300 K and a time constant of 1 ps; lipids and water were coupled separately to the heat bath.
-Pressure was kept constant at 1 bar using a semi–isotropic Parrinello–Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 5.0 ps.
+Pressure was kept constant at 1 bar using a semi-isotropic Parrinello--Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 5.0 ps.
 A real space cut-off of 1.2 nm was employed for electrostatic interactions while the long-range part was evaluated using the PME method [https://doi.org/10.1063/1.464397, https://doi.org/10.1063/1.470117].
-A force-based switching function was used to switch the Lennard-Jones forces to zero over a range of 1−1.2 nm.
+A force-based switching function was used to switch the Lennard-Jones forces to zero over a range of 1--1.2 nm.
 All bonds with hydrogen atoms were constrained with the LINCS algorithm [https://doi.org/10.1002/(SICI)1096-987X(199709)18:12%3C1463::AID-JCC4%3E3.0.CO;2-H].
@@
 For POPC, some impropers were missing leading to non planar systems for double bonds (in particular the carbonyls of sn-1 and sn-2 as well as the double bond of the oleoyl chain), we thus added back those impropers.
-The resulting fixed files can be found on Zenodo (https://zenodo.org/record/3741793/files/popc_fixed.itp, https://zenodo.org/record/3725637/files/pope_fixed.itp).
+The resulting fixed files can be found on Zenodo {\url{https://zenodo.org/record/3741793/files/popc_fixed.itp}, \url{https://zenodo.org/record/3725637/files/pope_fixed.itp}).
 To check our fix, one simulation of pure POPC was also run and the order parameter compared to previous published results of ref [https://doi.org/10.1021/acs.jpcb.5b04878].
 We found very similar values.
-All the details of the procedure described here are described in https://www.dsimb.inserm.fr/~fuchs/project_Samuli/POPC_POPE/report_results_comparison.pdf.
+All the details of the procedure described here are described in \url{https://www.dsimb.inserm.fr/~fuchs/project_Samuli/POPC_POPE/report_results_comparison.pdf}.
 The data for pure POPC are available from [https://zenodo.org/record/3741793], for the POPC:POPE mixture from [https://zenodo.org/record/3725637].

### 2021-04-19 manuscriptPGPEsuppl.tex
Updates from Overleaf

--- 2021-04-06 manuscriptPGPEsuppl.tex
@@
 \author{\scriptsize Pavel Buslaev}
-\affiliation{\tiny Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, Finland}
-\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, Dolgoprudny, Russia}
+\affiliation{\tiny Nanoscience Center and Department of Chemistry, University of Jyväskylä, P.O. Box 35, 40014 Jyväskylä, Finland}
+\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, 141701 Dolgoprudny, Russia}
 
@@
 \author{Fernando Favela-Rosales}
-\affiliation[Tecnol\'{o}gico Nacional de M\'{e}xico]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}
+\affiliation[Tecnol\'{o}gico Nacional de M\'{e}xico]{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico, ITS Zacatecas Occidente, M\'{e}xico}
 
@@
 \author{Ivan Gushchin}
-\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, Dolgoprudny, Russia}
+\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, 141701 Dolgoprudny, Russia}
 
@@
 %  \centering
-  \includegraphics[width=0.8\textwidth]{./Figs/POPEexperiment.pdf}
+  \includegraphics[width=\textwidth]{./Figs/POPE_SI.pdf}
   \caption{\label{POPEspectra}
-    (A) Chemical structure of POPE with the labeling of headgroup and glycerol backbone carbons.
-    (B) INEPT spectra from POPE sample with the headgroup and glycerol backbone peaks labeled.
-    (C) 2D R-PDLF spectra
-    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
-    (E) Experimetal S-DROSS curves giving signs of the order parameters.
+    (A) Chemical structure of POPE with carbon labels.
+    (B) Refocused INEPT $^{13}$C spectrum of POPE MLVs. The peak assignment was based on previous work. (\todo{citation}) 
+    (C) Headgroup and glycerol backbone R-PDLF dipolar slices.
+The arrows indicate the splittings used to determine the C--H bond order parameters by using  $|S_{\rm{CH}}|=\Delta\nu/0.315/d_{\rm{CH}}$.
+The rigid coupling $d_{\rm{CH}}$ used was 22 kHz.
+    (D) Experimental S-DROSS curves giving signs of the order parameters measured.
   }
@@
 %  \centering
-  \includegraphics[width=0.8\textwidth]{./Figs/PGexpRPDLF.eps}
+  \includegraphics[width=0.8\textwidth]{./Figs/POPG_SI.pdf}
   \caption{\label{POPGspectra}
@@
     (B) INEPT spectra from POPG sample with the headgroup and glycerol backbone peaks labeled.
-    (C) 2D R-PDLF spectra
-    (D) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
-    (E) Experimetal S-DROSS curves giving signs of the order parameters.
+    (C) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
+    (D) Experimetal S-DROSS curves giving signs of the order parameters.
   }
@@
 
-We estimated the relative energies of each dihedral angle value
-with respect to the most probable value (lowest energy) 
-from the inverse Boltzmann formula $\Delta E(\theta) = -kT \left[\ln\left[p(\theta)\right]-\ln\left[p(\theta_0)\right] \right]$,
-where $p(\theta)$ is the dihedral angle distribution and $\theta_0$ is the most probable angle from MD simulation.
-The dihedral angle distributions are shown in Fig. \ref{DIHdists}
-and relative energies in Fig. \ref{structures} in the main text.
+%The dihedral angle distributions are shown in Fig. \ref{DIHdists}
+%and relative energies in Fig. \ref{structures} in the main text.
 
@@
 
-\noindent {\it POPE} \todo{Simulation details by M. Javanainen}.
+\noindent {\it POPE} A lipid bilayer, consisting of a total of 144 POPE molecules, distributed equally between the two leaflets was set using CHARMM-GUI \cite{lee2016charmm}.
+The bilayer was solvated by 5760 water molecules (40 per lipid).
+The random initial configuration and topologies were generated using the CHARMM-GUI web portal, which provides  GROMACS-compatible simulation input files \cite{lee16}.
+CHARMM36 lipid parameters \cite{klauda10} were used for POPE, whereas the CHARMM-specific TIP3P water model \cite{jorgensen1983comparison} was used for water.
+The bilayer was simulated for 500~ns using GROMACS 2018.6 at 310~K.
+
+The recommended simulation parameters for CHARMM36 force field in GROMACS were used \cite{lee16}.
+Namely, buffered Verlet lists were used to keep track of neighbouring atoms \cite{Pall13}.
+The Lennard-Jones potential was cut off at 1.2~nm with the forces switched to 0 between 1.0~nm and the cutoff value.
+The smooth PME algorithm \cite{darden93,essman95} was used to account for long-range electrostatics.
+The temperatures of the lipid and the solvent were separately coupled to a Nos\'{e}-Hoover thermostat \cite{nose84,hoover85} with a time constant of 1~ps and a target temperature of 310~K.
+The system was coupled to a semi-isotropic (isotropic on the membrane plane) Parrinello--Rahman barostat \cite{parrinello81} with a time constant of 5~ps, a reference pressure of 1~bar, and compressibility of 4.5$\times$10$^{-5}$~1/bar.
+The bonds with hydrogen atoms were constrained using p-LINCS \cite{hess07,hess97}.
+
+The simulation files are available at Ref.~\citenum{charmm36POPEfiles}.
 
@@
 
+\noindent{\it POPC:POPE mixture}
+A POPC/POPE bilayer with its lipids distributed evenly among the leaflets was generated by from a pure POPC bilayer by removing and renaming atoms in the head group region.
+The bilayer contained a total of 100 POPC and 100 POPE lipids, and it was solvated by 45 water molecules per lipid (for a total of 9000 water molecules).
+
+The Slipids force field \cite{jambeck12,jambeck12b,jambeck13} was used for lipids, and the TIP3P model \cite{jorgensen1983comparison} for water.
+A 300~ns-long simulation was performed using GROMACS 2019.4 \cite{abraham2015gromacs}.
+The simulation parameters were equal to those used for the POPC/POPG mixture with additional CaCl$_2$ (see below).
+
+The simulation data are available at Ref.~\citenum{POPC1POPE1slipidT298K}.
+
 %\noindent {\it POPC:POPG mixture with additional NaCl} \todo{Simulation details by A. Peon.
@@
 
-\noindent {\it POPC:POPG mixture with additional CaCl} \todo{Simulation details by M. Javanainen.}
+\noindent {\it POPC:POPG mixture with additional CaCl$_2$} 
+A lipid bilayer consisting of a total of 256 lipids (128 POPC + 128 POPG) spread equally between the two leaflets was generated using CHARMM-GUI \cite{lee16}.
+The membrane was solvated by 50 water molecules per lipid (a total of 12800), and 128 Na$^{+}$ counter ions for the POPG charges.
+The initial random configuration was equilibrated using the CHARMM-GUI protocol and using the CHARMM36 force field \cite{klauda10}.
+Next, additional ions were added to obtain initial CaCl$_2$ concentrations of 0, 100, 200, 500, or 1000~mM (0/0, 23/46, 46/92, 115/230, 230/460 Ca$^{2+}$/Cl$^{-}$ ions) The systems were energy-minimized and equilibrated using the Slipids force field \cite{jambeck12,jambeck12b,jambeck13} before production simulations at 298~K.
+The ion parameters by \citeauthor{dang06} were used \cite{dang06}.
+
+The neighbour lists with a cutoff of 1.0~nm were updated every 10 simulation steps.
+The smooth PME algorithm was used to calculate long-range electrostatics \cite{darden93,essman95}.
+The Lennard-Jones potential was cut off at 1.0~nm, and the dispersion corrections \cite{shirts07} were applied to energy and pressure.
+The stochastic velocity rescaling thermostat \cite{bussi07} with a time constant of 0.5~ps and a target temperature of 298~K was applied separately to lipids and the solvent.
+A constant pressure of 1~bar was maintained by a Parrinello--Rahman barostat \cite{parrinello81} with a time constant of 10~ps and compressibility of 4.5$\times$10$^{-5}$~1/bar.
+The pressure coupling was performed semi-isotropically with the two simulation box vectors aligned along the membrane plane considered isotropic.
+All bonds were constrained using the p-LINCS algorithm \cite{hess97,hess07}.
+
+The systems simulated for 1500~ns (CaCl$_2$-containing systems) or 500~ns (systems without CaCl$_2$).
+The simulations were performed using GROMACS 2019.4 \cite{abraham2015gromacs}, and the simulation files are available at Ref.~\citenum{slipidPOPC50POPG50T298K}.
 
@@
 
-\noindent {\it POPE} \todo{Simulation details by M. Javanainen} % and P. Fuchs.}
+\noindent {\it POPE}
+
+A bilayer patch with a total of 144 POPE molecules, distributed evenly between two leaflets, was created by reordering atoms in a final structure of a POPE simulation performed using the CHARMM36 force field.
+The bilayer was hydrated by 40 water molecules per lipid for a total of 5760 water molecules.
+The OPLS-based MacRog force field was used \cite{Kulig15b} for lipids and the TIP3P model \cite{jorgensen1983comparison} for water.
+
+Buffered Verlet lists \cite{Pall13} were used to keep track of neighbours.
+The Lennard-Jones potential was cut off at 1.0~nm, and dispersion corrections were applied to energy and pressure \cite{shirts07}.
+Smooth PME algorithm was used to calculate long-range electrostatics \cite{darden93,essman95}.
+The Nos\'{e}--Hoover thermostat was used to keep the temperatures of the lipids and water at 310~K.
+These groups were coupled separately, and a time constat of 0.4~ps was used.
+The Parrinello--Rahman barostat was used to keep the pressures in the membrane plane as well as normal to it constant at 1~bar.
+For the barostat, a time constant of 10~ps was used, and the membrane compressibility was set to 4.5$\times$10$^{-5}$~1/bar.
+All bonds were constrained using p-LINCS \cite{hess97,hess07}.
+
+The system was simulated for 500~ns using GROMACS 2019.2 \cite{abraham2015gromacs}, and the simulation data are available at Ref.~\citenum{MacRogPOPEfiles}.
 
@@
 For POPC, some impropers were missing leading to non planar systems for double bonds (in particular the carbonyls of sn-1 and sn-2 as well as the double bond of the oleoyl chain), we thus added back those impropers.
-The resulting fixed files can be found on Zenodo (https://zenodo.org/record/3741793/files/popc_fixed.itp, https://zenodo.org/record/3725637/files/pope_fixed.itp).
+The resulting fixed files can be found on Zenodo (https://zenodo.org/record/3741793/files/popc\_fixed.itp, https://zenodo.org/record/3725637/files/pope\_fixed.itp).
 To check our fix, one simulation of pure POPC was also run and the order parameter compared to previous published results of ref [https://doi.org/10.1021/acs.jpcb.5b04878].
 We found very similar values.
-All the details of the procedure described here are described in https://www.dsimb.inserm.fr/~fuchs/project_Samuli/POPC_POPE/report_results_comparison.pdf.
+All the details of the procedure described here are described in https://www.dsimb.inserm.fr/~fuchs/project\_Samuli/POPC\_POPE/report\_results\_comparison.pdf.
 The data for pure POPC are available from [https://zenodo.org/record/3741793], for the POPC:POPE mixture from [https://zenodo.org/record/3725637].
@@
 %% Please write description of your contribution here
+Performed the analysis of dihedrals and isomers of lipids. Analysed lipid structures from Protein Data Bank. Prepared panels for figures 2 and 4. Participated in discussions.
 
@@
 \noindent
-{\it Fernando Favela-Rosales}
+{\it Fernando Favela-Rosales} set up and performed DPPE, POPG and DPPG simulations with the Slipids Force Field.
 %% Please write description of your contribution here
@@
 \noindent
-{\it Ivan Gushchin}
+{\it Ivan Gushchin} supervised Pavel Buslaev and contributed to the analysis of lipid structures in the Protein Data Bank.
 %% Please write description of your contribution here
@@
 \noindent
-{\it Matti Javanainen}
+{\it Matti Javanainen} set up and performed simulations using CHARMM36, Slipids, and MacRog lipid models. He contributed to the organization of the manuscript.
 %% Please write description of your contribution here

### 2021-04-19 manuscriptPGPE.tex

@@
 \author{Pavel Buslaev}
-\affiliation{Nanoscience Center and Department of Chemistry, University of Jyv{\"a}skyl{\"a}, Finland}
-\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, Dolgoprudny, Russia}
+\affiliation{Nanoscience Center and Department of Chemistry, University of Jyväskylä, P.O. Box 35, 40014 Jyväskylä, Finland}
+\affiliation{Research Center for Molecular Mechanisms of Aging and Age-related Diseases, Moscow Institute of Physics and Technology, 141701 Dolgoprudny, Russia}
 
@@
 \author{Fernando Favela-Rosales}
-\affiliation{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico, Campus Zacatecas Occidente, M\'{e}xico}
+\affiliation{Departamento de Ciencias B\'{a}sicas, Tecnol\'{o}gico Nacional de M\'{e}xico - ITS Zacatecas Occidente, M\'{e}xico}
 
@@
   in probability distributions of conformations, but all lipid types can access almost any of the possible conformations.
-  Together with the analysis of protein-bound lipids from the protein data bank (PDB), this suggests that lipids can bind to proteins in wide range of conformations independently on the headgroup chemistry.
+  Together with the analysis of protein-bound lipids from the protein data bank (PDB), this suggests that lipids can bind to proteins in a wide range of conformations independently of their headgroup chemistry.
 Therefore, the selective adsorption of proteins to membranes is likely regulated by specific protein--lipid interactions rather than conformational restrictions of the bilayer lipids.
@@
 
+Relative energy costs for turning dihedral angles with respect to the most probable value (lowest energy) were estimated from the inverse Boltzmann formula $\Delta E(\theta) = -kT \left[\ln\left[p(\theta)\right]-\ln\left[p(\theta_0)\right] \right]$, where $p(\theta)$ is the dihedral angle distribution and $\theta_0$ is the most probable angle from MD simulation.
+
 %\clearpage
@@
 \subsection{Analysis of protein-bound lipid conformations}
-Lipid structures from Protein Data Bank (PDB, \url{http://www.rcsb.org/})
+Lipid structures from the Protein Data Bank (PDB, \cite{berman2000})
 were searched using PDBe REST API (\url{www.ebi.ac.uk/pdbe/pdbe-rest-api})
@@
 Heavy atom dihedral distributions from the lipid structures were calculated
-using the MDAnalysis python library \cite{agrawal11,gowers16} and
+using the MDAnalysis Python library \cite{agrawal11,gowers16} and
 Jupyter notebook available from \url{https://github.com/pbuslaev/scr/blob/master/PDB%20analysis.ipynb}.
-Only structures from X-ray and Cryo-EM with a resolution higher than 3.2~\AA were used in the analysis.
-Lipids from same PDB identifier with all dihedrals angles within 3 degrees were discarded from the analysis
-to avoid overweighting of lipid structures present, \textit{e.g.}, due to rotational symmetries.
+Only the structures determined using X-ray crystallography or Cryo-EM with a resolution higher than 3.2~\text{\AA} were used in the analysis.
+Some structures of multimeric proteins contained multiple lipids in the same conformation (analyzed dihedral angles within 3 degrees of each other).
+In that case, this conformation was counted once for the subsequent analyses
+to avoid its overweighting.
 
@@
 we searched for pairs having the last five dihedral angles (excluding g$_1$-g$_2$-g$_3$-O$_{g3}$) within
-30 degrees from structures with resolution higher than 2.5~\AA. The two most representative examples
+30 degrees of each other among the structures with resolution higher than 2.5~\text{\AA}. The two most representative examples
 were handpicked from the results.
@@
 The increased barriers for dihedral rotations may explain the more rigid headgroup structures in PS~\cite{browning80,buldt81}.
-Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.~\ref{structures}D)).
+Furthermore, the angle between headgroup dipole and membrane normal decreases in the order of PG $>$ PE  $>$ PC  $>$ PS (Fig.~\ref{structures}D).
 However, the differences between PC and PE in P-O$_\alpha$-C$_\alpha$-C$_\beta$ dihedral
@@
 Charged entities, such as lipids, proteins, surfactants, drugs, and ions incorporated in membranes can reorient the headgroup dipole in PC lipids and affect the order parameters of the lipid headgroups \cite{seelig87}.
-However, the detailed understanding the structural and energetic responses of lipids to membrane-bound electric charge is still lacking \cite{Semchyschyn04}.
+However, the detailed understanding of the structural and energetic responses of lipids to membrane-bound electric charge is still lacking \cite{Semchyschyn04}.
 
@@
     B) The structure of PE headgroup bound to cytochrome {\it bc}$_1$ complex (PDB ID: 1PP9 \cite{huang05})
-    with identical conformation as the PC headgroup bound to yeast sec14 (PDB ID: 3B7Q \cite{schaaf08}),
+    with identical conformation as the PC headgroup bound to yeast Sec14 (PDB ID: 3B7Q \cite{schaaf08}),
     and the structure of PG headgroup bound to bovine cytochrome {\it c} oxidase (PDB ID: 3AG3 \cite{muramoto10})
@@
 Our results suggest that flexible lipid headgroups can optimize the intermolecular interactions with proteins by binding
-in wide range of conformations.
+in a wide range of conformations.
 Therefore, the specific binding of lipids to proteins is not driven by the structural differences between headgroups.
 This is demonstrated in Fig.~\ref{dihedralsFROMpdb} B with two examples where different lipids bound to
-different proteins have identical conformations:
-PE in cytochrome {\it bc}$_1$ complex is similar to PC bound to yeast sec14,
+different proteins have almost identical headgroup conformations:
+PE in cytochrome {\it bc}$_1$ complex is similar to PC bound to yeast Sec14,
 and PG bound to bovine cytochrome {\it c} oxidase is similar to PC bound to pore-forming toxin (FraC).
@@
 
+P.B. was supportedby the Academy of Finland (Grant 311031)
+
+F.F.-R.
+acknowledges Tecnol\'{o}gico Nacional de M\'{e}xico Proyecto IT16C431, Direcci\'{o}n General de Asuntos del Personal Acad\'{e}mico (DGAPA) Programa de Apoyo a Proyectos de Investigaci\'{o}n e Innovaci\'{o}n Tecnol\’{o}gica (PAPIIT) IG100920, CONACyT Ciencia de Frontera 74884 for financial support and Miztli-Direcci\'{o}n de C\'{o}mputo y de Tecnolog\'{i}as de Informaci\'{o}n y Comunicaci\'{o}n (DGTIC) - Universidad Nacional Aut\'{o}noma de México (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing-time allocation.
+
+I.G. was supported by the Ministry of Science and Higher Education of the Russian Federation (agreement #075-00337-20-03, project FSMG-2020-0003).
+
 J.J.M. gratefully acknowledges financial support from the Carlsberg 

### 2021-04-19 manuscriptPGPEsuppl.tex

@@
 However, a number of errors were detected in the published files so we fixed them in the following way. 
-For POPE, the two aliphatic chains sn-1 and sn-2 were switched (in the file the lipid was in fact OPPE); one atom (named C27) was not connected to the previous atom in the aliphatic chain so bonds, angles, dihedrals, pairs were included to create the connection; two atoms were called C27 thus one of them was renamed C28.
-For POPC, some impropers were missing leading to non planar systems for double bonds (in particular the carbonyls of sn-1 and sn-2 as well as the double bond of the oleoyl chain), we thus added back those impropers.
-The resulting fixed files can be found on Zenodo (https://zenodo.org/record/3741793/files/popc\_fixed.itp, https://zenodo.org/record/3725637/files/pope\_fixed.itp).
-To check our fix, one simulation of pure POPC was also run and the order parameter compared to previous published results of ref [https://doi.org/10.1021/acs.jpcb.5b04878].
-We found very similar values.
-All the details of the procedure described here are described in https://www.dsimb.inserm.fr/~fuchs/project\_Samuli/POPC\_POPE/report\_results\_comparison.pdf.
+For POPE, the two aliphatic chains sn-1 and sn-2 were switched (in the file the lipid was in fact OPPE); 
+one atom (named C27) was not connected to the previous atom in the aliphatic chain so bonds, angles, dihedrals, 
+pairs were included to create the connection; two atoms were called C27 thus one of them was renamed C28. 
+For POPC, some impropers were missing leading to non planar systems for double bonds 
+(in particular the carbonyls of sn-1 and sn-2 as well as the double bond of the oleoyl chain), 
+we thus added back those impropers. The resulting fixed files can be found on 
+Zenodo (https://zenodo.org/record/3741793/files/popc\_fixed.itp, https://zenodo.org/record/3725637/files/pope\_fixed.itp). 
+To check our fix, one simulation of pure POPC was also run and the order parameter compared to previous published results 
+of ref [https://doi.org/10.1021/acs.jpcb.5b04878]. We found very similar values.
+All the details of the procedure described here are described 
+in https://www.dsimb.inserm.fr/~fuchs/project\_Samuli/POPC\_POPE/report\_results\_comparison.pdf.
+
 The data for pure POPC are available from [https://zenodo.org/record/3741793], for the POPC:POPE mixture from [https://zenodo.org/record/3725637].

### 2021-04-19 manuscriptPGPE.tex
Update pdfs

@@
 \author{Markus S. Miettinen}
+\affiliation{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
 % \affiliation[Max Planck Institute of Colloids and Interfaces]{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
-\affiliation{Department of Theory and Bio-Systems, Max Planck Institute of Colloids and Interfaces, 14424 Potsdam, Germany}
 
@@
 F.F.-R.
-acknowledges Tecnol\'{o}gico Nacional de M\'{e}xico Proyecto IT16C431, Direcci\'{o}n General de Asuntos del Personal Acad\'{e}mico (DGAPA) Programa de Apoyo a Proyectos de Investigaci\'{o}n e Innovaci\'{o}n Tecnol\’{o}gica (PAPIIT) IG100920, CONACyT Ciencia de Frontera 74884 for financial support and Miztli-Direcci\'{o}n de C\'{o}mputo y de Tecnolog\'{i}as de Informaci\'{o}n y Comunicaci\'{o}n (DGTIC) - Universidad Nacional Aut\'{o}noma de México (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing-time allocation.
+acknowledges Tecnol\'{o}gico Nacional de M\'{e}xico Proyecto IT16C431, Direcci\'{o}n General de Asuntos del Personal Acad\'{e}mico (DGAPA) Programa de Apoyo a Proyectos de Investigaci\'{o}n e Innovaci\'{o}n Tecnol\'{o}gica (PAPIIT) IG100920, CONACyT Ciencia de Frontera 74884 for financial support and Miztli-Direcci\'{o}n de C\'{o}mputo y de Tecnolog\'{i}as de Informaci\'{o}n y Comunicaci\'{o}n (DGTIC) - Universidad Nacional Aut\'{o}noma de México (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing-time allocation.
 
-I.G. was supported by the Ministry of Science and Higher Education of the Russian Federation (agreement #075-00337-20-03, project FSMG-2020-0003).
+I.G. was supported by the Ministry of Science and Higher Education of the Russian Federation (agreement \#075-00337-20-03, project FSMG-2020-0003).
 

### 2021-04-19 manuscriptPGPEsuppl.tex
Update pdfs

@@
     (A) Chemical structure of POPE with carbon labels.
-    (B) Refocused INEPT $^{13}$C spectrum of POPE MLVs. The peak assignment was based on previous work. (\todo{citation}) 
+    (B) Refocused INEPT $^{13}$C spectrum of POPE MLVs. The peak assignment was based on previous work. %(\todo{citation}) 
     (C) Headgroup and glycerol backbone R-PDLF dipolar slices.
@@
 The rigid coupling $d_{\rm{CH}}$ used was 22 kHz.
-    (D) Experimental S-DROSS curves giving signs of the order parameters measured.
-  }
+    (D) Experimental S-DROSS curves giving signs of the order parameters measured. }
   \todo{NMR figures in SI to be updated.}
@@
 \noindent
-{\it Am{\'e}lie Bacle} Amélie Bacle set up, performed and analysed POPC and POPC:POPE (1:1) simulations with the Berger Force Field.
+{\it Am{\'e}lie Bacle} Am{\'e}lie Bacle set up, performed and analysed POPC and POPC:POPE (1:1) simulations with the Berger Force Field.
 %% Please write description of your contribution here
@@
 \noindent
-{\it Patrick F.J. Fuchs} supervised Paula Milán Rodríguez, Amélie Bacle and Chris Papadopoulos, created the buildH software, contributed to many discussions.
+{\it Patrick F.J. Fuchs} supervised Paula Mil{\'a}n Rodr{\'i}guez, Am{\'e}lie Bacle and Chris Papadopoulos, created the buildH software, contributed to many discussions.
 %% Please write description of your contribution here
@@
 \noindent
-{\it Paula Milán Rodríguez} set up, performed and analysed POPC and POPC:POPE (1:1) simulations with the MacRog Force Field.
+{\it Paula Mil{\'a}n Rodr{\'i}guez} set up, performed and analysed POPC and POPC:POPE (1:1) simulations with the MacRog Force Field.
 She also fixed the initial force field (itp) files prior to the simulations.

### 2021-04-19 manuscriptPGPEsuppl.tex
Add author
Updates from Overleaf

--- 2021-04-19 manuscriptPGPEsuppl.tex
@@
 
+\clearpage
 \section{R-PDLF and SDROSS experiments}
@@
   \caption{\label{POPEspectra}
-    (A) Chemical structure of POPE with carbon labels.
-    (B) Refocused INEPT $^{13}$C spectrum of POPE MLVs. The peak assignment was based on previous work. (\todo{citation}) 
+    Solid-state NMR spectroscopy for determining the C--H bond order parameters, $S_{\rm{CH}}$, in the POPE headgroup and glycerol backbone.
+(A) Chemical structure of POPE showing the labels used to identify different carbons.
+    (B) Refocused INEPT $^{13}$C spectrum of POPE MLVs. 
+    (C) Headgroup and glycerol backbone R-PDLF dipolar slices.
+The arrows indicate the splittings used to determine the C--H bond order parameters by using  $|S_{\rm{CH}}|=\Delta\nu/0.315/d_{\rm{CH}}$.
+The rigid coupling $d_{\rm{CH}}$ used was 22 kHz.
+    (D) Experimental S-DROSS curves giving signs of the order parameters measured. Grey lines are noise-only profiles.}
+\end{figure}
+
+\begin{figure}[]
+%  \centering
+  \includegraphics[width=\textwidth]{./Figs/POPG_SI.pdf}
+  \caption{\label{POPGspectra}
+    Solid-state NMR spectroscopy for determining the C--H bond order parameters, $S_{\rm{CH}}$, in the POPG headgroup and glycerol backbone.
+(A) Chemical structure of POPG showing the labels used to identify different carbons.
+    (B) Refocused INEPT $^{13}$C spectrum of POPG MLVs.
     (C) Headgroup and glycerol backbone R-PDLF dipolar slices.
@@
     (D) Experimental S-DROSS curves giving signs of the order parameters measured.
-  }
-  \todo{NMR figures in SI to be updated.}
-\end{figure}
-
-\begin{figure}[]
-%  \centering
-  \includegraphics[width=0.8\textwidth]{./Figs/POPG_SI.pdf}
-  \caption{\label{POPGspectra}
-    (A) Chemical structure of POPG with the labeling of headgroup and glycerol backbone carbons.
-    (B) INEPT spectra from POPG sample with the headgroup and glycerol backbone peaks labeled.
-    (C) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
-    (D) Experimetal S-DROSS curves giving signs of the order parameters.
-  }
-\end{figure}
-
-\begin{figure}[]
-%  \centering
-  \includegraphics[width=0.8\textwidth]{./Figs/SIMPSON.pdf}
-  \caption{\label{POPGsimpson}
-    Simpson simulaton of S-DROSS curve of $\beta$-carbon of POPG.
+Grey lines are a set of slices taken from a region of the spectrum without peaks, i.e.
+grey lines are noise profiles, to highlight the signal-to-noise ratio of the measured modulations.
+Note that the S-DROSS modulation for $g_3$ is very close to the noise level.
+In this case, we have confirmed that the order parameter sign of $g_3$ was negative by performing a measurement with less points in the indirect dimension, more scans and using a higher MAS frequency of 8 kHz.
   }

### 2021-04-19 manuscriptPGPEsuppl.tex

--- 2021-04-19 manuscriptPGPEsuppl.tex
@@
 
+\author{Salla I. Virtanen}
+\affiliation{Institute of Biotechnology, University of Helsinki}
+
 %\author{O. H. Samuli Ollila}
@@
   \caption{\label{POPEspectra}
-    (A) Chemical structure of POPE with carbon labels.
-    (B) Refocused INEPT $^{13}$C spectrum of POPE MLVs. The peak assignment was based on previous work. %(\todo{citation}) 
+    Solid-state NMR spectroscopy for determining the C--H bond order parameters, $S_{\rm{CH}}$, in the POPE headgroup and glycerol backbone.
+(A) Chemical structure of POPE showing the labels used to identify different carbons.
+    (B) Refocused INEPT $^{13}$C spectrum of POPE MLVs. 
     (C) Headgroup and glycerol backbone R-PDLF dipolar slices.
@@
 The rigid coupling $d_{\rm{CH}}$ used was 22 kHz.
-    (D) Experimental S-DROSS curves giving signs of the order parameters measured. }
-  \todo{NMR figures in SI to be updated.}
+    (D) Experimental S-DROSS curves giving signs of the order parameters measured. Grey lines are noise-only profiles.}
 \end{figure}
@@
 %  \centering
-  \includegraphics[width=0.8\textwidth]{./Figs/POPG_SI.pdf}
+  \includegraphics[width=\textwidth]{./Figs/POPG_SI.pdf}
   \caption{\label{POPGspectra}
-    (A) Chemical structure of POPG with the labeling of headgroup and glycerol backbone carbons.
-    (B) INEPT spectra from POPG sample with the headgroup and glycerol backbone peaks labeled.
-    (C) Dipolar sliced from the  2D R-PDLF spectra with the resulting order parameters on top of figures.
-    (D) Experimetal S-DROSS curves giving signs of the order parameters.
-  }
-\end{figure}
-
-\begin{figure}[]
-%  \centering
-  \includegraphics[width=0.8\textwidth]{./Figs/SIMPSON.pdf}
-  \caption{\label{POPGsimpson}
-    Simpson simulaton of S-DROSS curve of $\beta$-carbon of POPG.
+    Solid-state NMR spectroscopy for determining the C--H bond order parameters, $S_{\rm{CH}}$, in the POPG headgroup and glycerol backbone.
+(A) Chemical structure of POPG showing the labels used to identify different carbons.
+    (B) Refocused INEPT $^{13}$C spectrum of POPG MLVs.
+    (C) Headgroup and glycerol backbone R-PDLF dipolar slices.
+The arrows indicate the splittings used to determine the C--H bond order parameters by using  $|S_{\rm{CH}}|=\Delta\nu/0.315/d_{\rm{CH}}$.
+The rigid coupling $d_{\rm{CH}}$ used was 22 kHz.
+    (D) Experimental S-DROSS curves giving signs of the order parameters measured.
+Grey lines are a set of slices taken from a region of the spectrum without peaks, i.e.
+grey lines are noise profiles, to highlight the signal-to-noise ratio of the measured modulations.
+Note that the S-DROSS modulation for $g_3$ is very close to the noise level.
+In this case, we have confirmed that the order parameter sign of $g_3$ was negative by performing a measurement with less points in the indirect dimension, more scans and using a higher MAS frequency of 8 kHz.
   }
@@
 
+\noindent
+{\it Salla I. Virtanen}
+%% Please write description of your contribution here
+
 \bibliography{refs.bib}

### 2021-04-19 manuscriptPGPEsuppl.tex
Updates from Overleaf

--- 2021-04-19 manuscriptPGPEsuppl.tex
@@
 Note that the S-DROSS modulation for $g_3$ is very close to the noise level.
-In this case, we have confirmed that the order parameter sign of $g_3$ was negative by performing a measurement with less points in the indirect dimension, more scans and using a higher MAS frequency of 8 kHz.
+In this case, we have confirmed that the order parameter sign of $g_3$ was negative by performing a measurement with less points in the indirect dimension, more scans and using a higher MAS frequency of 8 kHz (not shown).
   }
@@
 {\it Tiago M. Ferreira}
-%% Please write description of your contribution here
+performed and analysed the R-PDLF and S-DROSS experiments. Wrote p
 

### 2021-04-22 manuscriptPGPE.tex
Add author
Updates from Overleaf

@@
 \author{Am{\'e}lie Bacle}
-\affiliation{Laboratoire Coopératif "Lipotoxicity and Channelopathies – ConicMeds", Université de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, Franc }
+\affiliation{Laboratoire Coop{\'e}ratif "Lipotoxicity and Channelopathies – ConicMeds", Universit{\'e} de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, Franc }
 
@@
 \affiliation{Center for Research in Biological Chemistry and Molecular Materials (CiQUS), Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}
-\affiliation{CIQUP, Centro de Investigação em Química, Departamento de Química e Bioquímica, Faculdade de Ciências, Universidade do Porto, Porto, Portugal}
+\affiliation{CIQUP, Centro de Investigação em Qu{\'i}mica, Departamento de Qu{\'i}mica e Bioqu{\'i}mica, Faculdade de Ci{\^e}ncias, Universidade do Porto, Porto, Portugal}
 
@@
 \author{Patrick F.J. Fuchs}
-\affiliation{Sorbonne Université, Ecole Normale Supérieure, PSL University, CNRS, Laboratoire des Biomolécules (LBM), 75005 Paris, France}
-\affiliation{Université de Paris, UFR Sciences du Vivant, 75013, Paris, France}
+\affiliation{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), 75005 Paris, France}
+\affiliation{Universit{\'e} de Paris, UFR Sciences du Vivant, 75013, Paris, France}
 
@@
 
-\author{Paula Milan Rodriguez}
-\affiliation{Sorbonne Université, Ecole Normale Supérieure, PSL University, CNRS, Laboratoire des Biomolécules (LBM), 75005 Paris, France}
+\author{Paula Mil{\'a}n Rodr{\'i}guez}
+\affiliation{Sorbonne Universit{\'e}, Ecole Normale Sup{\'e}rieure, PSL University, CNRS, Laboratoire des Biomol{\'e}cules (LBM), 75005 Paris, France}
 
@@
 \author{Chris G. Papadopoulos}
-\affiliation{Université Paris-Saclay, CEA, CNRS, Institute for Integrative Biology of the Cell (I2BC), 91198 Gif-sur-Yvette, France}
+\affiliation{Universit{\'e} Paris-Saclay, CEA, CNRS, Institute for Integrative Biology of the Cell (I2BC), 91198 Gif-sur-Yvette, France}
 
@@
 \affiliation{Departamento de F{\'i}sica Aplicada, Facultade de F{\'i}sica, Universidade de Santiago de Compostela, E-15782 Santiago de Compostela, Spain}
+
+\author{Salla I. Virtanen}
+\affiliation{Institute of Biotechnology, University of Helsinki}
 
@@
 
-\maketitle %\maketitle must follow title, authors, abstract and \pacs
+\maketitle
 
@@
 but their relation to the conformational ensembles in bulk membranes in the liquid lamellar phase remains unclear \cite{marsh13b}.
-%In addition to the changes in lipid conformational ensembles upon binding to proteins,
-%also the experimentally measured response of lipid headgroup to membrane bound charges remains poorly understood 
-%due to the lack of suitable models to interpret the lipid conformational ensembles in liquid lamellar state \cite{Semchyschyn04}.
 Most accurate experimental information on conformational ensembles of lipids
-%in the physiologically relevant
 in liquid lamellar phase are typically derived from NMR experiments, particularly from 
@@
 All the studied lipids specifically bind to various proteins \cite{yeagle14}.
-%We use our results to elucidate also lipid--protein interactions and the effect of charges on lipid conformations.
 
@@
 suggest that our results can be used to 
-%the results from model systems could be used to
 understand the biological role of lipid headgroups in lipid--protein interactions.
@@
 design of phoshoplipid-specific antibodies \cite{vigant15}.
-%The lipid conformational ensembles in liquid lamellar state paves the way toward understanding the
-%specific binding of different lipid types to membrane proteins and how they regulate the protein function.
-
-%Notably, such measurements can be performed also on living cells \cite{gally81,scherer87,seelig90}.
 
@@
 
+The differences in headgroup conformation ensembles were analyzed calculating the distributions of the last six heavy atom dihedral angles, i.e., dihedrals O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$, P-O$_\alpha$-C$_\alpha$-C$_\beta$, O$_{g3}$-P-O$_\alpha$-C$_\alpha$, g$_3$-O$_{g3}$-P-O$_\alpha$, g$_2$-g$_3$-O$_{g3}$-P, and g$_1$-g$_2$-g$_3$-O$_{g3}$ labeled in Fig.
+\ref{HGorderParameters} C.
 Relative energy costs for turning dihedral angles with respect to the most probable value (lowest energy) were estimated from the inverse Boltzmann formula $\Delta E(\theta) = -kT \left[\ln\left[p(\theta)\right]-\ln\left[p(\theta_0)\right] \right]$, where $p(\theta)$ is the dihedral angle distribution and $\theta_0$ is the most probable angle from MD simulation.
@@
   %  \includegraphics[width=9.0cm]{../Figs/HGorderparametersPCPSPEPG.eps}
-   \includegraphics[width=18.0cm]{./Figs/figure1.pdf}
+   \includegraphics[width=18.0cm]{./Figs/figure_PG_PE.pdf}
    \caption{\label{HGorderParameters}
@@
 using the MDAnalysis Python library \cite{agrawal11,gowers16} and
-Jupyter notebook available from \url{https://github.com/pbuslaev/scr/blob/master/PDB%20analysis.ipynb}.
+Jupyter notebook available from {\it scripts/PDBanalysis.ipynb} folder in Ref~\citenum{NMRlipidsIVbgit}.
 Only the structures determined using X-ray crystallography or Cryo-EM with a resolution higher than 3.2~\text{\AA} were used in the analysis.
@@
 
-To demonstrate the existence of similar structures of different lipids bound to different proteins,
-we searched for pairs having the last five dihedral angles (excluding g$_1$-g$_2$-g$_3$-O$_{g3}$) within
+To demonstrate the existence of similar structures of different lipids bound to different proteins, we searched for pairs having the last five headgroup heavy atom dihedral angles (excluding g$_1$-g$_2$-g$_3$-O$_{g3}$) within
 30 degrees of each other among the structures with resolution higher than 2.5~\text{\AA}. The two most representative examples
@@
 To experimentally characterize the differences in headgroup conformational ensembles of lipids that are not
-bound to proteins in an electrostatically neutral cell membrane, we measured the C--H bond order parameters
+bound to proteins, we measured the C--H bond order parameters
 and their signs of POPG and POPE in the liquid lamellar phase, as we did previously for POPC and POPS \cite{ferreira13,ferreira16,antila19}.
@@
 
-Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.~\ref{HGorderParameters}D, where signs determined from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
+Experimental order parameters of POPC, POPE, POPG and POPS glycerol backbones and headgroups from this and previous studies are collected in Fig.~\ref{HGorderParameters}D, where the signs from $^{13}$C NMR experiments are used also for the $^2$H NMR data from the literature.
 The overall agreement of order parameters determined by different research teams and different techniques for the same lipid headgroup is excellent here and in previous studies \cite{botan15,ollila16,antila19}.
@@
 
+MD simulations could be used to interpret the lipid conformational ensembles from experiments due to their direct connection to the measured C-H bond order parameters.
+To find the best models for this purpose, we have evaluated all the available MD simulation force fields of PC~\cite{botan15}, PS~\cite{antila19}, PE (Fig.~\ref{HGorderParametersPE}), and PG (Fig.~\ref{HGorderParametersPOPG}) lipids against NMR data in the NMRlipids project.
+The comparison of hundreds of simulation trajectories simulated with approximately 20 different force fields reveal that none of the current parameters can correctly capture the lipid conformational ensemble and reproduce the experimental C-H bond order parameters of all segments.
+However, the best parameters (CHARMM36) correctly capture the essential differences between PC, PE, PG and PS headgroup order parameters (Fig.~\ref{structures} B).
+
 To understand the structural origin of distinct order parameters for PG and PS lipids,
-we first calculated the heavy atom dihedral angle distributions from simulations
-that best reproduce the differences between headgroups according to the quality
-evaluation in the SI and Refs. \cite{botan15,antila19} (Fig. \ref{DIHdists}).
+we first calculated the heavy atom dihedral angle distributions from the best performing simulations that reproduced the differences between headgroups~(Fig. \ref{DIHdists}).
 Then, we used the inverse Boltzmann formula to estimate the
@@
 in g$_3$-O$_{{\mathrm g}3}$-P-O$_\alpha$ dihedral was observed.
-Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameters occur in the choline part, while also changes in phosphate region may contribute in PS lipids.
+Therefore we suggest that the main differences between lipid headgroups leading to distinct order parameters occur in the headgroup end, while phosphate region remains similar in different lipids with the exception of PS.
 The increased barriers for dihedral rotations may explain the more rigid headgroup structures in PS~\cite{browning80,buldt81}.
@@
 
-In conclusion, all studied lipid headgroups sample very broad conformational ensembles in the liquid lamellar phase and the sampled dihedral angles are within approximately same ranges for all headgroup types.
+In conclusion, all studied lipid headgroups sample very broad conformational ensembles in the liquid lamellar phase and the sampled dihedral angles are within similar ranges for all headgroup types.
 % (Fig.
 \ref{structures}).
-Since the rotation of dihedral angles to almost any position bears relatively low free energy cost,
-%results suggest that all lipid headgroups are very flexible---
-%thereby being
-the lipid headgroups are able to adopt a wide range of conformations when interacting with proteins, ions, or other biomolecules.
+Since the rotation of dihedral angles to almost any position bears relatively low free energy cost, the lipid headgroups are able to adopt a wide range of conformations when interacting with proteins, ions, or other biomolecules.
 %Our results support the models proposing free rotation around phosphate group for PC lipids, which decouples the dynamics and conformations between acyl chains and headgroups \cite{klauda08c,antila21b}, and suggests that these models could be applied for all lipids containing the phosphorus group.
 %The wide range of observed conformations suggest that
-Therefore, the structures in lipid crystals \cite{buldt81,pascher92} play only a minor role, and models aiming to explain NMR data using only a few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in the liquid lamellar state.
+Therefore, the structures in lipid crystals \cite{buldt81,pascher92} are not representative of the liquid state, and models aiming to explain NMR data using only a few conformations \cite{seelig77c,davis83,Semchyschyn04,akutsu20} are not sufficient to capture the large conformational space of lipids in the liquid lamellar state.
+However, MD simulations with accurate force fields can be expected to capture the correct conformational ensembles of lipids.
 
@@
 
-To resolve lipid headgroup conformational ensembles in cell membranes bearing positive charge,
+To resolve lipid headgroup conformational ensembles in cellular membranes bearing positive charge,
 we calculated the heavy atom dihedral angle distributions from 
@@
 The bound Ca$^{2+}$ ion to PC headgroup leads to similar decrease in trans state probability for g$_3$-O$_{g3}$-P-O$_\alpha$ dihedral
-in the most realistic MD simulations models
+in the most realistic MD simulation models
 as observed for cationic surfactants
@@
 allowing them to bind proteins in various different conformations.
-To test this prediction, we analyzed the protein-bound lipid conformations
+To test this prediction, we analyzed the conformations of all protein-bound lipids 
 from structures deposited in the PDB \cite{berman00}.
-We found 311 PC, 394 PE, 154 PG, and 35 PS protein-bound lipid conformations
-%presenting lipids that are tightly bound to proteins in fixed conformations
-that were determined as a part of protein structure using crystallography or cryo-EM.
+We found 311 PC, 394 PE, 154 PG, and 35 PS lipid conformations that were bound to different types of proteins and were determined as a part of protein structure using crystallography or cryo-EM.
+The full list of structures with bound lipids is available from {\it Data/PDBdistributions/pdbtable.pdf} in Ref.~\cite{NMRlipidsIVbgit}.
 
@@
 are realistic also in protein-bound states.
-%The %range of available angles for
-%O$_\alpha$-C$_\alpha$-C$_\beta$-N/C$_\gamma$ and
-%g$_1$-g$_2$-g$_3$-O$_{g_3}$ dihedrals are even less restricted in protein bound state
-%than in liquid lamellar phase where $cis$ and $anti$ eclipsed state were not present (Fig.~\ref{structures}). 
-%Major differences between different lipid headgroups are not observed when bound to proteins.
-%Slight preference for the trans state in O$_\alpha$-C$_\alpha$-C$_\beta$-/C$_\gamma$ dihedral of PC,
-%and for positive angles in P-O$_\alpha$-C$_\alpha$-C$_\gamma$ dihedral of PS lipids with respect
-%to other lipids could be present in the data, but the statistics is not sufficient for conclusions.
-%Therefore, the differences in conformational ensembles between different
-%lipids in liquid lamellar state in Fig.~\ref{structures} are not seen in the protein bound states.
 
@@
 adapt to various binding sites to optimize the intermolecular lipid--protein interactions.
-We tested this prediction by analyzing the conformations of lipids that are tightly bound to proteins in from the PDB.
+We tested this prediction by analyzing the conformations of lipids that are tightly bound to proteins in the PDB.
 Indeed, also protein--bound lipids exhibit wide range of conformations without significant 
@@
 On the other hand, broad conformational ensembles in bulk bilayers suggest that lipid crystal structures
-play a minor role and that the entropic cost of lipid binding may be significant.
+play a minor role while the entropic cost of lipid binding may be significant.
 Finally, we demonstrated the power of open access MD simulation data from the NMRlipids open collaboration
@@
 F.F.-R.
-acknowledges Tecnol\'{o}gico Nacional de M\'{e}xico Proyecto IT16C431, Direcci\'{o}n General de Asuntos del Personal Acad\'{e}mico (DGAPA) Programa de Apoyo a Proyectos de Investigaci\'{o}n e Innovaci\'{o}n Tecnol\'{o}gica (PAPIIT) IG100920, CONACyT Ciencia de Frontera 74884 for financial support and Miztli-Direcci\'{o}n de C\'{o}mputo y de Tecnolog\'{i}as de Informaci\'{o}n y Comunicaci\'{o}n (DGTIC) - Universidad Nacional Aut\'{o}noma de México (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing-time allocation.
+acknowledges Tecnol\'{o}gico Nacional de M\'{e}xico Proyecto IT16C431, Direcci\'{o}n General de Asuntos del Personal Acad\'{e}mico (DGAPA) Programa de Apoyo a Proyectos de Investigaci\'{o}n e Innovaci\'{o}n Tecnol\'{o}gica (PAPIIT) IG100920, CONACyT Ciencia de Frontera 74884 for financial support and Miztli-Direcci\'{o}n de C\'{o}mputo y de Tecnolog\'{i}as de Informaci\'{o}n y Comunicaci\'{o}n (DGTIC) - Universidad Nacional Aut\'{o}noma de M{\'e}xico (UNAM) (Project LANCAD-UNAM-DGTIC-057) facilities for computing-time allocation.
 

### 2021-04-22 manuscriptPGPEsuppl.tex
Updates from Overleaf

@@
 \usepackage{microtype}
-\usepackage{xr}
-\externaldocument{manuscriptPGPE}
+\usepackage{xr-hyper}
 \usepackage{soul} % for highlights with \hl{} 
-
 \usepackage{hyperref} % to make TOC titles clickable
@@
 
+\makeatletter
+\newcommand*{\addFileDependency}[1]{% argument=file name and extension
+  \typeout{(#1)}
+  \@addtofilelist{#1}
+  \IfFileExists{#1}{}{\typeout{No file #1.}}
+}
+\makeatother
+
+\newcommand*{\myexternaldocument}[1]{%
+    \externaldocument{#1}%
+    \addFileDependency{#1.tex}%
+    \addFileDependency{#1.aux}%
+}
+
+\myexternaldocument{manuscriptPGPE}
+
 \author{\scriptsize Am{\'e}lie Bacle}
@@
 %  \centering
-  \includegraphics[width=\textwidth]{./Figs/POPE_SI.pdf}
+  \includegraphics[width=0.9\textwidth]{./Figs/POPE_SI.pdf}
   \caption{\label{POPEspectra}
@@
     (C) Headgroup and glycerol backbone R-PDLF dipolar slices.
-The arrows indicate the splittings used to determine the C--H bond order parameters by using  $|S_{\rm{CH}}|=\Delta\nu/0.315/d_{\rm{CH}}$.
+The arrows indicate the splittings used to determine the C--H bond order parameters by using  $|S_{\rm{CH}}|=\Delta\nu/(0.315 \times d_{\rm{CH}})$.
 The rigid coupling $d_{\rm{CH}}$ used was 22 kHz.
-    (D) Experimental S-DROSS curves giving signs of the order parameters measured. Grey lines are noise-only profiles.}
+    (D) Experimental S-DROSS curves giving signs of the order parameters measured.
+Grey lines are a set of slices taken from a region of the spectrum without peaks, i.e.
+grey lines are noise profiles, to highlight the signal-to-noise ratio of the measured modulations.}
 \end{figure}
@@
     (C) Headgroup and glycerol backbone R-PDLF dipolar slices.
-The arrows indicate the splittings used to determine the C--H bond order parameters by using  $|S_{\rm{CH}}|=\Delta\nu/0.315/d_{\rm{CH}}$.
+The arrows indicate the splittings used to determine the C--H bond order parameters by using  $|S_{\rm{CH}}|=\Delta\nu/(0.315 \times d_{\rm{CH}})$.
 The rigid coupling $d_{\rm{CH}}$ used was 22 kHz.
     (D) Experimental S-DROSS curves giving signs of the order parameters measured.
-Grey lines are a set of slices taken from a region of the spectrum without peaks, i.e.
-grey lines are noise profiles, to highlight the signal-to-noise ratio of the measured modulations.
 Note that the S-DROSS modulation for $g_3$ is very close to the noise level.
@@
 
-The quality of PE and PG headgroup conformational ensembles in different simulations against NMR experiments is evaluated in figures \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG} using C-H bond order parameters as in our previous studies for PC and PS lipids \cite{botan15,antila19}.
+The quality of PE and PG headgroup conformational ensembles in different simulations are evaluated against NMR experiments in figures \ref{HGorderParametersPE} and \ref{HGorderParametersPOPG} using C-H bond order parameters as in our previous studies for PC and PS lipids \cite{botan15,antila19}.
 Conclusions are the same for all lipids: None of the force fields correctly captures the lipid headgroup conformational ensembles, but CHARMM36 gives results closest to experiments.
@@
 headgroup dipole tilts more parallel to the membrane plane after incorporation
-of negative charges into the membrane~\cite{seelig87, scherer87,antila18}.
+of negative charges into the membrane~\cite{seelig87, scherer87,antila19}.
 The response of PC headgroup order parameters to the addition of PE or PG lipids
@@
 \clearpage
-\section{Dihedral angle distributions and the analysis of relative energies}
-
-\subsection{Dihedral angles and relative energies of PC, PE, PG and PS headgroups}
+\section{Dihedral angle distributions}
+
+\subsection{Dihedral angles of PC, PE, PG and PS headgroups}
 
@@
   \caption{\label{DIHdists}
-    Heavy atom dihedral angle distributions from CHARMM36 simulations that correctly capture the
-    order parameter differences between the force fields.
+    Heavy atom dihedral angle distributions from CHARMM36 simulations that correctly capture the order parameter differences between lipid headgroups.
   }
@@
       		& 144	& 5760  %&0    
-		& 310  & 500          & 400          & \cite{charmm36POPEfiles} \\
+		& 310  & 500          & 400          & \citenum{charmm36POPEfiles} \\
       POPE & CHARMM36 \cite{klauda10}           %& 0      
       		& 500       & 25000 %& 0   
-		&  310  & 500 & 100 & \cite{POPEcharmm} \\
+		&  310  & 500 & 100 & \citenum{POPEcharmm} \\
 %   POPE	& CHARMM36 \cite{klauda10}           %& 0.11
@@
       		& 336	& 15254 %&0    
-		& 310  & 2$\times$200 & 2$\times$100 & \cite{charmm36uaPOPEfiles}  \\
+		& 310  & 2$\times$200 & 2$\times$100 & \citenum{charmm36uaPOPEfiles}  \\
       \hline
@@
       		& 288 	& 9386  %&0    
-		& 336  & 200 & 100 & \cite{slipidsDPPEfiles}  \\
+		& 336  & 200 & 100 & \citenum{slipidsDPPEfiles}  \\
       POPE	& Slipids \cite{jambeck12b} %&0    
       		& 336	& ?     %&0    
-		& 310  & 2$\times$200 &  2$\times$100 & \cite{slipidsPOPEfiles}  \\
+		& 310  & 2$\times$200 &  2$\times$100 & \citenum{slipidsPOPEfiles}  \\
       POPE	& Slipids \cite{jambeck12b}            %& 0
 		& 500 & 25000 %& 0   
-		&  310  & 500 & 100 & \cite{POPEslipids} \\
+		&  310  & 500 & 100 & \citenum{POPEslipids} \\
 %      POPE  & Slipids / {\AA}qvist \cite{jambeck12b,aqvist90}  %& 0.11
@@
           	& 128	& 3655  %&0    
-		& 342  & 2$\times$500 & 2$\times$400 & \cite{gromosCKPdppe} \\
+		& 342  & 2$\times$500 & 2$\times$400 & \citenum{gromosCKPdppe} \\
 %      POPE  & GROMOS-CKP    \cite{??}      %&0
@@
           	& 500	& 25000 %&0    
-		& 310  & 500 & 100 & \cite{gromosCKPpopeT310} \\
+		& 310  & 500 & 100 & \citenum{gromosCKPpopeT310} \\
 %      POPE  & GROMOS-CKP    \cite{??}      &0.11 & 500	& 25000 &50   & 310  & 500 & 100 & \cite{gromosCKPpopeT310150mMNaCl} \\
@@
           	& 128	& 3552     %&0    
-		& 313  & 2$\times$200 & 2$\times$100 & \cite{gromos43a1s3POPEfiles}  \\
+		& 313  & 2$\times$200 & 2$\times$100 & \citenum{gromos43a1s3POPEfiles}  \\
       \hline
@@
       		& 128	& 3328     %&0    
-		& 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaWvdWPOPEfiles} \\
+		& 303  & 2$\times$200 & 2$\times$100 & \citenum{OPLSuaWvdWPOPEfiles} \\
       POPE	& OPLS-UA \cite{??}            %&0
       		& 128	& 3328     %&0    
-		& 303  & 2$\times$200 & 2$\times$100 & \cite{OPLSuaPOPEfiles} \\
+		& 303  & 2$\times$200 & 2$\times$100 & \citenum{OPLSuaPOPEfiles} \\
       \hline
@@
       		& 144	& 5760     %&0    
-		& 310  & 500 & 350 & \cite{MacRogPOPEfiles} \\
+		& 310  & 500 & 350 & \citenum{MacRogPOPEfiles} \\
 %   POPE	& OPLS-MacRog \cite{rog16}     &0    & 128	& 5120     &0    & 300  & 500 & 300 & \cite{MacRogPOPEfilesT300K} \\
@@
       		& 128	& 3552  %&0    
-		& 303  & 2$\times$200 & 2$\times$100 & \cite{bergerPOPEfiles}  \\
+		& 303  & 2$\times$200 & 2$\times$100 & \citenum{bergerPOPEfiles}  \\
       POPE	& Berger-POPE-2018 \cite{??}      %&0    
       		& 128	& 3552  %&0    
-		& 303  & 2$\times$200 & 2$\times$100 & \cite{berger2POPEfiles}  \\
+		& 303  & 2$\times$200 & 2$\times$100 & \citenum{berger2POPEfiles}  \\
 %   DOPE  & Berger-POPE-2004 \cite{??}       &0    & 128	& 4789  &0    & 271  & 2$\times$200 & 2$\times$100 & \cite{bergerDOPEfiles}  \\
@@
       		& 500 & 25000 %& 0  
-		&  310  & 500 & 100 & \cite{POPElipid17} \\
+		&  310  & 500 & 100 & \citenum{POPElipid17} \\
 %      POPE             & Lipid17 \cite{gould18} & 0.11   & 500 & 25000 & 50  &  310  & 500 & 100 & \cite{POPElipid17150mMNaCl} \\
@@
       % some footnotes are not visible in typeset-MS (pdf)
-      lipid/counter-ions & force field for lipids / ions & NaCl (M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
-      \hline
-      POPG/K$^+$  & CHARMM36 \cite{venable13}       &0         & 118& 4110   &0    & 298  & 100 & 100 & \cite{CHARMM36popg}  \\
+      lipid/counter-ions & force field for lipids / ions &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      \hline
+      POPG/K$^+$  & CHARMM36 \cite{venable13}        & 118& 4110    & 298  & 100 & 100 & \citenum{CHARMM36popg}  \\
 %      POPG             & CHARMM36 \cite{venable13}        & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGcharmm150mMNaCl} \\
-      POPG             & CHARMM36 \cite{venable13}        & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGcharmm} \\
-      \hline
-      POPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 10664   &0     & 298  & 250 & 100 & \cite{slipidsPOPGfiles} \\
-      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232  &0     & 314  & 200 & 100 & \cite{slipidsDPPGfiles} \\
-      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    &0         & 288 	& 11232   &0     & 298  & 400 & 100 & \cite{slipidsDPPGfilesT298K} \\
-      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0    & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGslipids} \\
+      POPG             & CHARMM36 \cite{venable13}                & 500 & 25000  &  310  & 500 & 100 & \citenum{POPGcharmm} \\
+      \hline
+      POPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}            & 288 	& 10664   & 298  & 250 & 100 & \citenum{slipidsPOPGfiles} \\
+      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}           & 288 	& 11232   & 314  & 200 & 100 & \citenum{slipidsDPPGfiles} \\
+      DPPG/Na$^+$  & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}           & 288 	& 11232   & 298  & 400 & 100 & \citenum{slipidsDPPGfilesT298K} \\
+      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}      & 500 & 25000  &  310  & 500 & 100 & \citenum{POPGslipids} \\
 %      POPG         & Slipids / {\AA}qvist \cite{jambeck13,aqvist90}    & 0.11    & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGslipids150mMNaCl} \\
       \hline
-      POPG             & LIPID17 / Dang \cite{gould18,smith94,dang06}         & 0              & 500 & 25000 & 0   &  310  & 500 & 100 & \cite{POPGlipid17} \\
+      POPG             & LIPID17 / Dang \cite{gould18,smith94,dang06}     & 500 & 25000  &  310  & 500 & 100 & \citenum{POPGlipid17} \\
 %      POPG             & LIPID17 \cite{gould18}         & 0.11           & 500 & 25000 & 49  &  310  & 500 & 100 & \cite{POPGlipid17150mMNaCl} \\
       \hline
-      POPG             & GROMOS-CKP \cite{??}         & 0              & 500 & 25000 & 0  &  310  & 500 & 100 & \cite{POPGgromosCKP} \\
+      POPG             & GROMOS-CKP \cite{??}         & 500 & 25000 &  310  & 500 & 100 & \citenum{POPGgromosCKP} \\
 %      POPG             & GROMOS-CKP \cite{??}         & 0.11           & 500 & 25000 & 49 &  310  & 500 & 100 & \cite{POPGgromosCKP150mMNaCl} \\
@@
       % some footnotes are not visible in typeset-MS (pdf)
-      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      lipid/counter-ions & force field for lipids / ions & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
       \hline
@@
 %      POPC:POPG (7:3)        & CHARMM36 \cite{??}        & 0.11      & 0  & 350     & ?     & ?   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36NaCl}  \\
-      POPC                   & CHARMM36 \cite{klauda10}        & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
-      POPC:POPG (7:3)        & CHARMM36 \cite{klauda10,venable13}        & 0         & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
-      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
-      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
-      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
-      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}        &0          & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
-      \hline
-      POPC             & CHARMM36 \cite{klauda10}        &0          & 0  & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
-      POPC:POPE (1:1)  & CHARMM36 \cite{klauda10,venable13}        &0          & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
-      \hline
-      POPC             & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPCmacrog300K} \\
-      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0          & 0  & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1macrogT300K} \\
-      \hline
-      POPC             & Slipid \cite{jambeck12b}     &0          & 0  & 512 & 23943 & 0  &  298  & 170 & 100 & \cite{POPCslipid298K} \\
-      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 0  & 128 & 5120  & 0  &  298  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\
-     \hline
-      POPC                   & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCgromosCKPT310K}  \\
-      POPC:POPG (7:3)        & GROMOS-CKP / ?? \cite{??,??}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG3gromosCKPT310K} \\
-     \hline
-      POPC                   & Slipid \cite{jambeck12b}  & 0         & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
-      POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90}  & 0         & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
-      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0         & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC                   & CHARMM36 \cite{klauda10}            & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCcharmmT310K}  \\
+      POPC:POPG (7:3)        & CHARMM36 \cite{klauda10,venable13}    & 0  & 350     & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG1charmm36}  \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}     & 0  & 150:150 & 31500 & 0   &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG5050} \\ 
+      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}     & 0.1 & 150:150 & 31329 & 57 &  298  & 400 & 300 & \cite{CHARMM36POPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & CHARMM36 \cite{klauda10,venable13}     & 1.08 & 150:150 & 29766 & 578  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG50501000mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}    & 0  & 350:88 & 26280 & 0  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}    & 0.1 & 350:88 & 26280 & 47  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG4010100mMCaCl} \\
+      POPC:POPG (4:1)        & CHARMM36 \cite{klauda10,venable13}    & 1.0 & 350:88 & 24927 & 451  &  298  & 500 & 400 & \cite{CHARMM36POPCPOPG40101000mMCaCl} \\
+      \hline
+      POPC             & CHARMM36 \cite{klauda10}        &0          & 256 & 8704 & 0  &  300  & 300 & 250 & \cite{POPCcharmm300K} \\
+      POPC:POPE (1:1)  & CHARMM36 \cite{klauda10,venable13}         & 0  & 128 & 8704 & 0  &  300  & 300 & 250 & \cite{POPC1POPE1charmm36} \\
+      \hline
+      POPC             & OPLS-MacRog \cite{rog16}     &0           & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPCmacrog300K} \\
+      POPC:POPE (1:1)  & OPLS-MacRog \cite{rog16}     &0           & 128 & 5120 & 0  &  300  & 500 & 300 & \cite{POPC1POPE1macrogT300K} \\
+      \hline
+      POPC             & Slipid \cite{jambeck12b}     &0          & 512 & 23943 & 0  &  298  & 170 & 100 & \cite{POPCslipid298K} \\
+      POPC:POPE (1:1)  & Slipid \cite{jambeck12b}     &0          & 128 & 5120  & 0  &  298  & 500 & 300 & \cite{POPC1POPE1slipidT298K} \\
+     \hline
+      POPC                   & GROMOS-CKP / ?? \cite{??,??}  & 0  & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCgromosCKPT310K}  \\
+      POPC:POPG (7:3)        & GROMOS-CKP / ?? \cite{??,??}  & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{POPC7POPG3gromosCKPT310K} \\
+     \hline
+      POPC                   & Slipid \cite{jambeck12b}  & 0      & 500     & 25000 & 0   &  310  & 500 & 100 & \cite{POPCslipid301K}  \\
+      POPC:POPG (7:3)        & Slipid / {\AA}qvist \cite{jambeck12b,aqvist90} & 0  & 350:150 & 25000 & 0   &  310  & 500 & 100 & \cite{slipidPOPC70POPG30T310K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0  & 128:128 & 12800 & 0   &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0.1  & 128:128 & 12800 & 23  &  298  & 500 & 400 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0.2  & 128:128 & 12800 & 46  &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 0.5  & 128:128 & 12800 & 115 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
+      POPC:POPG (1:1)        & Slipid / Dang \cite{jambeck12b,jambeck2012another,smith94,dang06} & 1.0  & 128:128 & 12800 & 230 &  298  & 1500 & 500 & \cite{slipidPOPC50POPG50T298K} \\
     \end{tabular}
@@
       % some footnotes are not visible in typeset-MS (pdf)
-      lipid/counter-ions & force field for lipids / ions & NaCl (M) & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
-      \hline
-      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
-      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
-      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
-      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
-      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 0.1& 150:150 & 31401 & 57 &  298  & 718 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
-      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}        &0          & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
-      \hline
-      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 350:88 & 26265 & 0  &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020} \\
-      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020100mMCaCl} \\
-      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 350:88 & 24840 & 475 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG80201000mMCaCl} \\
-      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
-      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG5050100mMCaCl} \\
-      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0          & 1.0& 150:150 & 29865 & 569 &  298  & 600 & 400 & \cite{Lipid17eccPOPCPOPG50501000mMCaCl} \\
-      \hline
-      POPC             & Berger \cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  &0  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
-      POPC:POPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
+      lipid/counter-ions & force field for lipids / ions & CaCl$_2$\,(M) &  \footnote{Number of lipid molecules with largest mole fraction}N$_{\rm l}$   &  \footnote{Number of water molecules}N$_{\rm w}$   & \footnote{Number of additional cations}N$_{\rm c}$  & \footnote{Simulation temperature}T (K)  & \footnote{Total simulation time}t$_{{\rm sim}}$(ns) & \footnote{Time used for analysis}t$_{{\rm anal}}$ (ns) &   \footnote{Reference for simulation files}files\\
+      \hline
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}       & 0  & 350:88 & 26265 & 0  &  298  & 400 & 350 & \cite{Lipid17POPCPOPG8020} \\
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}    & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 250 & \cite{Lipid17POPCPOPG8020100mMCaCl} \\
+      POPC:POPG (4:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}    & 1.0& 350:88 & 24840 & 475 &  298  & 1200 & 200 & \cite{Lipid17POPCPOPG80201000mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}   & 0  & 150:150 & 31572 & 0  &  298  & 320 & 200 & \cite{Lipid17POPCPOPG5050} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}   & 0.1& 150:150 & 31401 & 57 &  298  & 718 & 198 & \cite{Lipid17POPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17 / Dang \cite{gould18,smith94,dang06}  & 1.0& 150:150 & 29865 & 569 &  298  & 720 & 200 & \cite{Lipid17POPCPOPG50501000mMCaCl} \\
+      \hline
+      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     &0     & 350:88 & 26265 & 0  &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020} \\
+      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 0.1& 350:88 & 26124 & 47 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG8020100mMCaCl} \\
+      POPC:POPG (4:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 1.0& 350:88 & 24840 & 475 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG80201000mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 0  & 150:150 & 31572 & 0  &  298  & 347.8 & 333 & \cite{Lipid17eccPOPCPOPG5050} \\
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 0.1& 150:150 & 29865 & 54 &  298  & 400 & 300 & \cite{Lipid17eccPOPCPOPG5050100mMCaCl} \\
+      POPC:POPG (1:1)        & Lipid17ecc / ECC-ions \cite{pluharova14,kohagen16,martinek18}     & 1.0& 150:150 & 29865 & 569 &  298  & 600 & 400 & \cite{Lipid17eccPOPCPOPG50501000mMCaCl} \\
+      \hline
+      POPC             & Berger-POPC-??
+\cite{??} \todoi{This is probable not plain berger, correct force filed should be described.}  & 0  & 256 & 10240 & 0  &  300  & 300 & 200 & \cite{POPCberger300K} \\
+      POPC:POPE (1:1)  & Berger-POPE-04~\cite{devries04}  & 0  & 128 & 11008 & 0  &  300  & 300 & 200 & \cite{POPC1POPE1berger} \\
       %POPC:DOPE (1:1)  & Berger \cite{??}  \todoi{This is probable not plain berger, correct force filed should be described.} &0          & 0  & 128 & 10240 & 0  &  300  & 300 & 200 & \cite{POPC1DOPE1berger} \\
@@
   \end{minipage}
-  \todo{Citation and description for ''Berger'' model?} \\
+  \todo{Citation and description for Berger-POPC-?? model?} \\
 \end{table*}
@@
 
-\noindent {\it POPE} A lipid bilayer, consisting of a total of 144 POPE molecules, distributed equally between the two leaflets was set using CHARMM-GUI \cite{lee2016charmm}.
+\noindent {\it POPE} A lipid bilayer, consisting of a total of 144 POPE molecules, distributed equally between the two leaflets was set using CHARMM-GUI \cite{lee16}.
 The bilayer was solvated by 5760 water molecules (40 per lipid).
 The random initial configuration and topologies were generated using the CHARMM-GUI web portal, which provides  GROMACS-compatible simulation input files \cite{lee16}.
-CHARMM36 lipid parameters \cite{klauda10} were used for POPE, whereas the CHARMM-specific TIP3P water model \cite{jorgensen1983comparison} was used for water.
+CHARMM36 lipid parameters \cite{klauda10} were used for POPE, whereas the CHARMM-specific TIP3P water model \cite{jorgensen83} was used for water.
 The bilayer was simulated for 500~ns using GROMACS 2018.6 at 310~K.
@@
 The bonds with hydrogen atoms were constrained using p-LINCS \cite{hess07,hess97}.
-
 The simulation files are available at Ref.~\citenum{charmm36POPEfiles}.
@@
 
-\noindent {\it POPC:POPE mixtures} Data is available at \cite{POPCcharmm300K,POPC1POPE1charmm36}.
-A pure POPC system and a 50:50 POPC:POPE mixture were built and equilibrated using CHARMM-GUI [https://doi.org/10.1021/acs.jctc.5b00935].
+\noindent {\it POPC:POPE mixtures} 
+A pure POPC system and a 50:50 POPC:POPE mixture were built and equilibrated using CHARMM-GUI~\cite{lee16}.
 They contained 256 lipids (for the mixture 64 lipids per leaflet for each species) and 34 water molecules per lipid.
@@
 The first 50 ns were discarded for the analysis.
-The simulations were run with the GROMACS 2016.4 [https://doi.org/10.1016/j.softx.2015.06.001] version.
-The v-rescale thermostat [http://dx.doi.org/10.1063/1.2408420] was used with a temperature of 300 K and a time constant of 1 ps; lipids and water were coupled separately to the heat bath.
-Pressure was kept constant at 1 bar using a semi-isotropic Parrinello--Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 5.0 ps.
-A real space cut-off of 1.2 nm was employed for electrostatic interactions while the long-range part was evaluated using the PME method [https://doi.org/10.1063/1.464397, https://doi.org/10.1063/1.470117].
+The simulations were run with the GROMACS 2016.4~\cite{abraham2015gromacs} version.
+The v-rescale thermostat~\cite{bussi07} was used with a temperature of 300 K and a time constant of 1 ps; lipids and water were coupled separately to the heat bath.
+Pressure was kept constant at 1 bar using a semi-isotropic Parrinello--Rahman barostat~\cite{parrinello81} with a time constant of 5.0 ps.
+A real space cut-off of 1.2 nm was employed for electrostatic interactions while the long-range part was evaluated using the PME method~\cite{darden93,essman95}.
 A force-based switching function was used to switch the Lennard-Jones forces to zero over a range of 1--1.2 nm.
-All bonds with hydrogen atoms were constrained with the LINCS algorithm [https://doi.org/10.1002/(SICI)1096-987X(199709)18:12%3C1463::AID-JCC4%3E3.0.CO;2-H].
-Water molecules were kept rigid with the SETTLE algorithm [https://doi.org/10.1002/jcc.540130805].
-The simulation files are available from ref 64 [https://doi.org/10.5281/zenodo.1306800] (pure POPC) and 65 [https://doi.org/10.5281/zenodo.1306821] (POPC:POPE mixture).
+All bonds with hydrogen atoms were constrained with the LINCS algorithm~\cite{hess07,hess97}.
+Water molecules were kept rigid with the SETTLE algorithm~\cite{miyamoto92}.
+The simulation files are available from Ref.~\citenum{POPCcharmm300K} (pure POPC) and~\citenum{POPC1POPE1charmm36} (POPC:POPE mixture).
 
@@
 The bilayer contained a total of 100 POPC and 100 POPE lipids, and it was solvated by 45 water molecules per lipid (for a total of 9000 water molecules).
-
-The Slipids force field \cite{jambeck12,jambeck12b,jambeck13} was used for lipids, and the TIP3P model \cite{jorgensen1983comparison} for water.
+The Slipids force field \cite{jambeck12,jambeck12b,jambeck13} was used for lipids, and the TIP3P model \cite{jorgensen83} for water.
 A 300~ns-long simulation was performed using GROMACS 2019.4 \cite{abraham2015gromacs}.
 The simulation parameters were equal to those used for the POPC/POPG mixture with additional CaCl$_2$ (see below).
-
 The simulation data are available at Ref.~\citenum{POPC1POPE1slipidT298K}.
@@
 
-\noindent {\it POPC:POPE mixtures} Data is available at \cite{POPCberger300K,POPC1POPE1berger}. 
-Two systems were simulated using the Berger force field [https://doi.org/10.1016/S0006-3495(97)78845-3], pure POPC and a mixture 50:50 POPC:POPE.
-For POPE, we additionnally used the de Vries modification implementing a repulsion of the hydrogen atoms located on the amino group of ethanolamine [https://doi.org/10.1021/jp0366926].
+\noindent {\it POPC:POPE mixtures} 
+Two systems were simulated using the Berger force field~\cite{Berger97}, pure POPC and a mixture 50:50 POPC:POPE.
+For POPE, we additionnally used the de Vries modification implementing a repulsion of the hydrogen atoms located on the amino group of ethanolamine~\cite{devries04}.
 Starting from a PDB file of a pure POPC system with Berger atom names, a 50:50 POPC:POPE mixture was built by mutating randomly methyl groups to hydrogens (POPC -> POPE).
@@
 The production simulations were run for 300 ns with a time step of 2 fs.
-The first 100 ns were discarded for the analysis (@Samuli: this sentence may be removed if the analysis was done on a different time window).
-The simulations were run with GROMACS 4.5.3 [https://doi.org/10.1093/bioinformatics/btt055] version.
-The v-rescale thermostat [http://dx.doi.org/10.1063/1.2408420] was used with a temperature of 300 K and a time constant of 0.1 ps; lipids and water were coupled separately to the heat bath.
-Pressure was kept constant at 1 bar using a semi–isotropic Parrinello–Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 4.0 ps.
+The first 100 ns were discarded for the analysis.
+The simulations were run with GROMACS 4.5.3~\cite{pronk13} version.
+The v-rescale thermostat~\cite{bussi07} was used with a temperature of 300 K and a time constant of 0.1 ps; lipids and water were coupled separately to the heat bath.
+Pressure was kept constant at 1 bar using a semi–isotropic Parrinello–Rahman barostat~\cite{parrinello81} with a time constant of 4.0 ps.
 A real space cut-off of 1.0 nm was employed for van der Waals and electrostatic interactions.
-The long-range part of electrostatic interactions was evaluated using the PME method [https://doi.org/10.1063/1.464397, https://doi.org/10.1063/1.470117] with a grid spacing of 0.12 nm and an interpolation order of 4.
-All bonds with hydrogen atoms were constrained with the LINCS algorithm [https://doi.org/10.1002/(SICI)1096-987X(199709)18:12%3C1463::AID-JCC4%3E3.0.CO;2-H].
-Water molecules were kept rigid with the SETTLE algorithm [https://doi.org/10.1002/jcc.540130805].
-The simulation files are available from ref 91 [https://doi.org/10.5281/zenodo.1402417] (pure POPC) and 92 [https://doi.org/10.5281/zenodo.1402449] (POPC:POPE mixture).
+The long-range part of electrostatic interactions was evaluated using the PME method~\cite{darden93,essman95} with a grid spacing of 0.12 nm and an interpolation order of 4.
+All bonds with hydrogen atoms were constrained with the LINCS algorithm~\cite{hess07,hess97}.
+Water molecules were kept rigid with the SETTLE algorithm~\cite{miyamoto92}.
+The simulation files are available from
+Ref.~\citenum{POPCberger300K} (pure POPC) and~\citenum{POPC1POPE1berger} (POPC:POPE mixture).
 
@@
 
-\noindent {\it POPE}
-
-A bilayer patch with a total of 144 POPE molecules, distributed evenly between two leaflets, was created by reordering atoms in a final structure of a POPE simulation performed using the CHARMM36 force field.
+\noindent {\it POPE} A bilayer patch with a total of 144 POPE molecules, distributed evenly between two leaflets, was created by reordering atoms in a final structure of a POPE simulation performed using the CHARMM36 force field.
 The bilayer was hydrated by 40 water molecules per lipid for a total of 5760 water molecules.
-The OPLS-based MacRog force field was used \cite{Kulig15b} for lipids and the TIP3P model \cite{jorgensen1983comparison} for water.
+The OPLS-based MacRog force field was used \cite{Kulig15b} for lipids and the TIP3P model \cite{jorgensen83} for water.
 
@@
 These groups were coupled separately, and a time constat of 0.4~ps was used.
-The Parrinello--Rahman barostat was used to keep the pressures in the membrane plane as well as normal to it constant at 1~bar.
+The Parrinello--Rahman barostat~\cite{parrinello81} was used to keep the pressures in the membrane plane as well as normal to it constant at 1~bar.
 For the barostat, a time constant of 10~ps was used, and the membrane compressibility was set to 4.5$\times$10$^{-5}$~1/bar.
@@
 
-\noindent {\it POPC:POPE mixtures}
-
-Force field details
-Simulations of POPC:POPE mixtures using the OPLS-MacRog force field [https://doi.org/10.1021/jp5016627] were performed. 
-The initial force field parameter files in GROMACS format were taken from ref [https://doi.org/10.1016/j.dib.2016.03.067]. 
+\noindent {\it POPC:POPE mixtures} The initial force field parameter files in GROMACS format were taken from Ref.~\citenum{rog16}. 
 However, a number of errors were detected in the published files so we fixed them in the following way. 
@@
 (in particular the carbonyls of sn-1 and sn-2 as well as the double bond of the oleoyl chain), 
-we thus added back those impropers. The resulting fixed files can be found on 
-Zenodo (https://zenodo.org/record/3741793/files/popc\_fixed.itp, https://zenodo.org/record/3725637/files/pope\_fixed.itp). 
-To check our fix, one simulation of pure POPC was also run and the order parameter compared to previous published results 
-of ref [https://doi.org/10.1021/acs.jpcb.5b04878]. We found very similar values.
+we thus added back those impropers. The resulting fixed files can be found from Refs.~\citenum{POPCmacrog300K} (popc\_fixed.itp) and~\citenum{POPC1POPE1macrogT300K} (pope\_fixed.itp). 
+To check our fix, one simulation of pure POPC was also run and the order parameter compared to previous published results~\cite{botan15}. We found very similar values.
 All the details of the procedure described here are described 
-in https://www.dsimb.inserm.fr/~fuchs/project\_Samuli/POPC\_POPE/report\_results\_comparison.pdf.
-
-The data for pure POPC are available from [https://zenodo.org/record/3741793], for the POPC:POPE mixture from [https://zenodo.org/record/3725637].
-
-Simulation details
-Two systems of pure POPC and 50:50 POPC:POPE mixture were built using CHARMM-GUI [https://doi.org/10.1021/acs.jctc.5b00935].
+in file {\it scratch/report\_results\_comparison.pdf} in Ref.~\citenum{NMRlipidsIVbgit}.
+
+Two systems of pure POPC and 50:50 POPC:POPE mixture were built using CHARMM-GUI~\cite{lee16}.
 They contained 128 lipids and 5120 TIP3 water molecules.
@@
 The production simulations were run for 500 ns with a time step of 2 fs.
-The first 200 ns were discarded for the analysis (@Samuli: this sentence may be removed if the analysis was done on a different time window).
-The simulations were run withthe GROMACS 2018.5 [https://doi.org/10.1016/j.softx.2015.06.001] version.
-The v-rescale thermostat [http://dx.doi.org/10.1063/1.2408420] was used with a temperature of 300 K and a time constant of 0.1 ps; lipids and water were coupled separately to the heat bath.
-Pressure was kept constant at 1 bar using a semi–isotropic Parrinello–Rahman barostat [https://doi.org/10.1103/PhysRevLett.45.1196] with a time constant of 4.0 ps.
-Long-range electrostatic interactions were calculated using the PME method [https://doi.org/10.1063/1.464397, https://doi.org/10.1063/1.470117].
+The first 200 ns were discarded for the analysis.
+The simulations were run with the GROMACS 2018.5~\cite{abraham2015gromacs} version.
+The v-rescale thermostat~\cite{bussi07} was used with a temperature of 300 K and a time constant of 0.1 ps; lipids and water were coupled separately to the heat bath.
+Pressure was kept constant at 1 bar using a semi–isotropic Parrinello–Rahman barostat~\cite{parrinello81} with a time constant of 4.0 ps.
+Long-range electrostatic interactions were calculated using the PME method~\cite{darden93,essman95}.
 A real space cut-off of 1.0 nm was employed with a grid spacing of 0.1 nm in the reciprocal space.
 Lennard-Jones potentials were cut off at 1.0 nm.
-All covalent bonds in lipids were constrained using the LINCS algorithm [https://doi.org/10.1002/(SICI)1096-987X(199709)18:12%3C1463::AID-JCC4%3E3.0.CO;2-H].
-Water molecules were kept rigid with the SETTLE algorithm [https://doi.org/10.1002/jcc.540130805].
+All covalent bonds in lipids were constrained using the LINCS algorithm~\cite{hess07,hess97}.
+Water molecules were kept rigid with the SETTLE algorithm~\cite{miyamoto92}.
 The simulation files are available from ref 66 [https://doi.org/10.5281/zenodo.3741793] (pure POPC) and 67 [https://doi.org/10.5281/zenodo.3725637] (POPC:POPE mixture).
+
+The data for pure POPC are available from Refs.~\citenum{POPCmacrog300K}, and for the POPC:POPE mixture from Ref.~\citenum{POPC1POPE1macrogT300K}.
 
@@
 %% Please write description of your contribution here
-Performed the analysis of dihedrals and isomers of lipids. Analysed lipid structures from Protein Data Bank. Prepared panels for figures 2 and 4. Participated in discussions.
+performed the analysis of dihedrals and isomers of lipids. Analysed lipid structures from Protein Data Bank. Prepared panels for figures 2 and 4. Participated in discussions.
 
@@
 {\it Rebeca Garc{\'i}a Fandi{\~n}o}
-%% Please write description of your contribution here
-
-\noindent
-{\it Fernando Favela-Rosales} set up and performed DPPE, POPG and DPPG simulations with the Slipids Force Field.
+designed and supervised the molecular dynamics simulations carried out by Antonio Pe{\'o}n, and contributed to some discussions.
+%% Please write description of your contribution here
+
+\noindent
+{\it Fernando Favela-Rosales} 
+set up and performed DPPE, POPG and DPPG simulations with the Slipids Force Field.
 %% Please write description of your contribution here
@@
 {\it Tiago M. Ferreira}
-performed and analysed the R-PDLF and S-DROSS experiments. Wrote p
+was responsible for the solid-state NMR experiments/figures and took part in writing the manuscript.
 
@@
 {\it {\'A}ngel Pi{\~n}eiro}
+Created the opAAUA\_prod.py code for the calculation of order parameters. Contributed to some discussions. 
 %% Please write description of your contribution here

### 2021-04-23 manuscriptPGPE.tex
Updates from Overleaf

@@
 
-\title{NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids} %Title of paper
+\title{Conformational plasticity of phospholipid headgroups in simulations and experiments} %Title of paper
 
@@
 \author{Am{\'e}lie Bacle}
-\affiliation{Laboratoire Coop{\'e}ratif "Lipotoxicity and Channelopathies – ConicMeds", Universit{\'e} de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, Franc }
+\affiliation{Laboratoire Coop{\'e}ratif "Lipotoxicity and Channelopathies - ConicMeds", Universit{\'e} de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }
 
@@
 %\url{github.com/NMRlipids/NMRlipidsIVotherHGs}. 
-Simulation details, accessibility information and quality evaluation of more than 70 systems simulated for this work are in the supplementary information.
+Simulation details, accessibility information and quality evaluation of almost 100 combinations of lipid headgroup and force field simulated for this work are in the supplementary information.
 
@@
 
-%Despite the difficulties in simulations to capture the headgroup conformations in mixtures involving charged lipids,
 In conclusion, the structural response of lipid headgroups to 
-%experimentally observed changes in headgroup order parameters upon addition of
 membrane-bound charges arise
-%from relatively small changes in conformational ensembles. These changes eventuate
 from mild changes in dihedral angle distributions, rather than restriction of lipids to
@@
 %
-MJ thanks CSC -- IT Center for Science for computational resources and the Emil Aaltonen foundation for financial support.s
+MJ thanks CSC -- IT Center for Science for computational resources and the Emil Aaltonen foundation for financial support.
     %     Put your acknowledgments here.
@@
 
+O.H.S.O, A.M.K, and S.I.V acknowledge CSC -- IT Center for Science for computational resources and Academy of Finland (315596 and 319902) for financial support.
+
 \end{acknowledgments}

### 2021-04-23 manuscriptPGPEsuppl.tex

@@
 \author{\scriptsize Am{\'e}lie Bacle}
-\affiliation{Laboratoire Coop\'{e}ratif "Lipotoxicity and Channelopathies – ConicMeds", Universit\'{e} de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }
+\affiliation{Laboratoire Coop\'{e}ratif "Lipotoxicity and Channelopathies - ConicMeds", Universit\'{e} de Poitiers, 1 rue Georges Bonnet, 86000 Poitiers, France }
 
@@
 
-\title{ Supporting Information:\\ NMRlipids IV: Headgroup \& glycerol backbone structures, and cation binding in bilayers with PE and PG lipids }
+\title{ Supporting Information:\\Conformational plasticity of phospholipid headgroups in simulations and experiments}
 
@@
 Water molecules were kept rigid with the SETTLE algorithm~\cite{miyamoto92}.
-The simulation files are available from ref 66 [https://doi.org/10.5281/zenodo.3741793] (pure POPC) and 67 [https://doi.org/10.5281/zenodo.3725637] (POPC:POPE mixture).
 
@@
 \noindent
-{\it Am{\'e}lie Bacle} Am{\'e}lie Bacle set up, performed and analysed POPC and POPC:POPE (1:1) simulations with the Berger Force Field.
+{\it Am{\'e}lie Bacle} set up, performed and analysed POPC and POPC:POPE (1:1) simulations with the Berger Force Field.
 %% Please write description of your contribution here

### 2021-04-27 manuscriptPGPEsuppl.tex
Update manuscriptPGPEsuppl.tex

--- 2021-04-19 manuscriptPGPEsuppl.tex
@@
 \author{Antonio Pe{\'o}n}
-\affiliation[]{Spain}
+\affiliation{CIQUP, Centro de Investigação em Qu\'{i}mica, Departamento de Qu\'{i}mica e Bioqu\'{i}mica, Faculdade de Ciências, Universidade do Porto, Porto, Portugal}
 
@@
 Note that the S-DROSS modulation for $g_3$ is very close to the noise level.
-In this case, we have confirmed that the order parameter sign of $g_3$ was negative by performing a measurement with less points in the indirect dimension, more scans and using a higher MAS frequency of 8 kHz (not shown).
+In this case, we have confirmed that the order parameter sign of $g_3$ was negative by performing a measurement with less points in the indirect dimension, more scans and using a higher MAS frequency of 8 kHz.
   }
@@
   %\todo{Which ion model is used in \cite{POPEcharmm150mMNaCl}?} \\
-  \todo{Citation for GROMOS-CKP?} \\
+  \todo{Citation for GROMOS-CKP?  } \\
+  \todo{Citation for GROMOS-CKP POPE AND POPG  http://pubs.acs.org/doi/abs/10.1021/jp207013v  } \\
   \todo{Citation for GROMOS 43A1-S3?} \\
@@
     \todo{Citation and ion model for GROMOS-CKP?} \\
+    \todo{Citation and ion model for GROMOS-CKP POPC https://pubs.acs.org/doi/abs/10.1021/ct3003157 } \\
 \end{table*}
@@
 
-%\noindent {\it POPE with additional NaCl} \todo{Simulation details by A. Peon.}
+%\noindent {\it POPE with additional NaCl} \todo{Simulation details by A.
+Peon.}.
+A pure POPE system was built with CHARMM-GUI Membrane builder [https://doi.org/10.1021/acs.jctc.5b00935].
+A symmetric lipid distribution was used.
+The bilayer contains 500 lipids and 50 CHARMM TIP3P Water model per lipid.
+The simulated system was minimized, equilibrated and production run.
+Production simulation was 500-ns-long, which the last 100 ns were analysed.
+For production simulation, a Leap-Frog algorithm was applied for integrating Newton’s equations of motion, and the LINCS algorithm was used to constrain all covalent bonds, allowing a time step of 2 fs.
+For water, the SETTLE method was applied.
+The simulation was performed in the NPT ensemble where the temperature was kept constant with the V-rescale thermostat (JCP 126, 014101, https://doi.org/10.1063/1.2408420) with water together with eventual ions and lipids coupled to separate thermostats with a coupling constant of 0.1 ps.
+The pressure was kept constant with a semi-isotropic scheme.
+The Parrinello–Rahman barostat was used to keep the pressure at 1 bar with a coupling constant of 1 ps and a compressibility of 4.6 × 10–5 bar–1.
+Long-range electrostatic interactions with a real-space cutoff of 1.0 nm were treated by a particle mesh Ewald scheme (PME) with a Fourier spacing of 0.12 nm and a fourth-order interpolation to the Ewald mesh.
+van der Waals interactions were treated with a Lennard-Jones (LJ) potential with a cutoff of 1.0 nm.
+Long-range dispersion correction to the pressure and energy was added.
 
@@
 
-%\noindent {\it POPG with additional NaCl} \todo{Simulation details by A. Peon.}
+%\noindent {\it POPG with additional NaCl} \todo{Simulation details by A.
+Peon.} A pure POPG system was built with CHARMM-GUI Membrane builder [https://doi.org/10.1021/acs.jctc.5b00935].
+A symmetric lipid distribution was used.
+The bilayer contains 500 lipids and 50 CHARMM TIP3P Water model per lipid.
+The simulated system was minimized, equilibrated and production run.
+The production simulation was 500-ns-long, which the last 100 ns were analysed.
+For production simulation, a Leap-Frog algorithm was applied for integrating Newton’s equations of motion, and the LINCS algorithm was used to constrain all covalent bonds, allowing a time step of 2 fs.
+For water, the SETTLE method was applied.
+The simulation was performed in the NPT ensemble where the temperature was kept constant with the V-rescale thermostat (JCP 126, 014101, https://doi.org/10.1063/1.2408420) with water together with eventual ions and lipids coupled to separate thermostats with a coupling constant of 0.1 ps.
+The pressure was kept constant with a semi-isotropic scheme.
+The Parrinello–Rahman barostat was used to keep the pressure at 1 bar with a coupling constant of 1 ps and a compressibility of 4.6 × 10–5 bar–1.
+Long-range electrostatic interactions with a real-space cutoff of 1.0 nm were treated by a particle mesh Ewald scheme (PME) with a Fourier spacing of 0.12 nm and a fourth-order interpolation to the Ewald mesh.
+van der Waals interactions were treated with a Lennard-Jones (LJ) potential with a cutoff of 1.0 nm.
+Long-range dispersion correction to the pressure and energy was added.
 
@@
 
-\noindent {\it POPC and POPC:POPG (7:3) mixture} \todo{Simulation details by A. Peon.}
+\noindent {\it POPC and POPC:POPG (7:3) mixture} \todo{Simulation details by A.
+Peon.} A pure POPC and POPC:POPG (7:3) mixture were built with CHARMM-GUI Membrane builder [https://doi.org/10.1021/acs.jctc.5b00935].
+Bilayer contains 500 lipids and 50 CHARMM TIP3P Water model per lipid.
+All simulated systems were minimized, equilibrated and production run.
+Production simulations were 500-ns-long, which the last 100 ns were analysed.
+For production simulation, a Leap-Frog algorithm was applied for integrating Newton’s equations of motion, and the LINCS algorithm was used to constrain all covalent bonds, allowing a time step of 2 fs.
+For water, the SETTLE method was applied.
+All simulations were performed in the NPT ensemble where the temperature was kept constant with the V-rescale thermostat (JCP 126, 014101, https://doi.org/10.1063/1.2408420) with water together with eventual ions and lipids coupled to separate thermostats with a coupling constant of 0.1 ps.
+The pressure was kept constant with a semi-isotropic scheme.
+The Parrinello–Rahman barostat was used to keep the pressure at 1 bar with a coupling constant of 1 ps and a compressibility of 4.6 × 10–5 bar–1.
+Long-range electrostatic interactions with a real-space cutoff of 1.0 nm were treated by a particle mesh Ewald scheme (PME) with a Fourier spacing of 0.12 nm and a fourth-order interpolation to the Ewald mesh.
+van der Waals interactions were treated with a Lennard-Jones (LJ) potential with a cutoff of 1.0 nm.
+Long-range dispersion correction to the pressure and energy was added.
 
@@
 \noindent {\it POPE} Data is available at \cite{gromosCKPpopeT310}. \todo{Simulation details by  A. Peon.}
+A pure POPE system was built with CHARMM-GUI Membrane builder [https://doi.org/10.1021/acs.jctc.5b00935].
+A symmetric lipid distribution was used.
+The bilayer contains 500 lipids and 50 SPC Water model per lipid.
+GROMOS-CKP force field was used for lipid and GROMOS54A7 for the rest.
+The simulated system was minimized, equilibrated and production run.
+The production simulation was 500-ns-long, which the last 100 ns were analysed.
+For production simulation, a Leap-Frog algorithm was applied for integrating Newton’s equations of motion, and the LINCS algorithm was used to constrain all covalent bonds, allowing a time step of 2 fs.
+For water, the SETTLE method was applied.
+The simulation was performed in the NPT ensemble where the temperature was kept constant with the V-rescale thermostat (JCP 126, 014101, https://doi.org/10.1063/1.2408420) with water together with eventual ions and lipids coupled to separate thermostats with a coupling constant of 0.1 ps.
+The pressure was kept constant with a semi-isotropic scheme.
+The Parrinello–Rahman barostat was used to keep the pressure at 1 bar with a coupling constant of 1 ps and a compressibility of 4.6 × 10–5 bar–1.
+Long-range electrostatic interactions with a real-space cutoff of 1.0 nm were treated by a particle mesh Ewald scheme (PME) with a Fourier spacing of 0.12 nm and a fourth-order interpolation to the Ewald mesh.
+van der Waals interactions were treated with a Lennard-Jones (LJ) potential with a cutoff of 1.0 nm.
+Long-range dispersion correction to the pressure and energy was added.
 
@@
 \noindent {\it POPG} \todo{Simulation details by A. Peon.}
-
+A pure L-POPG system was built with CHARMM-GUI Membrane builder [https://doi.org/10.1021/acs.jctc.5b00935].
+A symmetric lipid distribution was used.
+The bilayer contains 500 lipids (http://pubs.acs.org/doi/abs/10.1021/jp207013v) and 50 SPC Water model per lipid with 500 sodium ions.
+GROMOS-CKP force field was used for lipid and GROMOS54A7 for the rest.
+The simulated system was minimized, equilibrated and production run.
+The production simulation was 500-ns-long, which the last 100 ns were analysed.
+For production simulation, a Leap-Frog algorithm was applied for integrating Newton’s equations of motion, and the LINCS algorithm was used to constrain all covalent bonds, allowing a time step of 2 fs.
+For water, the SETTLE method was applied.
+The simulation was performed in the NPT ensemble where the temperature was kept constant with the V-rescale thermostat (JCP 126, 014101, https://doi.org/10.1063/1.2408420) with water together with eventual ions and lipids coupled to separate thermostats with a coupling constant of 0.1 ps.
+The pressure was kept constant with a semi-isotropic scheme.
+The Parrinello–Rahman barostat was used to keep the pressure at 1 bar with a coupling constant of 1 ps and a compressibility of 4.6 × 10–5 bar–1.
+Long-range electrostatic interactions with a real-space cutoff of 1.0 nm were treated by a particle mesh Ewald scheme (PME) with a Fourier spacing of 0.12 nm and a fourth-order interpolation to the Ewald mesh.
+van der Waals interactions were treated with a Lennard-Jones (LJ) potential with a cutoff of 1.0 nm.
+Long-range dispersion correction to the pressure and energy was added.
 \noindent {\it POPC:POPG mixture} \todo{Simulation details by A. Peon.}
@@
 \noindent {\it POPE} \todo{Simulation details by A. Peon.}
+From a pure POPE system built with CHARMM-GUI was used to create the Amber input files.
+Amber prmtop and inpcrd files were obtained with LEaP from AmberTools 16 distribution (http://ambermd.org/).
+Then, in order to create the Gromacs input files, ParmEd (v 3.0.0, Swails J.
+ParmEd [Internet].
+https://github.com/ParmEd/ParmEd; https://dx.doi.org/10.1007/s10822-016-9977-1) was used to convert the Amber prmtop and inpcrd files into Gromacs top and gro files.
+The bilayer contains 500 lipids and 50 TIP3P waters per lipid.
+The simulated system was minimized, equilibrated and production run.
+Production simulation was 500-ns-long, which the last 100 ns were analysed.
+For production simulation, a Leap-Frog algorithm was applied for integrating Newton’s equations of motion, and the LINCS algorithm was used to constrain all covalent bonds, allowing a time step of 2 fs.
+For water, the SETTLE method was applied.
+The simulation was performed in the NPT ensemble where the temperature was kept constant with the V-rescale thermostat (JCP 126, 014101, https://doi.org/10.1063/1.2408420) with water together with eventual ions and lipids coupled to separate thermostats with a coupling constant of 0.1 ps.
+The pressure was kept constant with a semi-isotropic scheme.
+The Parrinello–Rahman barostat was used to keep the pressure at 1 bar with a coupling constant of 1 ps and a compressibility of 4.6 × 10–5 bar–1.
+Long-range electrostatic interactions with a real-space cutoff of 1.0 nm were treated by a particle mesh Ewald scheme (PME) with a Fourier spacing of 0.12 nm and a fourth-order interpolation to the Ewald mesh.
+van der Waals interactions were treated with a Lennard-Jones (LJ) potential with a cutoff of 1.0 nm.
+Long-range dispersion correction to the pressure and energy was added.
 
 \noindent {\it POPG} \todo{Simulation details by A. Peon.}
+From a pure POPG system built with CHARMM-GUI was used to create the Amber input files.
+Amber prmtop and inpcrd files were obtained with LEaP from AmberTools 16 distribution (http://ambermd.org/).
+Then, in order to create the Gromacs input files, ParmEd (v 3.0.0, Swails J.
+ParmEd [Internet].
+https://github.com/ParmEd/ParmEd; https://dx.doi.org/10.1007/s10822-016-9977-1) was used to convert the Amber prmtop and inpcrd files into Gromacs top and gro files.
+The bilayer contains 500 lipids and 50 TIP3P waters per lipid and 500 sodium ions.
+The simulated system was minimized, equilibrated and production run.
+Production simulation was 500-ns-long, which the last 100 ns were analysed.
+For production simulation, a Leap-Frog algorithm was applied for integrating Newton’s equations of motion, and the LINCS algorithm was used to constrain all covalent bonds, allowing a time step of 2 fs.
+For water, the SETTLE method was applied.
+The simulation was performed in the NPT ensemble where the temperature was kept constant with the V-rescale thermostat (JCP 126, 014101, https://doi.org/10.1063/1.2408420) with water together with eventual ions and lipids coupled to separate thermostats with a coupling constant of 0.1 ps.
+The pressure was kept constant with a semi-isotropic scheme.
+The Parrinello–Rahman barostat was used to keep the pressure at 1 bar with a coupling constant of 1 ps and a compressibility of 4.6 × 10–5 bar–1.
+Long-range electrostatic interactions with a real-space cutoff of 1.0 nm were treated by a particle mesh Ewald scheme (PME) with a Fourier spacing of 0.12 nm and a fourth-order interpolation to the Ewald mesh.
+van der Waals interactions were treated with a Lennard-Jones (LJ) potential with a cutoff of 1.0 nm.
+Long-range dispersion correction to the pressure and energy was added.
 
@@
 {\it Tiago M. Ferreira}
-performed and analysed the R-PDLF and S-DROSS experiments. Wrote p
+%% Please write description of your contribution here
 
@@
 \noindent
-{\it Antonio Pe{\'o}n}
+{\it Antonio Pe{\'o}n} set up, performed and analysed POPC, POPE, POPG and POPC:POPG (7:3) simulations with the CHARMM36, SLIPIDS, GROMOS-CKP and Lipid17 Force Fields.
 %% Please write description of your contribution here
